### 2016-03-14 writing/tex/sections/coincidence.tex
Add dominating rules, fix code

\section{Coincidences amongst families of mesh patterns and classical patterns}
One interesting question to ask about permutation patterns considers when a
pattern may be avoided by, or contained in, arbitrary permutations. Two
patterns \(\pi\) and \(\sigma\) are said to be \emph{coincident} if the set
of permutations that avoid \(\pi\) is the same as the set of permutations that
avoid \(\sigma\). This extends to sets of patterns as well as single patterns.

We consider the avoidance sets, \(\av{p,q}\) where \(p\) is a classical pattern
of length \(3\) and \(q\) is a mesh pattern of length \(2\) in order to
establish some rules about when these two sets give the same avoidance set.

We first define some operations on mesh patterns.
\begin{definition}
Given a pattern \(p\), let \(\addp[D]{p}{(a,b)}\) be the operation that returns
a mesh pattern equivalent to placing a point in the center of box \((a,b)\) in
\(p\), with shading defined by \(D\subseteq\{N,E,S,W\}\).
\end{definition}
The set \(D\) defines the shading by indicating that the boxes in the cardinal
directions in \(D\) next to the point are shaded in the resulting pattern.
Since there is no ambiguity we let \(\addpe[D]\) be equivalent to \(\addp[D]{\varepsilon}{(0,0)}\).
This operation fails if the box \((a,b)\) is in the mesh set of \(p\).

\begin{example}
    The result of adding a single point to the empty permutation for each cardinal direction.
 \begin{align*}
     \addpe[\{N\}] = \textpattern{}{1}{0/1,1/1} &\qquad\addpe[\{E\}] =\textpattern{}{1}{1/1,1/0} \qquad \\
     \addpe[\{S\}] = \textpattern{}{1}{0/0,1/0} &\qquad\addpe[\{W\}] =\textpattern{}{1}{0/0,0/1} \qquad
\end{align*}
    A more complex example for \texttt{add\_point}
\begin{equation*}
\addp[\{N,E\}]{\pattern{}{2,1,3}{0/3}}{(2,3)} = \pattern{}{2,1,4,3}{0/3,0/4,2/4,3/4,3/3}
\end{equation*}
\end{example}

\begin{definition}
    Given a pattern \(p\), define \(\addd{p}{(a,b)}\), and \(\addi{p}{(a,b)}\),
    as the operations that return a mesh pattern equivalent to placing an
    decrease, or increase, in the center of box \((a,b)\) in \(p\).
\end{definition}

\begin{example}
    \begin{align*}
        \addie = \pattern{}{1,2}{0/1,1/0,1/1,1/2,2/1}\\
        \addde = \pattern{}{2,1}{0/1,1/0,1/1,1/2,2/1}
    \end{align*}
\end{example}

We now attempt to fully classify coincidences in families characterised by avoidance
of a classical pattern of length \(3\) and a mesh pattern of length \(2\).

It can be easily seen that in order to classify set equivalences one need only
consider coincidences within the family of mesh patterns with the same underlying
classical pattern, this is due to the fact that \(\perm{2,1}\in\av{\mperm{1,2}{R}}\)
and \(\perm{1,2}\in\av{\mperm{2,1}{R}}\) for all mesh-sets \(R\). So
\(\av{\mperm{1,2}{R}}\setminus \av{\mperm{2,1}{S}}\neq \emptyset\ \forall
R,S\in [0,2]\times [0,2]\), and hence the sets are disjoint.

We know that there are a total of \(512\) mesh-sets for each underlying classical
pattern. By use of the shading lemma, simultaneous shading lemma, and
one special case, we can reduce the number of equivalence classes to \(220\).

\subsection{Equivalence classes of Av(\{321, (21, \textit{R})\}).}

\begin{proposition}[First Dominating Pattern Rule]
    \label{prop:dom1}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\) such that
    \(\setsize{\pi} \le \setsize{\sigma} + 1\), the sets \(\av{\{\pi,m_1\}}\) and
    \(\av{\{\pi,m_2\}}\) are coincident if

    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item \(\pi \preceq \addp{\sigma}{(a,b)}\)
    \end{enumerate}
\end{proposition}
\begin{proof}
\end{proof}

\subsection{Equivalence classes of Av(\{231, (21, \textit{R})\}).}

\begin{lemma}
    Given a mesh pattern \(m =(\sigma, R)\), where the box \((a,b)\) is not
    in \(R\), and a dominating classical pattern \(\pi = (\pi,\emptyset)\) if
    \(\pi \preceq \addi{\sigma}{(a,b)}\)(\(\pi \preceq \addd{\sigma}{(a,b)}\))
    then in any occurrence of \(m\) in a permutation \(\varrho\) the region
    corresponding to the box \((a,b)\) can only contain an increasing
    (decreasing) subsequence of \(\varrho\).
\end{lemma}
The proof is analogous to the proof of Proposition~\ref{prop:dom1}.

\begin{proposition}[Second Dominating Pattern Rule]
    \label{prop:dom2}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\) such that
    \(\setsize{\pi} \le \setsize{\sigma} + 2\), the sets \(\av{\{\pi,m_1\}}\) and
    \(\av{\{\pi,m_2\}}\) are coincident if

    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item   \begin{enumerate}
                    \item \(\pi \preceq \addi{\sigma}{(a,b)}\) and
                        \begin{enumerate}
                            \item \((a+1,b) \in \sigma\) and \\
                                \((x,b-1)\in R \implies (x,b) \in R \) (where \(x\neq a,a+1\)) and\\
                                  \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b-1,b\)).
                            \item \((a,b+1) \in \sigma\) and \\
                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
                                  \((a-1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
                        \end{enumerate}
                    \item \(\pi \preceq \addd{\sigma}{(a,b)}\) and
                        \begin{enumerate}
                            \item \((a+1,b+1) \in \sigma\) and \\
                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a,a+1\)) and\\
                                  \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
                            \item \((a,b) \in \sigma\) and \\
                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
                                  \((a-1,y)\in R \implies (a,y) \in R\)  (where \(y\neq b-1,b\)).
                        \end{enumerate}
                \end{enumerate}
    \end{enumerate}
\end{proposition}
\begin{proof}
\end{proof}

### 2016-03-14 writing/tex/sections/introduction.tex
Add dominating rules, fix code

\section{Introduction}
\subsection{What is a Permutation?}
In \emph{The Art of Computer Programming} Donald Knuth defines
A \emph{permutation of \(n\) objects} is an arrangement of \(n\) distinct
objects in a row. When considering permutations we can consider them as
occuring on the set \(\nrange{n} = \setrange{1}{n}\), therefore a permutation
is a \emph{bijection} \(\pi:\nrange{n} \mapsto \nrange{n}\). We can write a
permutation \(\pi\) in two line notation
\begin{equation*}
\pi = \twoline{1,2,\dotso,n}{\pi(1),\pi(2),\dotso,\pi(n)}
\end{equation*}
However, the most frequent notation used in computer science is
\emph{one-line notation}, in this form we drop the top line of the two line
notation so are left with the following.
\begin{equation*}
\pi = \perm{\pi(1),\pi(2),\dotso,\pi(n)}
\end{equation*}

\begin{example}
There are 6 permutations on \(\nrange{3}\).
\begin{equation*}
\perm{1,2,3},\perm{1,3,2},\perm{2,1,3},\perm{2,3,1}, \perm{3,1,2}, \perm{3,2,1}
\end{equation*}
\end{example}

We can display a permutation on a \emph{figure} in order to give a graphical
representation of the permutation. In such a figure we let the \(x\)-axis
denote the index in the permutation, and the \(y\)-axis denotes the values of
\(\pi(x)\). The figure of the permutation \(\pi = \perm{2,3,1}\) is shown below
\begin{equation*}
    \pattern{}{2,3,1}{}
\end{equation*}

The class of all permutations of length \(n\) is \(\mathfrak{S}_n\) and
the class has size \(n!\). The class of all permutations is
\(\mathfrak{S}=\bigcup_{i=0}^{\infty}\mathfrak{S}_i\).

\subsection{Classical Permutation Patterns}
Classical Permutation Patterns began to be studied as a result of Knuth's
statements about stack-sorting in [The Art of Computer Programming].

\begin{definition}{Order isomorphism.}
    Two sequences \(\range{\alpha_1,\alpha_2}{\alpha_n}\) and
    \(\range{\beta_1,\beta_2}{\beta_n}\) are said to be \emph{order isomorphic}
    if they share the same relative order,i.e.,\(\alpha_r<\alpha_s\) if and
    only if \(\beta_r<\beta_s\).
\end{definition}

A permutation \(\pi\) us said to \emph{contain} the permutation \(\sigma\) of
length \(k\) as a pattern (denoted \(\sigma \preceq \pi\)) if there is some
an increasing subsequence \(\range{i_1,i_2}{i_n}\) such that the sequence
\(\range{\pi(i_1),\pi(i_2)}{\pi(i_k)}\) is order isomorphic to
\(\range{\sigma(1),\sigma(2)}{\sigma(k)}\). If \(\pi\) does not contain
\(\sigma\), we say that \(\pi\) \emph{avoids} \(\sigma\).

For example the permutation \(\pi = \perm{2,4,1,5,3}\) contains the pattern
\(\sigma = \perm{2,3,1}\), since the second, fourth and fifth entries
(\(4,5,\) and \(3\)) share the same relative order as the entries of \(\sigma\).
This can be seen graphically below, the points order isomorphic to \(\sigma\)
are highlighted.
\begin{equation*}
    \pattern{2,4,5}{2,4,1,5,3}{}
\end{equation*}

We denote the set of permutations of length \(n\) avoiding a pattern \(\sigma\)
as \(\Av_n(\sigma)\) and \(\av{\sigma} = \bigcup_{i=0}^{\infty}\Av_i(\sigma)\).

\subsection{Mesh Patterns}
Mesh Patterns were introduced by Brändén and Claesson to capture explicit expansions
for certain permutation statistics. They are a natural extension of Classical
permutation patterns. A \emph{mesh-pattern} is a pair
\begin{equation*}
    p = (\tau,R)\text{ with } \tau \in \mathfrak{S} \text{ and } R \subseteq
    [0,k]\times [0,k].
\end{equation*}
By this definition the empty permutation \(\varepsilon\) as a mesh pattern consisting solely of the
box \((0,0)\).

The figure for a mesh pattern looks similar to that for a classical pattern with the
addition that we shade the unit square with bottom corner \((i,j)\) for each \((i,j) \in R\):
\begin{equation*}
    \pattern{}{ 2, 1, 3 }{ 0/1, 0/2, 1/0, 1/1, 2/1, 2/2 }
\end{equation*}

We define containment, and avoidance, of the pattern \(p\) in the permutation
\(\pi\) on mesh patterns analogously to classical containment, and avoidance,
of \(\tau\) in \(\pi\) with the additional restrictions on the relative
position of the occurence of \(\tau\) in \(\pi\). These restrictions say that
the shaded regions of the figure above contain no points from \(\pi\).

\begin{example}
    The pattern \(p=\mperm{2,1,3}{\{(0,1),(0,2),(1,0),(1,1),(2,1),(2,2)\}}=
    \textpattern{}{ 2, 1, 3 }{ 0/1, 0/2, 1/0, 1/1, 2/1, 2/2 }\) is contained in
    \(\pi = \perm{3,4,2,1,5}\) but is not contained in \(\sigma = \perm{4,2,3,1,5}\).
\end{example}
\begin{proof}
    Let us consider the figure for the permutation \(\pi\) we only need to find one occurence.
    \begin{equation*}
        \pattern{1,3,5}{3,4,2,1,5}{0/2,0/3,0/4,
                                   1/0,1/1,1/2,
                                   2/0,2/1,2/2,
                                   3/2,3/3,3/4,
                                   4/2,4/3,4/4}
    \end{equation*}
    We have found an occurence of the pattern \(p\) in \(\pi\) and therefore
    \(\pi\) contains \(p\).

    Now we consider the figure for the permutation \(\sigma\). This permutation
    avoids \(p\) since for every occurence of the classical pattern \(\perm{2,1,3}\)
    there is at least one point in one of the shaded boxes. Consider the subsequence
    \(315\) in \(\sigma\), this is an occurence of \(\perm{2,1,3}\) but not the mesh
    pattern since the points with values \(4\) and \(2\) are in the shaded areas.
    This is shown in the figure below.

    \begin{equation*}
        \pattern{3,4,5}{4,2,3,1,5}{0/1,0/2,0/3,0/4,
                                   1/1,1/2,1/3,1/4,
                                   2/1,2/2,2/3,2/4,
                                   3/0,3/1,3/2,
                                   4/1,4/2,4/3,4/4}
    \end{equation*}
    This is true for all occurences of \(\perm{2,1,3}\) in \(\sigma\) and
    therefore \(\sigma\) avoids \(p\).
\end{proof}

We denote the avoidance sets for mesh patterns in the same way as for
classical patterns.

\begin{note}
    \label{not:classmesh}
    Classical patterns are just a subclass of mesh patterns where the mesh
    set \(R\) is empty. The classical pattern \(\pi\) can be represented
    by a mesh pattern as \((\pi,\emptyset)\).
\end{note}

### 2016-03-14 src/helper_funcs.py

@@ -53,13 +53,26 @@ def get_representatives(classes):
 def get_coincidences(representatives, length, perm):
     counters = [MyCounter() for representative in representatives]
 
-    for p in AvoidanceClass(length,perm):
+    for permlength in range(1,length+1):
+        for p in AvoidanceClass(permlength,perm):
+            for i,v in enumerate(representatives):
+                counters[i] + p.avoids(v)
+
+    coinc = [(p1,p2) for p1,p2 in itertools.combinations(enumerate(counters),2) if p1[1] == p2[1]]
+
+    return coinc
+
+def get_all_coincidences(representatives, length):
+    counters = [MyCounter() for representative in representatives]
+
+    for p in Permutations(length):
         for i,v in enumerate(representatives):
             counters[i] + p.avoids(v)
 
     coinc = [(p1,p2) for p1,p2 in itertools.combinations(enumerate(counters),2) if p1[1] == p2[1]]
 
     return coinc
+
 
 def coincidences_to_dictionary(coincidences):
     coincdict = {}

### 2016-03-14 src/main.py

@@ -1,0 +1,57 @@
+from permuta import MeshPattern
+import sys
+from classes import *
+from helper_funcs import *
+from p_rule import *
+
+def stderrwrite(string,end="\n"):
+    sys.stderr.write(string+end)
+    sys.stderr.flush()
+
+classes2 = [map(MeshPattern.reverse, x) for x in classes]
+stderrwrite("Classes created")
+
+p_231 = Permutation([2,3,1])
+p_321 = Permutation([3,2,1])
+
+reprs_12 = get_representatives(classes)
+reprs_21 = get_representatives(classes2)
+stderrwrite("Representatives created")
+
+meshpd_12 = mpdict(classes)
+meshpd_21 = mpdict(classes2)
+
+stderrwrite("Computing Coincidences 1 of 4")
+coincs231_12 = get_coincidences(reprs_12, 8, p_231)
+stderrwrite("Computing Coincidences 2 of 4")
+coincs321_12 = get_coincidences(reprs_12, 8, p_321)
+stderrwrite("Computing Coincidences 3 of 4")
+coincs231_21 = get_coincidences(reprs_21, 8, p_231)
+stderrwrite("Computing Coincidences 4 of 4")
+coincs321_21 = get_coincidences(reprs_21, 8, p_321)
+stderrwrite("Coincidences done")
+
+coincsdict231_12 = coincidences_to_dictionary(coincs231_12)
+coincsdict321_12 = coincidences_to_dictionary(coincs321_12)
+coincsdict231_21 = coincidences_to_dictionary(coincs231_21)
+coincsdict321_21 = coincidences_to_dictionary(coincs321_21)
+
+p_coincs231_12 = p_rule_coincidences(meshpd_12, p_231)
+p_coincs321_12 = p_rule_coincidences(meshpd_12, p_321)
+p_coincs231_21 = p_rule_coincidences(meshpd_21, p_231)
+p_coincs321_21 = p_rule_coincidences(meshpd_21, p_321)
+
+p2_coincs231_12 = p_2_rule_coincidences(meshpd_12, p_231)
+p2_coincs321_12 = p_2_rule_coincidences(meshpd_12, p_321)
+p2_coincs231_21 = p_2_rule_coincidences(meshpd_21, p_231)
+p2_coincs321_21 = p_2_rule_coincidences(meshpd_21, p_321)
+
+unexplained231_12 = funexplained(p_coincs231_12, coincsdict231_12)
[…]

### 2016-03-14 src/p_rule.py

@@ -2,7 +2,7 @@ import itertools
 import sys
 from permuta import *
 
-def p_rule(mp,p=Permutation([2,3,1])):
+def p_rule(mp,p):
     possibilities = []
     for box in itertools.product(range(len(mp)+1),range(len(mp)+1)):
         if not box in mp.mesh:
@@ -11,7 +11,7 @@ def p_rule(mp,p=Permutation([2,3,1])):
                 possibilities.append(mp.shade(box))
     return possibilities
 
-def p_rule_checker(class0,class1,classlist,p=Permutation([2,3,1])):
+def p_rule_checker(class0,class1,classlist,p):
     # we only need one between class p-rule coincidence
     for mp1 in classlist[class0]:
         prule = p_rule(mp1,p)
@@ -24,13 +24,13 @@ def p_rule_checker(class0,class1,classlist,p=Permutation([2,3,1])):
         sys.stderr.flush()
         return None
 
-def p_rule_check_base_class(class0,classlist,coincdict,p=Permutation([2,3,1])):
+def p_rule_check_base_class(class0,classlist,coincdict,p):
     instdict = {}
     for classi in coincdict[class0]:
         instdict[classi] = p_rule_checker(class0,classi,classlist,p)
     return instdict
 
-def p_rule_coincidences(meshpattdict,p=Permutation([2,3,1])):
+def p_rule_coincidences(meshpattdict,p):
     thiscoincdict = {k:[] for k in range(len(set(meshpattdict.values())))}
     mpperm = meshpattdict.keys()[0].perm
     for mp in MeshPatterns(len(meshpattdict.keys()[0].perm)):
@@ -62,43 +62,43 @@ def funexplained(pcoincsdict, coincsdict):
 
     return {sorted(list(k))[0]:(sorted(tuple(k)),map(sorted,map(tuple,v))) for k,v in thisshets.iteritems() if not all(b == k for b in v)}
 
-def p_rule231(mp): # ASSUME 231 for now will fix later, maybe, now works for most basic
-    possibilities = []
-    p = Permutation([2,3,1])
-    for (x,y) in mp.non_pointless_boxes(): # otherwise there are no points to slide
-        if not (x,y) in mp.mesh:
-            thismp = mp.add_increase((x,y))
-            if thismp.perm.contains(p) and (x+1,y-1) not in mp.mesh:
-                # we don't care about the box immediately right of box we're moving points from
[…]

### 2016-03-14 src/start.py

@@ -1,7 +1,0 @@
-import os
-import sys
-from permuta import *
-
-sys.path.append(os.path.abspath("/Users/murraytannock/Documents/Programming/python/patterns/TheShading/the_shading_algorithm/"))
-
-from tsa5_eq import tsa5_coincident

### 2016-04-03 src/main.py
Update src to allow quick run and start of wilf-equivs

@@ -3,6 +3,7 @@ import sys
 from classes import *
 from helper_funcs import *
 from p_rule import *
+from wilf import *
 
 def stderrwrite(string,end="\n"):
     sys.stderr.write(string+end)
@@ -55,3 +56,11 @@ resolved231_12 = resolve_fundict(unexplained231_12, p2_coincs231_12)
 resolved321_12 = resolve_fundict(unexplained321_12, p2_coincs321_12)
 resolved231_21 = resolve_fundict(unexplained231_21, p2_coincs231_21)
 resolved321_21 = resolve_fundict(unexplained321_21, p2_coincs321_21)
+
+# need to prune by symmetries
+#
+reprs231 = [classes[x[0]][0] for x in coincsdict231_12.values()] + [classes[x[0]][0] for x in coincsdict231_21.values()]
+reprs321 = [classes[x[0]][0] for x in coincsdict321_21.values()] #+ [classes[x[0]][0] for x in coincsdict321_12.values()]
+
+wilf231 = [WilfCounter(Permutation([2,3,1]),x ) for x in reprs231]
+wilf321 = [WilfCounter(Permutation([3,2,1]),x ) for x in reprs321]

### 2016-04-03 src/p_rule.py

@@ -7,7 +7,7 @@ def p_rule(mp,p):
     for box in itertools.product(range(len(mp)+1),range(len(mp)+1)):
         if not box in mp.mesh:
             thismp = mp.add_point(box)
-            if thismp.perm == p:
+            if thismp.perm.contains(p):
                 possibilities.append(mp.shade(box))
     return possibilities
 

### 2016-04-03 src/wilf.py

@@ -1,0 +1,26 @@
+from permuta import *
+import pprint
+
+class WilfCounter(object):
+
+    def __init__(self, perm, mperm):
+        self.perm = perm
+        self.mperm= mperm
+        self.record = {}
+
+    def modify(self, checkperm):
+        if checkperm.avoids(self.mperm):
+            checklen = len(checkperm)
+            self.record[checklen] = self.record.get(checklen,0) + 1
+
+    def __repr__(self):
+        return "WilfCounter(%s,%s)" % (self.perm, repr(self.mperm))
+
+    def __str__(self):
+        return pprint.pformat(self.record)
+
+    def __eq__(self, other):
+        if isinstance(other, WilfCounter):
+            return self.record == other.record
+        else:
+            raise TypeError("Cannot compare WilfCounter with %s" % type(other))

### 2016-04-03 writing/tex/sections/coincidence.tex
last writing update, writing in another place, may yet move

@@
         \item \(R_1 \triangle R_2 = \{(a,b)\}\)
-        \item \(\pi \preceq \addp{\sigma}{(a,b)}\)
+        \item \(\pi \preceq \addp{\sigma}{(a,b)}\)\label{prop:dom1:cont}
     \end{enumerate}
 \end{proposition}
+In order to prove this proposition we must first make the following note.
+
+\begin{note}
+    \label{not:downcmesh}
+    Let \(R^\prime \subseteq R\). Then any occurence of \((\tau, R)\) in a permutation
+    is an occurence of \((\tau,R^\prime)\).
+\end{note}
 \begin{proof}
+    Assume without meaningful loss of generality that \(R^\prime\) is a proper
+    subset of \(R\).
+
+    Consider an occurence of \((\tau, R)\) in a permutation \(\sigma\), obviously
+    this corresponds to an occurrence of \(\tau\) in \(\sigma\). Now consider
+    the mesh sets \(R\) and \(R^\prime\), since \(R^\prime \subseteq R\) then
+    there are more restrictions on where points are in an occurrence of \((\tau,R)\).
+    Namely, for every shaded box in \(R\) the corresponding region in \(\sigma\)
+    must contain no points, since \(R^\prime\) has less shading than \(R\) there
+    exists a region in the occurence of \((\tau,R)\) in \(\sigma\) that is
+    now devoid of restrictions. However, by removing restrictions we cannot
+    make an occurence become not an occurence, and therefore the same occurrence
+    of \(\tau \) in \(\sigma\) is now an occurence of \((\tau, R^\prime)\).
+
+\end{proof}
+\begin{proof}[Proof of Proposition~\ref{prop:dom1}]
+
+    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\), or equivalently
+    \(\mathfrak{S}\setminus \av{\{\pi,m_1\}} = \mathfrak{S}\setminus \av{\{\pi,m_1\}}\).
+
+    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
+
+    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\)
+    by Note~\ref{not:downcmesh} any of these occurrences is also an occurence
+    of \(m_1\). This proves that every occurence of \(m_2\) is also an
+    occurence of \(m_1\) and \(\mathfrak{S}\setminus \av{\{\pi,m_2\}}
+    \subseteq \mathfrak{S}\setminus \av{\{\pi,m_1\}}\).
+
+    Now we consider a permutation \(\omega^\prime\) containing an occurence of
+    \(m_1\). Consider placing a point in the region corresponding to the box
+    \((a,b)\), regardless of where in this region we place the point by
+    condition~\ref{prop:dom1:cont} of the Proposition we create an occurrence
+    of \(\pi\), therefore there can be no points in this region, which
+    could have been represented in the mesh set \(R_1\) by adding the box
+    \((a,b)\), and therefore every occurrence of \(m_1\) is in fact an
+    occurrence of \(m_2\).\qedhere
+
 \end{proof}

### 2016-04-04 writing/latex/code.tex
Add writing with RU template

\chapter{Code}\label{cha:code}
You can put code in your document using the listings package, which is
loaded by default in \path{custom.tex}.  Be aware that the listings
package does not put code in your document if you are in draft mode
unless you set the \texttt{forcegraphics} option.

There is an example java (Listing~\ref{src:Data_Bus.java}) and XML
file (Listing~\ref{src:AndroidManifest.xml}).  Thanks to the
\texttt{url} package, you can typeset OSX and unix paths like this:
\path{/afs/rnd.ru.is/project/thesis-template}.  Windows paths:
\path{C:\windows\temp\ }.  You can also typeset them using the menukey
package, but it tends to delete the last separator and has other
complications.\footnote{The menukey package has issues with biblatex,
  read \path{custom.tex} for more information.}

If you are trying to include multiple different languages, you should
go read the documentation and set these up in \path{custom.tex}.  You
will save yourself a lot of effort, especially if you have to fix
anything.

%I have put the source code in the \directory{src/} folder.
\lstinputlisting[language=Java, firstline=1,
lastline=40, caption={Data\_Bus.java: Setting up the class.},
label={src:Data_Bus.java}]{src/Data_Bus.java}

\lstinputlisting[language={[android]XML}, firstline=1, lastline=20,
caption={AndroidManifest.xml: Configuration for the Android UI.},
label={src:AndroidManifest.xml}]{src/AndroidManifest.xml}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "msc-tannock-2016"
%%% End:

### 2016-04-04 writing/latex/coincidence.tex
Add writing with RU template

\chapter{Coincidences amongst families of mesh patterns and classical patterns}
One interesting question to ask about permutation patterns considers when a
pattern may be avoided by, or contained in, arbitrary permutations. Two
patterns \(\pi\) and \(\sigma\) are said to be \emph{coincident} if the set
of permutations that avoid \(\pi\) is the same as the set of permutations that
avoid \(\sigma\). This extends to sets of patterns as well as single patterns.

We consider the avoidance sets, \(\av{p,q}\) where \(p\) is a classical pattern
of length \(3\) and \(q\) is a mesh pattern of length \(2\) in order to
establish some rules about when these two sets give the same avoidance set.
We fix \(p\) in order to define the equivalences, we say that \(p\) is the
\emph{Dominating Pattern}. We fix \(p\) in order to calculate the coincidences.
We never consider occurrences with different \emph{Dominating Patterns}.

We first define some operations on mesh patterns.
\begin{definition}
Given a pattern \(p\), let \(\addp[D]{p}{(a,b)}\) be the operation that returns
a mesh pattern equivalent to placing a point in the center of box \((a,b)\) in
\(p\), with shading defined by \(D\subseteq\{N,E,S,W\}\).
\end{definition}
The set \(D\) defines the shading by indicating that the boxes in the cardinal
directions in \(D\) next to the point are shaded in the resulting pattern.
Since there is no ambiguity we let \(\addpe[D]\) be equivalent to \(\addp[D]{\varepsilon}{(0,0)}\).
This operation fails if the box \((a,b)\) is in the mesh set of \(p\).

\begin{example}
    The result of adding a single point to the empty permutation for each cardinal direction.
 \begin{align*}
     \addpe[\{N\}] = \textpattern{}{1}{0/1,1/1} &\qquad\addpe[\{E\}] =\textpattern{}{1}{1/1,1/0} \qquad \\
     \addpe[\{S\}] = \textpattern{}{1}{0/0,1/0} &\qquad\addpe[\{W\}] =\textpattern{}{1}{0/0,0/1} \qquad
\end{align*}
    A more complex example for \texttt{add\_point}
\begin{equation*}
\addp[\{N,E\}]{\pattern{}{2,1,3}{0/3}}{(2,3)} = \pattern{}{2,1,4,3}{0/3,0/4,2/4,3/4,3/3}
\end{equation*}
\end{example}

\begin{definition}
    Given a pattern \(p\), define \(\addd{p}{(a,b)}\), and \\ \(\addi{p}{(a,b)}\),
    as the operations that return a mesh pattern equivalent to placing an
    decrease, or increase, in the center of box \((a,b)\) in \(p\).
\end{definition}

\begin{example}
    \begin{align*}
        \addie = \pattern{}{1,2}{0/1,1/0,1/1,1/2,2/1}\\
        \addde = \pattern{}{2,1}{0/1,1/0,1/1,1/2,2/1}
    \end{align*}
\end{example}

We now attempt to fully classify coincidences in families characterised by avoidance
of a classical pattern of length \(3\) and a mesh pattern of length \(2\).

It can be easily seen that in order to classify set equivalences one need only
consider coincidences within the family of mesh patterns with the same underlying
classical pattern, this is due to the fact that \(\perm{2,1}\in\av{\mperm{1,2}{R}}\)
and \(\perm{1,2}\in\av{\mperm{2,1}{R}}\) for all mesh-sets \(R\). So
\(\av{\mperm{1,2}{R}}\setminus \av{\mperm{2,1}{S}}\neq \emptyset\ \forall
R,S\in [0,2]\times [0,2]\), and hence the sets are disjoint.

We know that there are a total of \(512\) mesh-sets for each underlying classical
pattern. By use of the shading lemma, simultaneous shading lemma, and
one special case, we can reduce the number of equivalence classes to \(220\).

\section{Equivalence classes of Av(\{321, (21, \textit{R})\}).}
Through expermentation we discover that there are a total of 29 equivalence
classes of mesh patterns with underlying classical pattern \(\perm{2,1}\).

\begin{proposition}[First Dominating Pattern Rule]
    \label{prop:dom1}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\) such that
    \(\setsize{\pi} \le \setsize{\sigma} + 1\), the sets \(\av{\{\pi,m_1\}}\) and
    \(\av{\{\pi,m_2\}}\) are coincident if

    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item \(\pi \preceq \addp{\sigma}{(a,b)}\)\label{prop:dom1:cont}
    \end{enumerate}
\end{proposition}
In order to prove this proposition we must first make the following note.

\begin{note}
    \label{not:downcmesh}
    Let \(R^\prime \subseteq R\). Then any occurence of \((\tau, R)\) in a permutation
    is an occurence of \((\tau,R^\prime)\).
\end{note}
\begin{proof}
    Assume without meaningful loss of generality that \(R^\prime\) is a proper
    subset of \(R\).

    Consider an occurence of \((\tau, R)\) in a permutation \(\sigma\), obviously
    this corresponds to an occurrence of \(\tau\) in \(\sigma\). Now consider
    the mesh sets \(R\) and \(R^\prime\), since \(R^\prime \subseteq R\) then
    there are more restrictions on where points are in an occurrence of \((\tau,R)\).
    Namely, for every shaded box in \(R\) the corresponding region in \(\sigma\)
    must contain no points, since \(R^\prime\) has less shading than \(R\) there
    exists a region in the occurence of \((\tau,R)\) in \(\sigma\) that is
    now devoid of restrictions. However, by removing restrictions we cannot
    make an occurence become not an occurence, and therefore the same occurrence
    of \(\tau \) in \(\sigma\) is now an occurence of \((\tau, R^\prime)\).
\end{proof}
\begin{proof}[Proof of \PropositionRef{prop:dom1}]
    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).

    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).

    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\)
    by \NoteRef{not:downcmesh} any of these occurrences is also an occurence
    of \(m_1\). This proves that every occurence of \(m_2\) is also an
    occurence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).

    Now we consider a permutation \(\omega^\prime\in \av{\pi}\),
    suppose we have an occurrence of \(m_1\). Consider placing a point in the
    region corresponding to the box \((a,b)\), regardless of where in this
    region we place the point by condition~\ref{prop:dom1:cont} of the
    Proposition we create an occurrence of \(\pi\), therefore there can be no
    points in this region, which could have been represented in the mesh set
    \(R_1\) by adding the box \((a,b)\). Hence every occurrence of
    \(m_1\) is in fact an occurrence of \(m_2\), and we have that
    \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).

    Taking both directions of the containment we can therefore draw the
    conclusion that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).
\end{proof}

By using this rule we completely capture the equivalence classes of
\(\av{\{\perm{3,2,1}, \mperm{2,1}{R}\}}\).

This rule is understood very easily by seeing it in graphical form. In the
pattern in \FigureRef{fig:rule1} we can gain shading in three boxes since if
there is a point in any of these boxes we would gain an occurrence of the
dominating pattern \(\perm{3,2,1}\).

\begin{figure}[htb]
    \begin{center}
    \begin{tikzpicture}[scale=\picscale]
        \modpattern{}{2,1}{}
    \end{tikzpicture}
        \raisebox{2ex}{\(\mapsto\)}
    \begin{tikzpicture}[scale=\picscale]
        \filldraw (0.5,2.5) circle (3pt);
        \filldraw (1.5,1.5) circle (3pt);
        \filldraw (2.5,0.5) circle (3pt);
        \modpattern{}{2,1}{0/2,1/1,2/0}
    \end{tikzpicture}

        \caption{Visual Depiction of first dominating pattern rule.}
        \label{fig:rule1}
    \end{center}
\end{figure}

\section{Equivalence classes of Av(\{231, (21, \textit{R})\}).}
By application of \PropositionRef{prop:dom1} we obtain \(43\) equivalence
classes. Experimentation shows that there are in fact \(39\) equivalence
classes, for example the following two patterns are coincident in
\(\av{\perm{2,3,1}}\) but this is not explained by \PropositionRef{prop:dom1}.
\begin{equation*}
    \pattern{}{2,1}{0/0,0/2,1/0,2/1} \text{ and } \pattern{}{2,1}{0/0,0/2,1/0,1/1,2/1}
\end{equation*}

Consider the box \((1,1)\), this is the box that we need to find reasoning
to allow shading. If this box were to be empty then we could shade it freely,
consider what happens if there are points in this box.  If there is any
increase in this box then we would have an occurrence of \(\perm{2,3,1}\),
however, since we are in \(\av{\perm{2,3,1}}\) this is not possible. This box
must contain a decreasing subsequence. This gives rise to the following lemma:

\begin{lemma}
    \label{lem:incdec}
    Given a mesh pattern \(m =(\sigma, R)\), where the box \((a,b)\) is not
    in \(R\), and a dominating classical pattern \(\pi = (\pi,\emptyset)\) if
    \(\pi \preceq \addi{\sigma}{(a,b)}\)\\(\(\pi \preceq \addd{\sigma}{(a,b)}\))
    then in any occurrence of \(m\) in a permutation \(\varrho\) the region
    corresponding to the box \((a,b)\) can only contain an increasing
    (decreasing) subsequence of \(\varrho\).
\end{lemma}
The proof is analogous to the proof of \PropositionRef{prop:dom1}.

Going back to our example mesh patterns
\begin{equation*}
    \begin{tikzpicture}[scale=\picscale]
        \modpattern{}{2,1}{0/0,0/2,1/0,2/1}
        \filldraw (1.66,1.33) circle (3pt);
        \filldraw (1.33,1.66) circle (3pt);
    \end{tikzpicture}
\end{equation*}
We know that the box \((1,1)\) contains an decreasing subsequence. The top point
of this decrease can be chosen to act as the second point in the mesh pattern,
and therefore there are no points between the first point and the new second point.
Hence, we can shade this box as it is guaranteed to be empty. This shows that our
two example patterns are coincident.

This result generalises into the following rule for categorising equivalences
of mesh patterns in cases where there is a dominating classical pattern.

\begin{proposition}[Second Dominating Pattern Rule]
    \label{prop:dom2}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\) such that
    \(\setsize{\pi} \le \setsize{\sigma} + 2\), the sets \(\av{\{\pi,m_1\}}\) and
    \(\av{\{\pi,m_2\}}\) are coincident if

    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item   \begin{enumerate}
                \item\label{prop:dom2:condc} \(\pi \preceq \addi{\sigma}{(a,b)}\) and
                        \begin{enumerate}
                            \item \((a+1,b) \in \sigma\) and \((a+1,b-1)\notin R\) and \\
                                \((x,b-1)\in R \implies (x,b) \in R \) (where \(x\neq a,a+1\)) and\\
                                  \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b-1,b\)).
                            \item \((a,b+1) \in \sigma\) and \((a-1,b+1)\notin R\) and \\
                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
                                  \((a-1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
                        \end{enumerate}
                    \item \(\pi \preceq \addd{\sigma}{(a,b)}\) and
                        \begin{enumerate}
                            \item \((a+1,b+1) \in \sigma\) and \((a+1,b+1)\notin R\) and \\
                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a,a+1\)) and\\
                                  \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
                            \item \((a,b) \in \sigma\) and \((a-1,b-1)\notin R\) and \\
                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
                                  \((a-1,y)\in R \implies (a,y) \in R\)  (where \(y\neq b-1,b\)).
                        \end{enumerate}
                \end{enumerate}
    \end{enumerate}
\end{proposition}
\begin{proof}
    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).

    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).

    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\)
    by \NoteRef{not:downcmesh} any of these occurrences is also an occurence
    of \(m_1\). This proves that every occurence of \(m_2\) is also an
    occurence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).

    We will consider taking the first branch of every choice.
    Now consider a permutation in \(\omega^\prime \in \av{\pi}\). Suppose
    \(\omega^\prime\) contains \(m_1\) consider the region corresponding to
    \((a,b)\) in \(R_1\).

    If the region is empty, then we can freely shade the corresponding
    box \((a,b)\) in \(m_1\) and hence have an occurence of \(m_2\).

    Now consider if the region is non-empty, by \LemmaRef{lem:incdec} and
    condition~\ref{prop:dom2:condc} of the proposition this region must contain
    a decreasing subsequence.
    We can choose the topmost point in the region to replace the corresponding
    point in the mesh pattern and move the points from the subsequence into
    the box southeast of the point. The other conditions allow this to be done
    without moving points into regions that were shaded.Hence there are no
    points in the region corresponding to the box \((a,b)\) in the mesh
    pattern, and therefore we can shade this region. This implies that every
    occurrence of \(m_1\) in \(\av{\pi}\) is in fact an occurrence of \(m_2\)
    so \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).

    Similar arguments satisfy the remainder of the branches.

\end{proof}
\begin{figure}[ht]
\centering
\begin{tikzpicture}[scale = 0.75]

     % Make the grid
     \draw[step=1cm,gray!50] (2/3,2/3) grid (5+1/3,5+1/3);

     % Make black cross in the middle
     \draw[-,very thick] (3,2/3) -- (3,5+1/3);
     \draw[-,very thick] (2/3,3) -- (5+1/3,3);
     \draw[-,very thick,gray] (4,2) -- (2,2) -- (2,4) -- (4,4) -- (4,2);

     \fill[black] (3,3) circle (0.1);
     \fill[red] (2,3) circle (0.1);

     % Draw the dots on the side
     \foreach \x in {1,2,4,5}{
         \draw[dotted, gray!50] (\x, 0) -- (\x, 2/3);
         \draw[dotted, gray!50] (\x, 5+1/3) -- (\x, 6);
         \draw[dotted, gray!50] (0,\x) -- (2/3,\x);
         \draw[dotted, gray!50] (5+1/3,\x) -- (6,\x);
        }
     % Draw black dots next to the cross
     \draw[dotted, black, very thick] (3, 0) -- (3, 2/3);
     \draw[dotted, black, very thick] (3, 5+1/3) -- (3, 6);
     \draw[dotted, black, very thick] (0, 3) -- (2/3, 3);
     \draw[dotted, black, very thick] (5+1/3, 3) -- (6, 3);

     \draw[->-=.5,thick] (3,3) to[bend right=15](2,4);

     \draw[->, thick] (5,3) -- (5,4);
     \draw[->, thick] (1,3) -- (1,4);
     \draw[->, thick] (3,5) -- (2,5);
     \draw[->, thick] (3,1) -- (2,1);

     \node [black,right] at (6,3) {$b$};
     \node [black,below] at (2,0) {$a$};

\end{tikzpicture}
    \caption{An illustration of how lines and points move by this proposition.}
\end{figure}

This proposition essentially states that we slide all of the points in the box we desire
to shade diagonally, and chose the topmost/bottommost point to replace the original
point in the mesh pattern.

By taking the First Dominating Pattern Rule and the Second Dominating Pattern
rule together coincidences of classes of the form \(\av{\{\perm{1,3,2},\mperm{2,1}{R}\}}\)
are fully explained, obtaining 39 equivalence classes of mesh patterns.

\section{Equivalence classes of Av(\{231, (12, \textit{R})\}).}
When considering the equivalence classes of \(\av{\perm{2,3,1},\mperm{1,2}{R}}\)
we first apply the two Dominating Pattern rules previously established.
Starting from \(220\) classes, application of the first Dominating Pattern rule
gives \(85\) classes. Following this with the second Dominating Pattern rule
reduces the number of classes to \(59\). However we know that there are patterns
where the coincidences are not explained by the rules given above.

For example the patterns
\begin{equation*}
    \pattern{}{1,2}{0/2,2/0,2/1} \text{ and } \pattern{}{1,2}{0/2,1/0,2/0,2/1}
\end{equation*}
are experimentally coincident. This coincidence is not explained by our rules,
it is necessary to attempt to capture these coincidences by establishing more
rules.

In order to rigorously establish this coincidence we need to consider what would
happen if we were to choose a point in the box \((1,0)\) that we would like to shade.
In order to get a chance of getting the shading we want consider choosing the
rightmost point in the box.
This gives us the following mesh pattern.
\begin{equation*}
    \pattern{}{2,1,3}{0/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
By application of the \PropositionRef{prop:dom1} we then achieve the following
mesh pattern
\begin{equation*}
    \pattern{2,3}{2,1,3}{0/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
If we look at the highlighted points we see that the subpattern is an occurrence
of the mesh pattern that we originally desired. This gives rise to the following
rule:

\begin{proposition}[Third Dominating Pattern Rule]
    \label{prop:dom3}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\)\footnote{The
    permutation \(\pi\) may be the empty permutation}, the sets
    \(\av{\{\pi,m_1\}}\) and \(\av{\{\pi,m_2\}}\) are coincident if
    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item \(\addp[D]{\mperm{\sigma}{R_1}}{(a,b)}\) is coincident with a
            mesh pattern containing an occurrence of \(\mperm{\sigma}{R_2}\).
    \end{enumerate}
\end{proposition}
\begin{proof}
    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).

    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).

    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\)
    by \NoteRef{not:downcmesh} any of these occurrences is also an occurence
    of \(m_1\). This proves that every occurence of \(m_2\) is also an
    occurence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).

    Now consider a permutation in \(\av{\pi}\)
\end{proof}

\section{Equivalence classes of Av(\{321, (12, \textit{R})\}).}
When considering equivalences of mesh patterns with underlying classical
pattern \(\perm{1,2}\) in \(\av{\perm{3,2,1}}\) application of the previously
established rules give no coincidences. Through experimentation we discover
that there are \(7\) equivalence classes which can be explained through the
use of two different lines of reasoning. Since the number of coincidences is
so small we will reason for these equivalences without attempting to generalise
into concrete rules.

Intuitively it is easy to see why our previous rules have no
power here. There is nowhere that it is possible to add a single point to
gain an occurrence of \(\pi = \perm{3,2,1}\). It is also impossible to have
a position where addition of an increase, or decrease, provides extra
shading power.

The patterns \fxnote{Is this dom3?}
\begin{equation*}
    \pattern{}{1,2}{0/0, 0/2, 1/1, 1/2, 2/1} \text { and } \pattern{}{1,2}{0/0, 0/1, 0/2, 1/1, 1/2, 2/1}
\end{equation*}
are equivalent in \(\av{\perm{3,2,1}}\). (There are 3 symmetries of these
patterns that are also equivalent to each other by the same reasoning.)

Consider the contents of the box \((0,1)\), by \LemmaRef{lem:incdec} it
must contain an increasing subsequence. If there is only one point in the box
we can choose this to replace the \(1\) in the mesh pattern to get the
required shading. If there is more than one point then choosing the
two leftmost points give us the following mesh pattern.
\begin{equation*}
    \pattern{3,4}{2,3,1,4}{0/0, 0/1, 0/2, 0/3, 0/4,
                            1/0, 1/1, 1/2, 1/3, 1/4,
                            2/0, 2/1,2/2, 2/4,
                            3/1, 3/2, 3/3, 3/4,
                            4/1, 4/2, 4/3}
\end{equation*}
Where the two highlighted points are the original two points.
Now if we take the other two points as the points in our
mesh permutation then we get an occurrence of the pattern
we originally desired, and hence the two patterns are coincident.

The other reasoning applies to the patterns
\begin{equation*}
    \pattern{}{1,2}{0/1,0/2,2/0,2/1} \text{ and } \pattern{}{1,2}{0/1,0/2,1/1,2/0,2/1}
\end{equation*}
which are coincident by expermentation.

In order to explain this we need to consider the middle box, and
any points that it may contain. If the box is empty, then the coincidence
is trivially true. Hence, consider if there is a point in the box. This
gives us a mesh pattern
\begin{equation*}
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=0.5, baseline=(current bounding box.center)]
        \modpattern[4]{}{1,2,3}{0/1,0/2,0/3,3/0,3/1,3/2}
        \draw (1.5,3.5) node {\(A\)};
        \draw (2.5,0.5) node {\(B\)};
    \end{tikzpicture}}
\end{equation*}

Where at most one of \(A\) or \(B\) contains an increasing subsequence.
If both are empty we can \emph{a posteriori} make our point in the center
box be either the leftmost, or the rightmost. Now consider if the box \(A\)
were to contain any points, we now know that the box labelled \(B\) contains
no points and can be shaded. By previously established rules we can also obtain
other shadings, leading to the following mesh pattern.
\begin{equation*}
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=0.5, baseline=(current bounding box.center)]
        \modpattern[4]{}{1,2,3}{0/1,0/2,0/3,3/0,3/1,3/2}
        \filldraw (1.5,3.5) circle (4 pt);
        \shadeboxes{2/0,2/1}
    \end{tikzpicture}}
\end{equation*}
Now we can \emph{a posteriori} say that our point in the center box was the
rightmost, giving
\begin{equation*}
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=0.5, baseline=(current bounding box.center)]
        \modpattern[4]{1,2}{1,2,3}{0/1,0/2,0/3,3/0,3/1,3/2}
        \filldraw (1.5,3.5) circle (4 pt);
        \shadeboxes{1/1,1/2,2/0,2/1}
    \end{tikzpicture}}
\end{equation*}
This gives shading that when we choose the highlighted points we
get an occurrence of the required mesh pattern, and as such the
two patterns are coincident.

This argument applies to another pair of classes. Therefore in
total in \(\av{\perm{3,2,1}, \mperm{1,2}{R}}\) there are \(213\) equivalence
classes.

### 2016-04-04 writing/latex/glossary.tex
Add writing with RU template

%% $URL$
%% $Id$
%% Glossary term definitions
%% Note, you will need to reference them in the text after defining them

\newglossaryentry{glossary}{name=glossary, description={a place to put all the terms that confuse people}}
\newglossaryentry{ohm}{name=ohm,symbol={\ensuremath{\Omega}}, description=unit of electrical resistance}  % Do not break this with newlines!
\newglossaryentry{dissertation}{name=dissertation, description={A large pile of paper that professors use to prop up shelves.}}
\newacronym{rfid}{RFID}{Radio Frequency IDentification}

\newacronym{led}{LED}{light-emitting diodes}

\newacronym{auv}{AUV}{Autonomous Undersea Vehicle}
%% Reference manual: http://ctan.uib.no/macros/latex/contrib/glossaries/glossaries-user.pdf

%% You can also define most entries near where they are used, but this may cause problems.  See manual chapter 4.8
%% Be careful about putting entries on Chapter, Section, etc.  See manual Chapter 6.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "msc-tannock-2016"
%%% End:

### 2016-04-04 writing/latex/instructions.tex
Add writing with RU template

\documentclass[projectreport,online,forcegraphics,miktex,draft]{ruthesis}
\usepackage[backend=biber,bibencoding=utf8,style=ieee]{biblatex}
\addbibresource{references.bib}
\InputIfExists{custom.tex}
\setTitle{Writing a Thesis in \LaTeX{}}
%% Because datetime is broken, we always have to set a \whensigned{day}{MLmonth}{year}
\whensigned{12}{\MLdec}{2016}
\author{Joseph Timothy Foley}
\setCourse{Final Project Reports, Thesis, and Dissertations}
\setSchool{\MLSSE{} \& \MLSCS{}}

%% Making glossaries needs perl installed
%% Usually only Windows needs this installed separately
%% http://www.activestate.com/activeperl
\RequirePackage[acronym,toc]{glossaries}
%% nomain: no main glossary
%% acronym: acronym glossary
%% toc: add entry in Table of Contents
%% automake: try to run makeindex using \write18 shell escape
%% without automake, run "makeglossaries" to process them
%% or the right makeindex options
\newglossary[slg]{symbols}{sym}{sbl}{List of Symbols}
\makeglossaries{}  % intitialize the glossary.  Must be in the main file!!!
\input{glossary}  % load glossary definitions in preamble

%\RequirePackage{imakeidx}
%\makeindex[intoc]  % add index to toc

\finalifforcegraphics{hyperref} %hyperlinks even in draft mode
\usepackage[hidelinks]{hyperref}

\begin{document}
\frontmatter{}
\frontcover{}
\enableindents{}
\starttables{} % setup formatting
%% TOC, list of figures and list of tables are required
\tableofcontents{}%%RUM: "Table of contents"
\thesislistoffigures{}%%RUM: "List of figures"
\thesislistoftables{}%%RUM: "List of tables"
\starttext{}
%\coverchapter{List of drawings and enclosed material}
%RUM: "List of drawings and enclosed material, e.g. CD(as appropriate)"

\listoffixmes{}
% if using fixme package, lists what needs to be done

\chapter{Getting started}
\section{Introduction}
Before using the template in \path{DEGREE-NAME-YEAR.tex}, the author
recommends reading the (in progress) guide to successful projects,
including \LaTeX{} at
\url{http://afs.rnd.ru.is/project/htgaru/trunk/how-to-get-around-projects.pdf}.
Full instructions on setting up \LaTeX{} and SubVersion are there.
Additionally, as issues are discovered, they are added to the
Frequently Asked Questions section in that document.

Text will be double spaced when in draft mode so that there is room
for reviewers to write comments.  Extra blank pages are also removed.
If you want to check formatting in a more final form, switch to
proofing mode.  This text in gray will disappear when you are not in
draft mode.  It is better to use FiXme's as a reminder mechanism,
though.  See Section~\ref{sec:fixme}

\textbf{WARNING} It is very very important that you print the generated PDF at 100\% scaling.   
On many computers, printing is set to adjust the size to fit the page.
Do not do this, because it will mess up the sizes of the font and the placement of the letters.
This becomes very evident when you print it out and the cover window does not line up with your title!

\section{Citing properly}
Any content from external sources must be properly cited.
If any part of this document is plagiarized (from another group,
the internet, or anywhere else) or your references are not properly
cited, you will be in a lot of trouble.

This is in compliance with Reykjavík University's Code of
Ethics\footnote{\url{http://en.ru.is/the-university/ru-code-of-ethics/}
  -- see Item 8 and General Rules on Study and Examinations}
\footnote{\url{http://en.ru.is/studies/study--and-examination-rules/}
  -- see Article 4.5} \footnote{see also the document "Your Work -
  Your Contribution" \url{http://www.ru.is/hugverk}, available only in
  Icelandic}. For your reference, here is a link to guidelines
defining plagiarism and discussing how to avoid it (from the
University of Oxford):

\url{http://www.ox.ac.uk/students/academic/goodpractice/about/}

This template uses BibLaTeX and Biber to produce the Bibliography.
This is because they have better support for accented and Icelandic
characters in UTF-8.  They also do more checking on the bibliography
to look for problems.  Unfortunately, this means that if you have a
semi-broken references.bib, Biber will refuse to produce a
Bibliography.  This is a feature because it is telling you that you
need to fix the problems.  BibTeX will also give errors, but will
continue to generate a Bibliography.  This can cause problems later.
For more information read
\url{http://afs.rnd.ru.is/project/htgaru/trunk/how-to-get-around-projects.pdf}.

\section{Document Structure}
The default document structure is the style called IMRAD:\@ Introduction, Methods, Resarch, And, Discussion.  
This is also known as the American Psychological Association (APA) format and is commonly used in the sciences.\cite{web:uta.fi:imrad}.
The template's current editor believes that this format is a good starting point, but may not be appropriate for some engineering works.  
In those cases, following a more technical report style may be easier: Introduction, Background, Design, Implementation, Analysis,  Conclusion.
For some works, this is still too confining.
In those cases, the editor recommends the Introduction, Body, Conclusion (IBC) format where the body is a series of IMRAD or other structures.
Each section of the body should ask a question and answer it relating to the topic.
If you are writing a thesis that is primarily a collection of papers, then the body section would consist of these papers.

\subsection{Reports for classes}
If you are writing a report for a class using this template, you probably want to get rid of a lot of the extra pages.
The set used in these instructions are a good place to start.
See Listing~\ref{src:minimal-document} for a quick example.
\lstinputlisting[language={[LaTeX]TeX}, caption={Minimal document example},label={src:minimal-document}]{minimal-document.tex}

\chapter{\LaTeX{} techniques}
\section{Fitting things on the page\label{sec:fitting}}
If you see little black boxes on the right in draft mode, this is
\LaTeX{} trying to tell you that it couldn't make it fit nicely.

\section{FixMe\label{sec:fixme} reminders}
\fxnote{This is a note.  It will disappear in final mode, which makes it a safe place to put reminders.}
\fxwarning{This is someting more important that needs to be corrected, but will also disappear in final mode.}
\fxerror{This is something very important to be fixed.  This note will disappear in final mode.}
%\fxfatal{This is a fatal issue that must be fixed before submission (final).  
If any of these still exist except for note when we are not in draft mode, \LaTeX{} will have errors during processing.
If the Fatal error is still there, \LaTeX{} will refuse to produce the document\footnote{This is a {\em feature\/} so that you don't submit the document without fixing really important issues.}.
The Fatal error has been commented out in this document so that we can successfully generate documents but it can be seen in the document's source.

\subsection{Indexes, glossaries, and abbreviations}
This section demonstrates indexes and glossaries.\index{glossary}
Note that you may get errors in this section if you turn the index and glossary options off.
In order to generate the glossary, you will need to run \verb|makeglossaries|, which requires the Perl language to be installed.  On windows, you can get a perl interpreter at \url{http://www.activestate.com/activeperl}
See the \verb|README.TXT| for details.

You might expect to find the term \gls{led} in a \gls{glossary}.
For maximum safety, the glossary and acronyms should be declared using newglossaryentry before \verb|\begin{document}| which is done automatically if you put them into \verb|glossary.tex|.
This is {\em highly recommended\/} in case you reorder your text.
The author did his \gls{dissertation} on \gls{rfid} security.  
You can even have symbols, such as the one for \gls{ohm} which is \glssymbol{ohm}

\subsection{Examples}
Some examples of how to use features of \LaTeX{} follow.

\subsubsection{Equations and references}
\begin{figure}
\centering
\includegraphics[height=20mm]{ru-logo}
\caption{The logo of Reykjavík University}\label{fig:ru-logo}
\end{figure}
Figures and Tables need to be properly formatted and referenced in the
text.  Number figures/tables consecutively, include captions, and
refer to the figures/tables in the text
e.g. Figure~\ref{fig:ru-logo}. Equations need to be numbered
consecutively as well.  Equations need to have each of their variables
defined when they are first used or redefined.  If you need to refer
to particular places in the document, use numbered references.  Note
that the \verb|\ref{BLAH}| and the \verb|\label{BLAH}| must match.
When using \verb|\label{}|, it has to be AFTER the item you are trying
to point to without spaces or inside of its declaration.  It is often
safest to put inside e.g.:
\begin{verbatim}\caption{This is a caption\label{fig:mycaption}}\end{verbatim}

As an example of an equation, in Equation~\ref{eq:freq} is the
relationship between angular frequency and hertz:
\begin{equation}
f = \frac{1}{T} = \frac{1}{2\pi\omega}\label{eq:freq}
\end{equation}
where $f$ is frequency in \si{\hertz}, $T$ is period in \si{\second},
and $\omega$ is angular frequency in \si{\radian\per\second}.
Resistance is in units of \si{\ohm}.  This example uses \verb|\si|
from the siunitx package to format the units correctly.  If you want
to format it with a value, it uses \verb|\SI{value}{unit}| e.g.:
\verb|\SI{2.99E8}{\meter\per\second}| becomes
\SI{2.99E8}{\meter\per\second}

As an example of referring to a particular part of the document, see
Section~\ref{sec:fitting} on page~\pageref{sec:fitting} for information
on how to fit text on a page.

\subsubsection{Chemical Formulas}
Chemical formulas are best formatted by the mhchem package (included
in the custom.tex by default).
\href{http://mirrors.ctan.org/macros/latex/contrib/mhchem/mhchem.pdf}{Click
  here for the documentation}.  This example is disabled by default
because it breaks due to a change in \LaTeX3 that hasn't propogated to
all distributions.

Here is an example in reaction~\ref{react:so4}:

\reaction[react:so4]{SO4^2- + Ba^2+ -> BaSO4 v}

\subsubsection{Tables}
When making tables, there are some very important rules for making
them look professional.  See
\url{http://ctan.uib.no/macros/latex/contrib/booktabs/booktabs.pdf}
For an example of a simple table, see Table~\ref{tab:example}.  Some general
rules are to always avoid \verb|\hrule| and vertical separator lines.

\begin{table}\centering
\caption{Listing of real name to username\label{tab:example}}
\begin{tabular}{lll}\toprule
Firstname &Lastname &username\\\midrule
Joseph &Foley &foley\\\bottomrule
\end{tabular}
\end{table}

\subsection{Mutilingual support}
In the main file, you have already seen the ML macros used to create multi-lingual documents.
You may want to read more about the babel and polyglossia packages to do even more.
Quotation marks can be made smart by using \enquote{encite}.
This will change the quotation marks to follow whatever the current language is set to.
The Context Sensite Quotes (csquote) package is full of useful macros of this sort. 
One caveat: don't use the shorcut "` and '" in the macros nor in the text\footnote{This is the shortcute for German style quotations, which is the same as Icelandic.}.
It doesn't work due to interactions with the multilanguage macros.
If you need to use those quotes, you can switch languages temporarily,  \selectlanguage{icelandic}
"`Like this'" \resetlanguage{}

\subsection{URLs and paths}
URLs are easy to use in both the bibliography and text.  In your
\path{.bib} file, make sure that you only put URLs into the \path{url}
field.  Putting it anywhere else is likely to cause it to go off the
page.  If you are using the \path{hyperref} package, then these will
be clickable.  You can put links in the text with
\verb|\url| like this: \url{http://www.ru.is}.  If you need to
specify a location of a file, use the
\verb|\path| similarly: \path{C:\Documents and Settings\ }.  The only
difference is that paths are not clickable.

\subsection{Hints}
When placing tables and figures, {\em do not\/} force \LaTeX{} to
place them in a given place.  Doing so almost always causes problems
later.  It is better to give hints \verb|[htb]| than to
force it \verb|!h]|.

Add linebreaks at least at the end of every sentence.  This is
especially important if you are using SVN or collaborating with
others.  Everything is done by line number, so if an error or
issue is on line 5 which happens to be a page of text, good luck
finding the problem.

Along those same lines, if you are not storing your thesis in version
control (SVN, git, etc) or on a backed up drive, you are asking for
trouble.
\subsection{Debugging}
When things go wrong, it is sometimes hard to figure out what to do.
The first step is always to go look at the log file.  If you are
making changes and they are not showing up, it is likely that you are
viewing an old file.  Delete the \path{.pdf} and \path{.aux} and try
again.  On some platforms, the package manager can get out of date.
Make sure you have updated/synchronized before running \path{pdflatex} again.

%% ---------------------------------------------------------------
\printbibliography{} %%RUM: "References"

%% If appendices are needed, uncomment the following line
%% and include the appendices in separate files
\appendix{}%%RUM: "Appendicies (as appropriate)
\input{code} % as an example, perhaps some of your code

%\backmatter{} % Sections after this don't get numbers
%% We prefer that all elements be numbered

%%%%%%%%%%%%% SHOW GLOSSARY %%%%%%%%%%%%%%%%%
%% Glossary, optional.  A good idea on longer documents
%% Remember to run "makeglossaries <DEGREE-NAME-YEAR>"
%%   to update the glossary
\printglossary{}%%RUM: Not mentioned
\printacronyms{}%%RUM: Not mentioned

%% Glossary debugging code
%% Uncomment this to figure out of one of the glossary entries is broken
%% by putting all of the entries in the glossary.
%\ifthenelse{\boolean{debug}}{%
%  \glsaddall{} %put all entries in
%  \printglossaries{}}{}% dump them all
%% This command dumps all of the glossaries

%%%%%%%%%%%%% SHOW INDEX %%%%%%%%%%%%%%%%%%
%% Index, optional.  A good idea on longer documents

% You can put instructions at the beginning of the index:
%\renewcommand{\preindexhook}{%
%  The first page number is usually, but not always,
%  the primary reference to the indexed topic.\vskip\onelineskip}

%% You may have to run "makeindex" to have it be generated
%% Depending upon which package you chose.
%% 
\clearforchapter{}
\printindex{}%%RUM: Not mentioned

\backcover{}%%RUM: "Back cover (standard)
\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2016-04-04 writing/latex/introduction.tex
Add writing with RU template

--- 2016-03-14 writing/tex/sections/introduction.tex
@@
-\section{Introduction}
-\subsection{What is a Permutation?}
+\chapter{Introduction\label{cha:introduction}}
+\section{What is a Permutation?}
 In \emph{The Art of Computer Programming} Donald Knuth defines
@@
 
-\subsection{Classical Permutation Patterns}
+\section{Classical Permutation Patterns}
 Classical Permutation Patterns began to be studied as a result of Knuth's
@@
 
-\subsection{Mesh Patterns}
+\section{Mesh Patterns}
 Mesh Patterns were introduced by Brändén and Claesson to capture explicit expansions
@@
 \begin{equation*}
-    p = (\tau,R)\text{ with } \tau \in \mathfrak{S} \text{ and } R \subseteq
+    p = (\tau,R)\text{ with } \tau \in \mathfrak{S}_k \text{ and } R \subseteq
     [0,k]\times [0,k].
@@
 \end{note}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "msc-tannock-2016"
+%%% End:

### 2016-04-04 writing/latex/minimal-document.tex
Add writing with RU template

--- 2016-04-04 writing/latex/instructions.tex
@@
 \InputIfExists{custom.tex}
-\setTitle{Writing a Thesis in \LaTeX{}}
-%% Because datetime is broken, we always have to set a \whensigned{day}{MLmonth}{year}
-\whensigned{12}{\MLdec}{2016}
-\author{Joseph Timothy Foley}
-\setCourse{Final Project Reports, Thesis, and Dissertations}
-\setSchool{\MLSSE{} \& \MLSCS{}}
-
-%% Making glossaries needs perl installed
-%% Usually only Windows needs this installed separately
-%% http://www.activestate.com/activeperl
-\RequirePackage[acronym,toc]{glossaries}
-%% nomain: no main glossary
-%% acronym: acronym glossary
-%% toc: add entry in Table of Contents
-%% automake: try to run makeindex using \write18 shell escape
-%% without automake, run "makeglossaries" to process them
-%% or the right makeindex options
-\newglossary[slg]{symbols}{sym}{sbl}{List of Symbols}
-\makeglossaries{}  % intitialize the glossary.  Must be in the main file!!!
-\input{glossary}  % load glossary definitions in preamble
-
-%\RequirePackage{imakeidx}
-%\makeindex[intoc]  % add index to toc
-
+\setTitle{My Title}
+\setDocumentType{Final Report}
+\author{Firstname1~Lastname1, Firstname1~Lastname2}
+\setCourse{T-411-MECH}
+\setSchool{\MLSSE{}}
+\whensigned{17}{\MLdec}{2015}
 \finalifforcegraphics{hyperref} %hyperlinks even in draft mode
 \usepackage[hidelinks]{hyperref} 
-
 \begin{document}
@@
 \starttables{} % setup formatting
-%% TOC, list of figures and list of tables are required
-\tableofcontents{}%%RUM: "Table of contents"
-\thesislistoffigures{}%%RUM: "List of figures"
-\thesislistoftables{}%%RUM: "List of tables"
+\tableofcontents{}
+\thesislistoffigures{}
+\thesislistoftables{}
 \starttext{}
-%\coverchapter{List of drawings and enclosed material}
-%RUM: "List of drawings and enclosed material, e.g. CD(as appropriate)"
-
 \listoffixmes{}
-% if using fixme package, lists what needs to be done
-
-\chapter{Getting started}
-\section{Introduction}
-Before using the template in \path{DEGREE-NAME-YEAR.tex}, the author
-recommends reading the (in progress) guide to successful projects,
-including \LaTeX{} at
-\url{http://afs.rnd.ru.is/project/htgaru/trunk/how-to-get-around-projects.pdf}.
-Full instructions on setting up \LaTeX{} and SubVersion are there.
-Additionally, as issues are discovered, they are added to the
-Frequently Asked Questions section in that document.
-
-Text will be double spaced when in draft mode so that there is room
-for reviewers to write comments.  Extra blank pages are also removed.
-If you want to check formatting in a more final form, switch to
-proofing mode.  This text in gray will disappear when you are not in
-draft mode.  It is better to use FiXme's as a reminder mechanism,
-though.  See Section~\ref{sec:fixme}
-
-\textbf{WARNING} It is very very important that you print the generated PDF at 100\% scaling.   
-On many computers, printing is set to adjust the size to fit the page.
-Do not do this, because it will mess up the sizes of the font and the placement of the letters.
-This becomes very evident when you print it out and the cover window does not line up with your title!
-
-\section{Citing properly}
-Any content from external sources must be properly cited.
-If any part of this document is plagiarized (from another group,
-the internet, or anywhere else) or your references are not properly
-cited, you will be in a lot of trouble.
-
-This is in compliance with Reykjavík University's Code of
-Ethics\footnote{\url{http://en.ru.is/the-university/ru-code-of-ethics/}
-  -- see Item 8 and General Rules on Study and Examinations}
-\footnote{\url{http://en.ru.is/studies/study--and-examination-rules/}
-  -- see Article 4.5} \footnote{see also the document "Your Work -
-  Your Contribution" \url{http://www.ru.is/hugverk}, available only in
-  Icelandic}. For your reference, here is a link to guidelines
-defining plagiarism and discussing how to avoid it (from the
-University of Oxford):
-
-\url{http://www.ox.ac.uk/students/academic/goodpractice/about/}
-
-This template uses BibLaTeX and Biber to produce the Bibliography.
-This is because they have better support for accented and Icelandic
-characters in UTF-8.  They also do more checking on the bibliography
-to look for problems.  Unfortunately, this means that if you have a
-semi-broken references.bib, Biber will refuse to produce a
-Bibliography.  This is a feature because it is telling you that you
-need to fix the problems.  BibTeX will also give errors, but will
-continue to generate a Bibliography.  This can cause problems later.
-For more information read
-\url{http://afs.rnd.ru.is/project/htgaru/trunk/how-to-get-around-projects.pdf}.
-
-\section{Document Structure}
-The default document structure is the style called IMRAD:\@ Introduction, Methods, Resarch, And, Discussion.  
-This is also known as the American Psychological Association (APA) format and is commonly used in the sciences.\cite{web:uta.fi:imrad}.
-The template's current editor believes that this format is a good starting point, but may not be appropriate for some engineering works.  
-In those cases, following a more technical report style may be easier: Introduction, Background, Design, Implementation, Analysis,  Conclusion.
-For some works, this is still too confining.
-In those cases, the editor recommends the Introduction, Body, Conclusion (IBC) format where the body is a series of IMRAD or other structures.
-Each section of the body should ask a question and answer it relating to the topic.
-If you are writing a thesis that is primarily a collection of papers, then the body section would consist of these papers.
-
-\subsection{Reports for classes}
-If you are writing a report for a class using this template, you probably want to get rid of a lot of the extra pages.
-The set used in these instructions are a good place to start.
-See Listing~\ref{src:minimal-document} for a quick example.
-\lstinputlisting[language={[LaTeX]TeX}, caption={Minimal document example},label={src:minimal-document}]{minimal-document.tex}
-
-\chapter{\LaTeX{} techniques}
-\section{Fitting things on the page\label{sec:fitting}}
-If you see little black boxes on the right in draft mode, this is
-\LaTeX{} trying to tell you that it couldn't make it fit nicely.
-
-\section{FixMe\label{sec:fixme} reminders}
-\fxnote{This is a note.  It will disappear in final mode, which makes it a safe place to put reminders.}
-\fxwarning{This is someting more important that needs to be corrected, but will also disappear in final mode.}
-\fxerror{This is something very important to be fixed.  This note will disappear in final mode.}
-%\fxfatal{This is a fatal issue that must be fixed before submission (final).  
-If any of these still exist except for note when we are not in draft mode, \LaTeX{} will have errors during processing.
-If the Fatal error is still there, \LaTeX{} will refuse to produce the document\footnote{This is a {\em feature\/} so that you don't submit the document without fixing really important issues.}.
-The Fatal error has been commented out in this document so that we can successfully generate documents but it can be seen in the document's source.
-
-\subsection{Indexes, glossaries, and abbreviations}
-This section demonstrates indexes and glossaries.\index{glossary}
-Note that you may get errors in this section if you turn the index and glossary options off.
-In order to generate the glossary, you will need to run \verb|makeglossaries|, which requires the Perl language to be installed.
-On windows, you can get a perl interpreter at \url{http://www.activestate.com/activeperl}
-See the \verb|README.TXT| for details.
-
-You might expect to find the term \gls{led} in a \gls{glossary}.
-For maximum safety, the glossary and acronyms should be declared using newglossaryentry before \verb|\begin{document}| which is done automatically if you put them into \verb|glossary.tex|.
-This is {\em highly recommended\/} in case you reorder your text.
-The author did his \gls{dissertation} on \gls{rfid} security.  
-You can even have symbols, such as the one for \gls{ohm} which is \glssymbol{ohm}
-
-\subsection{Examples}
-Some examples of how to use features of \LaTeX{} follow.
-
-\subsubsection{Equations and references}
-\begin{figure}
-\centering
-\includegraphics[height=20mm]{ru-logo}
-\caption{The logo of Reykjavík University}\label{fig:ru-logo}
-\end{figure}
-Figures and Tables need to be properly formatted and referenced in the
-text.  Number figures/tables consecutively, include captions, and
-refer to the figures/tables in the text
-e.g. Figure~\ref{fig:ru-logo}. Equations need to be numbered
-consecutively as well.  Equations need to have each of their variables
-defined when they are first used or redefined.  If you need to refer
-to particular places in the document, use numbered references.  Note
-that the \verb|\ref{BLAH}| and the \verb|\label{BLAH}| must match.
-When using \verb|\label{}|, it has to be AFTER the item you are trying
-to point to without spaces or inside of its declaration.  It is often
-safest to put inside e.g.:
-\begin{verbatim}\caption{This is a caption\label{fig:mycaption}}\end{verbatim}
-
-As an example of an equation, in Equation~\ref{eq:freq} is the
-relationship between angular frequency and hertz:
-\begin{equation}
-f = \frac{1}{T} = \frac{1}{2\pi\omega}\label{eq:freq}
-\end{equation}
-where $f$ is frequency in \si{\hertz}, $T$ is period in \si{\second},
-and $\omega$ is angular frequency in \si{\radian\per\second}.
-Resistance is in units of \si{\ohm}.  This example uses \verb|\si|
-from the siunitx package to format the units correctly.  If you want
-to format it with a value, it uses \verb|\SI{value}{unit}| e.g.:
-\verb|\SI{2.99E8}{\meter\per\second}| becomes
-\SI{2.99E8}{\meter\per\second}
-
-As an example of referring to a particular part of the document, see
-Section~\ref{sec:fitting} on page~\pageref{sec:fitting} for information
-on how to fit text on a page.
-
-\subsubsection{Chemical Formulas}
-Chemical formulas are best formatted by the mhchem package (included
-in the custom.tex by default).
-\href{http://mirrors.ctan.org/macros/latex/contrib/mhchem/mhchem.pdf}{Click
-  here for the documentation}.  This example is disabled by default
-because it breaks due to a change in \LaTeX3 that hasn't propogated to
-all distributions.
-
-Here is an example in reaction~\ref{react:so4}:
-
-\reaction[react:so4]{SO4^2- + Ba^2+ -> BaSO4 v}
-
-\subsubsection{Tables}
-When making tables, there are some very important rules for making
-them look professional.  See
-\url{http://ctan.uib.no/macros/latex/contrib/booktabs/booktabs.pdf}
-For an example of a simple table, see Table~\ref{tab:example}.  Some general
-rules are to always avoid \verb|\hrule| and vertical separator lines.
-
-\begin{table}\centering
-\caption{Listing of real name to username\label{tab:example}}
-\begin{tabular}{lll}\toprule
-Firstname &Lastname &username\\\midrule
-Joseph &Foley &foley\\\bottomrule
-\end{tabular}
-\end{table}
-
-\subsection{Mutilingual support}
-In the main file, you have already seen the ML macros used to create multi-lingual documents.
-You may want to read more about the babel and polyglossia packages to do even more.
-Quotation marks can be made smart by using \enquote{encite}.
-This will change the quotation marks to follow whatever the current language is set to.
-The Context Sensite Quotes (csquote) package is full of useful macros of this sort. 
-One caveat: don't use the shorcut "` and '" in the macros nor in the text\footnote{This is the shortcute for German style quotations, which is the same as Icelandic.}.
-It doesn't work due to interactions with the multilanguage macros.
-If you need to use those quotes, you can switch languages temporarily,  \selectlanguage{icelandic}
-"`Like this'" \resetlanguage{}
-
-\subsection{URLs and paths}
-URLs are easy to use in both the bibliography and text.  In your
-\path{.bib} file, make sure that you only put URLs into the \path{url}
-field.  Putting it anywhere else is likely to cause it to go off the
-page.  If you are using the \path{hyperref} package, then these will
-be clickable.  You can put links in the text with
-\verb|\url| like this: \url{http://www.ru.is}.  If you need to
-specify a location of a file, use the
-\verb|\path| similarly: \path{C:\Documents and Settings\ }.  The only
-difference is that paths are not clickable.
-
-\subsection{Hints}
-When placing tables and figures, {\em do not\/} force \LaTeX{} to
-place them in a given place.  Doing so almost always causes problems
-later.  It is better to give hints \verb|[htb]| than to
-force it \verb|!h]|.
-
-Add linebreaks at least at the end of every sentence.  This is
-especially important if you are using SVN or collaborating with
-others.  Everything is done by line number, so if an error or
-issue is on line 5 which happens to be a page of text, good luck
-finding the problem.
-
-Along those same lines, if you are not storing your thesis in version
-control (SVN, git, etc) or on a backed up drive, you are asking for
-trouble.
-\subsection{Debugging}
-When things go wrong, it is sometimes hard to figure out what to do.
-The first step is always to go look at the log file.  If you are
-making changes and they are not showing up, it is likely that you are
-viewing an old file.  Delete the \path{.pdf} and \path{.aux} and try
-again.  On some platforms, the package manager can get out of date.
-Make sure you have updated/synchronized before running \path{pdflatex} again.
-
-%% ---------------------------------------------------------------
-\printbibliography{} %%RUM: "References"
-
-%% If appendices are needed, uncomment the following line
-%% and include the appendices in separate files
-\appendix{}%%RUM: "Appendicies (as appropriate)
+\chapter{Some Chapter}
+\section{Some section}
+\subsection{Some Subsection}
+Some text goes here with a citation~\cite{adams84fish}.
+\printbibliography{}
+\appendix{}
 \input{code} % as an example, perhaps some of your code
-
-%\backmatter{} % Sections after this don't get numbers
-%% We prefer that all elements be numbered
-
-%%%%%%%%%%%%% SHOW GLOSSARY %%%%%%%%%%%%%%%%%
-%% Glossary, optional.  A good idea on longer documents
-%% Remember to run "makeglossaries <DEGREE-NAME-YEAR>"
-%%   to update the glossary
-\printglossary{}%%RUM: Not mentioned
-\printacronyms{}%%RUM: Not mentioned
-
-%% Glossary debugging code
-%% Uncomment this to figure out of one of the glossary entries is broken
-%% by putting all of the entries in the glossary.
-%\ifthenelse{\boolean{debug}}{%
-%  \glsaddall{} %put all entries in
-%  \printglossaries{}}{}% dump them all
-%% This command dumps all of the glossaries
-
-%%%%%%%%%%%%% SHOW INDEX %%%%%%%%%%%%%%%%%%
-%% Index, optional.  A good idea on longer documents
-
-% You can put instructions at the beginning of the index:
-%\renewcommand{\preindexhook}{%
-%  The first page number is usually, but not always,
-%  the primary reference to the indexed topic.\vskip\onelineskip}
-
-%% You may have to run "makeindex" to have it be generated
-%% Depending upon which package you chose.
-%% 
 \clearforchapter{}
-\printindex{}%%RUM: Not mentioned
-
-\backcover{}%%RUM: "Back cover (standard)
+\printindex{}
+\backcover{}
 \end{document}

### 2016-04-04 writing/latex/wilf.tex
Add writing with RU template

\chapter{Wilf-equivalences under dominating classical patterns}

Another question often asked in thefield of permutation patterns is that
of Wilf-equivalence. Two patterns \(\pi\) and \(\sigma\) are said to be
Wilf-equivalent if their avoidance sets have the same size at each
length. More formally

\begin{definition}[Wilf-equivalence]
    Two patterns \(\pi\) and \(\sigma\) are said to be \emph{Wilf-equivalent}
    if for all \(k_{} \ge 0\), \(\size{\av[k]{\pi}} = \size{\av[k]{\sigma}}\).

    Two sets of permutation patterns \(R\) and \(S\) are are
    \emph{Wilf-equivalent} if for all \(k_{} \ge 0\),
    \(\size{\av[k]{R}} = \size{\av[k]{S}}\).
\end{definition}

Wilf-equivalence is of interest as if two permutation classes are enumerated
in the same way then there should exist a bijection between them, and therefore
any other combinatorial object that they represent.

There are a number of symmetries we can use when examining Wilf-equivalences
to reduce the amount of work, it can be easily seen that the reverse, complement
and inverse operations preserve enumeration, and therefore these classes are trivially
Wilf-equivalent.
Since we are always considering Wilf-equivalences in the set \(\av{\pi}\) we
must only use these symmetries when they preserve the dominating pattern.

### 2016-04-04 writing/latex/coincidence.tex
Finish set coincidence section

@@
     We can choose the topmost point in the region to replace the corresponding
-    point in the mesh pattern and move the points from the subsequence into
+    point in the mesh pattern and the points from the subsequence are now in
     the box southeast of the point. The other conditions allow this to be done
-    without moving points into regions that were shaded.Hence there are no
+    without points being present in regions that were shaded. Hence there are no
     points in the region corresponding to the box \((a,b)\) in the mesh
@@
     occurrence of \(m_1\) in \(\av{\pi}\) is in fact an occurrence of \(m_2\)
-    so \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).
+    so \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).
 
@@
 \begin{equation*}
-    \pattern{2,3}{2,1,3}{0/3,2/0,2/1,3/0,3/1,3/2}
+    \pattern{2,3}{2,1,3}{0/3,1/3,2/0,2/1,3/0,3/1,3/2}
 \end{equation*}
@@
         \item \(R_1 \triangle R_2 = \{(a,b)\}\)
-        \item \(\addp[D]{\mperm{\sigma}{R_1}}{(a,b)}\) is coincident with a
-            mesh pattern containing an occurrence of \(\mperm{\sigma}{R_2}\).
+        \item\label{prop:dom3:condocc} \(\addp[D]{\mperm{\sigma}{R_1}}{(a,b)}\)
+            is coincident with a mesh pattern containing an occurrence of
+            \(\mperm{\sigma}{R_2}\).
     \end{enumerate}
@@
 
-    Now consider a permutation in \(\av{\pi}\)
+    Now consider a permutation \(\varrho\) in \(\av{\pi}\) that contains an
+    occurrence of \(m_1\). If the region corresponding to the box \((a,b)\)
+    is empty then we have an occurrence of \(m_2\). If the region is non-empty
+    then by condition~\ref{prop:dom3:condocc}  of the proposition there exists
+    a direction such that there exists an occurrence of a mesh pattern of
+    length one longer than \(m_1\) in this position. This mesh pattern is
+    coincident with another mesh pattern. This mesh pattern contains an
+    occurrence of \(m_2\) so every occurrence of \(m_1\) is also an occurrence
+    of \(m_2\).  Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the
+    two patterns are coincident.
 \end{proof}
@@
 
-The patterns \fxnote{Is this dom3?}
+The patterns
 \begin{equation*}
@@
 \begin{equation*}
-    \pattern{}{1,2}{0/1,0/2,2/0,2/1} \text{ and } \pattern{}{1,2}{0/1,0/2,1/1,2/0,2/1}
+    m_1 = \pattern{}{1,2}{0/1,0/2,2/0,2/1} \text{ and } m_2 = \pattern{}{1,2}{0/1,0/2,1/1,2/0,2/1}
 \end{equation*}
@@
 
-In order to explain this we need to consider the middle box, and
-any points that it may contain. If the box is empty, then the coincidence
-is trivially true. Hence, consider if there is a point in the box. This
-gives us a mesh pattern
+In order to prove this coincidence we will proceed by mathematical
+induction on the number of points in the middle box we call this number \(n\).
+\begin{description}
+    \item [Base Case \((n=0)\):] The base case hold since we can freely shade the
+        box if it contains no points.
+    \item [Inductive Hypothesis \((n=k)\):] Suppose that the we can find an occurrence
+        of the second pattern if we have an occurrence of the first with \(k\) points
+        in the middle box.
+    \item [Inductive Step \((n=k+1)\)] Suppose that we have \((k+1)\) points in
+        the middle box. Choose the bottom most point in the middle box, this gives a
+        mesh pattern equivalent to
      \begin{equation*}
-    \raisebox{0.6ex}{
-    \begin{tikzpicture}[scale=0.5, baseline=(current bounding box.center)]
-        \modpattern[4]{}{1,2,3}{0/1,0/2,0/3,3/0,3/1,3/2}
-        \draw (1.5,3.5) node {\(A\)};
-        \draw (2.5,0.5) node {\(B\)};
-    \end{tikzpicture}}
+        \begin{tikzpicture}[scale=0.5]
+            \modpattern[5]{}{1,2,3}{0/1,0/2,0/3,1/1,2/1,3/0,3/1,3/2}
+            \draw (2.5,0.5) node {\(X\)};
+        \end{tikzpicture}
     \end{equation*}
-
-Where at most one of \(A\) or \(B\) contains an increasing subsequence.
-If both are empty we can \emph{a posteriori} make our point in the center
-box be either the leftmost, or the rightmost. Now consider if the box \(A\)
-were to contain any points, we now know that the box labelled \(B\) contains
-no points and can be shaded. By previously established rules we can also obtain
-other shadings, leading to the following mesh pattern.
+        Now we need to consider the box labelled \(X\) if this box is empty then
+        we have an occurrence of \(m_2\) and are done. If this box contains
+        any points then we gain some extra shading on the mesh pattern
+        due to the dominating pattern
      \begin{equation*}
-    \raisebox{0.6ex}{
-    \begin{tikzpicture}[scale=0.5, baseline=(current bounding box.center)]
-        \modpattern[4]{}{1,2,3}{0/1,0/2,0/3,3/0,3/1,3/2}
-        \filldraw (1.5,3.5) circle (4 pt);
-        \shadeboxes{2/0,2/1}
-    \end{tikzpicture}}
+        \begin{tikzpicture}[scale=0.5]
+            \modpattern[5]{2,3}{1,2,3}{0/1,0/2,0/3,1/1,1/2,1/3,2/1,3/0,3/1,3/2}
+            \filldraw (2.5,0.5) circle (4 pt);
+        \end{tikzpicture}
     \end{equation*}
-Now we can \emph{a posteriori} say that our point in the center box was the
-rightmost, giving
-\begin{equation*}
-    \raisebox{0.6ex}{
-    \begin{tikzpicture}[scale=0.5, baseline=(current bounding box.center)]
-        \modpattern[4]{1,2}{1,2,3}{0/1,0/2,0/3,3/0,3/1,3/2}
-        \filldraw (1.5,3.5) circle (4 pt);
-        \shadeboxes{1/1,1/2,2/0,2/1}
-    \end{tikzpicture}}
-\end{equation*}
-This gives shading that when we choose the highlighted points we
-get an occurrence of the required mesh pattern, and as such the
-two patterns are coincident.
-
-This argument applies to another pair of classes. Therefore in
+        The two highlighted points form an occurrence of \(m_1\) with \(k\)
+        points in the middle box, and thus by the Inductive Hypothesis we
+        are done.
+\end{description}
+By induction we have that every occurrence of \(m_1\) is an occurrence of \(m_2\)
+and by \NoteRef{not:downcmesh} every occurrence of \(m_2\) is an occurrence of
+\(m_1\) so the two patterns are coincident.
+This argument applies to another two pairs of classes. Therefore in
 total in \(\av{\perm{3,2,1}, \mperm{1,2}{R}}\) there are \(213\) equivalence

### 2016-04-06 src/main.py
Add start of Wilf-Equivalences

@@ -57,10 +57,24 @@ resolved321_12 = resolve_fundict(unexplained321_12, p2_coincs321_12)
 resolved231_21 = resolve_fundict(unexplained231_21, p2_coincs231_21)
 resolved321_21 = resolve_fundict(unexplained321_21, p2_coincs321_21)
 
-# need to prune by symmetries
-#
-reprs231 = [classes[x[0]][0] for x in coincsdict231_12.values()] + [classes[x[0]][0] for x in coincsdict231_21.values()]
-reprs321 = [classes[x[0]][0] for x in coincsdict321_21.values()] #+ [classes[x[0]][0] for x in coincsdict321_12.values()]
+wilf_equivs231_12 = wilf_equivs231(classes, meshpd_12)
+wilf_equivs321_12 = wilf_equivs321(classes, meshpd_12)
+wilf_equivs231_21 = wilf_equivs231(classes2, meshpd_21)
+wilf_equivs321_21 = wilf_equivs321(classes2, meshpd_21)
 
-wilf231 = [WilfCounter(Permutation([2,3,1]),x ) for x in reprs231]
-wilf321 = [WilfCounter(Permutation([3,2,1]),x ) for x in reprs321]
+pre_test_wilf231_12 = resolve_all_prewilf_classes(coincsdict231_12, wilf_equivs231_12)
+pre_test_wilf321_12 = resolve_all_prewilf_classes(coincsdict321_12, wilf_equivs321_12)
+pre_test_wilf231_21 = resolve_all_prewilf_classes(coincsdict231_21, wilf_equivs231_21)
+pre_test_wilf321_21 = resolve_all_prewilf_classes(coincsdict321_21, wilf_equivs321_21)
+
+wreprs231_12 = [sorted(classes[sorted(list(x))[0]])[0] for x in pre_test_wilf231_12]
+wreprs321_12 = [sorted(classes[sorted(list(x))[0]])[0] for x in pre_test_wilf321_12]
+wreprs231_21 = [sorted(classes[sorted(list(x))[0]])[0] for x in pre_test_wilf231_21]
+wreprs321_21 = [sorted(classes[sorted(list(x))[0]])[0] for x in pre_test_wilf321_21]
+
+wreprs231 = wreprs231_12 + wreprs231_21
+wreprs321 = wreprs321_12 + wreprs231_21
+
[…]

### 2016-04-06 src/wilf.py

@@ -24,3 +24,50 @@ class WilfCounter(object):
             return self.record == other.record
         else:
             raise TypeError("Cannot compare WilfCounter with %s" % type(other))
+
+
+def wilf_equivs231(classes,meshpd):
+    wilf_equiv231 = set()
+    for classn, thisclass in enumerate(classes):
+        currset = frozenset(map(lambda x: meshpd[x.reverse().complement().inverse()], thisclass)) | frozenset(map(lambda x: meshpd[x],thisclass))
+        if len(currset) > 1:
+            wilf_equiv231.add(currset)
+    return wilf_equiv231
+
+def wilf_equivs321(classes,meshpd):
+    wilf_equiv321 = set()
+    for classn, thisclass in enumerate(classes):
+        currset = (frozenset(map(lambda x: meshpd[x.reverse().complement().inverse()], thisclass)) |
+                    frozenset(map(lambda x: meshpd[x],thisclass)) |
+                    frozenset(map(lambda x: meshpd[x.reverse().complement()], thisclass)) |
+                    frozenset(map(lambda x: meshpd[x.inverse()], thisclass)))
+        if len(currset) > 1:
+            wilf_equiv321.add(currset)
+    return wilf_equiv321
+
+def find_containing_set(setset, val):
+    iset =  [aset for aset in setset if val in aset]
+    if len(iset) == 1:
+        return iset[0]
+    else:
+        return set([val])
+
+def resolve_prewilf_class(classnum, coincssets, wilfsets):
+    to_resolve = set([classnum])
+    fully_resolved = set()
+    thisset = set()
+    while len(to_resolve) > 0:
+        currnum = to_resolve.pop()
+        fully_resolved.add(currnum)
+        currcoincs = find_containing_set(coincssets, currnum)
+        currwilfs = find_containing_set(wilfsets,currnum)
+        thisset = thisset | currcoincs | currwilfs
+        to_resolve = thisset - fully_resolved
+    return frozenset(thisset)
+
+def resolve_all_prewilf_classes(coincsdict, wilfsets):
+    coincssets = set(map(frozenset, coincsdict.values()))
+    s = set()
+    for classnum in range(220):
+        s.add(resolve_prewilf_class(classnum, coincssets, wilfsets))
+    return s

### 2016-04-15 src/helper_funcs.py
Add partition function and catalan function, autowilf

@@ -1,6 +1,7 @@
 import json
 from permuta import *
 import itertools
+from math import factorial
 
 def load_set(filename):
     patts = []
@@ -98,3 +99,20 @@ def mpdict(classes):
 
 def countclasses(coincidict):
     return 220 - sum(map(len,coincidict.values())) + len(coincidict)
+
+def catalan(n):
+    return factorial(2*n)/(factorial(n+1)*factorial(n))
+
+def partition(perm):
+    last = 0
+    currblock = -1
+    partition = []
+    for value in perm:
+        if value > last:
+            partition.append([])
+            currblock += 1
+        partition[currblock].append(value)
+        last = value
+    return partition
+
+

### 2016-04-15 src/main.py

@@ -69,12 +69,31 @@ pre_test_wilf321_21 = resolve_all_prewilf_classes(coincsdict321_21, wilf_equivs3
 
 wreprs231_12 = [sorted(classes[sorted(list(x))[0]])[0] for x in pre_test_wilf231_12]
 wreprs321_12 = [sorted(classes[sorted(list(x))[0]])[0] for x in pre_test_wilf321_12]
-wreprs231_21 = [sorted(classes[sorted(list(x))[0]])[0] for x in pre_test_wilf231_21]
-wreprs321_21 = [sorted(classes[sorted(list(x))[0]])[0] for x in pre_test_wilf321_21]
+wreprs231_21 = [sorted(classes2[sorted(list(x))[0]])[0] for x in pre_test_wilf231_21]
+wreprs321_21 = [sorted(classes2[sorted(list(x))[0]])[0] for x in pre_test_wilf321_21]
 
-wreprs231 = wreprs231_12 + wreprs231_21
-wreprs321 = wreprs321_12 + wreprs231_21
+wreprs231 = set(wreprs231_12 + wreprs231_21)
+wreprs321 = set(wreprs321_12 + wreprs321_21)
 
-wilfcounters231 = [WilfCounter(Permutation(2,3,1), mperm) for mperm in wreprs231]
-wilfcounters321 = [WilfCounter(Permutation(3,2,1), mperm) for mperm in wreprs321]
+wilfcounters231 = [WilfCounter(Permutation([2,3,1]), mperm) for mperm in wreprs231]
+wilfcounters321 = [WilfCounter(Permutation([3,2,1]), mperm) for mperm in wreprs321]
 
+for permlen in range(11):
+    for perm in PermutationsAvoiding231(permlen):
+        for counter in wilfcounters231:
+            counter.modify(perm)
+    for perm in PermutationsAvoiding321(permlen):
+        for counter in wilfcounters321:
+            counter.modify(perm)
+    stderrwrite("Length %d complete" % permlen)
+
+wilfcounts231 = list(set([tuple(counter.record.values()) for counter in wilfcounters231]))
+wilfcounts321 = list(set([tuple(counter.record.values()) for counter in wilfcounters321]))
+
+
+wilf_equivs231 = get_wilf_sets(wilfcounts231, wilfcounters231,
+                               pre_test_wilf231_12, pre_test_wilf231_21,
+                               meshpd_12, meshpd_21)
[…]

### 2016-04-15 src/wilf.py

@@ -50,7 +50,7 @@ def find_containing_set(setset, val):
     if len(iset) == 1:
         return iset[0]
     else:
-        return set([val])
+        return frozenset([val])
 
 def resolve_prewilf_class(classnum, coincssets, wilfsets):
     to_resolve = set([classnum])
@@ -71,3 +71,16 @@ def resolve_all_prewilf_classes(coincsdict, wilfsets):
     for classnum in range(220):
         s.add(resolve_prewilf_class(classnum, coincssets, wilfsets))
     return s
+
+def get_wilf_sets(wilfcounts, wilfcounters, preequiv_12, preequiv_21, mp12, mp21):
+    superset = set()
+    for wilfseq in wilfcounts:
+        thisset = frozenset()
+        equivreprs = [x.mperm for x in wilfcounters if x.record == dict(enumerate(wilfseq))]
+        ureprs = [x for x in equivreprs if x.perm.perm == [1,2]]
+        dreprs = [x for x in equivreprs if x.perm.perm == [2,1]]
+        ureprclasses = [(find_containing_set(preequiv_12, mp12[x]), '12') for x in ureprs]
+        dreprclasses = [(find_containing_set(preequiv_21, mp21[x]), '21') for x in dreprs]
+        thisset = thisset | frozenset(ureprclasses) | frozenset(dreprclasses)
+        superset.add(thisset)
+    return superset

### 2016-04-15 writing/latex/coincidence.tex
start wilf equivalences

@@
-\chapter{Coincidences amongst families of mesh patterns and classical patterns}
+\chapter{Coincidences amongst families of mesh patterns and classical patterns}\label{chap:coincs}
 One interesting question to ask about permutation patterns considers when a

### 2016-04-15 writing/latex/wilf.tex
start wilf equivalences

@@
-\chapter{Wilf-equivalences under dominating classical patterns}
+\chapter{Wilf-equivalences under dominating  patterns}
 
@@
 
+Coincident pattern classes are also Wilf-equivalent. This is due to the fact that
+if \(\av[k]{S} = \av[k]{R}\) then obviously \(\size{\av[k]{R}} = \size{\av[k]{S}}\).
+Coincidence is therefore a stronger equivalence condition than Wilf-equivalence.
+
 There are a number of symmetries we can use when examining Wilf-equivalences
@@
 Wilf-equivalent.
-Since we are always considering Wilf-equivalences in the set \(\av{\pi}\) we
+
+\begin{align*}
+    \rev{\textpattern{}{2,3,1}{}} &= \textpattern{}{1,3,2}{}\\
+    \com{\textpattern{}{2,3,1}{}} &= \textpattern{}{2,1,3}{}\\
+    \inv{\textpattern{}{2,3,1}{}} &= \textpattern{}{3,1,2}{}
+\end{align*}
+
+The group of symmetries on permutations is isomorphic to the dihedral group
+of order \(8\), the group of symmetries of a square. Composition of the above
+symmetries gives the remaining 5 symmetries. If we consider generators of the
+group the operations \emph{reverse-inverse} and \emph{reverse} correspond to
+the generators of the dihedral group.
+
+Since we are always considering Wilf-equivalences in the set \(\av{S}\) we
 must only use these symmetries when they preserve the dominating pattern.
+
+Throughout this section we will consider Wilf-equivalences of patterns
+whilst avoiding the \emph{Dominating Pattern} \(\perm{2,3,1}\). We will
+use \(\mathcal{C}\) to denote the set of these avoiders and \(C(x)\) will
+be the usual Catalan generating function satisfying \(C(x) = 1 + C(x)^2\).
+This is easy to see by structural decomposition around the maximum, as
+shown in \FigureRef{fig:decompmax}
+
+\begin{figure}[!ht]
+    \centering
+    \decompmax{0/1,1/0}{A}{}{}{B}
+    \caption{Structural decomposition of a typical avoider of 231}
+    \label{fig:decompmax}
+\end{figure}
+
+The elements to the left of the maximum, \(A\), have the structure of a \(231\)
+avoiding permutation, and the elements to the right of the maximum, \(B\), have the
+structure of a \(\perm{2,3,1}\) avoiding permutation. Furthermore, all the
+elements in \(A\) lie below all of the elements in \(B\).
+
+We can also decompose a permutation avoiding \(\perm{2,3,1}\) around the leftmost
+point, giving a similar figure.
+
+\section{Wilf-classes with patterns of length 1.}
+When considering the mesh patterns of length \(2\) it will be useful to know
+the Wilf-equivalence classes of the mesh patterns of length \(1\) inside
+\(\av{\perm{2,3,1}}\).
+
+It can be seen that the patterns in the following set are set equivalent.
+\begin{equation*}
+    \left\{
+        \begin{array}{c}
+        \textpattern{}{1}{},
+        \textpattern{}{1}{0/0},
+        \textpattern{}{1}{1/0},
+        \textpattern{}{1}{0/0,1/0},
+        \textpattern{}{1}{0/1},\\
+        \textpattern{}{1}{0/0,0/1},
+        \textpattern{}{1}{1/1},
+        \textpattern{}{1}{1/0,1/1},
+        \textpattern{}{1}{0/1,1/1}
+    \end{array}
+    \right\}
+\end{equation*}
+This is due to the fact that every permutation except the empty
+permutation must contain an occurrence of all of these patterns.
+
+The pattern \(\textpattern{}{1}{0/0,1/0,0/1,1/1}\) is in its own
+Wilf-class since the only permutation containing this pattern is
+the permutation \(\perm{1}\). The avoiders of this pattern therefore
+have generating function \(E(x) = C(x)-x\).
+
+The avoiders of the pattern \(p_1 = \textpattern{}{1}{0/1,1/0}\) can be
+decomposed around the maximum element in order to give the following
+structural decomposition.
+\begin{equation*}
+    \mathcal{F} = \varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{F}}{}{}{\mathcal{C}\setminus\varepsilon}
+\end{equation*}
+Since if the upper right section was empty the maximum would create an
+occurrence of the pattern, however no points in this section can create
+an occurrence since the maximum lies in a region corresponding to the
+shading in \(p_1\). The lower right region however can create occurrences
+of \(p_1\) and therefore must also avoid \(p_1\) as well as \(\perm{2,3,1}\).
+This gives the generating function of avoiders to be the function \(F(x)\)
+satisfying.
+\begin{equation*}
+    F(x) = 1 + xF(x)(C(x)-1)
+\end{equation*}
+
+It can be shown by use of \PropositionRef{prop:dom2} that the patterns
+\textpattern{}{1}{0/0,1/1} and \(p_2 = \textpattern{}{1}{0/0,0/1,1/1}\) are
+set-equivalent. Consider the decomposition of an avoider of \(p_2\) in
+\(\av{\perm{2,3,1}}\) around the maximum
+\begin{equation*}
+    \mathcal{G}_1 =\varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
+\end{equation*}
+This can be explained succinctly by the fact that a permutation containing
+\(p_2\) starts with it's maximum, by not allowing the left part of the
+\(\perm{2,3,1}\) avoider to be empty we prevent an occurrence from ever happening.
+
+Consider \(p_3 = \textpattern{}{1}{0/1,1/0,1/1}\), avoiding this pattern
+means that a permutation does not end with it's maximum. We can perform a similar
+decomposition as before to get
+\begin{equation*}
+    \mathcal{G}_2 =\varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\mathcal{C}\setminus\varepsilon}
+\end{equation*}
+
+Now consider \(p_4 = \textpattern{}{1}{0/0,0/1,1/0}\), the avoiders of this
+pattern are permutations that do not start with their minimum. In this case we
+perform the decomposition around the leftmost element
+\begin{equation*}
+    \mathcal{G}_3 =\varepsilon \sqcup \decompleft{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
+\end{equation*}
+All of these classes have the same generating function namely
+\begin{equation*}
+    G(x) = 1 + xC(x)(C(x)-1)
+\end{equation*}
+
+There is one pattern of length \(1\) still to consider. The pattern
+\(p_5 = \textpattern{}{1}{0/0,1/0,1/1}\) is avoided by all permutations
+that do not end in their minimum. Considering the standard decomposition
+of a \(\perm{2,3,1}\) avoider around the maximum we can see that an avoider
+of \(p_5\) must fit into the following form precisely once.
+\begin{equation*}
+    \mathcal{H} = \varepsilon \sqcup
+    \underbrace{\decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}}_\text{
+        Minimum comes before the maximum.
+    } \sqcup
+    \underbrace{\decompmax{0/0,0/1,1/0}{}{}{}{\mathcal{H}}}_\text{
+        Minimum is after the maximum, cannot be last.
+    }
+\end{equation*}
+Therefore this particular class has generating function \(H(x)\) satisfying
+\begin{equation*}
+    H(x) = 1 + xC(x)(C(x)-1) + xH(x)
+\end{equation*}
+
+\section{Wilf-classes with patterns of length 2}
+By use of the set equivalences from \ChapterRef{chap:coincs} we know there are
+at most \(95\) Wilf-equivalence classes.
+
+In order to consider symmetries we must only take the symmetries that preserve
+the pattern \(\perm{2,3,1}\). If we take any of the symmetries alone the
+permutation is different. The only symmetry that preserves the pattern
+\(\perm{2,3,1}\) is that of \emph{reverse-complement-inverse}. Using this
+set of symmetries to merge classes gives us \(61\) classes of trivial
+Wilf-equivalences.
+
+Computing avoiders up to length \(10\) gives us \(23\) Wilf-classes,
+of which \(13\) are non-trivial.
+
+When considering explanations of Wilf-equivalences we consider how the
+permutations correspond to set-partitions.
+
+\begin{note}
+    \label{not:setequiv}
+    The avoiders of the pattern \(\mperm{2,3,1}{\{(1,0),(1,1),(1,2),(1,3)\}}\)
+    in \(\mathfrak{S}_n\) are in one-to-one correspondence with partitions of
+    \(\nrange{n}\). [Anders, Generalized Pattern Avoidance]
+\end{note}
+\begin{proof}
+    Let \(\pi\) be a permutation in \(\av[n]{\perm{2,3,1}}\) take the permutation
+    in one-line notation and insert a dash between each ascent in \(\pi\). This
+    corresponds to set partitions where the blocks are the elements between the
+    dashes, the blocks are listed in increasing order of their least element,
+    with the elements written in each block in descending order.
+\end{proof}
+
+\begin{example}
+    Given the permutation \(\pi = \perm{5,4,2,1,3,9,6,8,7}\) this corresponds to the
+    partition \(\{\{1,2,4,5\},\{3\},\{6,9\},\{8,7\}\}\).
+\end{example}
+
+We are looking at permutations in \(\av{\perm{2,3,1}}\), all of these permutations
+also avoid the vincular pattern in \NoteRef{not:setequiv}.
+
+\subsection{}
+The set/symmetry classes containing the following patterns are
+Wilf-equivalent in \(\av{\perm{2,3,1}}\) but are not set equivalent or
+symmetries of each other
+\begin{equation*}
+    m_1 = \pattern{}{1,2}{0/0,0/2,
+                    1/0,1/1,
+                    2/0,2/1,2/2}
+    \text{ and }
+    m_2 = \pattern{}{1,2}{0/0,
+                    1/0,1/1,1/2,
+                    2/0,2/1,2/2}
+\end{equation*}
+In this case it is better to consider the containers of the patterns
+instead of the avoiders due to the amount of shadings in the mesh.
+
+We look at the containers of the pattern \(m_1\), there can only ever
+be one occurrence of this pattern in a permutation corresponding to
+the last point in the permutation and the minimum. The only form that
+points in either of the two boxes can take is a decreasing sequence.
+For a permutation of length \(k\) if we fix the number of points in
+one of the boxes the number of points in the other box is determined.
+Therefore we can have any number of points from \(\setrange{0}{k-2}\)
+points in the bottom box. Therefore there are \(k-1\) containers of
+length \(k\). These permutations correspond to set partitions of \(k\)
+points into exactly two non-overlapping parts partitioned by the first
+element and the minimum.
+
+Now consider the containers of \(m_2\), we know that the unshaded region
+must contain a decreasing subsequence, with the point corresponding to
+the \(1\) in the mesh pattern. This decreasing subsequence has \(k-1\)
+points, we can put the point corresponding to the \(2\) above any of these
+points and therefore there are \(k-1\) containers of length \(k\).
+
+Therefore these two patterns have been shown to have the same number of
+avoiders of length \(k\) for all \(k\) and therefore all Wilf-equivalent.
+
+\subsection{}
+The classes containing the following patterns are Wilf-equivalent
+when avoiders are considered in \(\av{\perm{2,3,1}}\)
+\begin{align}
+    m_1 = \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/1} &\text{ and }
+    m_2 = \pattern{}{2,1}{0/0,0/1,1/1,1/2,2/0,2/1} \label{eq:singleud}\\
+    \text{ and }
+    m_3 = \pattern{}{2,1}{0/1,0/2,1/0,1/1,1/2,2/1} &\text{ and }
+    m_4 = \pattern{}{2,1}{0/1,0/2,1/0,1/1,1/2,2/0} \label{eq:other}
+\end{align}
+
+First we consider the coincidence in two parts then consolidate these parts.
+Consider the coincidence shown in \eqref{eq:singleud}. The easiest way to
+show that these are equinumerous is to consider the containers in the realm
+of set partitions.
+
+Due to the shading we know the following about the points corresponding
+to the points in the patterns.
+\begin{itemize}
+    \item The point corresponding to the first point in both patterns
+        must lie in the first block of our set partition (there are
+        no points southwest from it in the permutation).
+    \item The point corresponding to the second point in both patterns
+        is a block bottom (there are no points southeast of it in
+        the permutation).
+    \item The block containing the point corresponding to the second
+        point in both patterns contains only the point (it is a singleton
+        block).
+\end{itemize}
+This tells us that an occurence of the patterns must happen when there
+is a singleton block occurring after the first block. The difference
+between the patterns is in the underlying classical pattern. This
+means that permtuations containing \(m_1\) correspond to set partitions
+with a singleton block with value one higher than some element in the
+block containing \(1\). The permutations containing \(m_2\) correspond
+to the set partitions containing a block with block bottom having value
+one lower than some element in the block containing \(1\) and if this
+block is not the block containing \(1\) then it is a singleton block.
+
+Consider an avoider of \(\perm{2,3,1}\) and \(m_3\). We can perform
+the decomposition around the maximum
+\begin{equation*}
+    \mathcal{I}_1 = \varepsilon \sqcup
+    \decompmax{1/0,0/1}{\mathcal{I}_1}{}{}{\mathcal{G}_1}
+\end{equation*}
+Since only the first point in the top right region can
+create an occurrence of \(m_3\) if and only if it is the
+maximum in this region we must avoid starting with the maximum.
+
+Looking at avoiders of \(\perm{2,3,1}\) and \(m_4\) we
+can perform the same decomposition around the maximum to get
+\begin{equation*}
+    \mathcal{I}_2 = \varepsilon \sqcup
+    \decompmax{1/0,0/1}{\mathcal{I}_2}{}{}{\mathcal{G}_3}
+\end{equation*}
+Since an occurrence of \(m_4\) can never occur in the
+top right region, and could only occur between the first point
+in the region and the maximum, if and only if this first point is the minimum.
+Since both \(\mathcal{G}_1\) and \(\mathcal{G}_3\) have the same
+enumeration, \(\mathcal{I}_1\) and \(\mathcal{I}_2\) must also have
+the same enumeration and are therefore Wilf-equivalent.
+
+Now we must consolidate these two subclasses. In order to do this we must
+consider the decomposition around the leftmost point of a permutation in
+\(\av{\perm{2,3,1},m_1}\) we gain the following.
+\begin{equation*}
+    \mathcal{I}_3 = \varepsilon \sqcup
+    \decompleft{1/0,0/1}{\mathcal{I}_3}{}{}{\mathcal{G}_3}
+\end{equation*}
+It is therefore obvious that avoiders of \(m_1\) and avoiders of \(m_4\)
+have the same enumeration, and therefore all four patterns are Wilf-equivalent
+in \(\av{\perm{2,3,1}}\) with generating function satisfying
+\begin{equation*}
+    I(x) = 1 + xI(x)G(x)
+\end{equation*}
+
+\subsection{}
+The classes containing the following patterns are Wilf-equivalent
+in \(\av{\perm{2,3,1}}\)
+\begin{equation*}
+    m_1 = \pattern{}{1,2}{0/0,0/1,0/2,
+                    1/0,1/1,1/2,
+                    2/0,2/1,2/2}
+    \text{ and }
+    m_2 = \pattern{}{2,1}{0/0,0/1,0/2,
+                    1/0,1/1,1/2,
+                    2/0,2/1,2/2}
+\end{equation*}
+It is obvious that these two are Wilf-equivalent since the only
+permutations that contain these patterns are \(\perm{1,2}\) and
+\(\perm{2,1}\) respectively, therefore the avoiders of these patterns
+are counted by the Catalan numbers at all lengths except for length
+\(2\) where there is precisely \(1\) avoider.
+
+\subsection{}
+
+The classes containing the following patterns are Wilf-equivalent
+in \(\av{\perm{2,3,1}}\)
+\begin{equation*}
+    m_1 = \pattern{}{2,1}{0/0,0/1,0/2,
+                    1/1,1/2,
+                    2/0,2/2}
+    \text{ and }
+    m_2 = \pattern{}{1,2}{0/1,0/2,
+                    1/0,1/2,
+                    2/0,2/1,2/2}
+\end{equation*}
+
+First consider the structure of an avoider of \(m_1\) and \(\perm{2,3,1}\)
+we can perform the usual structural decomposition of an avoider
+of \(\perm{2,3,1}\) where we consider decomposition around the maximum.
+Any permutation in \(\av{\perm{2,3,1},m_1}\) is in the following set
+precisely once.
+\begin{equation*}
+    \mathcal{J}_1 = \varepsilon \sqcup
+    \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}} \sqcup
+    \decompmax{0/1,1/0,0/0}{}{}{}{\mathcal{F}}
+\end{equation*}
+
+Now consider the decomposition around the maximum of a permutation in
+\(\av{\perm{2,3,1},m_2}\) This fits into the following set.
+\begin{equation*}
+    \mathcal{J}_2 = \varepsilon \sqcup
+    \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\mathcal{C}\setminus\varepsilon} \sqcup
+    \decompmax{0/1,1/0,1/1}{\mathcal{F}}{}{}{}
+\end{equation*}
+Therefore both of these sets of avoiders are enumerated in the same manner
+having generating function satisfying
+\begin{equation*}
+    J(x) = 1 + xC(x)(C(x)-1) + xF(x)
+\end{equation*}
+
+\subsection{}
+Consider the containers of the patterns
+\begin{equation*}
+    \pattern{}{1,2}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
+    \pattern{}{1,2}{1/1,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
+    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,1/1,2/1,2/2},
+    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,1/1,2/2},
+    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,1/1},
+    \pattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0,2/1,2/2}
+\end{equation*}
+
+For each of these patterns there is precisely one occurrence
+in any permutation containing the pattern. Now consider
+the points in the free box in each case. Each of these regions
+must contain an avoider of \(\perm{2,3,1}\) that is of length
+\(n-2)\). Therefore these classes are all Wilf-equivalent
+and the number of length \(n\) avoiders is
+\begin{equation*}
+    K_n = C_n - C_{n-2}
+\end{equation*}
+for \(n \ge 2\) where \(C_n\) is the \(n\)th Catalan number, the number of
+\(\perm{2,3,1}\) avoiders of length \(n\).
+
+\subsection{}
+Now consider the containers of the patterns
+\begin{equation*}
+    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
+    \pattern{}{1,2}{0/0,1/1,0/2,1/0,1/2,2/0,2/1,2/2},
+    \pattern{}{1,2}{0/0,0/1,1/1,1/0,1/2,2/0,2/1,2/2}
+\end{equation*}
+Each of these patterns again occurs precisely once in
+any containing permutation. However this time when considering
+the free box we need to take into consideration \LemmaRef{lem:incdec}
+and so the empty box can only contain a decreasing subsequence. There
+is precisely one decreasing subsequence at every length, and so there
+is exactly one container of each pattern at each length. The three
+patterns are Wilf-equivalent and have \(C_n - 1\) avoiders of
+length \(n\) for all \(n \ge 2\).
+
+\subsection{}
+The containers of the following patterns can only
+have exactly one occurrence.
+\begin{equation*}
+    \pattern{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
+    \pattern{}{2,1}{0/0,0/1,1/0,1/1,1/2,2/0,2/2}
+    \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/1,2/2}
+    \pattern{}{1,2}{0/1,0/2,1/0,1/1,2/0,2/1,2/2}
+    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1}
+    \pattern{}{1,2}{0/0,0/1,1/0,1/2,2/0,2/1,2/2}
+\end{equation*}
+Once again we consider the free boxes, obviously for every pattern
+except the first the two boxes are independent, and one contains
+any avoider of \(\perm{2,3,1}\) and the other must contain a
+decreasing sequence by \LemmaRef{lem:incdec}.
+Let us consider the first pattern on it's own. In order to
+avoid \(\perm{2,3,1}\) across the free boxes we can add some additional
+restrictions
+\begin{equation*}
+    \raisebox{0.6ex}{
+    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
+        \fill[pattern=north east lines, pattern color=black!75] (1.5, 0) rectangle +(0.5,1);
+        \fill[pattern=north east lines, pattern color=black!75] (1.0, 2) rectangle +(0.5,1);
+        \draw (1.5,0) -- (1.5,1)
+              (1.5,2) -- (1.5,3);
+        \modpattern[2]{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
+    \end{tikzpicture}
+    }
+\end{equation*}
+Now we can see that the top free box must contain a decreasing sequence,
+and the bottom must contain an avoider of \(\perm{2,3,1}\) and these
+two do not interact in any manner. The containers of this pattern are counted
+the same as the other patterns, and due to this they are Wilf-equivalent.
+
+\subsection{}
+\subsection{}
+The classes containing the following patterns are Wilf-equivalent
+\begin{equation*}
+    \pattern{}{1,2}{} \text{ and }
+    \pattern{}{2,1}{}
+\end{equation*}
+This is true since the only avoiders of these patterns are the decreasing
+sequence and the increasing sequence respectively, and both of these
+avoid \(\perm{2,3,1}\) in all cases. There is therefore \(1\) avoider at
+every length.
+
+\subsection{}
+\subsection{}
+\begin{equation*}
+    p_1 = \pattern{}{1,2}{1/0,1/1,1/2,2/0,2/1} \text{ and }
+    p_2 = \pattern{}{2,1}{0/1,1/0,1/1,1/2,2/1}
+\end{equation*}
+Let \(\mathcal{L}_1\) be the set of avoiders of \(p_1\), by structural decomposition
+around the leftmost point we have

### 2016-04-17 writing/latex/wilf.tex
Wilf-equivs

@@
 must contain an avoider of \(\perm{2,3,1}\) that is of length
-\(n-2)\). Therefore these classes are all Wilf-equivalent
+\(n-2\). Therefore these classes are all Wilf-equivalent
 and the number of length \(n\) avoiders is
@@
 restrictions
-\begin{equation*}
+\begin{equation}
     \raisebox{0.6ex}{
@@
     }
-\end{equation*}
+    \label{eqn:splitboxes}
+\end{equation}
 Now we can see that the top free box must contain a decreasing sequence,
@@
 \subsection{}
+The classes containing the following patterns are Wilf-equivalent
+\begin{equation*}
+    \begin{aligned}
+        m_1 = \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/2} &\text{ and }
+        m_2 = \pattern{}{1,2}{0/0,0/1,1/0,2/0,2/1,2/2} \\
+        \text{ and }
+        m_3 = \pattern{}{1,2}{0/2,1/0,1/1,2/0,2/1,2/2} &\text{ and }
+        m_4 = \pattern{}{1,2}{0/1,1/0,1/1,2/0,2/1,2/2}
+    \end{aligned}
+\end{equation*}
+
+First consider the decomposition of avoiders of \(m_1\) in \(\av{\perm{2,3,1}}\)
+around the maximum.
+\begin{equation*}
+    \mathcal{L}_1 =
+    \varepsilon \sqcup
+    \decompmax{0/1,1/0}{\scriptstyle \mathcal{G}_1\setminus\varepsilon}{}{}{\mathcal{C}}
+    \sqcup
+    \decompmax{0/0,0/1,1/0}{}{}{}{\mathcal{L}_1}
+\end{equation*}
+
+Now we decompose the avoiders of \(m_2\) around the leftmost point
+\begin{equation*}
+    \mathcal{L}_2 =
+    \varepsilon \sqcup
+    \decompleft{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}\setminus\varepsilon}
+    \sqcup
+    \decompleft{0/1,1/0,1/1}{\mathcal{L}_2}{}{}{}
+\end{equation*}
+This gives us two generating functions satisfying the following pair of equations
+\begin{align}
+    L_1(x) &= 1 + xC(x)(G(x)-1) +xL_1(x) \label{eqn:genL1}\\
+    \text{ and }L_2(x) &= 1 + x(C(x)-1)^2 + xL_2(x) \label{eqn:genL2}
+\end{align}
+In order for these two functions to give the same value it is necessary to show that
+\eqref{eqn:genL1} and \eqref{eqn:genL2} are equal, this occurs if and only if
+\(C(x)(G(x)-1) = (C(x)-1)^2\).
+\begin{align*}
+        (C(x)-1)^2 &= C(x)(G(x)-1)\\
+        \shortintertext{Using the definition of \(G\) and \(C\)}
+        x^2C^4(x) &=  (1+xC(x))(xC^2(x)-xC(x))\\
+        xC^3(x)&=C(x)-1+xC^3(x)-xC^2(x)\\
+        0 &= C(x) - 1 -xC^2(x)\\
+        C(x) &= 1 +xC^2(x)
+\end{align*}
+The final line is always satisfied, and therefore the two generating functions
+are equal.
+
+Now we look at the other patterns. In particular note that any container of these
+patterns can contain the pattern precisely once, \(m_2\) specifies the
+minimum and last point, \(m_3\) and \(m_4\) both use the last point and the
+previous block bottom (in the set partition context).
+
+Looking at the structure of a container of \(m_3\) in \(\av{\perm{2,3,1}}\)
+\begin{equation*}
+    \raisebox{0.6ex}{
+        \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
+            \modpattern[2]{}{1,2}{0/2,1/0,1/1,2/0,2/1,2/2}
+            \draw (0.5,0.5) node {\(b_1\)};
+            \draw (0.5,1.5) node {\(b_2\)};
+            \draw (1.5,2.5) node {\(b_3\)};
+        \end{tikzpicture}
+    }
+\end{equation*}
+The boxes \(b_2\) and \(b_3\) must contain a decreasing sequence by
+\LemmaRef{lem:incdec}. The box labelled \(b_1\) must contain an avoider
+of \(\perm{2,3,1}\). However note that the points in this box can have
+interaction with any points in box \(b_2\). If there is just one point
+in \(b_2\) then any points in \(b_1\) to the left of this point must
+be lower than any points to the right of this point. By extension,
+if \(b_2\) contains a decreasing sequence with \(k\) points, there
+are \(k+1\) non-interacting avoiders of \(\perm{2,3,1}\) in \(b_3\).
+
+Now in \(m_2\) and \(m_4\) containers we can use the same method as in
+\eqref{eqn:splitboxes} to separate the two decreasing sequences in the
+free boxes in the top row, and the mixing happens in the same manner
+as in a container of \(m_3\).
+We now have that \(m_2,m_3\) and \(m_4\) have the same number of containers
+so are Wilf-equivalent, and that \(m_1\) and \(m_2\) have the same generating
+function. So all four classes are Wilf-equivalent.
 \subsection{}
@@
 \subsection{}
+The patterns
+\begin{equation*}
+    \pattern{}{1,2}{0/0,0/2,1/0,1/2,2/0,2/2},
+    \pattern{}{1,2}{1/1,2/1,1/0,1/2,2/0,2/2},
+    \pattern{}{1,2}{0/1,1/0,1/2,2/0,2/2},
+    \pattern{}{1,2}{0/1,0/2,1/1,1/2,2/0}, \text{ and }
+    \pattern{}{1,2}{0/0,0/1,0/2,1/0,1/1,1/2}
+\end{equation*}
+are Wilf-equivalent.
+If \(\mathcal{M}_1\) is the set of avoiders of
+\textpattern{}{1,2}{0/0,0/2,1/0,1/2,2/0,2/2}, then by the structural
+decomposition around the maximum we have
+\begin{equation*}
+    \mathcal{M}_1 = \varepsilon \sqcup
+    \decompmax{0/0,0/1,1/0}{}{}{}{\mathcal{C}}
+\end{equation*}
+This is because \(\mathcal{M}_1\) is the set of permutations
+who have their minimum occur after their maximum.
+
+The second pattern occurs if the last element is higher than
+the penultimate element. This can only occur if the last element is
+in a single block in the set partition context, In order to construct
+a avoider of length \(n\) we can take any avoider \(\perm{2,3,1}\)
+of length \(n-1\) and insert the new maximum into the last block.
+This ensures that the last block is never a singleton. This means
+that these permutations are also counted by \(M(x) = 1+xC(x)\).
+
+The avoiders of the third pattern can be decomposed by the maximum
+to give
+\begin{equation*}
+    \mathcal{M}_3 = \varepsilon \sqcup
+    \decompmax{0/1,1/0}{\mathcal{F}}{}{}{\mathcal{M}_3}
+\end{equation*}
+The generating function derived satisfies \(M_3(X) = 1 + xF(x)M_3(x)\).
+The fourth pattern can be decomposed around the maximum in a similar
+manner.
+\begin{equation*}
+    \mathcal{M}_4 = \varepsilon \sqcup
+    \decompmax{0/1,1/0}{\mathcal{M}_4}{}{}{\mathcal{F}}
+\end{equation*}
+
+Finally considering the last pattern, the only way we can construct an
+avoider is to take any \(\perm{2,3,1}\) avoider and add a new minimum
+at the start of the permutation. Adding a new leftmost point in any other
+position would either create an occurrence of \(\perm{2,3,1}\) or the pattern.
+Therefore this is also counted by \(M(x) = 1+xC(x)\).
 \subsection{}
@@
 \end{equation*}
-Let \(\mathcal{L}_1\) be the set of avoiders of \(p_1\), by structural decomposition
+Let \(\mathcal{N}_1\) be the set of avoiders of \(p_1\), by structural decomposition
 around the leftmost point we have
+\begin{equation*}
+    \mathcal{N}_1 = \varepsilon \sqcup
+    \decompleft{0/1,1/0}{\mathcal{N}_1}{}{}{\mathcal{N}^\prime_1}
+\end{equation*}
+Where \(\mathcal{N}^\prime_1\) is a permutation avoiding \(\perm{2,3,1}, m_1\)
+and \(\textpattern{}{1}{0/0,0/1,1/1}\)
+Now consider the decomposition of a permutation in \(\mathcal{N}^\prime_1\) it can
+once again be decomposed around the leftmost point
+\begin{equation*}
+    \mathcal{N}^\prime_1 = \varepsilon \sqcup
+    \decompleft{0/1,1/0}{\scriptstyle \mathcal{N}_1\setminus\varepsilon}{}{}{\mathcal{N}^\prime_1}
+\end{equation*}
+This is a complete decomposition of avoiders of \(p_1\). Now we
+look at an avoider of \(p_2\), this time decomposition is around the maximum
+\begin{equation*}
+    \mathcal{N}_2 = \varepsilon \sqcup
+    \decompleft{0/1,1/0}{\mathcal{N}_2}{}{}{\mathcal{N}^\prime_2}
+\end{equation*}
+
+Again we use the same method of decomposition of a permutation in \(\mathcal{N}^\prime_2\)
+\begin{equation*}
+    \mathcal{N}^\prime_2 = \varepsilon \sqcup
+    \decompleft{0/1,1/0}{\scriptstyle \mathcal{N}_2\setminus\varepsilon}{}{}{\mathcal{N}^\prime_2}
+\end{equation*}
+
+This gives us a generating function \(N(x)\) satisfying
+\begin{align}
+    N(x) &= 1 + xN(x)N^\prime(x) \label{eqn:Ngen}\\
+    N(x) &= 1 + x(N(x)-1)N^\prime(x)\label{eqn:NprimeGen}
+\end{align}
+
+Solving \EquationRef{eqn:NprimeGen} for \(N^\prime(x)\) and substituting into
+\EquationRef{eqn:Ngen} gives us the fact that a the generating function for
+\(N(x)\) satisfies
+\begin{equation}
+    N(x) = xN^2(x) - x(N(x) - 1) + 1
+\end{equation}

### 2016-04-17 src/main.py

@@ -78,7 +78,7 @@ wreprs321 = set(wreprs321_12 + wreprs321_21)
 wilfcounters231 = [WilfCounter(Permutation([2,3,1]), mperm) for mperm in wreprs231]
 wilfcounters321 = [WilfCounter(Permutation([3,2,1]), mperm) for mperm in wreprs321]
 
-for permlen in range(11):
+for permlen in range(10):
     for perm in PermutationsAvoiding231(permlen):
         for counter in wilfcounters231:
             counter.modify(perm)

### 2016-04-20 writing/latex/coincidence.tex
Finish wilf, start appendixes

@@
 \section{Equivalence classes of Av(\{321, (21, \textit{R})\}).}
-Through expermentation we discover that there are a total of 29 equivalence
+Through experimentation we discover that there are a total of 29 equivalence
 classes of mesh patterns with underlying classical pattern \(\perm{2,1}\).
@@
 \end{equation*}
-which are coincident by expermentation.
+which are coincident by experimentation.
 

### 2016-04-20 writing/latex/wilf.tex
Finish wilf, start appendixes

@@
 satisfying.
-\begin{equation*}
-    F(x) = 1 + xF(x)(C(x)-1)
-\end{equation*}
+\begin{align*}
+    F(x) &= 1 + xF(x)(C(x)-1) \\
+    \shortintertext{Solving for \(F\) gives}
+    F(x) &= \frac{1}{1+x-xC(x)}
+\end{align*}
 
@@
 In order for these two functions to give the same value it is necessary to show that
-\eqref{eqn:genL1} and \eqref{eqn:genL2} are equal, this occurs if and only if
+\eqref{eqn:genL1} and \eqref{eqn:genL2} are equal, this occurs if
 \(C(x)(G(x)-1) = (C(x)-1)^2\).
 \begin{align*}
-        (C(x)-1)^2 &= C(x)(G(x)-1)\\
-        \shortintertext{Using the definition of \(G\) and \(C\)}
-        x^2C^4(x) &=  (1+xC(x))(xC^2(x)-xC(x))\\
-        xC^3(x)&=C(x)-1+xC^3(x)-xC^2(x)\\
-        0 &= C(x) - 1 -xC^2(x)\\
-        C(x) &= 1 +xC^2(x)
+    &&(C(x)-1)^2 &= C(x)(G(x)-1)\\
+        &\Leftrightarrow& x^2C^4(x) &=  xC(x)(C(x)-1)C(x) &&\text{By definition of \(G\) and \(C\)}\\
+        &\Leftrightarrow& x^2C^4(x)&=xC^3(x)-xC^2(x)\\
+        &\Leftrightarrow& xC^2(x) &= C(x) -1 &&\text{Divide by \(xC^2(x)\)}\\
+        &\Leftrightarrow& C(x) &= 1 +xC^2(x)
 \end{align*}
-The final line is always satisfied, and therefore the two generating functions
-are equal.
+The final line is always satisfied since it is the form of \(C(x)\), and
+therefore the two generating functions are equal.
 
@@
 so are Wilf-equivalent, and that \(m_1\) and \(m_2\) have the same generating
-function. So all four classes are Wilf-equivalent.
+function so all four classes are Wilf-equivalent.
 \subsection{}
@@
 
-The second pattern occurs if the last element is higher than
-the penultimate element. This can only occur if the last element is
-in a single block in the set partition context, In order to construct
-a avoider of length \(n\) we can take any avoider \(\perm{2,3,1}\)
-of length \(n-1\) and insert the new maximum into the last block.
-This ensures that the last block is never a singleton. This means
+% If \(\mathcal{M}_2\) is the set of avoiders of
+% \textpattern{}{1,2}{1/1,2/1,1/0,1/2,2/0,2/2}, then by structural
+% decomposition around the maximum we have
+% \begin{equation*}
+%     \mathcal{M}_2 = \varepsilon \sqcup
+%     \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\scriptstyle\mathcal{M}_1\setminus\varepsilon}
+%     \sqcup \tikz{\filldraw (0,0) circle (3pt);}
+% \end{equation*}
+% Alternatively,
+The pattern \textpattern{}{1,2}{1/1,2/1,1/0,1/2,2/0,2/2} occurs if
+the last element is higher than the penultimate element. This can only occur
+if the last element is in a single block in the set partition context, In order
+to construct a avoider of length \(n\) we can take any avoider of
+\(\perm{2,3,1}\) of length \(n-1\) and insert the new maximum into the last
+block.  This ensures that the last block is never a singleton. This means
 that these permutations are also counted by \(M(x) = 1+xC(x)\).
@@
 \end{equation*}
-The generating function derived satisfies \(M_3(X) = 1 + xF(x)M_3(x)\).
+The generating function derived satisfies \(M_3(x) = 1 + xF(x)M_3(x)\).
 The fourth pattern can be decomposed around the maximum in a similar
@@
 Therefore this is also counted by \(M(x) = 1+xC(x)\).
+
+We need to show that the generating function \(M_3(x)\) is the same as \(M(x)\)
+\begin{equation*}
+    \begin{aligned}
+        M_3(x) &= 1 + xF(x)M_3(x) \\
+        &= \frac{1}{1-xF(x)} && \text{Solving for \(M_3(x)\)}\\
+        &= \frac{1}{1 - \frac{x}{1+x-xC(x)}} &&\text{Substituting for \(F(x)\)}\\
+        &= \frac{1-xC(x) +x}{1-xC(x)}\\
+        &= 1+xC(x)
+    \end{aligned}
+\end{equation*}
+We have that \(M_3(x) = 1+xC(x)= M(x) \) so all four patterns are Wilf-equivalent.
+
+\subsection{}
+The patterns
+\begin{equation*}
+    \begin{aligned}
+        m_1 = \pattern{}{1,2}{0/0,0/1,1/0,1/2,2/1} \text{ and }
+        m_2 = \pattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0}\\
+        \text{ and }
+        m_3 = \pattern{}{1,2}{0/1,0/2,1/0,1/2,2/0,2/2} \text{ and }
+        m_4 = \pattern{}{2,1}{1/0,1/1,2/1,2/2}\\
+    \end{aligned}
+\end{equation*}
+Can be shown to be Wilf-equivalent. First we consider a container of \(m_2\) in \(\av{\perm{2,3,1}}\)
+\begin{equation*}
+    \raisebox{0.6ex}{
+    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
+        \fill[pattern=north east lines, pattern color=black!75] (2.5, 1) rectangle +(0.5,1);
+        \fill[pattern=north east lines, pattern color=black!75] (2.0, 2) rectangle +(0.5,1);
+        \draw (2.5,1) -- (2.5,3);
+        \draw (1.5,0.5) node {\(b_1\)};
+        \draw (2.25,1.5) node {\(\scriptstyle b_2\)};
+        \draw (2.75,2.5) node {\(\scriptstyle b_3\)};
+        \modpattern[2]{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0}
+    \end{tikzpicture}
+    }
+\end{equation*}
+We can choose the lowest occurrence of \(m_2\) without loss of generality.
+The region corresponding to \(b_1\) must avoid the pattern \(\textpattern{}{1}{0/1,1/0}\)
+as well as \(\perm{2,1,3}\). The regions corresponding to \(b_2\) and \(b_3\) must
+now contain avoiders of \(\perm{2,3,1}\), these regions cannot mix in order to avoid
+\(\perm{2,3,1}\). Since we already have an occurrence of
+\(m_2\) we do not need to care about creating more occurrences so there are no
+other conditions on these boxes.
+
+Now looking at a container of \(m_3\) in \(\av{\perm{2,3,1}}\)
+\begin{equation*}
+    \raisebox{0.6ex}{
+    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
+        \fill[pattern=north east lines, pattern color=black!75] (1, 1.5) rectangle +(1,0.5);
+        \fill[pattern=north east lines, pattern color=black!75] (2, 1) rectangle +(1,0.5);
+        \draw (1,1.5) -- (3,1.5);
+        \draw (0.5,0.5) node {\(b_1\)};
+        \draw (1.5,1.25) node {\(\scriptstyle b_2\)};
+        \draw (2.5,1.75) node {\(\scriptstyle b_3\)};
+        \modpattern[2]{}{1,2}{0/1,0/2,1/0,1/2,2/0,2/2}
+    \end{tikzpicture}
+    }
+\end{equation*}
+We consider the leftmost occurrence of \(m_3\), the region corresponding
+to \(b_1\) must avoid the pattern \(\textpattern{}{1}{0/1,1/0}\) as well as
+\(\perm{2,3,1}\) once more. The regions corresponding to \(b_2\) and \(b_3\)
+must avoid \(\perm{2,3,1}\) and as in a container of \(m_2\) these regions cannot
+mix, as doing so would lead to an occurrence of \(\perm{2,3,1}\).
+Therefore both of these sets of containers are enumerated in the same way.
+
+Now we find a structural decomposition for an avoider of \(m_2\).
+Decomposing around the maximum we see the set of avoiders of \(m_2\)
+have the form
+\begin{equation*}
+    \mathcal{N}_2 = \varepsilon \sqcup
+    \decompmax{0/1,1/0}{\scriptstyle\mathcal{N}_2\setminus\varepsilon}{}{}{\mathcal{C}}
+    \sqcup
+    \decompmax{0/0,0/1,1/0}{}{}{}{\mathcal{F}}
+\end{equation*}
+We can decompose an avoider of \(m_1\) in \(\av{\perm{2,3,1}}\) around the leftmost point
+in a similar manner
+\begin{equation*}
+    \mathcal{N}_1 = \varepsilon \sqcup
+    \decompleft{0/1,1/0}{\scriptstyle\mathcal{N}_1\setminus\varepsilon}{}{}{\mathcal{C}}
+    \sqcup
+    \decompleft{0/0,0/1,1/0}{}{}{}{\mathcal{F}}
+\end{equation*}
+These two decompositions tell us that these two patterns are Wilf-equivalent
+and have generating function
+\begin{equation}
+    N_1 = 1 + x(N_1(x)-1)C(x) +xF(x)
+    \label{eqn:ngen}
+\end{equation}
+Now consider an avoider of \(m_4\) decomposed around the maximum
+\begin{equation*}
+    \mathcal{N}_4 = \varepsilon \sqcup
+    \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\mathcal{O}}
+\end{equation*}
+Where \(\mathcal{O}\) is the permutations avoiding \(\perm{2,3,1},m_4\) and
+\(p = \textpattern{}{1}{0/0,0/1,1/1}\) since if the subsequence in this box
+were to start with the maximum then this point and the maximum would create
+an occurrence of \(m_4\).
+Now consider decomposition of a permutation in \(\mathcal{O}\) around its
+leftmost point.
+\begin{equation*}
+    \mathcal{O} = \varepsilon \sqcup
+    \decompleft{0/1,1/0}{\mathcal{C}}{}{}{\scriptstyle\mathcal{N}_4\setminus\varepsilon}
+\end{equation*}
+This gives us the generating function
+\begin{equation*}
+    O(x) = 1 + xC(x)(N_4(x)-1)
+\end{equation*}
+Now we can construct the following for \(N_4\)
+\begin{equation}
+    N_4(x) = 1 +xC(x)(xC(x)(N_4(x)-1)+1)
+    \label{eqn:ngen2}
+\end{equation}
+
+All that remains to show Wilf-equivalence is to show that \EquationRef{eqn:ngen} and
+\EquationRef{eqn:ngen2} are the same generating function.
+First solve \EquationRef{eqn:ngen2} for \(N_4(x)\)
+\begin{equation}
+    \begin{aligned}
+        N_4(x) &= 1+xC(x)(xC(x)(N_4(x)-1)+1)\\
+        &= 1 + x^2C^2(x)N_4(x) -x^2C^2(x) +xC(x)\\
+        &= 1 + \frac{xC(x)}{1 - x^2C^2(x)}\\
+        &= 1 + \frac{xC(x)}{(1-xC(x))(1+xC(x))} && \text{Difference of squares}\\
+        N_4(x)&= 1 + \frac{xC^2(x)}{1+xC(x)} && C(x) = \frac{1}{1-xC(x)}
+    \end{aligned}
+\end{equation}
+
+Now we solve \EquationRef{eqn:ngen} for \(N_1(x)\)
+\begin{equation}
+    \begin{aligned}
+        N_1(x)&= 1+x(N_1(x)-1)C(x)+xF(x)\\
+        &= 1+xN_1(x)C(x)-xC(x)+\frac{x}{1+x-xC(x)} &&\text{Substitution of \(F(x)\)}\\
+        N_1(x)(1-xC(x))&= \frac{x^2C^2(x) - (x^2+2x)C(x) + 2x+1}{1+x-xC(x)}\\
+        N_1(x) &= \frac{x^2C^2(x) - (x^2+2x)C(x) + x+1+x}{x^2C^2(x) - (x^2+2x)C(x) +x+1}\\
+        &= 1 +\frac{x}{x^2C^2(x) - (x^2+2x)C(x) +x+1}\\
+        &= 1 + \frac{x}{1-x^2C(x)-xC(x)} && xC^2(x) = C(x)-1\\
+        &= 1 + \frac{xC^2(x)}{C^2(x) - xC^3(x)(x+1)}\\
+        &= 1 + \frac{xC^2(x)}{C(x) - xC^2(x) + xC(x)} && xC^2(x) = C(x)-1\\
+        N_1(x)&= 1 + \frac{xC^2(x)}{1+xC(x)} && C(x) = 1+xC^2(x)\\
+    \end{aligned}
+\end{equation}
+We have shown that \(N_1\) and \(N_4\) are indeed the same generating function,
+and we have that the classes containing these four patterns are Wilf-equivalent.
+
 \subsection{}
 \begin{equation*}
-    p_1 = \pattern{}{1,2}{1/0,1/1,1/2,2/0,2/1} \text{ and }
-    p_2 = \pattern{}{2,1}{0/1,1/0,1/1,1/2,2/1}
+    m_1 = \pattern{}{1,2}{1/0,1/1,1/2,2/0,2/1} \text{ and }
+    m_2 = \pattern{}{2,1}{0/1,1/0,1/1,1/2,2/1}
 \end{equation*}
-Let \(\mathcal{N}_1\) be the set of avoiders of \(p_1\), by structural decomposition
+Let \(\mathcal{P}_1\) be the set of avoiders of \(m_1\), by structural decomposition
 around the leftmost point we have
 \begin{equation*}
-    \mathcal{N}_1 = \varepsilon \sqcup
-    \decompleft{0/1,1/0}{\mathcal{N}_1}{}{}{\mathcal{N}^\prime_1}
+    \mathcal{P}_1 = \varepsilon \sqcup
+    \decompleft{0/1,1/0}{\mathcal{P}_1}{}{}{\mathcal{P}^\prime_1}
 \end{equation*}
-Where \(\mathcal{N}^\prime_1\) is a permutation avoiding \(\perm{2,3,1}, m_1\)
+Where \(\mathcal{P}^\prime_1\) is a permutation avoiding \(\perm{2,3,1}, m_1\)
 and \(\textpattern{}{1}{0/0,0/1,1/1}\)
-Now consider the decomposition of a permutation in \(\mathcal{N}^\prime_1\) it can
+Now consider the decomposition of a permutation in \(\mathcal{P}^\prime_1\) it can
 once again be decomposed around the leftmost point
 \begin{equation*}
-    \mathcal{N}^\prime_1 = \varepsilon \sqcup
-    \decompleft{0/1,1/0}{\scriptstyle \mathcal{N}_1\setminus\varepsilon}{}{}{\mathcal{N}^\prime_1}
+    \mathcal{P}^\prime_1 = \varepsilon \sqcup
+    \decompleft{0/1,1/0}{\scriptstyle \mathcal{P}_1\setminus\varepsilon}{}{}{\mathcal{P}^\prime_1}
 \end{equation*}
-This is a complete decomposition of avoiders of \(p_1\). Now we
-look at an avoider of \(p_2\), this time decomposition is around the maximum
+This is a complete decomposition of avoiders of \(m_1\). Pow we
+look at an avoider of \(m_2\), this time decomposition is around the maximum
 \begin{equation*}
-    \mathcal{N}_2 = \varepsilon \sqcup
-    \decompleft{0/1,1/0}{\mathcal{N}_2}{}{}{\mathcal{N}^\prime_2}
+    \mathcal{P}_2 = \varepsilon \sqcup
+    \decompleft{0/1,1/0}{\mathcal{P}_2}{}{}{\mathcal{P}^\prime_2}
 \end{equation*}
 
-Again we use the same method of decomposition of a permutation in \(\mathcal{N}^\prime_2\)
+Again we use the same method of decomposition of a permutation in \(\mathcal{P}^\prime_2\)
 \begin{equation*}
-    \mathcal{N}^\prime_2 = \varepsilon \sqcup
-    \decompleft{0/1,1/0}{\scriptstyle \mathcal{N}_2\setminus\varepsilon}{}{}{\mathcal{N}^\prime_2}
+    \mathcal{P}^\prime_2 = \varepsilon \sqcup
+    \decompleft{0/1,1/0}{\scriptstyle \mathcal{P}_2\setminus\varepsilon}{}{}{\mathcal{P}^\prime_2}
 \end{equation*}
 
-This gives us a generating function \(N(x)\) satisfying
+This gives us a generating function \(P(x)\) satisfying
 \begin{align}
-    N(x) &= 1 + xN(x)N^\prime(x) \label{eqn:Ngen}\\
-    N(x) &= 1 + x(N(x)-1)N^\prime(x)\label{eqn:NprimeGen}
+    P(x) &= 1 + xP(x)P^\prime(x) \label{eqn:Pgen}\\
+    P(x) &= 1 + x(P(x)-1)P^\prime(x)\label{eqn:PprimeGen}
 \end{align}
 
-Solving \EquationRef{eqn:NprimeGen} for \(N^\prime(x)\) and substituting into
-\EquationRef{eqn:Ngen} gives us the fact that a the generating function for
-\(N(x)\) satisfies
+Solving \EquationRef{eqn:PprimeGen} for \(P^\prime(x)\) and substituting into
+\EquationRef{eqn:Pgen} gives us the fact that a the generating function for
+\(P(x)\) satisfies
 \begin{equation}
-    N(x) = xN^2(x) - x(N(x) - 1) + 1
+    P(x) = xP^2(x) - x(P(x) - 1) + 1
 \end{equation}
+
+\subsection{}
+\begin{gather*}
+        \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,1/2,2/2},
+        \pattern{}{1,2}{0/1,0/2,1/0,1/1,1/2,2/0,2/2},
+        \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,1/2,2/1},
+        \pattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0,2/1},
+        \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,1/2,2/0},\\
+        \pattern{}{1,2}{0/0,0/1,0/2,1/2,2/0,2/1,2/2},
+        \pattern{}{1,2}{0/0,0/1,0/2,1/1,1/2,2/0,2/2},
+        \pattern{}{1,2}{0/0,0/1,1/0,1/1,1/2,2/0,2/1},
+        \pattern{}{1,2}{0/0,0/1,1/0,1/1,1/2,2/0,2/1}
+\end{gather*}
+
+In order to gain enumeration, consider decomposition of avoiders
+of \(\textpattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,1/2,2/2}\) around
+the maximum.
+\begin{equation*}
+    \mathcal{Q}_1 = \varepsilon \sqcup
+    \tikz{\filldraw (0,0) circle (3pt);}
+    \sqcup
+    \decompmax{0/1,1/0}{\mathcal{C\setminus\varepsilon}}{}{}{\mathcal{C}}
+\end{equation*}
+
+This gives us the following generating function
+\begin{equation}
+    Q(x) = 1 + x + xC(x)(C(x)-1) \label{eq:qgen}
+\end{equation}
+
+Now we consider decomposition of an avoider
+\textpattern{}{1,2}{0/1,0/2,1/0,1/1,1/2,2/0,2/2} around the maximum,
+this avoider must fit into the following form
+\begin{equation*}
+    \mathcal{Q}_2 = \varepsilon \sqcup
+    \decompmax{0/1,1/0}{\mathcal{G}_2}{}{}{\mathcal{C}}
+\end{equation*}
+This gives us the generating function
+\begin{equation*}
+    \begin{aligned}
+        Q_2(x) &= 1+xC(x)G(x)\\
+        &= 1+xC(x)(1+xC(x)(C(x)-1))\\
+        &= 1 +xC(x)(C(x)-xC(x)) && C(x) = 1+xC^2(x)\\
+        &= 1 +x + xC^2(x) - xC(x) \\
+        &= 1 +x + xC(x)(C(x)-1)
+    \end{aligned}
+\end{equation*}
+Therefore this generating function is the same as \EquationRef{eq:qgen}.
+We can decompose
+\textpattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,1/2,2/1},
+\textpattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0,2/1},
+\textpattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,1/2,2/0},
+\textpattern{}{1,2}{0/0,0/1,0/2,1/2,2/0,2/1,2/2},
+and \textpattern{}{1,2}{0/0,0/1,0/2,1/1,1/2,2/0,2/2} around the leftmost point
+into an avoider of one of the patterns with generating function \(G(x)\) and
+an avoider of \(\perm{2,3,1}\).
+
+Now decompose an avoider of \textpattern{}{1,2}{0/0,0/1,1/0,1/1,1/2,2/0,2/1}
+around the leftmost point.
+\begin{equation*}
+    \mathcal{Q}_3 = \varepsilon \sqcup
+    \decompleft{0/1,1/0}{\scriptstyle \mathcal{Q}\setminus\varepsilon}{}{}{\mathcal{C}}
+    \sqcup \decompleft{0/0,0/1,1/0}{}{}{}{\mathcal{G}_3}
+\end{equation*}
+This gives generating function \(Q_3(x)\) satisfying
+\begin{equation*}
+    \begin{aligned}
+        Q(x) &= 1 +xC(x)(P(x)-1) + xG(x)\\
+        &= C(x) - xC^2(x) + xC(x) +x^2C^3(x) - x^2C^2 && \text{Solving for \(Q(x)\)}\\
+        &= 1 + xC(x) +x^2C^3(x)-xC^2(x) && C = 1+xC^2(x)\\
+        &= 1 + x + x^2C^3(x) && xC^2(x) = C(x)-1\\
+        &= 1 + x + xC(x)(C(x)-1)
+    \end{aligned}
+\end{equation*}
+This is equivalent to \EquationRef{eq:qgen}, and therefore
+the classes containing all of these patterns are Wilf-equivalent.

### 2016-04-20 src/helper_funcs.py

@@ -2,6 +2,7 @@ import json
 from permuta import *
 import itertools
 from math import factorial
+import functools
 
 def load_set(filename):
     patts = []
@@ -115,4 +116,16 @@ def partition(perm):
         last = value
     return partition
 
+def meshpatt_to_tex(mp):
+    cj = ",".join
+    stm = functools.partial(map,str)
+    pstr = cj(stm(mp.perm))
+    mstr = cj("/".join(stm(pair)) for pair in mp.mesh)
+    outstr = "\\textpattern{{}}{{{}}}{{{}}}".format(pstr, mstr)
+    return outstr
 
+def pad_sequence(length, seq):
+    seql = list(seq)
+    while len(seql) < length:
+        seql.append(0)
+    return tuple(seql)

### 2016-04-20 src/main.py

@@ -78,7 +78,7 @@ wreprs321 = set(wreprs321_12 + wreprs321_21)
 wilfcounters231 = [WilfCounter(Permutation([2,3,1]), mperm) for mperm in wreprs231]
 wilfcounters321 = [WilfCounter(Permutation([3,2,1]), mperm) for mperm in wreprs321]
 
-for permlen in range(10):
+for permlen in range(11):
     for perm in PermutationsAvoiding231(permlen):
         for counter in wilfcounters231:
             counter.modify(perm)

### 2016-04-21 writing/latex/coincidence.tex
Add references, add appendages

@@
 We know that there are a total of \(512\) mesh-sets for each underlying classical
-pattern. By use of the shading lemma, simultaneous shading lemma, and
+pattern. By use of the previous results of the Shading
+Lemma\cite[Lemma~3.11]{DBLP:journals/combinatorics/HilmarssonJSVU15},
+Simultaneous Shading Lemma\cite[Lemma~7.6]{DBLP:journals/corr/ClaessonTU14}, and
 one special case, we can reduce the number of equivalence classes to \(220\).
@@
 \end{tikzpicture}
-    \caption{An illustration of how lines and points move by this proposition.}
+    \caption{If the conditions of \PropositionRef{prop:dom2} are satisfied the box
+    \((a-1,b)\) can be shaded.}
 \end{figure}
@@
 \end{proof}
+
+By application of this rule we can reduce the number of classes in
+\(\av{\{\perm{2,3,1},(\perm{1,2}, R)\}}\) to \(56\).
 

### 2016-04-21 writing/latex/introduction.tex
Add references, add appendages

@@
 \section{What is a Permutation?}
-In \emph{The Art of Computer Programming} Donald Knuth defines
+In \emph{The Art of Computer Programming}\cite[p.~45]{Knuth:1997:ACP:260999} Donald Knuth defines
 A \emph{permutation of \(n\) objects} is an arrangement of \(n\) distinct
@@
 \end{equation*}
+It is convenient to call the elements of the permutation \emph{points} when
+referring to these figures.
 
@@
 Classical Permutation Patterns began to be studied as a result of Knuth's
-statements about stack-sorting in [The Art of Computer Programming].
+statements about stack-sorting in \emph{The Art of Computer Programming}
+\cite[p.~243, Ex.~5,6]{Knuth:1997:ACP:260999}.
 
@@
 \section{Mesh Patterns}
-Mesh Patterns were introduced by Brändén and Claesson to capture explicit expansions
+Mesh Patterns were introduced by Brändén and Claesson\cite{journals/combinatorics/BrandenC11} to capture explicit expansions
 for certain permutation statistics. They are a natural extension of Classical

### 2016-04-21 src/scratchpad.py

@@ -1,0 +1,37 @@
+f = sys.stdout
+with open("classes.tex", "w") as f:
+    f.write("\\chapter{Equivalence classes of mesh patterns}\n")
+    f.write("\\section{Equivalence classes with classical pattern 12}\n")
+    f.write("\\begin{center}")
+    for classnum,thisclass in enumerate(classes):
+        f.write("\\begin{equation}\n\t\\begin{gathered}\n")
+        for index, texstr in enumerate(map(meshpatt_to_tex, thisclass)):
+            f.write("\t\t")
+            f.write(texstr)
+            if index % 10 == 9:
+                f.write("\\\\")
+            f.write("\n")
+        f.write("\t\\end{{gathered}}\n\t\\label{{cls:{}}}\n\end{{equation}}\n\n".format(classnum))
+    f.write("\\end{center}")
+
+sorted(map(list,sorted(map(reversed, (pad_sequence(9,x) for x in  wilfcounts231)))))
+
+write_to_clipboard(("\n".join((("{:5d},"*9).format(*reversed(seq)))[:-1] for seq in sorted(map(list,sorted(map(reversed, (pad_sequence(9,x) for x in  wilfcounts231))))))))
+
+def number_of_patts(seq, wilfcounters, mpdu, mpdd,  setup, setdown, uclasses, dclasses):
+    wilfreps = [f.mperm for f in filter(lambda x: x.record == dict(enumerate(a)), wilfcounters)]
+    uperms = filter(lambda x: x.perm == Permutation([1,2]), wilfreps)
+    dperms = filter(lambda x: x.perm == Permutation([2,1]), wilfreps)
+    ucl = [mpdu[x] for x in uperms]
+    dcl = [mpdd[x] for x in dperms]
+    if len(ucl) >0:
+        ucls = frozenset.union(*(find_containing_set(setup, num) for num in ucl))
+        usum = sum(len(uclasses[l]) for l in ucls)
+    else: usum = 0
+    if len(dcl) > 0:
+        dcls = frozenset.union(*(find_containing_set(setdown, num) for num in dcl))
+        dsum = sum(len(dclasses[l]) for l in dcls)
+    else:dsum = 0
+    return usum + dsum
+
+number_of_patts(a, wilfcounters231, meshpd_12, meshpd_21, pre_test_wilf231_12, pre_test_wilf231_21, classes, classes2)

### 2016-04-22 writing/latex/abstracteng.tex
Add abstract starting

A permutation is an arrangement of \(n\) objects.
Two mesh patterns are coincident if they are avoided by the same set of
permutations. In this thesis, the author provides sufficient conditions for
coincidence among mesh patterns, whilst also avoiding a longer classical
pattern. These conditions, along with two special cases are used to completely
classify coincidence amongst families containing a mesh pattern of length 2
and a classical pattern of length 3. The author then goes on to completely
classify Wilf-equivalence amongst mesh patterns of length 2 when we avoid the
classical pattern 231.

### 2016-04-22 writing/latex/confut.tex
Add abstract starting

\chapter{Conclusions and Future work}
From \ChapterRef{chap:coincs} it is can be seen that automatically classifying
coincidences of mesh patterns is a difficult task, establishing rules for longer
dominating patterns requires many more cases to be taken.

### 2016-04-22 writing/latex/wilf.tex
Add references, add appendages
Add abstract starting

@@
 the permutation \(\perm{1}\). The avoiders of this pattern therefore
-have generating function \(E(x) = C(x)-x\).
+have generating function \(\var(x) = C(x)-x\).\nextvar
 
@@
 \begin{equation*}
-    \mathcal{F} = \varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{F}}{}{}{\mathcal{C}\setminus\varepsilon}
+    \mathcal{\var} = \varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{\var}}{}{}{\mathcal{C}\setminus\varepsilon}
 \end{equation*}
@@
 of \(p_1\) and therefore must also avoid \(p_1\) as well as \(\perm{2,3,1}\).
-This gives the generating function of avoiders to be the function \(F(x)\)
+This gives the generating function of avoiders to be the function \(\var(x)\)
 satisfying.
 \begin{align*}
-    F(x) &= 1 + xF(x)(C(x)-1) \\
-    \shortintertext{Solving for \(F\) gives}
-    F(x) &= \frac{1}{1+x-xC(x)}
+    \var(x) &= 1 + x\var(x)(C(x)-1) \\
+    \shortintertext{Solving for \(\var\) gives}
+    \var(x) &= \frac{1}{1+x-xC(x)}
 \end{align*}
+
+Evauluation of this generating function gives the Fine numbers
+(\href{https://oeis.org/A000957}{OEIS: A000957}). This pattern is one of the
+quadrant marked mesh patterns studied by Remmel, Kitaev and TiefenbruckZ\cite{2012arXiv1201.6243K}.
 
@@
 \textpattern{}{1}{0/0,1/1} and \(p_2 = \textpattern{}{1}{0/0,0/1,1/1}\) are
-set-equivalent. Consider the decomposition of an avoider of \(p_2\) in
-\(\av{\perm{2,3,1}}\) around the maximum
+coincident. Consider the decomposition of an avoider of \(p_2\) in
+\(\av{\perm{2,3,1}}\) around the maximum\nextvar[\varullr]
 \begin{equation*}
-    \mathcal{G}_1 =\varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
+    \mathcal{\var}_1 =\varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
 \end{equation*}
@@
 \begin{equation*}
-    \mathcal{G}_2 =\varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\mathcal{C}\setminus\varepsilon}
+    \mathcal{\var}_2 =\varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\mathcal{C}\setminus\varepsilon}
 \end{equation*}
@@
 \begin{equation*}
-    \mathcal{G}_3 =\varepsilon \sqcup \decompleft{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
+    \mathcal{\var}_3 =\varepsilon \sqcup \decompleft{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
 \end{equation*}
@@
 \begin{equation*}
-    G(x) = 1 + xC(x)(C(x)-1)
+    \var(x) = 1 + xC(x)(C(x)-1)
 \end{equation*}
@@
     in \(\mathfrak{S}_n\) are in one-to-one correspondence with partitions of
-    \(\nrange{n}\). [Anders, Generalized Pattern Avoidance]
+    \(\nrange{n}\). \cite{DBLP:journals/ejc/Claesson01}
 \end{note}
@@
 We are looking at permutations in \(\av{\perm{2,3,1}}\), all of these permutations
-also avoid the vincular pattern in \NoteRef{not:setequiv}.
+also avoid the mesh pattern in \NoteRef{not:setequiv}.
 

### 2016-04-23 writing/latex/coincidence.tex
Getting closer

@@
-\chapter{Coincidences amongst families of mesh patterns and classical patterns}\label{chap:coincs}
+\chapter[Coincidence Classification]{Coincidences amongst families of mesh patterns and classical patterns}\label{chap:coincs}
 One interesting question to ask about permutation patterns considers when a
@@
 one special case, we can reduce the number of equivalence classes to \(220\).
+
+By discussion of a number of rules we will show that the number of coincidence
+classes follows the values shown in  \TableRef{tab:domclasses}.
+
+\begin{table}[htb]
+\begin{center}
+\begin{tabularx}{\textwidth}{c|Y|Y|Y|Y|}
+\cline{2-5}
+& \multicolumn{4}{|c|}{Dominating Pattern}\\
+\cline{2-5}
+& \multicolumn{2}{|c|}{\(\perm{2,3,1}\)}&\multicolumn{2}{|c|}{\(\perm{3,2,1}\)}\\
+\cline{2-5}
+& \(\perm{1,2}\)&\(\perm{2,1}\)&\(\perm{1,2}\)&\(\perm{2,1}\)\\
+\hline
+\multicolumn{1}{|r|}{No Dominating rule}&\(220\)&\(220\)&\(220\)&\(220\)\\
+\hline
+\multicolumn{1}{|r|}{First Dominating rule}&\(85\)&\(43\)&\(220\)&\(29\)\\
+\hline
+\multicolumn{1}{|r|}{Second Dominating rule}&\(59\)&\(39\)&\(220\)&\(29\)\\
+\hline
+\multicolumn{1}{|r|}{Third Dominating rule}&\(56\)&\(39\)&\(220\)&\(29\)\\
+\hline
+\multicolumn{1}{|r|}{Experimental class size}&\(56\)&\(39\)&\(213\)&\(29\)\\
+\hline
+\end{tabularx}
+\end{center}
+    \caption{Class number reduction by application of Dominating Rules}
+    \label{tab:domclasses}
+\end{table}
 

### 2016-04-23 writing/latex/confut.tex
Getting closer

@@
 coincidences of mesh patterns is a difficult task, establishing rules for longer
-dominating patterns requires many more cases to be taken.
+dominating patterns requires many more cases to be taken. It would however
+be interesting to consider the self application of the Dominating Rules to mesh
+patterns in order to try to capture some of the coincidences described in
+\textcite{DBLP:journals/combinatorics/HilmarssonJSVU15} and \textcite{journals/combinatorics/BrandenC11}.
+
+It is also possible to take sets of mesh patterns instead of a single mesh pattern
+when considering dominating rules, and expressing coincidence between these sets.
+Doing this may give nice enumerative results.
+
+It would be interesting to consider a systematic explanation of Wilf-equivalences
+amongst classes where \(\perm{3,2,1}\) is the dominating pattern using the
+construction presented in \cite[Sec.~12]{2015arXiv151203226B}, in order to directly
+reach enumeration and hopefully establish some of the non-trivial Wilf-equivalences
+between classes with different dominating patterns.

### 2016-04-23 writing/latex/introduction.tex
Getting closer

@@
 
+Knuth's statements were exercises in showing that the permutations avoiding the
+pattern \(\perm{2,3,1}\) completely categorise permutations that are sortable to
+the identity permutation using only a single stack, and that permutations avoiding the
+pattern \(\perm{3,2,1}\) completely categorise permutations that are sortable to
+the identity permutation using only a single queue with bypass.
+
 \section{Mesh Patterns}
-Mesh Patterns were introduced by Brändén and Claesson\cite{journals/combinatorics/BrandenC11} to capture explicit expansions
+Mesh Patterns were introduced by \textcite{journals/combinatorics/BrandenC11} to capture explicit expansions
 for certain permutation statistics. They are a natural extension of Classical
@@
 
+In the past people have studied different classes
+of permutations that can be described by mesh patterns. \textcite{babstein2000}
+considered \emph{vincular} patterns (also known as \emph{generalised} or \emph{dashed} patterns),
+those where two adjacent entries in the pattern must be adjacent in the permutation,
+\ie entire columns are shaded in the mesh.
+\textcite{MR2652101} look at classes of pattern where both columns and rows
+can be shaded, these are called \emph{bivincular} patterns.
+\emph{Bruhat-restricted} patterns were studied by \textcite{MR2264071} in order
+to establish necessary conditions for a Schubert variety to be Gorenstein.
+Mesh patterns also encompass a subset of \emph{barred} patterns introduced by
+\textcite{MR2716312}.
+
+Avoiding pairs of patterns of the same length with certain properties has
+also been studied in the past, \textcite{MR2178749} considered avoiding a
+pair of vincular patterns of length 3. \textcite{2015arXiv151203226B} study avoiding
+a vincular and a covincular pattern simultaneously in order to achieve some
+interesting counting results. However, very little work has been done on avoiding
+a mesh pattern and a classical pattern simultaneously, in this work we aim to
+establish some ground in this field by computing coincidences and Wilf-classes
+and calculating some of the enumerations of avoiders of a mesh pattern of length
+2 and a classical pattern of length 3.
+
 %%% Local Variables:

### 2016-04-23 src/scratchpad.py

@@ -18,6 +18,8 @@ sorted(map(list,sorted(map(reversed, (pad_sequence(9,x) for x in  wilfcounts231)
 
 write_to_clipboard(("\n".join((("{:5d},"*9).format(*reversed(seq)))[:-1] for seq in sorted(map(list,sorted(map(reversed, (pad_sequence(9,x) for x in  wilfcounts231))))))))
 
+write_to_clipboard(("\\\n".join((("{:5d},"*11 + " \&\t\& {classize}").format(*x[0],classize=x[1])) for x in f )))
+
 def number_of_patts(seq, wilfcounters, mpdu, mpdd,  setup, setdown, uclasses, dclasses):
     wilfreps = [f.mperm for f in filter(lambda x: x.record == dict(enumerate(a)), wilfcounters)]
     uperms = filter(lambda x: x.perm == Permutation([1,2]), wilfreps)
@@ -34,4 +36,11 @@ def number_of_patts(seq, wilfcounters, mpdu, mpdd,  setup, setdown, uclasses, dc
     else:dsum = 0
     return usum + dsum
 
-number_of_patts(a, wilfcounters231, meshpd_12, meshpd_21, pre_test_wilf231_12, pre_test_wilf231_21, classes, classes2)
+import subprocess
+
+def write_to_clipboard(output):
+    process = subprocess.Popen(
+        'pbcopy', env={'LANG': 'en_US.UTF-8'}, stdin=subprocess.PIPE)
+    process.communicate(output.encode('utf-8'))
+
+number_of_patts(a, wilfcounters321, meshpd_12, meshpd_21, pre_test_wilf321_12, pre_test_wilf321_21, classes, classes2)

### 2016-04-28 writing/latex/confut.tex
Add enumerations

@@
 dominating patterns requires many more cases to be taken. It would however
-be interesting to consider the self application of the Dominating Rules to mesh
+be interesting to consider the self application of the  third Dominating Rule to mesh
 patterns in order to try to capture some of the coincidences described in
 \textcite{DBLP:journals/combinatorics/HilmarssonJSVU15} and \textcite{journals/combinatorics/BrandenC11}.
+It is not possible to apply the first and second Dominating Rules to the pattern
+itself, since when applying the rules we consider containers of the pattern inside
+the avoiders of the dominating pattern. For example if we were to attempt to apply
+the first rule to the pattern \(\perm{1,2}\) then we would have to consider
+containers of \(\perm{1,2}\) inside \(\av{\perm{1,2}}\), and obviously this can
+never occur.
 
@@
 reach enumeration and hopefully establish some of the non-trivial Wilf-equivalences
-between classes with different dominating patterns.
+between classes with different dominating patterns. It is possible to show that
+the sets consisting of \textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1} and
+\(\perm{2,3,1}\), or \(\perm{3,2,1}\), are Wilf-equivalent.

### 2016-04-28 writing/latex/wilf.tex
Getting closer
Add enumerations

@@
-\chapter{Wilf-equivalences under dominating  patterns}
+\chapter[Wilf-Classification]{Wilf-equivalences under dominating  patterns}
 
@@
 
-Evauluation of this generating function gives the Fine numbers
-(\href{https://oeis.org/A000957}{OEIS: A000957}). This pattern is one of the
-quadrant marked mesh patterns studied by Remmel, Kitaev and TiefenbruckZ\cite{2012arXiv1201.6243K}.
+Evauluation of this generating function gives a the Fine numbers
+(\href{https://oeis.org/A000957}{OEIS: A000957}).
+\begin{equation*}
+    1, 0, 1, 2, 6, 18, 57, 186, 622, 2120, 7338,\dotsc
+\end{equation*}
+This pattern is one of the
+quadrant marked mesh patterns studied by Remmel, Kitaev and Tiefenbruck\cite{2012arXiv1201.6243K}.
 
@@
 \end{equation*}
+Enumeration of this generating function gives
+\begin{equation*}
+    1, 0, 1, 3, 9, 28, 90, 297, 1001, 3432, 11934,\dotsc
+\end{equation*}
 
@@
 \begin{equation*}
-    H(x) = 1 + xC(x)(C(x)-1) + xH(x)
+    H(x) = 1 + xC(x)(C(x)-1) + x(H(x)-1)
+\end{equation*}
+This generating function can be enumerated to give
+\begin{equation*}
+    	1, 0, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc
 \end{equation*}
@@
 avoiders of length \(k\) for all \(k\) and therefore all Wilf-equivalent.
+The avoiders have enumeration
+\begin{equation*}
+    1, 1, 1, 3, 11, 38, 127, 423, 1423, 4854, 16787,\dotsc
+\end{equation*}
 
@@
 \end{equation*}
+This can be enumerated to give the sequence
+\begin{equation*}
+    1, 1, 1, 2, 6, 19, 61, 200, 670, 2286, 7918,\dotsc
+\end{equation*}
+This corresponds to OEIS sequence number \href{https://oeis.org/A035929}{A035929}
+with offset 1.
 
@@
 \(2\) where there is precisely \(1\) avoider.
+Therefore there are
+\begin{equation*}
+    1, 1, 1, 5, 14, 42, 132, 429, 1430, 4862, 16796,\cdots
+\end{equation*}
 
@@
 \end{equation*}
+This generating function gives
+\begin{equation*}
+    1, 1, 1, 4, 11, 34, 108, 354, 1187, 4054, 14054,\dotsc
+\end{equation*}
 
@@
 for \(n \ge 2\) where \(C_n\) is the \(n\)th Catalan number, the number of
-\(\perm{2,3,1}\) avoiders of length \(n\).
+\(\perm{2,3,1}\) avoiders of length \(n\). This gives the sequence
+\begin{equation*}
+    1, 1, 1, 4, 12, 37, 118, 387, 1298, 4433, 15366,\dotsc
+\end{equation*}
 
@@
 length \(n\) for all \(n \ge 2\).
+This gives the sequence
+\begin{equation*}
+    1, 1, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc\tag{\href{https://oeis.org/A001453}{OEIS: A001453} offset 2}
+\end{equation*}
 
@@
 two do not interact in any manner. The containers of this pattern are counted
-the same as the other patterns, and due to this they are Wilf-equivalent.
+the same as the other patterns, and due to this they are Wilf-equivalent. The containers
+have generating function \(\frac{x^2C(x)}{(1-x)}\). Enumerating avoiders
+therefore gives us
+\begin{equation*}
+    1, 1, 1, 3, 10, 33, 109, 364, 1233, 4236, 14740,\dotsc
+\end{equation*}
 
@@
 function so all four classes are Wilf-equivalent.
+
+Evaluating the generating function \(L(x)\) gives us the enumeration
+\begin{equation*}
+    1, 1, 1, 2, 6, 20, 68, 233, 805, 2807, 9879,\dotsc
+\end{equation*}
+
 \subsection{}
@@
 \end{equation*}
-We have that \(M_3(x) = 1+xC(x)= M(x) \) so all four patterns are Wilf-equivalent.
+We have that \(M_3(x) = 1+xC(x)= M(x) \) so all four patterns are Wilf-equivalent
+ and have enumeration sequence
+ \begin{equation*}
+     1, 1, 1, 2, 5, 14, 42, 132, 429, 1430, 4862,\dotsc
+ \end{equation*}
+ This is an offset of the Catalan numbers.
 
@@
 and we have that the classes containing these four patterns are Wilf-equivalent.
+Evaluating the generating function \(N(x)\) gives
+\begin{equation*}
+    1, 1, 1, 3, 8, 24, 75, 243, 808, 2742, 9458,\dotsc \tag{\href{https://oeis.org/A001453}{OEIS: A001453}}
+\end{equation*}
 
@@
     P(x) &= 1 + xP(x)P^\prime(x) \label{eqn:Pgen}\\
-    P(x) &= 1 + x(P(x)-1)P^\prime(x)\label{eqn:PprimeGen}
+    P^\prime(x) &= 1 + x(P(x)-1)P^\prime(x)\label{eqn:PprimeGen}
 \end{align}
@@
 \end{equation}
-
+Evaluating \(P(x)\) gives us the sequence
+\begin{equation*}
+    1, 1, 1, 2, 4, 9, 21, 51, 127, 323, 835,\dotsc
+\end{equation*}
+This is an offset of the Motzkin numbers \href{https://oeis.org/A001006}{(OEIS: A001006)}.
 \subsection{}
@@
 the classes containing all of these patterns are Wilf-equivalent.
+The classes have enumeration
+\begin{equation*}
+    1, 1, 1, 3, 9, 28, 90, 297, 1001, 3432, 11934,\dotsc
+\end{equation*}
+This is an offset of OEIS Sequence number \href{http://oeis.org/A071724}{A071724}.

### 2016-04-28 src/scratchpad.py

@@ -44,3 +44,17 @@ def write_to_clipboard(output):
     process.communicate(output.encode('utf-8'))
 
 number_of_patts(a, wilfcounters321, meshpd_12, meshpd_21, pre_test_wilf321_12, pre_test_wilf321_21, classes, classes2)
+
+def patts_with_sequence(seq, wilfcounters, mpdu, mpdd,  setup, setdown, uclasses, dclasses):
+    wilfreps = [f.mperm for f in filter(lambda x: x.record == dict(enumerate(a)), wilfcounters)]
+    uperms = filter(lambda x: x.perm == Permutation([1,2]), wilfreps)
+    dperms = filter(lambda x: x.perm == Permutation([2,1]), wilfreps)
+    ucl = [mpdu[x] for x in uperms]
+    dcl = [mpdd[x] for x in dperms]
+    if len(ucl) >0:
+        ucls = frozenset.union(*(find_containing_set(setup, num) for num in ucl))
+    else:ucls=frozenset()
+    if len(dcl) > 0:
+        dcls = frozenset.union(*(find_containing_set(setdown, num) for num in dcl))
+    else:dcls=frozenset()
+    return [list(ucls),list(dcls)]

### 2016-04-29 writing/latex/coincidence.tex
Getting closer

@@
 of permutations that avoid \(\pi\) is the same as the set of permutations that
-avoid \(\sigma\). This extends to sets of patterns as well as single patterns.
+avoid \(\sigma\) \ie \(\av{\pi} = \av{\sigma}\). This extends to sets of patterns as well as single patterns.
 
@@
 establish some rules about when these two sets give the same avoidance set.
-We fix \(p\) in order to define the equivalences, we say that \(p\) is the
-\emph{Dominating Pattern}. We fix \(p\) in order to calculate the coincidences.
-We never consider occurrences with different \emph{Dominating Patterns}.
+We fix \(p\) in order to define the equivalences and say that \(p\) is the
+\emph{dominating pattern}.
 
@@
 \begin{definition}
-Given a pattern \(p\), let \(\addp[D]{p}{(a,b)}\) be the operation that returns
-a mesh pattern equivalent to placing a point in the center of box \((a,b)\) in
-\(p\), with shading defined by \(D\subseteq\{N,E,S,W\}\).
-\end{definition}
+Given a mesh pattern \(p\), let \(\addp[D]{p}{(a,b)}\) be the operation that returns
+a mesh pattern equivalent to placing a point in the center of box \((a,b)\), where \((a,b)\) is not the
+mesh set of \(p\) in
+the figure of \(p\), with shading defined by \(D\subseteq\{N,E,S,W\}\).
 The set \(D\) defines the shading by indicating that the boxes in the cardinal
 directions in \(D\) next to the point are shaded in the resulting pattern.
+\end{definition}
 Since there is no ambiguity we let \(\addpe[D]\) be equivalent to \(\addp[D]{\varepsilon}{(0,0)}\).
-This operation fails if the box \((a,b)\) is in the mesh set of \(p\).
 
@@
 \end{example}
+Formally, given a mesh pattern \(p = (\pi,R)\) \(\addp[D]{p}{(a,b)}\) gives a mesh
+pattern \(p^\prime = (\pi^\prime,R^\prime)\) defined by
+\begin{equation*}
+    \pi^\prime(i) = \begin{cases}
+    \pi(i) & \text{if } i\neq a+1 \text{ and } \pi(i)<b\\
+    \pi(i)+1 & \text{if } i\neq a+1 \text{ and } \pi(i)>b\\
+    b+1 & \text{if } i = a+1
+\end{cases}
+\end{equation*}
+and
+\begin{equation*}
+        R^\prime = \bigcup_{(i,j)\in R}r((i,j))
+\end{equation*}
+Where \(r((i,j))\) is defined by
+\begin{equation*}
+    r((i,j)) =
+    \begin{cases}
+        \{(i,j)\} & \text{if } i<a, j<b\\
+        \{(i,j),(i,j+1)\} & \text{if } i<a, j=b\\
+        \{(i,j+1)\} & \text{if } i<a, j>b\\
+        \{(i,j),(i+1,j)\} & \text{if } i=a, j<b\\
+        \{(i,j+1),(i+1,j+1)\} & \text{if } i=a, j>b\\
+        \{(i+1,j)\} & \text{if } i>a, j<b\\
+        \{(i+1,j),(i+1,j+1)\} & \text{if } i>a, j=b\\
+        \{(i+1,j+1)\} &\text{if } i>a, j>b\\
+    \end{cases}
+\end{equation*}
+
+If the shading set \(D\) is non-empty we can modify definition of the directions
+slightly
+\begin{equation*}
+    \begin{aligned}
+        N &= \{(a,b+1),(a+1,b+1)\} \\
+        E &= \{(a+1,b),(a+1,b+1)\} \\
+        S &= \{(a,b),(a+1,b)\} \\
+        W &= \{(a,b),(a,b+1)\}
+    \end{aligned}
+\end{equation*}
+And we add the union of the sets in \(D\) into the mesh set \(R^\prime\).
 
@@
     as the operations that return a mesh pattern equivalent to placing an
-    decrease, or increase, in the center of box \((a,b)\) in \(p\).
+    decrease, or increase, in the center of box \((a,b)\), where \((a,b)\) is not
+    in the mesh set of \(p\), in \(p\).
 \end{definition}
@@
 \end{example}
+We can define these formally as with \(\addp[D]{p}{(a,b)}\). Considering
+only adding the ascent, adding a descent is very similar
+\begin{equation*}
+    \addi{p}{(a,b)} = (\pi^\prime, R^\prime)
+\end{equation*}
+\begin{equation*}
+    \pi^\prime(i) = \begin{cases}
+    \pi(i) & \text{if } i\neq a+1,a+2 \text{ and } \pi(i)<b\\
+    \pi(i)+2 & \text{if } i\neq a+1,a+2 \text{ and } \pi(i)>b\\
+    b+1, b+2 & \text{if } i = a+1, a+2
+\end{cases}
+\end{equation*}
+and
+\begin{equation*}
+        R^\prime = \bigcup_{(i,j)\in R}r((i,j)) \cup 
+\end{equation*}
+Where \(r((i,j))\) is defined by
+\begin{equation*}
+    r((i,j)) =
+    \begin{cases}
+        \{(i,j)\} & \text{if } i<a, j<b\\
+        \{(i,j),(i,j+1),(i,j+2)\} & \text{if } i<a, j=b\\
+        \{(i,j+2)\} & \text{if } i<a, j>b\\
+        \{(i,j),(i+1,j),(i+2,j)\} & \text{if } i=a, j<b\\
+        \{(i,j+2),(i+1,j+2),(i+2,j+2)\} & \text{if } i=a, j>b\\
+        \{(i+2,j)\} & \text{if } i>a, j<b\\
+        \{(i+2,j),(i+2,j+1),(i+2,j+2)\} & \text{if } i>a, j=b\\
+        \{(i+2,j+2)\} &\text{if } i>a, j>b\\
+    \end{cases}
+\end{equation*}
 
@@
 classical pattern, this is due to the fact that \(\perm{2,1}\in\av{\mperm{1,2}{R}}\)
-and \(\perm{1,2}\in\av{\mperm{2,1}{R}}\) for all mesh-sets \(R\). So
-\(\av{\mperm{1,2}{R}}\setminus \av{\mperm{2,1}{S}}\neq \emptyset\ \forall
-R,S\in [0,2]\times [0,2]\), and hence the sets are disjoint.
+and \(\perm{1,2}\in\av{\mperm{2,1}{R}}\) for all mesh-sets \(R\).
 
 We know that there are a total of \(512\) mesh-sets for each underlying classical
-pattern. By use of the previous results of the Shading
-Lemma\cite[Lemma~3.11]{DBLP:journals/combinatorics/HilmarssonJSVU15},
-Simultaneous Shading Lemma\cite[Lemma~7.6]{DBLP:journals/corr/ClaessonTU14}, and
-one special case, we can reduce the number of equivalence classes to \(220\).
+pattern. By use of the previous results of \textcite{DBLP:journals/corr/ClaessonTU14}\footnote{
+The authors use the Simultaneous Shading Lemma, a closure result and one worked out special case.
+} the number of coincidence classes can be reduced to \(220\).
+% the Shading
+% Lemma\cite[Lemma~3.11]{DBLP:journals/combinatorics/HilmarssonJSVU15},
+% Simultaneous Shading Lemma\textcite[Lemma~7.6]{DBLP:journals/corr/ClaessonTU14}, and
+% one special case, we can reduce the number of equivalence classes to \(220\).
 
 By discussion of a number of rules we will show that the number of coincidence
-classes follows the values shown in  \TableRef{tab:domclasses}.
+classes follows the values shown in  \TableRef{tab:domclasses}. The experimental
+data is calculated on permutations up to length 11.
 
@@
 
-\section{Equivalence classes of Av(\{321, (21, \textit{R})\}).}
-Through experimentation we discover that there are a total of 29 equivalence
+\section{Coincidence classes of Av(\{321, (21, \textit{R})\}).}
+Through experimentation we discover that there are at most 29 coincidence
 classes of mesh patterns with underlying classical pattern \(\perm{2,1}\).
@@
 \end{note}
-\begin{proof}
-    Assume without meaningful loss of generality that \(R^\prime\) is a proper
-    subset of \(R\).
-
-    Consider an occurence of \((\tau, R)\) in a permutation \(\sigma\), obviously
-    this corresponds to an occurrence of \(\tau\) in \(\sigma\). Now consider
-    the mesh sets \(R\) and \(R^\prime\), since \(R^\prime \subseteq R\) then
-    there are more restrictions on where points are in an occurrence of \((\tau,R)\).
-    Namely, for every shaded box in \(R\) the corresponding region in \(\sigma\)
-    must contain no points, since \(R^\prime\) has less shading than \(R\) there
-    exists a region in the occurence of \((\tau,R)\) in \(\sigma\) that is
-    now devoid of restrictions. However, by removing restrictions we cannot
-    make an occurence become not an occurence, and therefore the same occurrence
-    of \(\tau \) in \(\sigma\) is now an occurence of \((\tau, R^\prime)\).
-\end{proof}
+% \begin{proof}
+%     Assume without meaningful loss of generality that \(R^\prime\) is a proper
+%     subset of \(R\).
+%
+%     Consider an occurence of \((\tau, R)\) in a permutation \(\sigma\), obviously
+%     this corresponds to an occurrence of \(\tau\) in \(\sigma\). Now consider
+%     the mesh sets \(R\) and \(R^\prime\), since \(R^\prime \subseteq R\) then
+%     there are more restrictions on where points are in an occurrence of \((\tau,R)\).
+%     Namely, for every shaded box in \(R\) the corresponding region in \(\sigma\)
+%     must contain no points, since \(R^\prime\) has less shading than \(R\) there
+%     exists a region in the occurence of \((\tau,R)\) in \(\sigma\) that is
+%     now devoid of restrictions. However, by removing restrictions we cannot
+%     make an occurence become not an occurence, and therefore the same occurrence
+%     of \(\tau \) in \(\sigma\) is now an occurence of \((\tau, R^\prime)\).
+% \end{proof}
 \begin{proof}[Proof of \PropositionRef{prop:dom1}]
@@
     Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
-
-    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\)
-    by \NoteRef{not:downcmesh} any of these occurrences is also an occurence
-    of \(m_1\). This proves that every occurence of \(m_2\) is also an
-    occurence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).
+    % Consider a permutation \(\omega\) that contains an occurrence of \(m_2\)
+    % by \NoteRef{not:downcmesh} any of these occurrences is also an occurence
+    % of \(m_1\). This proves that every occurence of \(m_2\) is also an
+    % occurence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).
+    Since \(R_1\) is a subset of \(R_2\), \NoteRef{not:downcmesh} states that
+    \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\)
 
@@
 
-By using this rule we completely capture the equivalence classes of
-\(\av{\{\perm{3,2,1}, \mperm{2,1}{R}\}}\).
+All coincidence classes of \(\av{\{\perm{3,2,1}, \mperm{2,1}{R}\}}\) can be
+explained by repeated application of \PropositionRef{prop:dom1}.
 
@@
 
-        \caption{Visual Depiction of first dominating pattern rule.}
+        \caption{Visual depiction of first dominating pattern rule.}
         \label{fig:rule1}
@@
 \begin{equation*}
-    \pattern{}{2,1}{0/0,0/2,1/0,2/1} \text{ and } \pattern{}{2,1}{0/0,0/2,1/0,1/1,2/1}
+    m_1 = \pattern{}{2,1}{0/0,0/2,1/0,2/1} \text{ and } m_2 = \pattern{}{2,1}{0/0,0/2,1/0,1/1,2/1}
 \end{equation*}
 
-Consider the box \((1,1)\), this is the box that we need to find reasoning
-to allow shading. If this box were to be empty then we could shade it freely,
-consider what happens if there are points in this box.  If there is any
+Consider an occurrence of \(m_1\) in a permutation in \(\av{\perm{2,3,1}}\), then if the region corresponding
+to the box \((1,1)\) is empty we have an occurrence of \(m_2\). Otherwise, if there is any
 increase in this box then we would have an occurrence of \(\perm{2,3,1}\),
 however, since we are in \(\av{\perm{2,3,1}}\) this is not possible. This box
-must contain a decreasing subsequence. This gives rise to the following lemma:
+must therefore contain a decreasing subsequence. This gives rise to the following lemma:
 
@@
 \end{equation*}
-We know that the box \((1,1)\) contains an decreasing subsequence. The top point
-of this decrease can be chosen to act as the second point in the mesh pattern,
+We know that the region corresponding to the box \((1,1)\) contains a decreasing subsequence. The top point
+of this decreasing subsequence can be chosen to act as the second point in the mesh pattern,
 and therefore there are no points between the first point and the new second point.
@@
 By taking the First Dominating Pattern Rule and the Second Dominating Pattern
-rule together coincidences of classes of the form \(\av{\{\perm{1,3,2},\mperm{2,1}{R}\}}\)
+Rule together coincidences of classes of the form \(\av{\{\perm{2,3,1},\mperm{2,1}{R}\}}\)
 are fully explained, obtaining 39 equivalence classes of mesh patterns.
@@
 \begin{equation*}
-    \pattern{}{1,2}{0/2,2/0,2/1} \text{ and } \pattern{}{1,2}{0/2,1/0,2/0,2/1}
+    m_1 = \pattern{}{1,2}{0/2,2/0,2/1} \text{ and } m_2 = \pattern{}{1,2}{0/2,1/0,2/0,2/1}
 \end{equation*}
@@
 
-In order to rigorously establish this coincidence we need to consider what would
-happen if we were to choose a point in the box \((1,0)\) that we would like to shade.
-In order to get a chance of getting the shading we want consider choosing the
-rightmost point in the box.
+Consider an occurrence of \(m_1\) in a permutation, if the region corresponding
+to the box \((1,0)\) is empty then we have an occurrence of \(m_2\). Now look at
+the case when this region is not empty. Consider choosing the rightmost point in
+region.
+% In order to rigorously establish this coincidence we need to consider what would
+% happen if we were to choose a point in the box \((1,0)\) that we would like to shade.
+% In order to get a chance of getting the shading we want consider choosing the
+% rightmost point in the box.
 This gives us the following mesh pattern.
@@
     Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
-    and a dominating classical pattern \(\pi = (\pi,\emptyset)\)\footnote{The
-    permutation \(\pi\) may be the empty permutation}, the sets
+    and a dominating classical pattern \(\pi = (\pi,\emptyset)\), the sets
     \(\av{\{\pi,m_1\}}\) and \(\av{\{\pi,m_2\}}\) are coincident if
@@
 \begin{equation*}
-    \pattern{}{1,2}{0/0, 0/2, 1/1, 1/2, 2/1} \text { and } \pattern{}{1,2}{0/0, 0/1, 0/2, 1/1, 1/2, 2/1}
+    m_1 = \pattern{}{1,2}{0/0, 0/2, 1/1, 1/2, 2/1} \text { and } m_2 \pattern{}{1,2}{0/0, 0/1, 0/2, 1/1, 1/2, 2/1}
 \end{equation*}
@@
 
-Consider the contents of the box \((0,1)\), by \LemmaRef{lem:incdec} it
-must contain an increasing subsequence. If there is only one point in the box
-we can choose this to replace the \(1\) in the mesh pattern to get the
-required shading. If there is more than one point then choosing the
-two leftmost points give us the following mesh pattern.
+Consider the region corresponding to the box \((0,1)\) in any occurrence of \(m_1\),
+in a permutation by \LemmaRef{lem:incdec} it must contain an increasing subsequence.
+If there is only one point in the region we can choose this to replace the \(1\) in
+the mesh pattern to get the required shading. If there is more than one point then
+choosing the two leftmost points give us the following mesh pattern.
 \begin{equation*}

### 2016-04-29 writing/latex/introduction.tex
Getting closer

@@
 We can display a permutation on a \emph{figure} in order to give a graphical
-representation of the permutation. In such a figure we let the \(x\)-axis
-denote the index in the permutation, and the \(y\)-axis denotes the values of
-\(\pi(x)\). The figure of the permutation \(\pi = \perm{2,3,1}\) is shown below
+representation of the permutation. In such a figure we plot the points
+\((i,\pi(i))\) in a Cartesian coordinate system. The figure of the permutation
+\(\pi = \perm{2,3,1}\) is shown below
 \begin{equation*}
@@
 \begin{definition}{Order isomorphism.}
-    Two sequences \(\range{\alpha_1,\alpha_2}{\alpha_n}\) and
-    \(\range{\beta_1,\beta_2}{\beta_n}\) are said to be \emph{order isomorphic}
-    if they share the same relative order,i.e.,\(\alpha_r<\alpha_s\) if and
+    Two substrings \(\alpha_1\alpha_2\dotsm\alpha_n\) and
+    \(\beta_1\beta_2\dotsm\beta_n\) are said to be \emph{order isomorphic}
+    if they share the same relative order, \ie, \(\alpha_r<\alpha_s\) if and
     only if \(\beta_r<\beta_s\).
@@
 
-A permutation \(\pi\) us said to \emph{contain} the permutation \(\sigma\) of
+A permutation \(\pi\) is said to \emph{contain} the permutation \(\sigma\) of
 length \(k\) as a pattern (denoted \(\sigma \preceq \pi\)) if there is some
-an increasing subsequence \(\range{i_1,i_2}{i_n}\) such that the sequence
-\(\range{\pi(i_1),\pi(i_2)}{\pi(i_k)}\) is order isomorphic to
-\(\range{\sigma(1),\sigma(2)}{\sigma(k)}\). If \(\pi\) does not contain
+subsequence \(i_1i_2\dotsm{}i_n\) such that the sequence
+\(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is order isomorphic to
+\(\sigma(1)\sigma(2)\dotsm\sigma(k)\). If \(\pi\) does not contain
 \(\sigma\), we say that \(\pi\) \emph{avoids} \(\sigma\).
@@
 For example the permutation \(\pi = \perm{2,4,1,5,3}\) contains the pattern
-\(\sigma = \perm{2,3,1}\), since the second, fourth and fifth entries
-(\(4,5,\) and \(3\)) share the same relative order as the entries of \(\sigma\).
+\(\sigma = \perm{2,3,1}\), since the second, fourth and fifth elements
+(\(453\)) are order isomorphic to \(\perm{231}\).
 This can be seen graphically below, the points order isomorphic to \(\sigma\)
@@
 \end{equation*}
-By this definition the empty permutation \(\varepsilon\) as a mesh pattern consisting solely of the
-box \((0,0)\).
 
+The set \(R\) is called the \emph{mesh-set} of the mesh pattern \(p\).
 The figure for a mesh pattern looks similar to that for a classical pattern with the
-addition that we shade the unit square with bottom corner \((i,j)\) for each \((i,j) \in R\):
+addition that we shade the unit square with bottom left corner \((i,j)\) for each \((i,j) \in R\):
 \begin{equation*}
@@
 position of the occurence of \(\tau\) in \(\pi\). These restrictions say that
-the shaded regions of the figure above contain no points from \(\pi\).
+no elements of \(\pi\) are allowed in the shaded regions of the figure.
 
@@
 
+Given a mesh pattern \(p=(\sigma,R)\) we say that \(\sigma\) is the \emph{underlying
+classical pattern} of \(p\).
 \begin{note}
     \label{not:classmesh}
-    Classical patterns are just a subclass of mesh patterns where the mesh
-    set \(R\) is empty. The classical pattern \(\pi\) can be represented
-    by a mesh pattern as \((\pi,\emptyset)\).
+    Classical patterns are a subclass of mesh patterns the classical pattern \(\pi\)
+    can be represented by a mesh pattern as \((\pi,\emptyset)\).
 \end{note}
@@
 those where two adjacent entries in the pattern must be adjacent in the permutation,
-\ie entire columns are shaded in the mesh.
+\ie \(R\) is a union of vertical strips.
 \textcite{MR2652101} look at classes of pattern where both columns and rows
@@
 Mesh patterns also encompass a subset of \emph{barred} patterns introduced by
-\textcite{MR2716312}.
+\textcite{MR2716312}, those with only one barred letter.
 

### 2016-04-30 writing/latex/abstracteng.tex
tidy up a bit

@@
 A permutation is an arrangement of \(n\) objects.
+Avoidance classes of classical permutation patterns
+capture many interesting properties, such as
+stack-sortability, and have links to many different combinatorial objects.
+Mesh patterns are an extension of classical patterns that allow additional
+restrictions to be placed on occurrences of the pattern.
 Two mesh patterns are coincident if they are avoided by the same set of
@@
 classify coincidence amongst families containing a mesh pattern of length 2
-and a classical pattern of length 3. The author then goes on to completely
-classify Wilf-equivalence amongst mesh patterns of length 2 when we avoid the
-classical pattern 231.
+and a classical pattern of length 3. Two patterns are Wilf-equivalent if they
+have the same number of avoiders at every length, the author completely
+Wilf-classifies mesh patterns of length 2 when avoiding the classical pattern 231.
+% Finally we attempt to show some non-trivial Wilf-equivalences between avoiders
+% of sets of the form \(231,m_1\) and \(231,m_2\)

### 2016-04-30 writing/latex/acknowledgements.tex
tidy up a bit

\begin{quotation}
So long, and thanks for all the fish.
\end{quotation}\sourceatright{Douglas Adams\cite{adams84fish}}
\vspace{\baselineskip}

I would like to thank my advisor, Henning Ulfarsson, for his time, support,
guidance and introducing me to a research topic that I greatly enjoy.

I would like to thank Michael Albert and Anders Claesson for the discussions that
prompted the topic of this thesis, as well as their valuable input and suggestions.

I would like to thank Christian Bean for proofreading my thesis, providing useful
feedback, and for the time spent staring at a whiteboard while I tried to check
my mathematical reasoning.

### 2016-04-30 writing/latex/code.tex
tidy up a bit

@@
 \chapter{Code}\label{cha:code}
-You can put code in your document using the listings package, which is
-loaded by default in \path{custom.tex}.  Be aware that the listings
-package does not put code in your document if you are in draft mode
-unless you set the \texttt{forcegraphics} option.
-
-There is an example java (Listing~\ref{src:Data_Bus.java}) and XML
-file (Listing~\ref{src:AndroidManifest.xml}).  Thanks to the
-\texttt{url} package, you can typeset OSX and unix paths like this:
-\path{/afs/rnd.ru.is/project/thesis-template}.  Windows paths:
-\path{C:\windows\temp\ }.  You can also typeset them using the menukey
-package, but it tends to delete the last separator and has other
-complications.\footnote{The menukey package has issues with biblatex,
-  read \path{custom.tex} for more information.}
-
-If you are trying to include multiple different languages, you should
-go read the documentation and set these up in \path{custom.tex}.  You
-will save yourself a lot of effort, especially if you have to fix
-anything.
-
-%I have put the source code in the \directory{src/} folder.
-\lstinputlisting[language=Java, firstline=1,
-lastline=40, caption={Data\_Bus.java: Setting up the class.},
-label={src:Data_Bus.java}]{src/Data_Bus.java}
-
-\lstinputlisting[language={[android]XML}, firstline=1, lastline=20,
-caption={AndroidManifest.xml: Configuration for the Android UI.},
-label={src:AndroidManifest.xml}]{src/AndroidManifest.xml}
+% \begin{minted}[linenos, breaklines]{python}
+% def boring(args=None):
+%     pass
+% \end{minted}
+\inputminted[linenos, breaklines, frame=lines, stripall]{python}{src/main.py}
+% \lstinputlisting[language=Python]{src/main.py}
+% You can put code in your document using the listings package, which is
+% loaded by default in \path{custom.tex}.  Be aware that the listings
+% package does not put code in your document if you are in draft mode
+% unless you set the \texttt{forcegraphics} option.
+%
+% There is an example java (Listing~\ref{src:Data_Bus.java}) and XML
+% file (Listing~\ref{src:AndroidManifest.xml}).  Thanks to the
+% \texttt{url} package, you can typeset OSX and unix paths like this:
+% \path{/afs/rnd.ru.is/project/thesis-template}.  Windows paths:
+% \path{C:\windows\temp\ }.  You can also typeset them using the menukey
+% package, but it tends to delete the last separator and has other
+% complications.\footnote{The menukey package has issues with biblatex,
+%   read \path{custom.tex} for more information.}
+%
+% If you are trying to include multiple different languages, you should
+% go read the documentation and set these up in \path{custom.tex}.  You
+% will save yourself a lot of effort, especially if you have to fix
+% anything.
+%
+% %I have put the source code in the \directory{src/} folder.
+% \lstinputlisting[language=Java, firstline=1,
+% lastline=40, caption={Data\_Bus.java: Setting up the class.},
+% label={src:Data_Bus.java}]{src/Data_Bus.java}
+%
+% \lstinputlisting[language={[android]XML}, firstline=1, lastline=20,
+% caption={AndroidManifest.xml: Configuration for the Android UI.},
+% label={src:AndroidManifest.xml}]{src/AndroidManifest.xml}
 

### 2016-04-30 writing/latex/coincidence.tex
tidy up a bit

@@
 of permutations that avoid \(\pi\) is the same as the set of permutations that
-avoid \(\sigma\) \ie \(\av{\pi} = \av{\sigma}\). This extends to sets of patterns as well as single patterns.
+avoid \(\sigma\), \ie \(\av{\pi} = \av{\sigma}\). This extends to sets of patterns as well as single patterns.
 
-We consider the avoidance sets, \(\av{p,q}\) where \(p\) is a classical pattern
+We consider the avoidance sets, \(\av{\{\pi,q\}}\) where \(\pi\) is a classical pattern
 of length \(3\) and \(q\) is a mesh pattern of length \(2\) in order to
 establish some rules about when these two sets give the same avoidance set.
-We fix \(p\) in order to define the equivalences and say that \(p\) is the
+We fix \(\pi\) in order to define the  coincidence and say that \(\pi\) is the
 \emph{dominating pattern}.
@@
 Given a mesh pattern \(p\), let \(\addp[D]{p}{(a,b)}\) be the operation that returns
-a mesh pattern equivalent to placing a point in the center of box \((a,b)\), where \((a,b)\) is not the
-mesh set of \(p\) in
-the figure of \(p\), with shading defined by \(D\subseteq\{N,E,S,W\}\).
+a mesh pattern equivalent to placing a point in the center of box \((a,b)\),
+where \((a,b)\) is not in the mesh set of \(p\) , with shading defined by
+\(D\subseteq\{N,E,S,W\}\).
 The set \(D\) defines the shading by indicating that the boxes in the cardinal
@@
 \end{align*}
-    A more complex example for \texttt{add\_point}
+    A more complex example for \texttt{add\_point} could be
 \begin{equation*}
@@
 \end{example}
+\begin{definition}
 Formally, given a mesh pattern \(p = (\pi,R)\) \(\addp[D]{p}{(a,b)}\) gives a mesh
-pattern \(p^\prime = (\pi^\prime,R^\prime)\) defined by
+pattern \(p^\prime = (\pi^\prime,R^\prime)\) with length \(\setsize{\pi}+1\) defined by
 \begin{equation*}
@@
 \begin{equation*}
-        R^\prime = \bigcup_{(i,j)\in R}r((i,j))
+        R^\prime = \bigcup_{(i,j)\in R}r((i,j)),
 \end{equation*}
-Where \(r((i,j))\) is defined by
+where \(r((i,j))\) is defined by
 \begin{equation*}
@@
 And we add the union of the sets in \(D\) into the mesh set \(R^\prime\).
+\end{definition}
 
@@
     Given a pattern \(p\), define \(\addd{p}{(a,b)}\), and \\ \(\addi{p}{(a,b)}\),
-    as the operations that return a mesh pattern equivalent to placing an
+    as the operations that return a mesh pattern equivalent to placing a
     decrease, or increase, in the center of box \((a,b)\), where \((a,b)\) is not
@@
 We can define these formally as with \(\addp[D]{p}{(a,b)}\). Considering
-only adding the ascent, adding a descent is very similar
+only adding the ascent, as adding a descent is very similar
 \begin{equation*}
@@
     \pi(i)+2 & \text{if } i\neq a+1,a+2 \text{ and } \pi(i)>b\\
-    b+1, b+2 & \text{if } i = a+1, a+2
+    b+j, b+j & \text{if } i = a+j, j\in\{1,2\}
 \end{cases}
@@
 \begin{equation*}
-        R^\prime = \bigcup_{(i,j)\in R}r((i,j)) \cup 
+        R^\prime = \bigcup_{(i,j)\in R}r((i,j)) \cup \text{crap to deal with strong ascent},
 \end{equation*}
-Where \(r((i,j))\) is defined by
+where \(r((i,j))\) is defined by
 \begin{equation*}
@@
 
-It can be easily seen that in order to classify set equivalences one need only
+It can be easily seen that in order to classify coincidences one need only
 consider coincidences within the family of mesh patterns with the same underlying
@@
 classes follows the values shown in  \TableRef{tab:domclasses}. The experimental
-data is calculated on permutations up to length 11.
+data in the last row of the table is calculated on permutations up to length 11.
 
@@
 \end{center}
-    \caption{Class number reduction by application of Dominating Rules}
+    \caption{Coincidence class number reduction by application of Dominating Rules}
     \label{tab:domclasses}
@@
 \section{Coincidence classes of Av(\{321, (21, \textit{R})\}).}
-Through experimentation we discover that there are at most 29 coincidence
-classes of mesh patterns with underlying classical pattern \(\perm{2,1}\).
+Through experimentation, up to permutations of length 11, we discover that there
+are at most 29 coincidence classes of mesh patterns with underlying classical
+pattern \(\perm{2,1}\).
 
@@
 
-    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
+    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
     % Consider a permutation \(\omega\) that contains an occurrence of \(m_2\)
@@
     Now we consider a permutation \(\omega^\prime\in \av{\pi}\),
-    suppose we have an occurrence of \(m_1\). Consider placing a point in the
+    containing an occurrence of \(m_1\). Consider placing a point in the
     region corresponding to the box \((a,b)\), regardless of where in this
@@
 All coincidence classes of \(\av{\{\perm{3,2,1}, \mperm{2,1}{R}\}}\) can be
-explained by repeated application of \PropositionRef{prop:dom1}.
+explained by application of \PropositionRef{prop:dom1}. % repeated?
 
-This rule is understood very easily by seeing it in graphical form. In the
+This rule can be understood very in graphical form. In the
 pattern in \FigureRef{fig:rule1} we can gain shading in three boxes since if
-there is a point in any of these boxes we would gain an occurrence of the
+there is a point in any of these boxes we would get an occurrence of the
 dominating pattern \(\perm{3,2,1}\).
@@
     \begin{tikzpicture}[scale=\picscale]
-        \modpattern{}{2,1}{}
+        \modpattern{}{2,1}{1/0,1/1,1/2,2/2}
     \end{tikzpicture}
@@
     \begin{tikzpicture}[scale=\picscale]
-        \filldraw (0.5,2.5) circle (3pt);
-        \filldraw (1.5,1.5) circle (3pt);
         \filldraw (2.5,0.5) circle (3pt);
-        \modpattern{}{2,1}{0/2,1/1,2/0}
+        \modpattern{}{2,1}{1/0,1/1,1/2,2/2,2/0}
     \end{tikzpicture}
@@
 
+Note that there are two natural extensions of this rule. We can replace \(\pi\)
+with a set of classical patterns, or we can consider \(\pi\) to be a mesh pattern.
+
 \section{Equivalence classes of Av(\{231, (21, \textit{R})\}).}
-By application of \PropositionRef{prop:dom1} we obtain \(43\) equivalence
-classes. Experimentation shows that there are in fact \(39\) equivalence
+By application of \PropositionRef{prop:dom1} we obtain \(43\) coincidence
+classes. Experimentation shows that there are in fact at most \(39\) coincidence
 classes, for example the following two patterns are coincident in
@@
 
-Consider an occurrence of \(m_1\) in a permutation in \(\av{\perm{2,3,1}}\), then if the region corresponding
+Consider an occurrence of \(m_1\) in a permutation in \(\av{\perm{2,3,1}}\), consisting
+of elements \(x\) and \(y\). If the region corresponding
 to the box \((1,1)\) is empty we have an occurrence of \(m_2\). Otherwise, if there is any
@@
 however, since we are in \(\av{\perm{2,3,1}}\) this is not possible. This box
-must therefore contain a decreasing subsequence. This gives rise to the following lemma:
+must therefore contain a (non-empty) decreasing subsequence.
+ This gives rise to the following lemma:
 
@@
 \end{equation*}
-We know that the region corresponding to the box \((1,1)\) contains a decreasing subsequence. The top point
-of this decreasing subsequence can be chosen to act as the second point in the mesh pattern,
-and therefore there are no points between the first point and the new second point.
-Hence, we can shade this box as it is guaranteed to be empty. This shows that our
-two example patterns are coincident.
+We know that the region corresponding to the box \((1,1)\) contains a decreasing subsequence.
+If we let \(z\) be the topmost point in this decereasing subsequence, then \(xz\)
+is an occurrence of \(m_2\). This shows that our two example patterns are coincident.
 
-This result generalises into the following rule for categorising equivalences
+This result generalises into the following rule for categorising coincidences
 of mesh patterns in cases where there is a dominating classical pattern.
@@
 
-    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
+    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
 
-    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\)
-    by \NoteRef{not:downcmesh} any of these occurrences is also an occurence
+    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\).
+    By \NoteRef{not:downcmesh} any of these occurrences is also an occurence
     of \(m_1\). This proves that every occurence of \(m_2\) is also an
@@
     We will consider taking the first branch of every choice.
-    Now consider a permutation in \(\omega^\prime \in \av{\pi}\). Suppose
-    \(\omega^\prime\) contains \(m_1\) consider the region corresponding to
+    Now consider a permutation \(\omega^\prime \in \av{\pi}\). Suppose
+    \(\omega^\prime\) contains \(m_1\) and consider the region corresponding to
     \((a,b)\) in \(R_1\).
 
-    If the region is empty, then we can freely shade the corresponding
-    box \((a,b)\) in \(m_1\) and hence have an occurence of \(m_2\).
+    If the region is empty, the occurrence of \(m_1\) is trivially an occurence
+    of \(m_2\).
 
@@
     Similar arguments satisfy the remainder of the branches.
-
 \end{proof}
@@
 Rule together coincidences of classes of the form \(\av{\{\perm{2,3,1},\mperm{2,1}{R}\}}\)
-are fully explained, obtaining 39 equivalence classes of mesh patterns.
+are fully explained, obtaining 39 coincidence classes of mesh patterns.
 
 \section{Equivalence classes of Av(\{231, (12, \textit{R})\}).}
-When considering the equivalence classes of \(\av{\perm{2,3,1},\mperm{1,2}{R}}\)
+When considering the coincidence classes of \(\av{\perm{2,3,1},\mperm{1,2}{R}}\)
 we first apply the two Dominating Pattern rules previously established.
@@
 % rightmost point in the box.
-This gives us the following mesh pattern.
+This gives us an occurrence of the following mesh pattern.
 \begin{equation*}
@@
 \end{equation*}
-By application of the \PropositionRef{prop:dom1} we then achieve the following
+By application of \PropositionRef{prop:dom1} we then achieve the following
 mesh pattern
@@
         \item\label{prop:dom3:condocc} \(\addp[D]{\mperm{\sigma}{R_1}}{(a,b)}\)
+            where \(D\in\{N,E,S,W\}\)
             is coincident with a mesh pattern containing an occurrence of
@@
     length one longer than \(m_1\) in this position. This mesh pattern is
-    coincident with another mesh pattern. This mesh pattern contains an
-    occurrence of \(m_2\) so every occurrence of \(m_1\) is also an occurrence
+    coincident with another mesh pattern that contains an
+    occurrence of \(m_2\). Hence, every occurrence of \(m_1\) leads to an occurrence
     of \(m_2\).  Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the
@@
 \section{Equivalence classes of Av(\{321, (12, \textit{R})\}).}
-When considering equivalences of mesh patterns with underlying classical
+When considering coincidences of mesh patterns with underlying classical
 pattern \(\perm{1,2}\) in \(\av{\perm{3,2,1}}\) application of the previously
 established rules give no coincidences. Through experimentation we discover
-that there are \(7\) equivalence classes which can be explained through the
+that there are \(7\) non-trivial coincidence classes (all others are singletons)
+which can be explained through the
 use of two different lines of reasoning. Since the number of coincidences is
-so small we will reason for these equivalences without attempting to generalise
+so small we will reason for these coincidences without attempting to generalise
 into concrete rules.
@@
 Consider the region corresponding to the box \((0,1)\) in any occurrence of \(m_1\),
-in a permutation by \LemmaRef{lem:incdec} it must contain an increasing subsequence.
+in a permutation. By \LemmaRef{lem:incdec} it must contain an increasing subsequence.
+If the region is empty then we have an occurrence of \(m_2\).
 If there is only one point in the region we can choose this to replace the \(1\) in
 the mesh pattern to get the required shading. If there is more than one point then
-choosing the two leftmost points give us the following mesh pattern.
+choosing the two leftmost points gives us the following mesh pattern.
 \begin{equation*}
@@
 Now if we take the other two points as the points in our
-mesh permutation then we get an occurrence of the pattern
+mesh pattern then we get an occurrence of the pattern
 we originally desired, and hence the two patterns are coincident.
@@
 In order to prove this coincidence we will proceed by mathematical
-induction on the number of points in the middle box we call this number \(n\).
+induction on the number of points in region corresponding to the middle box.
+We call this number \(n\).
 \begin{description}
-    \item [Base Case \((n=0)\):] The base case hold since we can freely shade the
+    \item [Base Case \((n=0)\):] The base case holds since we can freely shade the
         box if it contains no points.
-    \item [Inductive Hypothesis \((n=k)\):] Suppose that the we can find an occurrence
+    \item [Inductive Hypothesis \((n=k)\):] Suppose that we can find an occurrence
         of the second pattern if we have an occurrence of the first with \(k\) points
@@
     \item [Inductive Step \((n=k+1)\)] Suppose that we have \((k+1)\) points in
-        the middle box. Choose the bottom most point in the middle box, this gives a
-        mesh pattern equivalent to
+        the middle box. Choose the bottom most point in the middle box, this gives the
+        mesh pattern
      \begin{equation*}
@@
     \end{equation*}
-        Now we need to consider the box labelled \(X\) if this box is empty then
+        Now we need to consider the box labelled \(X\). If this box is empty then
         we have an occurrence of \(m_2\) and are done. If this box contains
@@
 \end{description}
-By induction we have that every occurrence of \(m_1\) is an occurrence of \(m_2\)
+By induction we have that every occurrence of \(m_1\) leads to an occurrence of \(m_2\)
 and by \NoteRef{not:downcmesh} every occurrence of \(m_2\) is an occurrence of
@@
 This argument applies to another two pairs of classes. Therefore in
-total in \(\av{\perm{3,2,1}, \mperm{1,2}{R}}\) there are \(213\) equivalence
+total in \(\av{\perm{3,2,1}, \mperm{1,2}{R}}\) there are \(213\) coincidence
 classes.

### 2016-04-30 writing/latex/introduction.tex
tidy up a bit

@@
 \section{What is a Permutation?}
-In \emph{The Art of Computer Programming}\cite[p.~45]{Knuth:1997:ACP:260999} Donald Knuth defines
-A \emph{permutation of \(n\) objects} is an arrangement of \(n\) distinct
-objects in a row. When considering permutations we can consider them as
+In \emph{The Art of Computer Programming} \cite[p.~45]{Knuth:1997:ACP:260999} Knuth states,
+\enquote{A \emph{permutation of \(n\) objects} is an arrangement of \(n\) distinct
+objects in a row}. When considering permutations we can consider them as
 occuring on the set \(\nrange{n} = \setrange{1}{n}\), therefore a permutation
-is a \emph{bijection} \(\pi:\nrange{n} \mapsto \nrange{n}\). We can write a
-permutation \(\pi\) in two line notation
+is a \emph{bijection} \(\pi:\nrange{n} \mapsto \nrange{n}\). A
+permutation \(\pi\) can be written in two-line notation
 \begin{equation*}
@@
 \emph{one-line notation}, in this form we drop the top line of the two line
-notation so are left with the following.
+notation and are left with the following.
 \begin{equation*}
@@
 \begin{example}
-There are 6 permutations on \(\nrange{3}\).
+The 6 permutations on \(\nrange{3}\), in one-line notation, are
 \begin{equation*}
@@
 
-The class of all permutations of length \(n\) is \(\mathfrak{S}_n\) and
-the class has size \(n!\). The class of all permutations is
-\(\mathfrak{S}=\bigcup_{i=0}^{\infty}\mathfrak{S}_i\).
+The set of all permutations of length \(n\) is \(\mathfrak{S}_n\) and
+has size \(n!\). The set of all permutations is
+\(\mathfrak{S}=\bigcup_{i=0}^{\infty}\mathfrak{S}_i\). Note that \(\mathfrak{S}_0\)
+has exactly one element, the empty permutation \(\varepsilon\). As a function
+this is equivalent to the unique bijection \(\emptyset \mapsto \emptyset\), and
+it's one-line representation is the empty string.
 
 \section{Classical Permutation Patterns}
-Classical Permutation Patterns began to be studied as a result of Knuth's
+Classical permutation patterns began to be studied as a result of Knuth's
 statements about stack-sorting in \emph{The Art of Computer Programming}
@@
 
-\begin{definition}{Order isomorphism.}
+\begin{definition}{(Order isomorphism.)}
     Two substrings \(\alpha_1\alpha_2\dotsm\alpha_n\) and
@@
 A permutation \(\pi\) is said to \emph{contain} the permutation \(\sigma\) of
-length \(k\) as a pattern (denoted \(\sigma \preceq \pi\)) if there is some
-subsequence \(i_1i_2\dotsm{}i_n\) such that the sequence
+length \(k\) as a \emph{pattern} (denoted \(\sigma \preceq \pi\)) if there is some
+subsequence \(i_1i_2\dotsm{}i_k\) such that the sequence
 \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is order isomorphic to
@@
 (\(453\)) are order isomorphic to \(\perm{231}\).
-This can be seen graphically below, the points order isomorphic to \(\sigma\)
+This can be seen graphically below, the subsequence order isomorphic to \(\sigma\)
 are highlighted.
@@
 Knuth's statements were exercises in showing that the permutations avoiding the
-pattern \(\perm{2,3,1}\) completely categorise permutations that are sortable to
-the identity permutation using only a single stack, and that permutations avoiding the
-pattern \(\perm{3,2,1}\) completely categorise permutations that are sortable to
-the identity permutation using only a single queue with bypass.
+pattern \(\perm{2,3,1}\) are precisely the permutations that are sortable to
+the identity permutation using a single stack, and that permutations avoiding the
+pattern \(\perm{3,2,1}\) are precisely the permutations that are sortable to
+the identity permutation using a single queue with bypass.
 
 \section{Mesh Patterns}
-Mesh Patterns were introduced by \textcite{journals/combinatorics/BrandenC11} to capture explicit expansions
-for certain permutation statistics. They are a natural extension of Classical
-permutation patterns. A \emph{mesh-pattern} is a pair
+Mesh patterns were introduced by \textcite{journals/combinatorics/BrandenC11} to
+capture explicit expansions for certain permutation statistics. They are a
+natural extension of classical permutation patterns. A \emph{mesh pattern} is a pair
 \begin{equation*}
@@
 
-The set \(R\) is called the \emph{mesh-set} of the mesh pattern \(p\).
-The figure for a mesh pattern looks similar to that for a classical pattern with the
+The set \(R\) is called the \emph{mesh set} of the mesh pattern \(p\).
+The figure for a mesh pattern looks similar to that of a classical pattern with the
 addition that we shade the unit square with bottom left corner \((i,j)\) for each \((i,j) \in R\):
@@
 position of the occurence of \(\tau\) in \(\pi\). These restrictions say that
-no elements of \(\pi\) are allowed in the shaded regions of the figure.
+no elements of \(\pi\) are allowed in the regions of the figure corresponding
+to shaded boxes in the mesh (see \cite[2]{journals/combinatorics/BrandenC11}
+for the precise definition).
 
@@
 \end{example}
-\begin{proof}
-    Let us consider the figure for the permutation \(\pi\) we only need to find one occurence.
+
+Let us consider the figure for the permutation \(\pi\). The subsequence \(325\) is
+an occurrence of \(213\) in the classical sense and the remaining points of \(\pi\)
+are not contained in the regions corresponding to the shaded boxes in \(p\).
 \begin{equation*}
@@
 \end{equation*}
-    We have found an occurence of the pattern \(p\) in \(\pi\) and therefore
+The subsequence \(325\) is therefore an occurence of the pattern \(p\) in \(\pi\) and
 \(\pi\) contains \(p\).
@@
 avoids \(p\) since for every occurence of the classical pattern \(\perm{2,1,3}\)
-    there is at least one point in one of the shaded boxes. Consider the subsequence
-    \(315\) in \(\sigma\), this is an occurence of \(\perm{2,1,3}\) but not the mesh
-    pattern since the points with values \(4\) and \(2\) are in the shaded areas.
-    This is shown in the figure below.
+there is at least one point in one of the shaded boxes. For example, consider the
+subsequence\(315\) in \(\sigma\), this is an occurence of \(\perm{2,1,3}\) but
+not the mesh pattern since the points with values \(4\) and \(2\) are in the
+regions corresponding to the boxes \((0,1)\) and \((0,2)\), which are shaded in
+\(p\). This is shown in the figure below.
 
@@
 therefore \(\sigma\) avoids \(p\).
-\end{proof}
 
 We denote the avoidance sets for mesh patterns in the same way as for
-classical patterns.
-
-Given a mesh pattern \(p=(\sigma,R)\) we say that \(\sigma\) is the \emph{underlying
+classical patterns. Given a mesh pattern \(p=(\sigma,R)\) we say that \(\sigma\) is the \emph{underlying
 classical pattern} of \(p\).
@@
     \label{not:classmesh}
-    Classical patterns are a subclass of mesh patterns the classical pattern \(\pi\)
-    can be represented by a mesh pattern as \((\pi,\emptyset)\).
+    Classical patterns can be thought of as a subset of mesh patterns: the
+    classical pattern \(\pi\) can be represented by a mesh pattern as
+    \((\pi,\emptyset)\).
 \end{note}
@@
 interesting counting results. However, very little work has been done on avoiding
-a mesh pattern and a classical pattern simultaneously, in this work we aim to
+a mesh pattern and a classical pattern simultaneously. In this work we aim to
 establish some ground in this field by computing coincidences and Wilf-classes

### 2016-04-30 writing/latex/src/Data_Bus.java

@@ -1,25 +1,0 @@
-package com.example.mycoolapp;
-
-import android.app.Activity;
-import android.content.Intent;
-import android.os.Bundle;
-import android.view.View;
-import android.widget.Button;
-import android.widget.TextView;
-
-// I am creating a comment that is very very long to demonstrate how the line wrapping system works.  You should see a symbol to annotate that it has been wrapped to the next line.
-public class Data_Bus extends Activity {
-	Button Next;
-	TextView textdisplay1, textdisplay2; 
-	
-	@Override
-	protected void onCreate(Bundle savedInstanceState) {
-		// TODO Auto-generated method stub
-		super.onCreate(savedInstanceState);
-		setContentView(R.layout.adc_databustest);
-		Next = (Button) findViewById (R.id.checkButton); 
-		textdisplay1 = (TextView) findViewById (R.id.tvTop);
-		textdisplay2 = (TextView) findViewById (R.id.tvBottom);
-		       
-	}
-}

### 2016-04-30 writing/latex/src/main.py

@@ -1,0 +1,99 @@
+from permuta import MeshPattern
+import sys
+from classes import *
+from helper_funcs import *
+from p_rule import *
+from wilf import *
+
+def stderrwrite(string,end="\n"):
+    sys.stderr.write(string+end)
+    sys.stderr.flush()
+
+classes2 = [map(MeshPattern.reverse, x) for x in classes]
+stderrwrite("Classes created")
+
+p_231 = Permutation([2,3,1])
+p_321 = Permutation([3,2,1])
+
+reprs_12 = get_representatives(classes)
+reprs_21 = get_representatives(classes2)
+stderrwrite("Representatives created")
+
+meshpd_12 = mpdict(classes)
+meshpd_21 = mpdict(classes2)
+
+stderrwrite("Computing Coincidences 1 of 4")
+coincs231_12 = get_coincidences(reprs_12, 11, p_231)
+stderrwrite("Computing Coincidences 2 of 4")
+coincs321_12 = get_coincidences(reprs_12, 11, p_321)
+stderrwrite("Computing Coincidences 3 of 4")
+coincs231_21 = get_coincidences(reprs_21, 11, p_231)
+stderrwrite("Computing Coincidences 4 of 4")
+coincs321_21 = get_coincidences(reprs_21, 11, p_321)
+stderrwrite("Coincidences done")
+
+coincsdict231_12 = coincidences_to_dictionary(coincs231_12)
+coincsdict321_12 = coincidences_to_dictionary(coincs321_12)
+coincsdict231_21 = coincidences_to_dictionary(coincs231_21)
+coincsdict321_21 = coincidences_to_dictionary(coincs321_21)
+
+p_coincs231_12 = p_rule_coincidences(meshpd_12, p_231)
+p_coincs321_12 = p_rule_coincidences(meshpd_12, p_321)
+p_coincs231_21 = p_rule_coincidences(meshpd_21, p_231)
+p_coincs321_21 = p_rule_coincidences(meshpd_21, p_321)
+
+p2_coincs231_12 = p_2_rule_coincidences(meshpd_12, p_231)
+p2_coincs321_12 = p_2_rule_coincidences(meshpd_12, p_321)
+p2_coincs231_21 = p_2_rule_coincidences(meshpd_21, p_231)
+p2_coincs321_21 = p_2_rule_coincidences(meshpd_21, p_321)
+
+unexplained231_12 = funexplained(p_coincs231_12, coincsdict231_12)
[…]

### 2016-05-01 writing/latex/coincidence.tex
Sentence ordering and autovars

@@
 
-We first define some operations on mesh patterns.
+It is useful to be able to modify a mesh pattern by adding points to an
+already existing mesh pattern. First adding a single point into a pattern.
 \begin{definition}
-Given a mesh pattern \(p\), let \(\addp[D]{p}{(a,b)}\) be the operation that returns
-a mesh pattern equivalent to placing a point in the center of box \((a,b)\),
-where \((a,b)\) is not in the mesh set of \(p\) , with shading defined by
-\(D\subseteq\{N,E,S,W\}\).
-The set \(D\) defines the shading by indicating that the boxes in the cardinal
-directions in \(D\) next to the point are shaded in the resulting pattern.
-\end{definition}
-Since there is no ambiguity we let \(\addpe[D]\) be equivalent to \(\addp[D]{\varepsilon}{(0,0)}\).
-
-\begin{example}
-    The result of adding a single point to the empty permutation for each cardinal direction.
- \begin{align*}
-     \addpe[\{N\}] = \textpattern{}{1}{0/1,1/1} &\qquad\addpe[\{E\}] =\textpattern{}{1}{1/1,1/0} \qquad \\
-     \addpe[\{S\}] = \textpattern{}{1}{0/0,1/0} &\qquad\addpe[\{W\}] =\textpattern{}{1}{0/0,0/1} \qquad
-\end{align*}
-    A more complex example for \texttt{add\_point} could be
-\begin{equation*}
-\addp[\{N,E\}]{\pattern{}{2,1,3}{0/3}}{(2,3)} = \pattern{}{2,1,4,3}{0/3,0/4,2/4,3/4,3/3}
-\end{equation*}
-\end{example}
-\begin{definition}
-Formally, given a mesh pattern \(p = (\pi,R)\) \(\addp[D]{p}{(a,b)}\) gives a mesh
+Given a mesh pattern \(p = (\pi,R)\) \(\addp[D]{p}{(a,b)}\) gives a mesh
 pattern \(p^\prime = (\pi^\prime,R^\prime)\) with length \(\setsize{\pi}+1\) defined by
@@
 
-\begin{definition}
-    Given a pattern \(p\), define \(\addd{p}{(a,b)}\), and \\ \(\addi{p}{(a,b)}\),
-    as the operations that return a mesh pattern equivalent to placing a
-    decrease, or increase, in the center of box \((a,b)\), where \((a,b)\) is not
-    in the mesh set of \(p\), in \(p\).
-\end{definition}
+Given a mesh pattern \(p\), \(\addp[D]{p}{(a,b)}\) is the operation that returns
+a mesh pattern equivalent to placing a point in the center of box \((a,b)\),
+where \((a,b)\) is not in the mesh set of \(p\) , with shading defined by
+\(D\subseteq\{N,E,S,W\}\).
+The set \(D\) defines the shading by indicating that the boxes in the cardinal
+directions in \(D\) next to the point are shaded in the resulting pattern.
+Since there is no ambiguity we let \(\addpe[D]\) be equivalent to \(\addp[D]{\varepsilon}{(0,0)}\).
 
 \begin{example}
+    The result of adding a single point to the empty permutation for each cardinal direction.
  \begin{align*}
-        \addie = \pattern{}{1,2}{0/1,1/0,1/1,1/2,2/1}\\
-        \addde = \pattern{}{2,1}{0/1,1/0,1/1,1/2,2/1}
+     \addpe[\{N\}] = \textpattern{}{1}{0/1,1/1} &\qquad\addpe[\{E\}] =\textpattern{}{1}{1/1,1/0} \qquad \\
+     \addpe[\{S\}] = \textpattern{}{1}{0/0,1/0} &\qquad\addpe[\{W\}] =\textpattern{}{1}{0/0,0/1} \qquad
 \end{align*}
+    A more complex example for \texttt{add\_point} could be
+\begin{equation*}
+\addp[\{N,E\}]{\pattern{}{2,1,3}{0/3}}{(2,3)} = \pattern{}{2,1,4,3}{0/3,0/4,2/4,3/4,3/3}
+\end{equation*}
 \end{example}
-We can define these formally as with \(\addp[D]{p}{(a,b)}\). Considering
-only adding the ascent, as adding a descent is very similar
-\begin{equation*}
-    \addi{p}{(a,b)} = (\pi^\prime, R^\prime)
-\end{equation*}
+
+It is also useful to think about adding an ascent, or descent, into a pattern
+
+\begin{definition}
+Considering
+only adding the ascent, as adding a descent is very similar.
+Given a mesh pattern \(p = (\pi, R)\) \(\adda[D]{p}{(a,b)}\) gives a mesh
+pattern \(p^\prime = (\pi^\prime,R^\prime)\) with length \(\setsize{\pi}+2\) defined by
 \begin{equation*}
@@
 \begin{equation*}
-        R^\prime = \bigcup_{(i,j)\in R}r((i,j)) \cup \text{crap to deal with strong ascent},
+        R^\prime = \{(a+1,b),(a,b),(a+1,b),(a+2,b),(a+1,b+2)\} \cup \bigcup_{(i,j)\in R}r((i,j)),
 \end{equation*}
@@
 \end{equation*}
+\end{definition}
+
+Given a pattern \(p\), \(\addd{p}{(a,b)}\), and \\ \(\adda{p}{(a,b)}\),
+are the operations that return a mesh pattern equivalent to placing a
+decrease, or increase, in the center of box \((a,b)\), where \((a,b)\) is not
+in the mesh set of \(p\), in \(p\).
+
+\begin{example}
+    \begin{align*}
+        \addae = \pattern{}{1,2}{0/1,1/0,1/1,1/2,2/1}\\
+        \addde = \pattern{}{2,1}{0/1,1/0,1/1,1/2,2/1}
+    \end{align*}
+
+    \noindent A more complex example is
+\begin{equation*}
+\adda{\pattern{}{2,3,1}{0/0,1/0,1/3,
+                        3/0,3/1,3/3}}{(1,1)}
+= \pattern{}{4,2,3,5,1}{0/0,
+                        1/0,1/2,1/5,
+                        2/0,2/1,2/2,2/3,2/5,
+                        3/0,3/2,3/5,
+                        5/0,5/1,5/2,5/3,5/5}
+\end{equation*}
+\end{example}
 
 We now attempt to fully classify coincidences in families characterised by avoidance
-of a classical pattern of length \(3\) and a mesh pattern of length \(2\).
+of a classical pattern of length \(3\) and a mesh pattern of length \(2\), that is
+finding and explaining all coincidences where \(\av{\{p,m\}} = \av{\{p,m^\prime\}}\).
 
@@
 
+From the table it can be seen that the rules established capture almost all
+coincidences. However, there are still some coincidences that are not able to
+be explained by the rules. This shows that complete coincidence classification
+of mesh patterns is a very difficult task, even when we have additional tools
+available.
 \section{Coincidence classes of Av(\{321, (21, \textit{R})\}).}
@@
     in \(R\), and a dominating classical pattern \(\pi = (\pi,\emptyset)\) if
-    \(\pi \preceq \addi{\sigma}{(a,b)}\)\\(\(\pi \preceq \addd{\sigma}{(a,b)}\))
+    \(\pi \preceq \adda{\sigma}{(a,b)}\)\\(\(\pi \preceq \addd{\sigma}{(a,b)}\))
     then in any occurrence of \(m\) in a permutation \(\varrho\) the region
@@
         \item   \begin{enumerate}
-                \item\label{prop:dom2:condc} \(\pi \preceq \addi{\sigma}{(a,b)}\) and
+                \item\label{prop:dom2:condc} \(\pi \preceq \adda{\sigma}{(a,b)}\) and
                         \begin{enumerate}

### 2016-05-01 writing/latex/wilf.tex
Sentence ordering and autovars

@@
 the permutation \(\perm{1}\). The avoiders of this pattern therefore
-have generating function \(\var(x) = C(x)-x\).\nextvar
+have generating function \(\var(x) = C(x)-x\).
 
+\nextvar
 The avoiders of the pattern \(p_1 = \textpattern{}{1}{0/1,1/0}\) can be
@@
 \begin{equation*}
-    \mathcal{\var} = \varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{\var}}{}{}{\mathcal{C}\setminus\varepsilon}
+    \scriptvar = \varepsilon \sqcup \decompmax{0/1,1/0}{\scriptvar}{}{}{\mathcal{C}\setminus\varepsilon}
 \end{equation*}
@@
 
+\nextvar[\varfine]
 It can be shown by use of \PropositionRef{prop:dom2} that the patterns
@@
 coincident. Consider the decomposition of an avoider of \(p_2\) in
-\(\av{\perm{2,3,1}}\) around the maximum\nextvar[\varullr]
+\(\av{\perm{2,3,1}}\) around the maximum
 \begin{equation*}
-    \mathcal{\var}_1 =\varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
+    \scriptvar_1 =\varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
 \end{equation*}
@@
 
+\nextvar[\varmaxl]
 There is one pattern of length \(1\) still to consider. The pattern
@@
 \begin{equation*}
-    \mathcal{H} = \varepsilon \sqcup
+    \scriptvar = \varepsilon \sqcup
     \underbrace{\decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}}_\text{
@@
     } \sqcup
-    \underbrace{\decompmax{0/0,0/1,1/0}{}{}{}{\mathcal{H}}}_\text{
+    \underbrace{\decompmax{0/0,0/1,1/0}{}{}{}{\scriptstyle{\scriptvar\setminus\varepsilon}}}_\text{
         Minimum is after the maximum, cannot be last.
@@
 \end{equation*}
-Therefore this particular class has generating function \(H(x)\) satisfying
+Therefore this particular class has generating function \(\var(x)\) satisfying
 \begin{equation*}
-    H(x) = 1 + xC(x)(C(x)-1) + x(H(x)-1)
+    \var(x) = 1 + xC(x)(C(x)-1) + x(\var(x)-1)
 \end{equation*}
@@
 \subsection{}
+\nextvar[\varminl]
 The set/symmetry classes containing the following patterns are
@@
 \begin{equation*}
-    \mathcal{I}_1 = \varepsilon \sqcup
-    \decompmax{1/0,0/1}{\mathcal{I}_1}{}{}{\mathcal{G}_1}
+    \scriptvar_1 = \varepsilon \sqcup
+    \decompmax{1/0,0/1}{\scriptvar_1}{}{}{\mathcal{\varmaxl}_1}
 \end{equation*}
@@
 \begin{equation*}
-    \mathcal{I}_2 = \varepsilon \sqcup
-    \decompmax{1/0,0/1}{\mathcal{I}_2}{}{}{\mathcal{G}_3}
+    \scriptvar_2 = \varepsilon \sqcup
+    \decompmax{1/0,0/1}{\scriptvar_2}{}{}{\mathcal{\varmaxl}_3}
 \end{equation*}
@@
 in the region and the maximum, if and only if this first point is the minimum.
-Since both \(\mathcal{G}_1\) and \(\mathcal{G}_3\) have the same
-enumeration, \(\mathcal{I}_1\) and \(\mathcal{I}_2\) must also have
+Since both \(\mathcal{\varmaxl}_1\) and \(\mathcal{\varmaxl}_3\) have the same
+enumeration, \(\scriptvar_1\) and \(\scriptvar_2\) must also have
 the same enumeration and are therefore Wilf-equivalent.
@@
 \begin{equation*}
-    \mathcal{I}_3 = \varepsilon \sqcup
-    \decompleft{1/0,0/1}{\mathcal{I}_3}{}{}{\mathcal{G}_3}
+    \scriptvar_3 = \varepsilon \sqcup
+    \decompleft{1/0,0/1}{\scriptvar_3}{}{}{\mathcal{\varmaxl}_3}
 \end{equation*}
@@
 \begin{equation*}
-    I(x) = 1 + xI(x)G(x)
+    \var(x) = 1 + x\var(x)\varmaxl(x)
 \end{equation*}
@@
 \subsection{}
+\nextvar
 The classes containing the following patterns are Wilf-equivalent
@@
 \begin{equation*}
-    \mathcal{J}_1 = \varepsilon \sqcup
+    \scriptvar_1 = \varepsilon \sqcup
     \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}} \sqcup
-    \decompmax{0/1,1/0,0/0}{}{}{}{\mathcal{F}}
+    \decompmax{0/1,1/0,0/0}{}{}{}{\mathcal{\varfine}}
 \end{equation*}
@@
 \begin{equation*}
-    \mathcal{J}_2 = \varepsilon \sqcup
+    \scriptvar_2 = \varepsilon \sqcup
     \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\mathcal{C}\setminus\varepsilon} \sqcup
-    \decompmax{0/1,1/0,1/1}{\mathcal{F}}{}{}{}
+    \decompmax{0/1,1/0,1/1}{\mathcal{\varfine}}{}{}{}
 \end{equation*}
@@
 \begin{equation*}
-    J(x) = 1 + xC(x)(C(x)-1) + xF(x)
+    \var(x) = 1 + xC(x)(C(x)-1) + x\varfine(x)
 \end{equation*}
@@
 \subsection{}
+\nextvar
 Consider the containers of the patterns
@@
 \begin{equation*}
-    K_n = C_n - C_{n-2}
+    \var_n = C_n - C_{n-2}
 \end{equation*}
@@
 \subsection{}
+\nextvar
 Now consider the containers of the patterns
@@
 \begin{equation*}
-    \mathcal{L}_1 =
+    \scriptvar_1 =
     \varepsilon \sqcup
-    \decompmax{0/1,1/0}{\scriptstyle \mathcal{G}_1\setminus\varepsilon}{}{}{\mathcal{C}}
+    \decompmax{0/1,1/0}{\scriptstyle \mathcal{\varmaxl}_1\setminus\varepsilon}{}{}{\mathcal{C}}
     \sqcup
-    \decompmax{0/0,0/1,1/0}{}{}{}{\mathcal{L}_1}
+    \decompmax{0/0,0/1,1/0}{}{}{}{\scriptvar_1}
 \end{equation*}
@@
 \begin{equation*}
-    \mathcal{L}_2 =
+    \scriptvar_2 =
     \varepsilon \sqcup
@@
     \sqcup
-    \decompleft{0/1,1/0,1/1}{\mathcal{L}_2}{}{}{}
+    \decompleft{0/1,1/0,1/1}{\scriptvar_2}{}{}{}
 \end{equation*}
@@
 \begin{align}
-    L_1(x) &= 1 + xC(x)(G(x)-1) +xL_1(x) \label{eqn:genL1}\\
-    \text{ and }L_2(x) &= 1 + x(C(x)-1)^2 + xL_2(x) \label{eqn:genL2}
+    \var_1(x) &= 1 + xC(x)(\varmaxl(x)-1) +x\var_1(x) \label{eqn:genL1}\\
+    \text{ and }\var_2(x) &= 1 + x(C(x)-1)^2 + x\var_2(x) \label{eqn:genL2}
 \end{align}
@@
 \eqref{eqn:genL1} and \eqref{eqn:genL2} are equal, this occurs if
-\(C(x)(G(x)-1) = (C(x)-1)^2\).
+\(C(x)(\varmaxl(x)-1) = (C(x)-1)^2\).
 \begin{align*}
-    &&(C(x)-1)^2 &= C(x)(G(x)-1)\\
-        &\Leftrightarrow& x^2C^4(x) &=  xC(x)(C(x)-1)C(x) &&\text{By definition of \(G\) and \(C\)}\\
+    &&(C(x)-1)^2 &= C(x)(\varmaxl(x)-1)\\
+        &\Leftrightarrow& x^2C^4(x) &=  xC(x)(C(x)-1)C(x) &&\text{By definition of \(\varmaxl\) and \(C\)}\\
         &\Leftrightarrow& x^2C^4(x)&=xC^3(x)-xC^2(x)\\
@@
 
-Evaluating the generating function \(L(x)\) gives us the enumeration
+Evaluating the generating function \(\var(x)\) gives us the enumeration
 \begin{equation*}
@@
 are Wilf-equivalent.
-If \(\mathcal{M}_1\) is the set of avoiders of
+If \(\scriptvar_1\) is the set of avoiders of
 \textpattern{}{1,2}{0/0,0/2,1/0,1/2,2/0,2/2}, then by the structural
@@
 \begin{equation*}
-    \mathcal{M}_1 = \varepsilon \sqcup
+    \scriptvar_1 = \varepsilon \sqcup
     \decompmax{0/0,0/1,1/0}{}{}{}{\mathcal{C}}
 \end{equation*}
-This is because \(\mathcal{M}_1\) is the set of permutations
+This is because \(\scriptvar_1\) is the set of permutations
 who have their minimum occur after their maximum.
@@
 block.  This ensures that the last block is never a singleton. This means
-that these permutations are also counted by \(M(x) = 1+xC(x)\).
+that these permutations are also counted by \(\var(x) = 1+xC(x)\).
 
@@
 \begin{equation*}
-    \mathcal{M}_3 = \varepsilon \sqcup
-    \decompmax{0/1,1/0}{\mathcal{F}}{}{}{\mathcal{M}_3}
+    \scriptvar_3 = \varepsilon \sqcup
+    \decompmax{0/1,1/0}{\mathcal{\varfine}}{}{}{\scriptvar_3}
 \end{equation*}
-The generating function derived satisfies \(M_3(x) = 1 + xF(x)M_3(x)\).
+The generating function derived satisfies \(\var_3(x) = 1 + x\varfine(x)\var_3(x)\).
 The fourth pattern can be decomposed around the maximum in a similar
@@
 \begin{equation*}
-    \mathcal{M}_4 = \varepsilon \sqcup
-    \decompmax{0/1,1/0}{\mathcal{M}_4}{}{}{\mathcal{F}}
+    \scriptvar_4 = \varepsilon \sqcup
+    \decompmax{0/1,1/0}{\scriptvar_4}{}{}{\mathcal{\varfine}}
 \end{equation*}
@@
 position would either create an occurrence of \(\perm{2,3,1}\) or the pattern.
-Therefore this is also counted by \(M(x) = 1+xC(x)\).
+Therefore this is also counted by \(\var(x) = 1+xC(x)\).
 
-We need to show that the generating function \(M_3(x)\) is the same as \(M(x)\)
+We need to show that the generating function \(\var_3(x)\) is the same as \(\var(x)\)
 \begin{equation*}
     \begin{aligned}
-        M_3(x) &= 1 + xF(x)M_3(x) \\
-        &= \frac{1}{1-xF(x)} && \text{Solving for \(M_3(x)\)}\\
-        &= \frac{1}{1 - \frac{x}{1+x-xC(x)}} &&\text{Substituting for \(F(x)\)}\\
+        \var_3(x) &= 1 + x\varfine(x)\var_3(x) \\
+        &= \frac{1}{1-x\varfine(x)} && \text{Solving for \(\var_3(x)\)}\\
+        &= \frac{1}{1 - \frac{x}{1+x-xC(x)}} &&\text{Substituting for \(\varfine(x)\)}\\
         &= \frac{1-xC(x) +x}{1-xC(x)}\\
@@
 \end{equation*}
-We have that \(M_3(x) = 1+xC(x)= M(x) \) so all four patterns are Wilf-equivalent
+We have that \(\var_3(x) = 1+xC(x)= \var(x) \) so all four patterns are Wilf-equivalent
  and have enumeration sequence
@@
 \subsection{}
+\nextvar
 The patterns
@@
 \begin{equation*}
-    \mathcal{N}_2 = \varepsilon \sqcup
-    \decompmax{0/1,1/0}{\scriptstyle\mathcal{N}_2\setminus\varepsilon}{}{}{\mathcal{C}}
+    \scriptvar_2 = \varepsilon \sqcup
+    \decompmax{0/1,1/0}{\scriptstyle\scriptvar_2\setminus\varepsilon}{}{}{\mathcal{C}}
     \sqcup
-    \decompmax{0/0,0/1,1/0}{}{}{}{\mathcal{F}}
+    \decompmax{0/0,0/1,1/0}{}{}{}{\mathcal{\varfine}}
 \end{equation*}
@@
 \begin{equation*}
-    \mathcal{N}_1 = \varepsilon \sqcup
-    \decompleft{0/1,1/0}{\scriptstyle\mathcal{N}_1\setminus\varepsilon}{}{}{\mathcal{C}}
+    \scriptvar_1 = \varepsilon \sqcup
+    \decompleft{0/1,1/0}{\scriptstyle\scriptvar_1\setminus\varepsilon}{}{}{\mathcal{C}}
     \sqcup
-    \decompleft{0/0,0/1,1/0}{}{}{}{\mathcal{F}}
+    \decompleft{0/0,0/1,1/0}{}{}{}{\mathcal{\varfine}}
 \end{equation*}
@@
 \begin{equation}
-    N_1 = 1 + x(N_1(x)-1)C(x) +xF(x)
+    \var_1 = 1 + x(\var_1(x)-1)C(x) +x\varfine(x)
     \label{eqn:ngen}
 \end{equation}
-Now consider an avoider of \(m_4\) decomposed around the maximum
+Now consider an avoider of \(m_4\) decomposed around the maximum\nextvar %subvar
 \begin{equation*}
-    \mathcal{N}_4 = \varepsilon \sqcup
-    \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\mathcal{O}}
+    \mathcal{\lastvar}_4 = \varepsilon \sqcup
+    \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\scriptvar}
 \end{equation*}
-Where \(\mathcal{O}\) is the permutations avoiding \(\perm{2,3,1},m_4\) and
+Where \(\scriptvar\) is the permutations avoiding \(\perm{2,3,1},m_4\) and
 \(p = \textpattern{}{1}{0/0,0/1,1/1}\) since if the subsequence in this box
@@
 an occurrence of \(m_4\).
-Now consider decomposition of a permutation in \(\mathcal{O}\) around its
+Now consider decomposition of a permutation in \(\scriptvar\) around its
 leftmost point.
 \begin{equation*}
-    \mathcal{O} = \varepsilon \sqcup
-    \decompleft{0/1,1/0}{\mathcal{C}}{}{}{\scriptstyle\mathcal{N}_4\setminus\varepsilon}
+    \scriptvar = \varepsilon \sqcup
+    \decompleft{0/1,1/0}{\mathcal{C}}{}{}{\scriptstyle\mathcal{\lastvar}_4\setminus\varepsilon}
 \end{equation*}
@@
 \begin{equation*}
-    O(x) = 1 + xC(x)(N_4(x)-1)
+    \var(x) = 1 + xC(x)(\lastvar_4(x)-1)
 \end{equation*}
-Now we can construct the following for \(N_4\)
+Now we can construct the following for \(\lastvar_4\)
 \begin{equation}
-    N_4(x) = 1 +xC(x)(xC(x)(N_4(x)-1)+1)
+    \lastvar_4(x) = 1 +xC(x)(xC(x)(\lastvar_4(x)-1)+1)
     \label{eqn:ngen2}
@@
 \EquationRef{eqn:ngen2} are the same generating function.
-First solve \EquationRef{eqn:ngen2} for \(N_4(x)\)
+First solve \EquationRef{eqn:ngen2} for \(\lastvar_4(x)\)
 \begin{equation}
     \begin{aligned}
-        N_4(x) &= 1+xC(x)(xC(x)(N_4(x)-1)+1)\\
-        &= 1 + x^2C^2(x)N_4(x) -x^2C^2(x) +xC(x)\\
+        \lastvar_4(x) &= 1+xC(x)(xC(x)(\lastvar_4(x)-1)+1)\\
+        &= 1 + x^2C^2(x)\lastvar_4(x) -x^2C^2(x) +xC(x)\\
         &= 1 + \frac{xC(x)}{1 - x^2C^2(x)}\\
         &= 1 + \frac{xC(x)}{(1-xC(x))(1+xC(x))} && \text{Difference of squares}\\
-        N_4(x)&= 1 + \frac{xC^2(x)}{1+xC(x)} && C(x) = \frac{1}{1-xC(x)}
+        \lastvar_4(x)&= 1 + \frac{xC^2(x)}{1+xC(x)} && C(x) = \frac{1}{1-xC(x)}
     \end{aligned}
@@
 
-Now we solve \EquationRef{eqn:ngen} for \(N_1(x)\)
+Now we solve \EquationRef{eqn:ngen} for \(\lastvar_1(x)\)
 \begin{equation}
     \begin{aligned}
-        N_1(x)&= 1+x(N_1(x)-1)C(x)+xF(x)\\
-        &= 1+xN_1(x)C(x)-xC(x)+\frac{x}{1+x-xC(x)} &&\text{Substitution of \(F(x)\)}\\
-        N_1(x)(1-xC(x))&= \frac{x^2C^2(x) - (x^2+2x)C(x) + 2x+1}{1+x-xC(x)}\\
-        N_1(x) &= \frac{x^2C^2(x) - (x^2+2x)C(x) + x+1+x}{x^2C^2(x) - (x^2+2x)C(x) +x+1}\\
+        \lastvar_1(x)&= 1+x(\lastvar_1(x)-1)C(x)+x\varfine(x)\\
+        &= 1+x\lastvar_1(x)C(x)-xC(x)+\frac{x}{1+x-xC(x)} &&\text{Substitution of \(\varfine(x)\)}\\
+        \lastvar_1(x)(1-xC(x))&= \frac{x^2C^2(x) - (x^2+2x)C(x) + 2x+1}{1+x-xC(x)}\\
+        \lastvar_1(x) &= \frac{x^2C^2(x) - (x^2+2x)C(x) + x+1+x}{x^2C^2(x) - (x^2+2x)C(x) +x+1}\\
         &= 1 +\frac{x}{x^2C^2(x) - (x^2+2x)C(x) +x+1}\\
@@
         &= 1 + \frac{xC^2(x)}{C(x) - xC^2(x) + xC(x)} && xC^2(x) = C(x)-1\\
-        N_1(x)&= 1 + \frac{xC^2(x)}{1+xC(x)} && C(x) = 1+xC^2(x)\\
+        \lastvar_1(x)&= 1 + \frac{xC^2(x)}{1+xC(x)} && C(x) = 1+xC^2(x)\\
     \end{aligned}
 \end{equation}
-We have shown that \(N_1\) and \(N_4\) are indeed the same generating function,
+We have shown that \(\lastvar_1\) and \(\lastvar_4\) are indeed the same generating function,
 and we have that the classes containing these four patterns are Wilf-equivalent.
-Evaluating the generating function \(N(x)\) gives
+Evaluating the generating function \(\lastvar(x)\) gives
 \begin{equation*}
@@
 \subsection{}
+\nextvar
 \begin{equation*}
@@
 \end{equation*}
-Let \(\mathcal{P}_1\) be the set of avoiders of \(m_1\), by structural decomposition
+Let \(\scriptvar_1\) be the set of avoiders of \(m_1\), by structural decomposition
 around the leftmost point we have
 \begin{equation*}
-    \mathcal{P}_1 = \varepsilon \sqcup
-    \decompleft{0/1,1/0}{\mathcal{P}_1}{}{}{\mathcal{P}^\prime_1}
+    \scriptvar_1 = \varepsilon \sqcup
+    \decompleft{0/1,1/0}{\scriptvar_1}{}{}{\scriptvar^\prime_1}
 \end{equation*}
-Where \(\mathcal{P}^\prime_1\) is a permutation avoiding \(\perm{2,3,1}, m_1\)
+Where \(\scriptvar^\prime_1\) is a permutation avoiding \(\perm{2,3,1}, m_1\)
 and \(\textpattern{}{1}{0/0,0/1,1/1}\)
-Now consider the decomposition of a permutation in \(\mathcal{P}^\prime_1\) it can
+Now consider the decomposition of a permutation in \(\scriptvar^\prime_1\) it can
 once again be decomposed around the leftmost point
 \begin{equation*}
-    \mathcal{P}^\prime_1 = \varepsilon \sqcup
-    \decompleft{0/1,1/0}{\scriptstyle \mathcal{P}_1\setminus\varepsilon}{}{}{\mathcal{P}^\prime_1}
+    \scriptvar^\prime_1 = \varepsilon \sqcup
+    \decompleft{0/1,1/0}{\scriptstyle \scriptvar_1\setminus\varepsilon}{}{}{\scriptvar^\prime_1}
 \end{equation*}
@@
 \begin{equation*}
-    \mathcal{P}_2 = \varepsilon \sqcup
-    \decompleft{0/1,1/0}{\mathcal{P}_2}{}{}{\mathcal{P}^\prime_2}
+    \scriptvar_2 = \varepsilon \sqcup
+    \decompleft{0/1,1/0}{\scriptvar_2}{}{}{\scriptvar^\prime_2}
 \end{equation*}
 
-Again we use the same method of decomposition of a permutation in \(\mathcal{P}^\prime_2\)
+Again we use the same method of decomposition of a permutation in \(\scriptvar^\prime_2\)
 \begin{equation*}
-    \mathcal{P}^\prime_2 = \varepsilon \sqcup
-    \decompleft{0/1,1/0}{\scriptstyle \mathcal{P}_2\setminus\varepsilon}{}{}{\mathcal{P}^\prime_2}
+    \scriptvar^\prime_2 = \varepsilon \sqcup
+    \decompleft{0/1,1/0}{\scriptstyle \scriptvar_2\setminus\varepsilon}{}{}{\scriptvar^\prime_2}
 \end{equation*}
 
-This gives us a generating function \(P(x)\) satisfying
+This gives us a generating function \(\var(x)\) satisfying
 \begin{align}
-    P(x) &= 1 + xP(x)P^\prime(x) \label{eqn:Pgen}\\
-    P^\prime(x) &= 1 + x(P(x)-1)P^\prime(x)\label{eqn:PprimeGen}
+    \var(x) &= 1 + x\var(x)\var^\prime(x) \label{eqn:Pgen}\\
+    \var^\prime(x) &= 1 + x(\var(x)-1)\var^\prime(x)\label{eqn:PprimeGen}
 \end{align}
 
-Solving \EquationRef{eqn:PprimeGen} for \(P^\prime(x)\) and substituting into
+Solving \EquationRef{eqn:PprimeGen} for \(\var^\prime(x)\) and substituting into
 \EquationRef{eqn:Pgen} gives us the fact that a the generating function for
-\(P(x)\) satisfies
+\(\var(x)\) satisfies
 \begin{equation}
-    P(x) = xP^2(x) - x(P(x) - 1) + 1
+    \var(x) = x\var^2(x) - x(\var(x) - 1) + 1
 \end{equation}
-Evaluating \(P(x)\) gives us the sequence
+Evaluating \(\var(x)\) gives us the sequence
 \begin{equation*}
@@
 \subsection{}
+\nextvar
 \begin{gather*}
@@
 \begin{equation*}
-    \mathcal{Q}_1 = \varepsilon \sqcup
+    \scriptvar_1 = \varepsilon \sqcup
     \tikz{\filldraw (0,0) circle (3pt);}
@@
 \begin{equation}
-    Q(x) = 1 + x + xC(x)(C(x)-1) \label{eq:qgen}
+    \var(x) = 1 + x + xC(x)(C(x)-1) \label{eq:qgen}
 \end{equation}
@@
 \begin{equation*}
-    \mathcal{Q}_2 = \varepsilon \sqcup
-    \decompmax{0/1,1/0}{\mathcal{G}_2}{}{}{\mathcal{C}}
+    \scriptvar_2 = \varepsilon \sqcup
+    \decompmax{0/1,1/0}{\mathcal{\varmaxl}_2}{}{}{\mathcal{C}}
 \end{equation*}
@@
     \begin{aligned}
-        Q_2(x) &= 1+xC(x)G(x)\\
+        \var_2(x) &= 1+xC(x)\varmaxl(x)\\
         &= 1+xC(x)(1+xC(x)(C(x)-1))\\
@@
 and \textpattern{}{1,2}{0/0,0/1,0/2,1/1,1/2,2/0,2/2} around the leftmost point
-into an avoider of one of the patterns with generating function \(G(x)\) and
+into an avoider of one of the patterns with generating function \(\varmaxl(x)\) and
 an avoider of \(\perm{2,3,1}\).
@@
 \begin{equation*}
-    \mathcal{Q}_3 = \varepsilon \sqcup
-    \decompleft{0/1,1/0}{\scriptstyle \mathcal{Q}\setminus\varepsilon}{}{}{\mathcal{C}}
-    \sqcup \decompleft{0/0,0/1,1/0}{}{}{}{\mathcal{G}_3}
+    \scriptvar_3 = \varepsilon \sqcup
+    \decompleft{0/1,1/0}{\scriptstyle \scriptvar_3\setminus\varepsilon}{}{}{\mathcal{C}}
+    \sqcup \decompleft{0/0,0/1,1/0}{}{}{}{\varmaxl_3}
 \end{equation*}
-This gives generating function \(Q_3(x)\) satisfying
+This gives generating function \(\var_3(x)\) satisfying
 \begin{equation*}
     \begin{aligned}
-        Q(x) &= 1 +xC(x)(P(x)-1) + xG(x)\\
-        &= C(x) - xC^2(x) + xC(x) +x^2C^3(x) - x^2C^2 && \text{Solving for \(Q(x)\)}\\
+        \var_3(x) &= 1 +xC(x)(\var_3(x)-1) + x\varmaxl(x)\\
+        &= C(x) - xC^2(x) + xC(x) +x^2C^3(x) - x^2C^2 && \text{Solving for \(\var_3(x)\)}\\
         &= 1 + xC(x) +x^2C^3(x)-xC^2(x) && C = 1+xC^2(x)\\

### 2016-05-03 writing/latex/coincidence.tex
Make Edits. Temp remove code because borked computer

@@
 \chapter[Coincidence Classification]{Coincidences amongst families of mesh patterns and classical patterns}\label{chap:coincs}
-One interesting question to ask about permutation patterns considers when a
+One interesting question to ask about permutation patterns is to consider when a
 pattern may be avoided by, or contained in, arbitrary permutations. Two
@@
 Through experimentation, up to permutations of length 11, we discover that there
-are at most 29 coincidence classes of mesh patterns with underlying classical
+are at least 29 coincidence classes of mesh patterns with underlying classical
 pattern \(\perm{2,1}\).
@@
 By application of \PropositionRef{prop:dom1} we obtain \(43\) coincidence
-classes. Experimentation shows that there are in fact at most \(39\) coincidence
+classes. Experimentation shows that there are in fact at least \(39\) coincidence
 classes, for example the following two patterns are coincident in
@@
             is coincident with a mesh pattern containing an occurrence of
-            \(\mperm{\sigma}{R_2}\).
+            \(\mperm{\sigma}{R_2}\) as a subpattern. 
     \end{enumerate}

### 2016-05-03 writing/latex/wilf.tex
Make Edits. Temp remove code because borked computer

@@
 
-Another question often asked in thefield of permutation patterns is that
+Another question often asked in the field of permutation patterns is that
 of Wilf-equivalence. Two patterns \(\pi\) and \(\sigma\) are said to be
@@
 The group of symmetries on permutations is isomorphic to the dihedral group
-of order \(8\), the group of symmetries of a square. Composition of the above
-symmetries gives the remaining 5 symmetries. If we consider generators of the
-group the operations \emph{reverse-inverse} and \emph{reverse} correspond to
-the generators of the dihedral group.
+of order \(8\), the group of symmetries of a square. \emph{Reverse-inverse} and
+\emph{reverse} correspond tothe generators of the dihedral group.
 
@@
 whilst avoiding the \emph{Dominating Pattern} \(\perm{2,3,1}\). We will
-use \(\mathcal{C}\) to denote the set of these avoiders and \(C(x)\) will
+use \(\mathcal{C}\) to denote \(\av{\perm{2,3,1}}\) and \(C(x)\) will
 be the usual Catalan generating function satisfying \(C(x) = 1 + C(x)^2\).
@@
 structure of a \(\perm{2,3,1}\) avoiding permutation. Furthermore, all the
-elements in \(A\) lie below all of the elements in \(B\).
+elements in \(A\) lie below all of the elements in \(B\). We call \(A\) the
+\emph{lower-left region} and \(B\) the \emph{upper-right region}.
 
@@
 the Wilf-equivalence classes of the mesh patterns of length \(1\) inside
-\(\av{\perm{2,3,1}}\).
+\(\av{\perm{2,3,1}}\), this means that we are sonsdering the set \(\av{\{231,p\}}\)
+where \(p\) is a mesh-pattern.
 
@@
 \nextvar
-The avoiders of the pattern \(p_1 = \textpattern{}{1}{0/1,1/0}\) can be
-decomposed around the maximum element in order to give the following
-structural decomposition.
+The pattern \(p_1 = \textpattern{}{1}{0/1,1/0}\) is one of the
+quadrant marked mesh patterns studied by Remmel,
+Kitaev and Tiefenbruck\cite{2012arXiv1201.6243K}. Alternatively we can enumerate
+avoiders by decomposing a typical avoider of \(p_1\) around the maximum element
+in order to give the following structural decomposition.
 \begin{equation*}
@@
 an occurrence since the maximum lies in a region corresponding to the
-shading in \(p_1\). The lower right region however can create occurrences
+shading in \(p_1\). The lower left region however can create occurrences
 of \(p_1\) and therefore must also avoid \(p_1\) as well as \(\perm{2,3,1}\).
@@
 
-Evauluation of this generating function gives a the Fine numbers
+Evauluation of this generating function gives the Fine numbers
 (\href{https://oeis.org/A000957}{OEIS: A000957}).
@@
 \end{equation*}
-This pattern is one of the
-quadrant marked mesh patterns studied by Remmel, Kitaev and Tiefenbruck\cite{2012arXiv1201.6243K}.
 
@@
 \(\perm{2,3,1}\) is that of \emph{reverse-complement-inverse}. Using this
-set of symmetries to merge classes gives us \(61\) classes of trivial
+symmetry to merge classes gives us \(61\) classes of trivial
 Wilf-equivalences.
@@
     \label{not:setequiv}
-    The avoiders of the pattern \(\mperm{2,3,1}{\{(1,0),(1,1),(1,2),(1,3)\}}\)
+    The avoiders of the pattern \(q = \mperm{2,3,1}{\{(1,0),(1,1),(1,2),(1,3)\}}\)
     in \(\mathfrak{S}_n\) are in one-to-one correspondence with partitions of
@@
 \begin{proof}
-    Let \(\pi\) be a permutation in \(\av[n]{\perm{2,3,1}}\) take the permutation
+    Let \(\pi\) be a permutation in \(\av[n]{q}\) take the permutation
     in one-line notation and insert a dash between each ascent in \(\pi\). This
@@
 
+The classes containing the following patterns are Wilf-equivalent
+\begin{equation*}
+    \pattern{}{1,2}{} \text{ and }
+    \pattern{}{2,1}{}
+\end{equation*}
+This is true since the only avoiders of these patterns are the decreasing
+sequence and the increasing sequence respectively, and both of these
+avoid \(\perm{2,3,1}\) in all cases. There is therefore \(1\) avoider at
+every length.
+
+\nextvar[\varminl]
+The classes containing the following patterns are Wilf-equivalent
+in \(\av{\perm{2,3,1}}\)
+\begin{equation*}
+    m_1 = \pattern{}{1,2}{0/0,0/1,0/2,
+                    1/0,1/1,1/2,
+                    2/0,2/1,2/2}
+    \text{ and }
+    m_2 = \pattern{}{2,1}{0/0,0/1,0/2,
+                    1/0,1/1,1/2,
+                    2/0,2/1,2/2}
+\end{equation*}
+It is obvious that these two are Wilf-equivalent since the only
+permutations that contain these patterns are \(\perm{1,2}\) and
+\(\perm{2,1}\) respectively, therefore the avoiders of these patterns
+are counted by the Catalan numbers at all lengths except for length
+\(2\) where there is precisely \(1\) avoider.
+Therefore the generating function is
+\begin{equation*}
+  \var(x) = C(x)-x^2
+\end{equation*}
+
 \subsection{}
-\nextvar[\varminl]
+\nextvar
+Consider the containers of the patterns
+\begin{equation*}
+    \pattern{}{1,2}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
+    \pattern{}{1,2}{1/1,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
+    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,1/1,2/1,2/2},
+    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,1/1,2/2},
+    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,1/1},
+    \pattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0,2/1,2/2}
+\end{equation*}
+
+For each of these patterns there is precisely one occurrence
+in any permutation containing the pattern. Now consider
+the points in the region corresponding to the unshaded box in each case. Each
+must contain an avoider of \(\perm{2,3,1}\) that is of length
+\(n-2\). Therefore these classes are all Wilf-equivalent
+and the number of length \(n\) avoiders is
+\begin{equation*}
+    \var_n = C_n - C_{n-2}
+\end{equation*}
+for \(n \ge 2\) where \(C_n\) is the \(n\)th Catalan number, the number of
+\(\perm{2,3,1}\) avoiders of length \(n\). This gives the sequence
+\begin{equation*}
+    1, 1, 1, 4, 12, 37, 118, 387, 1298, 4433, 15366,\dotsc
+\end{equation*}
+
+\subsection{}
+Now consider the containers of the patterns
+\begin{equation*}
+    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
+    \pattern{}{1,2}{0/0,1/1,0/2,1/0,1/2,2/0,2/1,2/2},
+    \pattern{}{1,2}{0/0,0/1,1/1,1/0,1/2,2/0,2/1,2/2}
+\end{equation*}
+Each of these patterns again occurs precisely once in
+any containing permutation. However this time when considering
+the region corresponding to the unshaded box we need to take into consideration
+\LemmaRef{lem:incdec}
+and so the empty box can only contain a decreasing subsequence. There
+is precisely one decreasing subsequence at every length, and so there
+is exactly one container of each pattern at each length. The three
+patterns are Wilf-equivalent and have \(C_n - 1\) avoiders of
+length \(n\) for all \(n \ge 2\).
+This gives the sequence
+\begin{equation*}
+    1, 1, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc\tag{\href{https://oeis.org/A001453}{OEIS: A001453} offset 2}
+\end{equation*}
+
+\subsection{}
+The containers of the following patterns
+have exactly one occurrence.
+\begin{equation*}
+    \pattern{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
+    \pattern{}{2,1}{0/0,0/1,1/0,1/1,1/2,2/0,2/2}
+    \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/1,2/2}
+    \pattern{}{1,2}{0/1,0/2,1/0,1/1,2/0,2/1,2/2}
+    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1}
+    \pattern{}{1,2}{0/0,0/1,1/0,1/2,2/0,2/1,2/2}
+\end{equation*}
+Once again we consider the regions corresponding to the unshaded regions, For each
+pattern
+except the first the two regions are independent, and one contains
+any avoider of \(\perm{2,3,1}\) and the other must contain a
+decreasing sequence by \LemmaRef{lem:incdec}.
+Let us consider the first pattern separately. In order to
+avoid \(\perm{2,3,1}\) across the regions corresponding to the unshaded boxes
+we can add some additional restrictions
+\begin{equation}
+    \raisebox{0.6ex}{
+    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
+        \fill[pattern=north east lines, pattern color=black!75] (1.5, 0) rectangle +(0.5,1);
+        \fill[pattern=north east lines, pattern color=black!75] (1.0, 2) rectangle +(0.5,1);
+        \draw (1.5,0) -- (1.5,1)
+              (1.5,2) -- (1.5,3);
+        \modpattern[2]{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
+    \end{tikzpicture}
+    }
+    \label{eqn:splitboxes}
+\end{equation}
+Now we can see that the region corresponding to the top free box must contain a decreasing sequence,
+and the bottom must contain an avoider of \(\perm{2,3,1}\) and these
+two do not interact in any manner. The containers of this pattern are counted
+the same as the other patterns, and due to this they are Wilf-equivalent in \(\av{231}\).
+The containers
+have generating function \(\frac{x^2C(x)}{(1-x)}\). Enumerating avoiders
+therefore gives us
+\begin{equation*}
+    1, 1, 1, 3, 10, 33, 109, 364, 1233, 4236, 14740,\dotsc
+\end{equation*}
+
+\subsection{}
 The set/symmetry classes containing the following patterns are
@@
 be one occurrence of this pattern in a permutation corresponding to
-the last point in the permutation and the minimum. The only form that
-points in either of the two boxes can take is a decreasing sequence.
+the last point in the permutation and the minimum.Consider an occurrence of \(m_1\),
+the points in the two regions corresponding to the the two boxes must form decreasing
+subsequences.
 For a permutation of length \(k\) if we fix the number of points in
@@
 Therefore these two patterns have been shown to have the same number of
-avoiders of length \(k\) for all \(k\) and therefore all Wilf-equivalent.
+avoiders of length \(k\) for all \(k\) and are Wilf-equivalent.
 The avoiders have enumeration
@@
 
-Due to the shading we know the following about the points corresponding
+Considering an occurrence of either of these patterns in a permutation by
+the mesh sets we know the following about the points corresponding
 to the points in the patterns.
@@
         the permutation).
-    \item The block containing the point corresponding to the second
-        point in both patterns contains only the point (it is a singleton
-        block).
+    \item If the region corresponding to box \((2,2)\) in an occurrence of
+    \(m_1\) is empty, then the point corresponding to the second point is
+    precisely the last block bottom. If the region corresponding to box \((0,1)\)
+    in an occurrence of \(m_2\) is empty, then the point corresponding to the
+    second point is precisely the first block bottom.
+    If these regions are non-empty then the block containing the point
+    corresponding to the second point in both patterns contains only the point
+    (it is a singleton block).
 \end{itemize}
@@
 Looking at avoiders of \(\perm{2,3,1}\) and \(m_4\) we
-can perform the same decomposition around the maximum to get
+can perform a similar decomposition around the maximum to get
 \begin{equation*}
@@
 \begin{equation*}
-    m_1 = \pattern{}{1,2}{0/0,0/1,0/2,
-                    1/0,1/1,1/2,
-                    2/0,2/1,2/2}
-    \text{ and }
-    m_2 = \pattern{}{2,1}{0/0,0/1,0/2,
-                    1/0,1/1,1/2,
-                    2/0,2/1,2/2}
-\end{equation*}
-It is obvious that these two are Wilf-equivalent since the only
-permutations that contain these patterns are \(\perm{1,2}\) and
-\(\perm{2,1}\) respectively, therefore the avoiders of these patterns
-are counted by the Catalan numbers at all lengths except for length
-\(2\) where there is precisely \(1\) avoider.
-Therefore there are
-\begin{equation*}
-    1, 1, 1, 5, 14, 42, 132, 429, 1430, 4862, 16796,\cdots
-\end{equation*}
-
-\subsection{}
-
-The classes containing the following patterns are Wilf-equivalent
-in \(\av{\perm{2,3,1}}\)
-\begin{equation*}
     m_1 = \pattern{}{2,1}{0/0,0/1,0/2,
@@
 of \(\perm{2,3,1}\) where we consider decomposition around the maximum.
-Any permutation in \(\av{\perm{2,3,1},m_1}\) is in the following set
-precisely once.
+% Any permutation in \(\av{\perm{2,3,1},m_1}\) is in the following set
+% precisely once.
+If \(\scriptvar_1\) set \(\av{\perm{2,3,1},m_1}\) then any permutation in
+\(\scriptvar_1\) either starts with a maximum or does not, giving us the
+decomposition
 \begin{equation*}
@@
 Now consider the decomposition around the maximum of a permutation in
-\(\av{\perm{2,3,1},m_2}\) This fits into the following set.
+\(\scriptvar_2 = \av{\perm{2,3,1},m_2}\), the permutation either ends with
+the maximum, or it does not, so we get
 \begin{equation*}
@@
 \nextvar
-Consider the containers of the patterns
-\begin{equation*}
-    \pattern{}{1,2}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
-    \pattern{}{1,2}{1/1,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
-    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,1/1,2/1,2/2},
-    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,1/1,2/2},
-    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,1/1},
-    \pattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0,2/1,2/2}
-\end{equation*}
-
-For each of these patterns there is precisely one occurrence
-in any permutation containing the pattern. Now consider
-the points in the free box in each case. Each of these regions
-must contain an avoider of \(\perm{2,3,1}\) that is of length
-\(n-2\). Therefore these classes are all Wilf-equivalent
-and the number of length \(n\) avoiders is
-\begin{equation*}
-    \var_n = C_n - C_{n-2}
-\end{equation*}
-for \(n \ge 2\) where \(C_n\) is the \(n\)th Catalan number, the number of
-\(\perm{2,3,1}\) avoiders of length \(n\). This gives the sequence
-\begin{equation*}
-    1, 1, 1, 4, 12, 37, 118, 387, 1298, 4433, 15366,\dotsc
-\end{equation*}
-
-\subsection{}
-\nextvar
-Now consider the containers of the patterns
-\begin{equation*}
-    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
-    \pattern{}{1,2}{0/0,1/1,0/2,1/0,1/2,2/0,2/1,2/2},
-    \pattern{}{1,2}{0/0,0/1,1/1,1/0,1/2,2/0,2/1,2/2}
-\end{equation*}
-Each of these patterns again occurs precisely once in
-any containing permutation. However this time when considering
-the free box we need to take into consideration \LemmaRef{lem:incdec}
-and so the empty box can only contain a decreasing subsequence. There
-is precisely one decreasing subsequence at every length, and so there
-is exactly one container of each pattern at each length. The three
-patterns are Wilf-equivalent and have \(C_n - 1\) avoiders of
-length \(n\) for all \(n \ge 2\).
-This gives the sequence
-\begin{equation*}
-    1, 1, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc\tag{\href{https://oeis.org/A001453}{OEIS: A001453} offset 2}
-\end{equation*}
-
-\subsection{}
-The containers of the following patterns can only
-have exactly one occurrence.
-\begin{equation*}
-    \pattern{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
-    \pattern{}{2,1}{0/0,0/1,1/0,1/1,1/2,2/0,2/2}
-    \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/1,2/2}
-    \pattern{}{1,2}{0/1,0/2,1/0,1/1,2/0,2/1,2/2}
-    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1}
-    \pattern{}{1,2}{0/0,0/1,1/0,1/2,2/0,2/1,2/2}
-\end{equation*}
-Once again we consider the free boxes, obviously for every pattern
-except the first the two boxes are independent, and one contains
-any avoider of \(\perm{2,3,1}\) and the other must contain a
-decreasing sequence by \LemmaRef{lem:incdec}.
-Let us consider the first pattern on it's own. In order to
-avoid \(\perm{2,3,1}\) across the free boxes we can add some additional
-restrictions
-\begin{equation}
-    \raisebox{0.6ex}{
-    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
-        \fill[pattern=north east lines, pattern color=black!75] (1.5, 0) rectangle +(0.5,1);
-        \fill[pattern=north east lines, pattern color=black!75] (1.0, 2) rectangle +(0.5,1);
-        \draw (1.5,0) -- (1.5,1)
-              (1.5,2) -- (1.5,3);
-        \modpattern[2]{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
-    \end{tikzpicture}
-    }
-    \label{eqn:splitboxes}
-\end{equation}
-Now we can see that the top free box must contain a decreasing sequence,
-and the bottom must contain an avoider of \(\perm{2,3,1}\) and these
-two do not interact in any manner. The containers of this pattern are counted
-the same as the other patterns, and due to this they are Wilf-equivalent. The containers
-have generating function \(\frac{x^2C(x)}{(1-x)}\). Enumerating avoiders
-therefore gives us
-\begin{equation*}
-    1, 1, 1, 3, 10, 33, 109, 364, 1233, 4236, 14740,\dotsc
-\end{equation*}
-
-\subsection{}
 The classes containing the following patterns are Wilf-equivalent
@@
 First consider the decomposition of avoiders of \(m_1\) in \(\av{\perm{2,3,1}}\)
-around the maximum.
+around the maximum, we have different conditions if we start with the maximum or not.
 \begin{equation*}
@@
 
-Now we decompose the avoiders of \(m_2\) around the leftmost point
+Now we decompose the avoiders of \(m_2\) around the leftmost point, we have similar
+conditions for starting withthe maximum
 \begin{equation*}
@@
 
-Looking at the structure of a container of \(m_3\) in \(\av{\perm{2,3,1}}\)
+Consider an occurrence of of \(m_3\) in \(\av{\perm{2,3,1}}\)
 \begin{equation*}
@@
 \end{equation*}
-The boxes \(b_2\) and \(b_3\) must contain a decreasing sequence by
+The regions \(b_2\) and \(b_3\) must contain a decreasing sequence by
 \LemmaRef{lem:incdec}. The box labelled \(b_1\) must contain an avoider
 of \(\perm{2,3,1}\). However note that the points in this box can have
-interaction with any points in box \(b_2\). If there is just one point
+interaction with any points in box \(b_2\). If there is a point
 in \(b_2\) then any points in \(b_1\) to the left of this point must
@@
 \eqref{eqn:splitboxes} to separate the two decreasing sequences in the
-free boxes in the top row, and the mixing happens in the same manner
+free regions in the top row, and the mixing happens in the same manner
 as in a container of \(m_3\).
@@
 \nextvar
-The classes containing the following patterns are Wilf-equivalent
-\begin{equation*}
-    \pattern{}{1,2}{} \text{ and }
-    \pattern{}{2,1}{}
-\end{equation*}
-This is true since the only avoiders of these patterns are the decreasing
-sequence and the increasing sequence respectively, and both of these
-avoid \(\perm{2,3,1}\) in all cases. There is therefore \(1\) avoider at
-every length.
-
-\subsection{}
 The patterns
@@
 
+Considering the last pattern, the only way we can construct an
+avoider is to take any \(\perm{2,3,1}\) avoider and add a new minimum
+at the start of the permutation. Adding a new leftmost point in any other
+position would either create an occurrence of \(\perm{2,3,1}\) or the pattern.
+Therefore this is also counted by \(\var(x) = 1+xC(x)\).
+
 The avoiders of the third pattern can be decomposed by the maximum
@@
 \end{equation*}
-
-Finally considering the last pattern, the only way we can construct an
-avoider is to take any \(\perm{2,3,1}\) avoider and add a new minimum
-at the start of the permutation. Adding a new leftmost point in any other
-position would either create an occurrence of \(\perm{2,3,1}\) or the pattern.
-Therefore this is also counted by \(\var(x) = 1+xC(x)\).
 
@@
 \subsection{}
-\nextvar
+\nextvar\nextvar
 The patterns
@@
 \end{equation*}
-Can be shown to be Wilf-equivalent. First we consider a container of \(m_2\) in \(\av{\perm{2,3,1}}\)
+Can be shown to be Wilf-equivalent. First we consider an occurrence of \(m_2\) in a
+permutation in \(\av{\perm{2,3,1}}\)
 \begin{equation*}
@@
 
-Now looking at a container of \(m_3\) in \(\av{\perm{2,3,1}}\)
+Now looking at an occurrence of \(m_3\) in \(\pi \in \av{\perm{2,3,1}}\)
 \begin{equation*}
@@
 \nextvar
+The following patterns are experimentally Wilf-equivalent up to length 10
 \begin{equation*}
@@
 \end{equation*}
-Let \(\scriptvar_1\) be the set of avoiders of \(m_1\), by structural decomposition
+Let \(\scriptvar_1\) be the set of avoiders of \(m_1\) in \(\av{231}\), by structural decomposition
 around the leftmost point we have
@@
 Where \(\scriptvar^\prime_1\) is a permutation avoiding \(\perm{2,3,1}, m_1\)
-and \(\textpattern{}{1}{0/0,0/1,1/1}\)
+and \(\textpattern{}{1}{0/0,0/1,1/0}\).
 Now consider the decomposition of a permutation in \(\scriptvar^\prime_1\) it can
@@
     \scriptvar_2 = \varepsilon \sqcup
-    \decompleft{0/1,1/0}{\scriptvar_2}{}{}{\scriptvar^\prime_2}
+    \decompleft{0/1,1/0}{\scriptvar^\prime_2}{}{}{\scriptvar_2}
 \end{equation*}
+
+Where \(\scriptvar^\prime_2\) is a permutation avoiding \(\perm{2,3,1}, m_2\)
+and \(\textpattern{}{1}{0/0,0/1,1/1}\).
 
@@
     \scriptvar^\prime_2 = \varepsilon \sqcup
-    \decompleft{0/1,1/0}{\scriptstyle \scriptvar_2\setminus\varepsilon}{}{}{\scriptvar^\prime_2}
+    \decompleft{0/1,1/0}{\scriptvar^\prime_2}{}{}{\scriptstyle \scriptvar_2\setminus\varepsilon}
 \end{equation*}
@@
 Now we consider decomposition of an avoider
-\textpattern{}{1,2}{0/1,0/2,1/0,1/1,1/2,2/0,2/2} around the maximum,
-this avoider must fit into the following form
+\textpattern{}{1,2}{0/1,0/2,1/0,1/1,1/2,2/0,2/2}  in \(\av{231}\) around the maximum,
+this avoider has form
 \begin{equation*}

### 2016-05-03 writing/latex/introduction.tex
Make Edits. Temp remove code because borked computer
Add definition of mpcontainment by Anders, add subpattern def

@@
 
-We can display a permutation on a \emph{figure} in order to give a graphical
-representation of the permutation. In such a figure we plot the points
-\((i,\pi(i))\) in a Cartesian coordinate system. The figure of the permutation
+We can display a permutation in a \emph{plot} in order to give a graphical
+representation of the permutation. In such a plot we display the points
+\((i,\pi(i))\) in a Cartesian coordinate system. The plot of the permutation
 \(\pi = \perm{2,3,1}\) is shown below
@@
 It is convenient to call the elements of the permutation \emph{points} when
-referring to these figures.
+referring to these plots.
 
@@
 capture explicit expansions for certain permutation statistics. They are a
-natural extension of classical permutation patterns. A \emph{mesh pattern} is a pair
+natural extension of classical permutation patterns.
+A \emph{mesh pattern} is a pair
 \begin{equation*}
-    p = (\tau,R)\text{ with } \tau \in \mathfrak{S}_k \text{ and } R \subseteq
+    p = (\pi,R)\text{ with } \tau \in \mathfrak{S}_k \text{ and } R \subseteq
     [0,k]\times [0,k].
@@
 The set \(R\) is called the \emph{mesh set} of the mesh pattern \(p\).
-The figure for a mesh pattern looks similar to that of a classical pattern with the
+The plot for a mesh pattern looks similar to that of a classical pattern with the
 addition that we shade the unit square with bottom left corner \((i,j)\) for each \((i,j) \in R\):
@@
 We define containment, and avoidance, of the pattern \(p\) in the permutation
-\(\pi\) on mesh patterns analogously to classical containment, and avoidance,
-of \(\tau\) in \(\pi\) with the additional restrictions on the relative
-position of the occurence of \(\tau\) in \(\pi\). These restrictions say that
-no elements of \(\pi\) are allowed in the regions of the figure corresponding
-to shaded boxes in the mesh (see \cite[2]{journals/combinatorics/BrandenC11}
-for the precise definition).
+\(\tau\) on mesh patterns analogously to classical containment, and avoidance,
+of \(\pi\) in \(\tau\) with the additional restrictions on the relative
+position of the occurence of \(\pi\) in \(\tau\). These restrictions say that
+no elements of \(\tau\) are allowed in the regions of the plot corresponding
+to shaded boxes in the mesh. Formally defined by \textcite{journals/combinatorics/BrandenC11},
+an \emph{occurrence} of \(p\) in \(\tau\) is a subset \(\omega\) of the plot of
+\(\tau, G(\tau) = \{(i,\tau(i)\mid i\in[1,n]\}\) such that there are order-preserving
+injections \(\alpha,\beta:[1,k]\mapsto[1,n]\) satisfying the following two conditions.
+
+\noindent Firstly, \(\omega\) is an occurrence of \(\pi\) in the classical sense
+  \begin{enumerate}[i.]
+    \item \(\omega = \{(\alpha(i),\beta(j)):(i,j)\in G(\pi)\}\)
+  \end{enumerate}\saveenum
+  Define \(R_{ij} = [\alpha(i)+1,\alpha(i+1)-1]\times[\beta(j)+1,\beta(j+1)-1]\)
+   for \(i,j\in[0,k]\) where \(\alpha(0)=\beta(0)=0\) and \(\alpha(k+1)=\beta(k+1)=n+1\). Then
+   the second condition is
+  \begin{enumerate}[i.]\resetenum
+  \item if \((i,j) \in R \text{ then } R_{ij} \cap G(\tau) = \emptyset\)
+\end{enumerate}
+We call \(R_{ij}\) the region corresponding to \((i,j)\).
+We define containment of a mesh pattern \(p\) in another mesh \(\kappa\) as above,
+with the additional condition that if \((i,j)\in R\text{ then } R_{ij}\) is
+contained in the mesh set of \(\kappa\), in this case we call \(p\) a \emph{subpattern}
+of \(\kappa\).
 
@@
 
-Let us consider the figure for the permutation \(\pi\). The subsequence \(325\) is
+Let us consider the plot for the permutation \(\pi\). The subsequence \(325\) is
 an occurrence of \(213\) in the classical sense and the remaining points of \(\pi\)
@@
 
-Now we consider the figure for the permutation \(\sigma\). This permutation
+Now we consider the plot for the permutation \(\sigma\). This permutation
 avoids \(p\) since for every occurence of the classical pattern \(\perm{2,1,3}\)
@@
 
+Two patterns are said to be \emph{coincident} if they avoid the same set of
+permutations and \emph{Wilf-equivalent} if they avoid the same number of permutations
+at every length.
+
 In the past people have studied different classes

### 2016-05-03 writing/latex/wilf.tex
Add definition of mpcontainment by Anders, add subpattern def

@@
 \nextvar
-The pattern \(p_1 = \textpattern{}{1}{0/1,1/0}\) is one of the
+The pattern \(p = \textpattern{}{1}{0/1,1/0}\) is one of the
 quadrant marked mesh patterns studied by Remmel,
@@
 an occurrence since the maximum lies in a region corresponding to the
-shading in \(p_1\). The lower left region however can create occurrences
-of \(p_1\) and therefore must also avoid \(p_1\) as well as \(\perm{2,3,1}\).
+shading in \(p\). The lower left region however can create occurrences
+of \(p\) and therefore must also avoid \(p\) as well as \(\perm{2,3,1}\).
 This gives the generating function of avoiders to be the function \(\var(x)\)
@@
 It can be shown by use of \PropositionRef{prop:dom2} that the patterns
-\textpattern{}{1}{0/0,1/1} and \(p_2 = \textpattern{}{1}{0/0,0/1,1/1}\) are
-coincident. Consider the decomposition of an avoider of \(p_2\) in
+\textpattern{}{1}{0/0,1/1} and \(q_1 = \textpattern{}{1}{0/0,0/1,1/1}\) are
+coincident. Consider the decomposition of an avoider of \(q_1\) in
 \(\av{\perm{2,3,1}}\) around the maximum
@@
 This can be explained succinctly by the fact that a permutation containing
-\(p_2\) starts with it's maximum, by not allowing the left part of the
+\(q_1\) starts with it's maximum, by not allowing the left part of the
 \(\perm{2,3,1}\) avoider to be empty we prevent an occurrence from ever happening.
 
-Consider \(p_3 = \textpattern{}{1}{0/1,1/0,1/1}\), avoiding this pattern
+Consider \(q_2 = \textpattern{}{1}{0/1,1/0,1/1}\), avoiding this pattern
 means that a permutation does not end with it's maximum. We can perform a similar
@@
 
-Now consider \(p_4 = \textpattern{}{1}{0/0,0/1,1/0}\), the avoiders of this
+Now consider \(q_3 = \textpattern{}{1}{0/0,0/1,1/0}\), the avoiders of this
 pattern are permutations that do not start with their minimum. In this case we
@@
 There is one pattern of length \(1\) still to consider. The pattern
-\(p_5 = \textpattern{}{1}{0/0,1/0,1/1}\) is avoided by all permutations
+\(r = \textpattern{}{1}{0/0,1/0,1/1}\) is avoided by all permutations
 that do not end in their minimum. Considering the standard decomposition
 of a \(\perm{2,3,1}\) avoider around the maximum we can see that an avoider
-of \(p_5\) must fit into the following form precisely once.
+of \(r\) must fit into the following form precisely once.
 \begin{equation*}

### 2016-05-04 writing/latex/wilf.tex
start edits r2 to wilf

@@
 Wilf-equivalent if their avoidance sets have the same size at each
-length. More formally
+length. More formally:
 
@@
     if for all \(k_{} \ge 0\), \(\size{\av[k]{\pi}} = \size{\av[k]{\sigma}}\).
-
     Two sets of permutation patterns \(R\) and \(S\) are are
@@
 
-Wilf-equivalence is of interest as if two permutation classes are enumerated
+Wilf-equivalence is of interest since if two permutation classes are enumerated
 in the same way then there should exist a bijection between them, and therefore
@@
 to reduce the amount of work, it can be easily seen that the reverse, complement
-and inverse operations preserve enumeration, and therefore these classes are trivially
+and inverse operations (see \FigureRef{fig:symm}) preserve enumeration, and therefore these classes are trivially
 Wilf-equivalent.
-
+\begin{figure}[!htb]
 \begin{align*}
@@
 \end{align*}
+\caption{The operations reverse, complement and inverse for the pattern 231}
+\label{fig:symm}
+\end{figure}
 
@@
 of order \(8\), the group of symmetries of a square. \emph{Reverse-inverse} and
-\emph{reverse} correspond tothe generators of the dihedral group.
+\emph{reverse} correspond can be taken as generators of the dihedral group.
 
 Since we are always considering Wilf-equivalences in the set \(\av{S}\) we
-must only use these symmetries when they preserve the dominating pattern.
+must only use these symmetries when they preserve the dominating pattern, if we
+were to allow other symmetries, then the equivalences calculated in the previous
+section do not necessarily hold.
 
 Throughout this section we will consider Wilf-equivalences of patterns
-whilst avoiding the \emph{Dominating Pattern} \(\perm{2,3,1}\). We will
+whilst avoiding the \emph{dominating pattern} \(\perm{2,3,1}\). We will
 use \(\mathcal{C}\) to denote \(\av{\perm{2,3,1}}\) and \(C(x)\) will
@@
 This is easy to see by structural decomposition around the maximum, as
-shown in \FigureRef{fig:decompmax}
+shown in \FigureRef{fig:decompmax}.
 
@@
     \decompmax{0/1,1/0}{A}{}{}{B}
-    \caption{Structural decomposition of a typical avoider of 231}
+    \caption{Structural decomposition of a non-empty avoider of 231}
     \label{fig:decompmax}
@@
 elements in \(A\) lie below all of the elements in \(B\). We call \(A\) the
-\emph{lower-left region} and \(B\) the \emph{upper-right region}.
+\emph{lower-left section} and \(B\) the \emph{upper-right section}.
 
@@
 the Wilf-equivalence classes of the mesh patterns of length \(1\) inside
-\(\av{\perm{2,3,1}}\), this means that we are sonsdering the set \(\av{\{231,p\}}\)
-where \(p\) is a mesh-pattern.
+\(\av{\perm{2,3,1}}\), this means that we are consdering the set \(\av{\{231,p\}}\)
+where \(p\) is a mesh-pattern of length 1.
 
-It can be seen that the patterns in the following set are set equivalent.
+The patterns in the following set are set equivalent,
 \begin{equation*}
@@
 \end{equation*}
-This is due to the fact that every permutation except the empty
+due to the fact that every permutation, except the empty
 permutation must contain an occurrence of all of these patterns.
@@
 Kitaev and Tiefenbruck\cite{2012arXiv1201.6243K}. Alternatively we can enumerate
-avoiders by decomposing a typical avoider of \(p_1\) around the maximum element
+avoiders by decomposing a non-empty avoider of \(p_1\) around the maximum element
 in order to give the following structural decomposition.
@@
 \end{equation*}
-Since if the upper right section was empty the maximum would create an
+Since if the upper-right section was empty the maximum would create an
 occurrence of the pattern, however no points in this section can create
 an occurrence since the maximum lies in a region corresponding to the
-shading in \(p\). The lower left region however can create occurrences
-of \(p\) and therefore must also avoid \(p\) as well as \(\perm{2,3,1}\).
+shading in \(p\). The lower-left section however can create occurrences
+of \(p\) and therefore must also avoid \(p\), as well as \(\perm{2,3,1}\).
 This gives the generating function of avoiders to be the function \(\var(x)\)
@@
 \textpattern{}{1}{0/0,1/1} and \(q_1 = \textpattern{}{1}{0/0,0/1,1/1}\) are
-coincident. Consider the decomposition of an avoider of \(q_1\) in
-\(\av{\perm{2,3,1}}\) around the maximum
+coincident. Consider the decomposition of a non-empty avoider of \(q_1\) in
+\(\av{\perm{2,3,1}}\) around the maximum:
 \begin{equation*}
@@
 This can be explained succinctly by the fact that a permutation containing
-\(q_1\) starts with it's maximum, by not allowing the left part of the
+\(q_1\) starts with it's maximum, by not allowing the lower-left section of the
 \(\perm{2,3,1}\) avoider to be empty we prevent an occurrence from ever happening.
@@
 \end{equation*}
-All of these classes have the same generating function namely
+All of these classes have the same generating function, namely
 \begin{equation*}
-    \var(x) = 1 + xC(x)(C(x)-1)
+    \var(x) = 1 + xC(x)(C(x)-1).
 \end{equation*}
@@
 \begin{equation*}
-    1, 0, 1, 3, 9, 28, 90, 297, 1001, 3432, 11934,\dotsc
+    1, 0, 1, 3, 9, 28, 90, 297, 1001, 3432, 11934,\dotsc \tag{\href{https://oeis.org/A000245}{OEIS: A000245} with offset}
 \end{equation*}
@@
 of a \(\perm{2,3,1}\) avoider around the maximum we can see that an avoider
-of \(r\) must fit into the following form precisely once.
+of \(r\) must fit into precisely on of the following twp forms.
 \begin{equation*}
@@
     \underbrace{\decompmax{0/0,0/1,1/0}{}{}{}{\scriptstyle{\scriptvar\setminus\varepsilon}}}_\text{
-        Minimum is after the maximum, cannot be last.
+        Minimum is after the maximum, minimum cannot be last.
     }
@@
 \end{equation*}
-This generating function can be enumerated to give
+This generating function is enumerated to give
 \begin{equation*}
-    	1, 0, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc
+    	1, 0, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc \tag{\href{https://oeis.org/A141364}{OEIS: A141364} }
 \end{equation*}
@@
 In order to consider symmetries we must only take the symmetries that preserve
-the pattern \(\perm{2,3,1}\). If we take any of the symmetries alone the
-permutation is different. The only symmetry that preserves the pattern
-\(\perm{2,3,1}\) is that of \emph{reverse-complement-inverse}. Using this
+the pattern \(\perm{2,3,1}\). The only symmetry that preserves the pattern
+\(\perm{2,3,1}\) is \emph{reverse-complement-inverse}. Using this
 symmetry to merge classes gives us \(61\) classes of trivial
-Wilf-equivalences.
+Wilf-equivalences, these Wilf-equivalences are explained by patterns being either
+coincident, or being the reverse-complement-inverse of a pattern.
 
-Computing avoiders up to length \(10\) gives us \(23\) Wilf-classes,
-of which \(13\) are non-trivial.
+Computing avoiders up to length \(10\) suggest that there are at least \(23\)
+Wilf-classes, of which \(13\) are non-trivial.
 
@@
     \label{not:setequiv}
-    The avoiders of the pattern \(q = \mperm{2,3,1}{\{(1,0),(1,1),(1,2),(1,3)\}}\)
+    The avoiders of the pattern \(q = \mperm{2,3,1}{\{(1,0),(1,1),(1,2),(1,3)\}}\),
+    \textpattern{}{2,3,1}{1/0,1/1,1/2,1/3},
     in \(\mathfrak{S}_n\) are in one-to-one correspondence with partitions of
-    \(\nrange{n}\). \cite{DBLP:journals/ejc/Claesson01}
+    \(\nrange{n}\). (\textcite[Prop.~2]{DBLP:journals/ejc/Claesson01})
 \end{note}
 \begin{proof}
-    Let \(\pi\) be a permutation in \(\av[n]{q}\) take the permutation
+    Let \(\pi\) be a permutation in \(\av[n]{q}\)
     in one-line notation and insert a dash between each ascent in \(\pi\). This
@@
     Given the permutation \(\pi = \perm{5,4,2,1,3,9,6,8,7}\) this corresponds to the
-    partition \(\{\{1,2,4,5\},\{3\},\{6,9\},\{8,7\}\}\).
+    partition \(\{\{5,4,2,1\},\{3\},\{9,6\},\{7,8\}\}\).
 \end{example}
 
+We call the least element in each block the \emph{block bottom}
+
 We are looking at permutations in \(\av{\perm{2,3,1}}\), all of these permutations
-also avoid the mesh pattern in \NoteRef{not:setequiv}.
+also avoid the mesh pattern in \NoteRef{not:setequiv}, \ie
+\(\av{\perm{2,3,1}} \subset \av{q}\).
 
@@
 avoid \(\perm{2,3,1}\) across the regions corresponding to the unshaded boxes
-we can add some additional restrictions
+we can add some additional restrictions, \ie all elements in the top region must
+be to the right of all elements in the bottom region.
 \begin{equation}
@@
 \subsection{}
-The set/symmetry classes containing the following patterns are
-Wilf-equivalent in \(\av{\perm{2,3,1}}\) but are not set equivalent or
+\nextvar
+The trivial Wilf-classes containing the following patterns are
+Wilf-equivalent in \(\av{\perm{2,3,1}}\) but are not coincident or
 symmetries of each other
@@
 be one occurrence of this pattern in a permutation corresponding to
-the last point in the permutation and the minimum.Consider an occurrence of \(m_1\),
+the last point in the permutation and the minimum. Consider an occurrence of \(m_1\),
 the points in the two regions corresponding to the the two boxes must form decreasing
@@
 avoiders of length \(k\) for all \(k\) and are Wilf-equivalent.
-The avoiders have enumeration
+The avoiders have general form
+\begin{equation*}
+  \var_k = C_k - (k-1), \var_0 = 1
+\end{equation*}
+and have enumeration
 \begin{equation*}
@@
 
-First we consider the coincidence in two parts then consolidate these parts.
-Consider the coincidence shown in \eqref{eq:singleud}. The easiest way to
-show that these are equinumerous is to consider the containers in the realm
-of set partitions.
+First we prove the Wilf-equivalence between \(m_1\) and \(m_2\) shown in
+\eqref{eq:singleud}. The easiest way to
+show that these are equinumerous is to consider the containers as set partitions.
 
-Considering an occurrence of either of these patterns in a permutation by
-the mesh sets we know the following about the points corresponding
-to the points in the patterns.
+Considering an occurrence of either of these patterns in a permutation  we know
+the following about the points corresponding to the points in the patterns.
 \begin{itemize}
     \item The point corresponding to the first point in both patterns
-        must lie in the first block of our set partition (there are
+        must lie in the first block of the set partition (there are
         no points southwest from it in the permutation).
@@
 between the patterns is in the underlying classical pattern. This
-means that permtuations containing \(m_1\) correspond to set partitions
+means that permutations containing \(m_1\) correspond to set partitions
 with a singleton block with value one higher than some element in the
@@
 one lower than some element in the block containing \(1\) and if this
-block is not the block containing \(1\) then it is a singleton block.
+block is not the block containing \(1\) then it is a singleton block. This
+proves that the containers of both of these patterns in \(\av{231}\) are
+equinumerous, and therefore so are their avoiders.
 
@@
 \begin{equation*}
-    1, 1, 1, 2, 6, 19, 61, 200, 670, 2286, 7918,\dotsc
+    1, 1, 1, 2, 6, 19, 61, 200, 670, 2286, 7918,\dotsc \tag{\href{https://oeis.org/A035929}{OEIS: A035929} offset 1}
 \end{equation*}
-This corresponds to OEIS sequence number \href{https://oeis.org/A035929}{A035929}
-with offset 1.
 
@@
 
-First consider the structure of an avoider of \(m_1\) and \(\perm{2,3,1}\)
-we can perform the usual structural decomposition of an avoider
+First consider the structure of an avoider of \(m_1\) and \(\perm{2,3,1}\).
+We can perform the usual structural decomposition of an avoider
 of \(\perm{2,3,1}\) where we consider decomposition around the maximum.
@@
 % precisely once.
-If \(\scriptvar_1\) set \(\av{\perm{2,3,1},m_1}\) then any permutation in
+If \(\scriptvar_1\) is the set \(\av{\perm{2,3,1},m_1}\) then any permutation in
 \(\scriptvar_1\) either starts with a maximum or does not, giving us the
@@
 First consider the decomposition of avoiders of \(m_1\) in \(\av{\perm{2,3,1}}\)
-around the maximum, we have different conditions if we start with the maximum or not.
+around the maximum. We have different conditions if we start with the maximum or not.
 \begin{equation*}
@@
 Now we decompose the avoiders of \(m_2\) around the leftmost point, we have similar
-conditions for starting withthe maximum
+conditions for starting with the maximum
 \begin{equation*}
@@
 if \(b_2\) contains a decreasing sequence with \(k\) points, there
-are \(k+1\) non-interacting avoiders of \(\perm{2,3,1}\) in \(b_3\).
+are \(k+1\) non-interacting avoiders of \(\perm{2,3,1}\) in \(b_1\).
 
@@
 \subsection{}
-\nextvar\nextvar
+\nextvar
 The patterns

### 2016-05-04 writing/latex/coincidence.tex
spell occurence correctly, add oeis seqs

@@
     \label{not:downcmesh}
-    Let \(R^\prime \subseteq R\). Then any occurence of \((\tau, R)\) in a permutation
-    is an occurence of \((\tau,R^\prime)\).
+    Let \(R^\prime \subseteq R\). Then any occurrence of \((\tau, R)\) in a permutation
+    is an occurrence of \((\tau,R^\prime)\).
 \end{note}
@@
 We know that the region corresponding to the box \((1,1)\) contains a decreasing subsequence.
-If we let \(z\) be the topmost point in this decereasing subsequence, then \(xz\)
+If we let \(z\) be the topmost point in this decreasing subsequence, then \(xz\)
 is an occurrence of \(m_2\). This shows that our two example patterns are coincident.
@@
     Consider a permutation \(\omega\) that contains an occurrence of \(m_2\).
-    By \NoteRef{not:downcmesh} any of these occurrences is also an occurence
-    of \(m_1\). This proves that every occurence of \(m_2\) is also an
-    occurence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).
+    By \NoteRef{not:downcmesh} any of these occurrences is also an occurrence
+    of \(m_1\). This proves that every occurrence of \(m_2\) is also an
+    occurrence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).
 
@@
 
-    If the region is empty, the occurrence of \(m_1\) is trivially an occurence
+    If the region is empty, the occurrence of \(m_1\) is trivially an occurrence
     of \(m_2\).

### 2016-05-04 writing/latex/confut.tex
Sentence ordering and autovars
spell occurence correctly, add oeis seqs

@@
 \chapter{Conclusions and Future work}
-From \ChapterRef{chap:coincs} it is can be seen that automatically classifying
+From \ChapterRef{chap:coincs} it can be seen that automatically classifying
 coincidences of mesh patterns is a difficult task, establishing rules for longer
 dominating patterns requires many more cases to be taken. It would however
-be interesting to consider the self application of the  third Dominating Rule to mesh
-patterns in order to try to capture some of the coincidences described in
-\textcite{DBLP:journals/combinatorics/HilmarssonJSVU15} and \textcite{journals/combinatorics/BrandenC11}.
+be interesting to consider the application of the third Dominating Rule to mesh
+patterns without any dominating pattern in order to try to capture some of the
+coincidences described in\textcite{DBLP:journals/combinatorics/HilmarssonJSVU15}
+ and \textcite{DBLP:journals/corr/ClaessonTU14}.
 It is not possible to apply the first and second Dominating Rules to the pattern
@@
 when considering dominating rules, and expressing coincidence between these sets.
-Doing this may give nice enumerative results.
+Doing this may give nice enumerative results. Coincidences between sets with
+multiple dominating patterns can also be considered, as well as sets where the
+dominating pattern is a mesh pattern.
 

### 2016-05-04 writing/latex/introduction.tex
spell occurence correctly, add oeis seqs

@@
 objects in a row}. When considering permutations we can consider them as
-occuring on the set \(\nrange{n} = \setrange{1}{n}\), therefore a permutation
+occurring on the set \(\nrange{n} = \setrange{1}{n}\), therefore a permutation
 is a \emph{bijection} \(\pi:\nrange{n} \mapsto \nrange{n}\). A
@@
 of \(\pi\) in \(\tau\) with the additional restrictions on the relative
-position of the occurence of \(\pi\) in \(\tau\). These restrictions say that
+position of the occurrence of \(\pi\) in \(\tau\). These restrictions say that
 no elements of \(\tau\) are allowed in the regions of the plot corresponding
@@
 \end{equation*}
-The subsequence \(325\) is therefore an occurence of the pattern \(p\) in \(\pi\) and
+The subsequence \(325\) is therefore an occurrence of the pattern \(p\) in \(\pi\) and
 \(\pi\) contains \(p\).
@@
 Now we consider the plot for the permutation \(\sigma\). This permutation
-avoids \(p\) since for every occurence of the classical pattern \(\perm{2,1,3}\)
+avoids \(p\) since for every occurrence of the classical pattern \(\perm{2,1,3}\)
 there is at least one point in one of the shaded boxes. For example, consider the
-subsequence\(315\) in \(\sigma\), this is an occurence of \(\perm{2,1,3}\) but
+subsequence\(315\) in \(\sigma\), this is an occurrence of \(\perm{2,1,3}\) but
 not the mesh pattern since the points with values \(4\) and \(2\) are in the
@@
 \end{equation*}
-This is true for all occurences of \(\perm{2,1,3}\) in \(\sigma\) and
+This is true for all occurrences of \(\perm{2,1,3}\) in \(\sigma\) and
 therefore \(\sigma\) avoids \(p\).

### 2016-05-04 writing/latex/wilf.tex
spell occurence correctly, add oeis seqs

@@
 the Wilf-equivalence classes of the mesh patterns of length \(1\) inside
-\(\av{\perm{2,3,1}}\), this means that we are consdering the set \(\av{\{231,p\}}\)
+\(\av{\perm{2,3,1}}\), this means that we are considering the set \(\av{\{231,p\}}\)
 where \(p\) is a mesh-pattern of length 1.
@@
 The pattern \(p = \textpattern{}{1}{0/1,1/0}\) is one of the
-quadrant marked mesh patterns studied by Remmel,
-Kitaev and Tiefenbruck\cite{2012arXiv1201.6243K}. Alternatively we can enumerate
+quadrant marked mesh patterns studied by \textcite{2012arXiv1201.6243K}.
+Alternatively we can enumerate
 avoiders by decomposing a non-empty avoider of \(p_1\) around the maximum element
@@
 
-Evauluation of this generating function gives the Fine numbers
+Evaluation of this generating function gives the Fine numbers
 (\href{https://oeis.org/A000957}{OEIS: A000957}).
@@
 of a \(\perm{2,3,1}\) avoider around the maximum we can see that an avoider
-of \(r\) must fit into precisely on of the following twp forms.
+of \(r\) must fit into precisely on of the following two forms.
 \begin{equation*}
@@
 \begin{equation*}
-    	1, 0, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc \tag{\href{https://oeis.org/A141364}{OEIS: A141364} }
+    	1, 0, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc \tag{\href{https://oeis.org/A141364}{OEIS: A141364}}
 \end{equation*}
@@
 \(\perm{2,3,1}\) is \emph{reverse-complement-inverse}. Using this
-symmetry to merge classes gives us \(61\) classes of trivial
+symmetry to reduces the number of Wilf-classes gives us \(61\) classes of trivial
 Wilf-equivalences, these Wilf-equivalences are explained by patterns being either
@@
 \begin{equation*}
-    1, 1, 1, 4, 12, 37, 118, 387, 1298, 4433, 15366,\dotsc
+    1, 1, 1, 4, 12, 37, 118, 387, 1298, 4433, 15366,\dotsc \tag{\(C_n\) - \href{https://oeis.org/A001453}{A001453} offset 2}
 \end{equation*}
@@
 \begin{equation*}
-    1, 1, 1, 3, 10, 33, 109, 364, 1233, 4236, 14740,\dotsc
+    1, 1, 1, 3, 10, 33, 109, 364, 1233, 4236, 14740,\dotsc \tag{\(C_n -\) \href{https://oeis.org/A014137}{A014137} offset 2}
 \end{equation*}
@@
 \begin{equation*}
-    1, 1, 1, 3, 11, 38, 127, 423, 1423, 4854, 16787,\dotsc
+    1, 1, 1, 3, 11, 38, 127, 423, 1423, 4854, 16787,\dotsc \tag{\(C_n -\) \href{https://oeis.org/A000027}{A000027} offset 2}
 \end{equation*}
@@
 \end{itemize}
-This tells us that an occurence of the patterns must happen when there
+This tells us that an occurrence of the patterns must happen when there
 is a singleton block occurring after the first block. The difference
@@
 
-First consider the structure of an avoider of \(m_1\) and \(\perm{2,3,1}\).
+First consider the structure of an avoider of \(m_1\) in \(\av{\perm{2,3,1}}\).
 We can perform the usual structural decomposition of an avoider
@@
 \end{equation*}
-
+Where \(\mathcal{\varfine} = \av{(231,\textpattern{}{1}{0/1,1/0})}\).
 Now consider the decomposition around the maximum of a permutation in
@@
 \begin{equation*}
-    1, 1, 1, 4, 11, 34, 108, 354, 1187, 4054, 14054,\dotsc
+    1, 1, 1, 4, 11, 34, 108, 354, 1187, 4054, 14054,\dotsc \tag{\(C_n - \)\href{https://oeis.org/A000958}{A000958} offset 2}
 \end{equation*}
@@
 \end{equation*}
-
+Where \(\mathcal{\varmaxl}_1 = \av{(231,\textpattern{}{1}{0/0,0/1,1/1})}\).
 Now we decompose the avoiders of \(m_2\) around the leftmost point, we have similar
@@
 \begin{equation*}
-    1, 1, 1, 2, 6, 20, 68, 233, 805, 2807, 9879,\dotsc
+    1, 1, 1, 2, 6, 20, 68, 233, 805, 2807, 9879,\dotsc\tag{\(C_n - \) \href{https://oeis.org/A014138}{A014138} offset 1}
 \end{equation*}
@@
 \end{equation*}
+Where \(\mathcal{\varfine} = \av{(231,\textpattern{}{1}{0/1,1/0})}\).
 The generating function derived satisfies \(\var_3(x) = 1 + x\varfine(x)\var_3(x)\).
@@
  \begin{equation*}
-     1, 1, 1, 2, 5, 14, 42, 132, 429, 1430, 4862,\dotsc
+     1, 1, 1, 2, 5, 14, 42, 132, 429, 1430, 4862,\dotsc \tag{\href{https://oeis.org/A000108}{OEIS: A000108} offset 1}
  \end{equation*}
@@
 
-Now we consider decomposition of an avoider
+Now we consider decomposition of an avoider of
 \textpattern{}{1,2}{0/1,0/2,1/0,1/1,1/2,2/0,2/2}  in \(\av{231}\) around the maximum,
@@
 \end{equation*}
+Where \(\mathcal{\varmaxl}_2 = \av{(231,\textpattern{}{1}{0/1,1/0,1/1})}\).
 This gives us the generating function
@@
 \end{equation*}
+Where \(\mathcal{\varmaxl}_3 = \av{(231,\textpattern{}{1}{0/0,0/1,1/0})}\).
 This gives generating function \(\var_3(x)\) satisfying
@@
 \begin{equation*}
-    1, 1, 1, 3, 9, 28, 90, 297, 1001, 3432, 11934,\dotsc
+    1, 1, 1, 3, 9, 28, 90, 297, 1001, 3432, 11934,\dotsc \tag{\href{https://oeis.org/A071724}{OEIS: A071724 with offset}}
 \end{equation*}
-This is an offset of OEIS Sequence number \href{http://oeis.org/A071724}{A071724}.

### 2016-05-04 writing/latex/confut.tex
edit r3

@@
 \chapter{Conclusions and Future work}
+\nextvar
 From \ChapterRef{chap:coincs} it can be seen that automatically classifying
@@
 when considering dominating rules, and expressing coincidence between these sets.
-Doing this may give nice enumerative results. Coincidences between sets with
-multiple dominating patterns can also be considered, as well as sets where the
+Doing this may give nice results. Coincidences between sets with
+multiple dominating patterns can also be considered, as this provides even more
+power to the rules discussed, these methods may be useful for classifying coincidence
+in large sets of pattern. It is also possible to apply these rules to sets where the
 dominating pattern is a mesh pattern.
@@
 reach enumeration and hopefully establish some of the non-trivial Wilf-equivalences
-between classes with different dominating patterns. It is possible to show that
-the sets consisting of \textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1} and
-\(\perm{2,3,1}\), or \(\perm{3,2,1}\), are Wilf-equivalent.
+between classes with different dominating patterns. For example, it is possible to show that
+the sets \(\scriptvar = \av{\left\{\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1},\perm{2,3,1}\right\}}\)\nextvar and
+\(\scriptvar = \av{\left\{\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1},\perm{3,2,1}\right\}}\), are Wilf-equivalent.
+This can be seen by considering an occurrence of the mesh pattern,
+\(p=\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1}\) in a permutation
+\begin{equation*}
+    \raisebox{0.6ex}{
+    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
+        \draw (0.5,1.5) node {\(b_1\)};
+        \draw (0.5,2.5) node {\( b_2\)};
+        \draw (2.5,2.5) node {\( b_3\)};
+        \modpattern[2]{2}{1,2}{0/0,1/0,1/1,1/2,2/0,2/1}
+    \end{tikzpicture}
+    }
+\end{equation*}
+
+If we are inside \(\av{\perm{2,3,1}}\) then any points in the region corresponding
+to the box \(b_1\) must be to the right of the points in the region corresponding to
+box \(b_2\), and must also form a decreasing subsequence by \LemmaRef{lem:incdec}.
+Furthermore, the points in the regions corresponding to \(b_2\) and
+\(b_3\) must form an avoider of \(\perm{2,3,1}\) with the indicated point.
+Therefore the containers of \(p\) in \(\av{\perm{2,3,1}}\) have generating function
+\(\lastvar(x) = \frac{x(C(x)-1)}{1-x}\).
+Now, if we are \(\av{\perm{3,2,1}}\) then any points in the region corresponding
+to the box \(b_1\) must be to the left of the points in the region corresponding to
+box \(b_2\), and must also form a increasing subsequence by \LemmaRef{lem:incdec}.
+Furthermore, the points in the regions corresponding to \(b_2\) and
+\(b_3\) must form an avoider of \(\perm{3,2,1}\) with the indicated point.
+Hence, the containers of \(p\) in \(\av{\perm{3,2,1}}\) have generating function
+\(\var(x) = \frac{x(C(x)-1)}{1-x}\) and both of these classes have the same
+enumeration.

### 2016-05-04 writing/latex/wilf.tex
edit r3

@@
 
-The classes containing the following patterns are Wilf-equivalent
+The classes containing following patterns are experimentally Wilf-equivalent up to length 10 in
+\(\av{\perm{2,3,1}}\)
 \begin{equation*}
@@
 \nextvar[\varminl]
-The classes containing the following patterns are Wilf-equivalent
-in \(\av{\perm{2,3,1}}\)
+The following patterns are experimentally Wilf-equivalent up to length 10 in
+\(\av{\perm{2,3,1}}\)
 \begin{equation*}
@@
 \nextvar
-Consider the containers of the patterns
+The following patterns are experimentally Wilf-equivalent up to length 10 in
+\(\av{\perm{2,3,1}}\)
 \begin{equation*}
@@
 \end{equation*}
-
+Consider containers of these patterns.
 For each of these patterns there is precisely one occurrence
@@
 \subsection{}
-Now consider the containers of the patterns
+The following patterns are experimentally Wilf-equivalent up to length 10 in
+\(\av{\perm{2,3,1}}\)
 \begin{equation*}
@@
 \end{equation*}
+Consider containers of these patterns.
 Each of these patterns again occurs precisely once in
@@
 \subsection{}
-The containers of the following patterns
-have exactly one occurrence.
+The following patterns are experimentally Wilf-equivalent up to length 10 in
+\(\av{\perm{2,3,1}}\)
 \begin{equation*}
@@
 \end{equation*}
+The containers of the patterns
+have exactly one occurrence.
 Once again we consider the regions corresponding to the unshaded regions, For each
@@
 \nextvar
-The trivial Wilf-classes containing the following patterns are
-Wilf-equivalent in \(\av{\perm{2,3,1}}\) but are not coincident or
-symmetries of each other
+The following patterns are experimentally Wilf-equivalent up to length 10 in
+\(\av{\perm{2,3,1}}\)
 \begin{equation*}
@@
 \nextvar
-The classes containing the following patterns are Wilf-equivalent
-when avoiders are considered in \(\av{\perm{2,3,1}}\)
+The following patterns are experimentally Wilf-equivalent up to length 10 in
+\(\av{\perm{2,3,1}}\)
 \begin{align}
@@
 \nextvar
-The classes containing the following patterns are Wilf-equivalent
-in \(\av{\perm{2,3,1}}\)
+The following patterns are experimentally Wilf-equivalent up to length 10 in
+\(\av{\perm{2,3,1}}\)
 \begin{equation*}
@@
 \nextvar
-The classes containing the following patterns are Wilf-equivalent
+The following patterns are experimentally Wilf-equivalent up to length 10 in
+\(\av{\perm{2,3,1}}\)
 \begin{equation*}
@@
 \nextvar
-The patterns
+The following patterns are experimentally Wilf-equivalent up to length 10 in
+\(\av{\perm{2,3,1}}\)
 \begin{equation*}
@@
 \end{equation*}
-are Wilf-equivalent.
+
 If \(\scriptvar_1\) is the set of avoiders of
@@
 \end{equation*}
-This is because \(\scriptvar_1\) is the set of permutations
-who have their minimum occur after their maximum.
+The lower-left section is empty because the minimum must occur after the maximum.
+These are counted by \(\var(x) = 1+xC(x)\)
+% This is because \(\scriptvar_1\) is the set of permutations
+% who have their minimum occur after their maximum.
 
@@
 if the last element is in a single block in the set partition context, In order
-to construct a avoider of length \(n\) we can take any avoider of
+to construct an avoider of length \(n\) we can take any avoider of
 \(\perm{2,3,1}\) of length \(n-1\) and insert the new maximum into the last
 block.  This ensures that the last block is never a singleton. This means
-that these permutations are also counted by \(\var(x) = 1+xC(x)\).
+that these permutations are also counted by \(\var(x)\).
 
@@
 avoider is to take any \(\perm{2,3,1}\) avoider and add a new minimum
-at the start of the permutation. Adding a new leftmost point in any other
-position would either create an occurrence of \(\perm{2,3,1}\) or the pattern.
-Therefore this is also counted by \(\var(x) = 1+xC(x)\).
+at the start of the permutation. Adding a new leftmost point with any other
+value would either create an occurrence of \(\perm{2,3,1}\) or the mesh pattern.
+Therefore these permutations are also counted by \(\var(x) = 1+xC(x)\).
 
@@
 
+So clearly \(\var_4(x) = \var_3(x)\).
 We need to show that the generating function \(\var_3(x)\) is the same as \(\var(x)\)
@@
 \end{equation*}
-Can be shown to be Wilf-equivalent. First we consider an occurrence of \(m_2\) in a
+ First we consider an occurrence of \(m_2\) in a
 permutation in \(\av{\perm{2,3,1}}\)
@@
 \end{equation*}
-We can choose the lowest occurrence of \(m_2\) without loss of generality.
-The region corresponding to \(b_1\) must avoid the pattern \(\textpattern{}{1}{0/1,1/0}\)
-as well as \(\perm{2,1,3}\). The regions corresponding to \(b_2\) and \(b_3\) must
-now contain avoiders of \(\perm{2,3,1}\), these regions cannot mix in order to avoid
+We can choose the lowest occurrence of \(m_2\) in the sense that the region
+corresponding to \(b_1\) must avoid the pattern \(\textpattern{}{1}{0/1,1/0}\)
+as well as \(\perm{2,3,1}\). The regions corresponding to \(b_2\) and \(b_3\) must
+now contain avoiders of \(\perm{2,3,1}\), all points in the region corresponding
+to \(b_2\) must be to the right of those in the region corresponding to \(b_3\)
 \(\perm{2,3,1}\). Since we already have an occurrence of
@@
 \end{equation*}
-We consider the leftmost occurrence of \(m_3\), the region corresponding
+We consider the leftmost occurrence of \(m_3\) in the sense that the region
+corresponding
 to \(b_1\) must avoid the pattern \(\textpattern{}{1}{0/1,1/0}\) as well as
 \(\perm{2,3,1}\) once more. The regions corresponding to \(b_2\) and \(b_3\)
-must avoid \(\perm{2,3,1}\) and as in a container of \(m_2\) these regions cannot
-mix, as doing so would lead to an occurrence of \(\perm{2,3,1}\).
+must avoid \(\perm{2,3,1}\) and as in a container of \(m_2\) the points in the
+region corresponding to the box containing \(b_2\) must be lower in value than
+all of those in the region corresponding to the box containing \(b_3\), as doing
+so would lead to an occurrence of \(\perm{2,3,1}\).
 Therefore both of these sets of containers are enumerated in the same way.
@@
 We can decompose an avoider of \(m_1\) in \(\av{\perm{2,3,1}}\) around the leftmost point
-in a similar manner
+in a similar manner:
 \begin{equation*}
@@
 \end{equation*}
-Where \(\scriptvar\) is the permutations avoiding \(\perm{2,3,1},m_4\) and
-\(p = \textpattern{}{1}{0/0,0/1,1/1}\) since if the subsequence in this box
+Here \(\scriptvar\) is the permutations avoiding \(\perm{2,3,1},m_4\) and
+\(p = \textpattern{}{1}{0/0,0/1,1/1}\), since if the subsequence in this box
 were to start with the maximum then this point and the maximum would create
@@
 \end{equation*}
-Now we can construct the following for \(\lastvar_4\)
+Now we get the following for \(\lastvar_4\)
 \begin{equation}
@@
 \nextvar
-The following patterns are experimentally Wilf-equivalent up to length 10
+The following patterns are experimentally Wilf-equivalent up to length 10 in
+\(\av{\perm{2,3,1}}\)
 \begin{equation*}
@@
 \end{equation*}
-Let \(\scriptvar_1\) be the set of avoiders of \(m_1\) in \(\av{231}\), by structural decomposition
+Let \(\scriptvar_1\) be the set of avoiders of \(m_1\) in \(\av{231}\). By structural decomposition
 around the leftmost point we have
@@
 and \(\textpattern{}{1}{0/0,0/1,1/0}\).
-Now consider the decomposition of a permutation in \(\scriptvar^\prime_1\) it can
+Now consider the decomposition of a permutation in \(\scriptvar^\prime_1\). It can
 once again be decomposed around the leftmost point
@@
 \end{equation*}
-This is a complete decomposition of avoiders of \(m_1\). Pow we
+This is a complete decomposition of avoiders of \(m_1\). Now we
 look at an avoider of \(m_2\), this time decomposition is around the maximum
@@
 and \(\textpattern{}{1}{0/0,0/1,1/1}\).
-
 Again we use the same method of decomposition of a permutation in \(\scriptvar^\prime_2\)
@@
 \end{equation*}
-
 This gives us a generating function \(\var(x)\) satisfying
@@
 \end{align}
-
 Solving \EquationRef{eqn:PprimeGen} for \(\var^\prime(x)\) and substituting into
-\EquationRef{eqn:Pgen} gives us the fact that a the generating function for
+\EquationRef{eqn:Pgen} gives us that the generating function for
 \(\var(x)\) satisfies
@@
 This is an offset of the Motzkin numbers \href{https://oeis.org/A001006}{(OEIS: A001006)}.
+
 \subsection{}
 \nextvar
+The following patterns are experimentally Wilf-equivalent up to length 10 in
+\(\av{\perm{2,3,1}}\)
 \begin{gather*}
@@
 In order to gain enumeration, consider decomposition of avoiders
-of \(\textpattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,1/2,2/2}\) around
+of first pattern, \(\textpattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,1/2,2/2}\), around
 the maximum.
@@
 
-Now we consider decomposition of an avoider of
-\textpattern{}{1,2}{0/1,0/2,1/0,1/1,1/2,2/0,2/2}  in \(\av{231}\) around the maximum,
-this avoider has form
+Now we consider decomposition of an avoider of the second pattern,
+\textpattern{}{1,2}{0/1,0/2,1/0,1/1,1/2,2/0,2/2},  in \(\av{231}\) around the maximum.
+This avoider has form
 \begin{equation*}
@@
 \end{equation*}
-This is equivalent to \EquationRef{eq:qgen}, and therefore
-the classes containing all of these patterns are Wilf-equivalent.
+This is equivalent to \EquationRef{eq:qgen}, and therefore these patterns are Wilf-equivalent.
 The classes have enumeration

### 2016-05-07 writing/latex/code.tex
Make Edits. Temp remove code because borked computer
add a little bit to the code section

@@
 \chapter{Code}\label{cha:code}
+    The exploratory work for this project was done mostly in the python
+programming language. The code can be obtained from \url{https://github.com/MurrayT/PRule-MSc-work/tree/master/src}
+and is split into the following files.
+
+The code relies on the the permuta python package available at \url{https://github.com/PermutaTriangle/Permuta}
 % \begin{minted}[linenos, breaklines]{python}
@@
 % \end{minted}
-\inputminted[linenos, breaklines, frame=lines, stripall]{python}{src/main.py}
+% \inputminted[linenos, breaklines, frame=lines, stripall]{python}{src/main.py}
 % \lstinputlisting[language=Python]{src/main.py}

### 2016-05-07 writing/latex/src/main.py

@@ -1,99 +1,0 @@
-from permuta import MeshPattern
-import sys
-from classes import *
-from helper_funcs import *
-from p_rule import *
-from wilf import *
-
-def stderrwrite(string,end="\n"):
-    sys.stderr.write(string+end)
-    sys.stderr.flush()
-
-classes2 = [map(MeshPattern.reverse, x) for x in classes]
-stderrwrite("Classes created")
-
-p_231 = Permutation([2,3,1])
-p_321 = Permutation([3,2,1])
-
-reprs_12 = get_representatives(classes)
-reprs_21 = get_representatives(classes2)
-stderrwrite("Representatives created")
-
-meshpd_12 = mpdict(classes)
-meshpd_21 = mpdict(classes2)
-
-stderrwrite("Computing Coincidences 1 of 4")
-coincs231_12 = get_coincidences(reprs_12, 11, p_231)
-stderrwrite("Computing Coincidences 2 of 4")
-coincs321_12 = get_coincidences(reprs_12, 11, p_321)
-stderrwrite("Computing Coincidences 3 of 4")
-coincs231_21 = get_coincidences(reprs_21, 11, p_231)
-stderrwrite("Computing Coincidences 4 of 4")
-coincs321_21 = get_coincidences(reprs_21, 11, p_321)
-stderrwrite("Coincidences done")
-
-coincsdict231_12 = coincidences_to_dictionary(coincs231_12)
-coincsdict321_12 = coincidences_to_dictionary(coincs321_12)
-coincsdict231_21 = coincidences_to_dictionary(coincs231_21)
-coincsdict321_21 = coincidences_to_dictionary(coincs321_21)
-
-p_coincs231_12 = p_rule_coincidences(meshpd_12, p_231)
-p_coincs321_12 = p_rule_coincidences(meshpd_12, p_321)
-p_coincs231_21 = p_rule_coincidences(meshpd_21, p_231)
-p_coincs321_21 = p_rule_coincidences(meshpd_21, p_321)
-
-p2_coincs231_12 = p_2_rule_coincidences(meshpd_12, p_231)
-p2_coincs321_12 = p_2_rule_coincidences(meshpd_12, p_321)
-p2_coincs231_21 = p_2_rule_coincidences(meshpd_21, p_231)
-p2_coincs321_21 = p_2_rule_coincidences(meshpd_21, p_321)
-
-unexplained231_12 = funexplained(p_coincs231_12, coincsdict231_12)
[…]

### 2016-05-09 writing/latex/abstracteng.tex
add Anders' suggestions

@@
 A permutation is an arrangement of \(n\) objects.
-Avoidance classes of classical permutation patterns
+Permutation classes of classical permutation patterns
 capture many interesting properties, such as
@@
 Two mesh patterns are coincident if they are avoided by the same set of
-permutations. In this thesis, the author provides sufficient conditions for
+permutations. We provide sufficient conditions for
 coincidence among mesh patterns, whilst also avoiding a longer classical
@@
 and a classical pattern of length 3. Two patterns are Wilf-equivalent if they
-have the same number of avoiders at every length, the author completely
+have the same number of avoiders at every length, we completely
 Wilf-classifies mesh patterns of length 2 when avoiding the classical pattern 231.

### 2016-05-09 writing/latex/coincidence.tex
add Anders' suggestions

@@
 
-If the shading set \(D\) is non-empty we can modify definition of the directions
+If the shading set \(D\) is non-empty we can modify the definition of the directions
 slightly
@@
 only adding the ascent, as adding a descent is very similar.
-Given a mesh pattern \(p = (\pi, R)\) \(\adda[D]{p}{(a,b)}\) gives a mesh
+Given a mesh pattern \(p = (\pi, R)\), \(\adda{p}{(a,b)}\) gives a mesh
 pattern \(p^\prime = (\pi^\prime,R^\prime)\) with length \(\setsize{\pi}+2\) defined by
@@
 All coincidence classes of \(\av{\{\perm{3,2,1}, \mperm{2,1}{R}\}}\) can be
-explained by application of \PropositionRef{prop:dom1}. % repeated?
+explained by application of \PropositionRef{prop:dom1}. By experimentation
+we see that there are at least \(29\) coincidence classes, and all of these
+coincidences are explained by this Proposition.
 
@@
     Consider a permutation \(\omega\) that contains an occurrence of \(m_2\).
-    By \NoteRef{not:downcmesh} any of these occurrences is also an occurrence
-    of \(m_1\). This proves that every occurrence of \(m_2\) is also an
-    occurrence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).
+    By \NoteRef{not:downcmesh},
+    %  any of these occurrences is also an occurrence
+    % of \(m_1\). This proves that every occurrence of \(m_2\) is also an
+    % occurrence of \(m_1\) and therefore
+     \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.
 

### 2016-05-09 writing/latex/introduction.tex
add Anders' suggestions

@@
 \enquote{A \emph{permutation of \(n\) objects} is an arrangement of \(n\) distinct
-objects in a row}. When considering permutations we can consider them as
+objects in a row}.
+When considering permutations we can consider them as
 occurring on the set \(\nrange{n} = \setrange{1}{n}\), therefore a permutation
@@
 \end{equation*}
-However, the most frequent notation used in computer science is
-\emph{one-line notation}, in this form we drop the top line of the two line
-notation and are left with the following.
+However,
+This
+% most frequent
+notation
+% used in computer science
+is called
+\emph{one-line notation}.
+% in this form we drop the top line of the two line
+% notation and are left with the following.
+In this form we write the entries of the permutation in order, and get
 \begin{equation*}
@@
 This can be seen graphically below, the subsequence order isomorphic to \(\sigma\)
-are highlighted.
+is highlighted.
 \begin{equation*}
@@
 
-We define containment, and avoidance, of the pattern \(p\) in the permutation
+We define containment (denoted \(p\preceq \pi\)), and avoidance, of the pattern \(p\) in the permutation
 \(\tau\) on mesh patterns analogously to classical containment, and avoidance,
@@
 there is at least one point in one of the shaded boxes. For example, consider the
-subsequence\(315\) in \(\sigma\), this is an occurrence of \(\perm{2,1,3}\) but
+subsequence \(315\) in \(\sigma\), this is an occurrence of \(\perm{2,1,3}\) but
 not the mesh pattern since the points with values \(4\) and \(2\) are in the

### 2016-05-12 writing/presentation/introduction.tex
Introduction for presentation

\section{Introduction}
\label{sec:Introduction}
\subsection{Permutations}
\label{sub:Permutations}

\begin{frame}{Permutations}
  A \emph{permutation} is a \emph{bijection}, \(\pi\), from the set
  \(\nrange{n} = \setrange{1}{n}\) to itself.

  More intuitively \textquote[Knuth]{A \emph{permutation of \(n\) objects} is
  an arrangement of \(n\) distinct objects in a row}.

  We write permutations in \emph{one-line notation}, writing the entries of
  the entries of the permutation in order \(\pi\)
  \begin{equation*}
    \pi = \perm{\pi(1),\pi(2),\dotso,\pi(n)}
  \end{equation*}
  \begin{example} The \(6\) permutations on \(\nrange{3}\) are
    \begin{equation*}
      123, 132, 213, 231, 312, 321
    \end{equation*}
  \end{example}
\end{frame}

\begin{frame}
  We can display a permutation in a \emph{plot} to give a graphical
  represention. We plot the points \((i,\pi(i))\) in a Cartesian coordinate
  system.
  \begin{figure}[htb]
    \centering
    \pattern{}{2,3,1}{}
    \caption{Plot of the permutation \(\perm{2,3,1}\)}
  \end{figure}
  In this setting we call the elements of the permutations \emph{points}.

  The set of all permutations of length \(n\) is \(\mathfrak{S}_n\) and
  has size \(n!\). The set of all permutations is
  \(\mathfrak{S}=\bigcup_{i=0}^{\infty}\mathfrak{S}_i\).
\end{frame}
\subsection{Classical Permutation Patterns}
\label{sub:Classical Permutation Patterns}

\begin{frame}{Classical Permutation Patterns}
  \emph{Classical permutation patterns} capture many interesting combinatorial
  objects and properties.

  \begin{definition}[Order Isomorphism]
    Two substrings \(\alpha_1\alpha_2\dotsm\alpha_n\) and
    \(\beta_1\beta_2\dotsm\beta_n\) are said to be \emph{order isomorphic}
    if \(\alpha_r<\alpha_s\) if and only if \(\beta_r<\beta_s\).
  \end{definition}
\end{frame}

\begin{frame}
  \begin{definition}
    A permutation \(\pi\) is said to \emph{contain} the \emph{classical
    permutation pattern} \(\sigma\)
    (denoted \(\sigma \preceq \pi\)) if there is some
    subsequence \(i_1i_2\dotsm{}i_k\) such that the sequence
    \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is order isomorphic to
    \(\sigma(1)\sigma(2)\dotsm\sigma(k)\).
  \end{definition}

  If \(\pi\) does not contain \(\sigma\) we say that \(\pi\) \emph{avoids}
  \(\sigma\).

  We the set of permutations of length \(n\) avoiding a pattern \(\sigma\) is
  denoted as \(\Av_n(\sigma)\) and
  \(\av{\sigma} = \bigcup_{i=0}^{\infty}\Av_i(\sigma)\).
\end{frame}

\begin{frame}
  \begin{example}
    The permutation \(\pi = \perm{2,4,1,5,3}\) contains the pattern
    \(\sigma = \perm{2,3,1}\)
    \begin{figure}[htb]
      \centering
      \pattern{2,4,5}{2,4,1,5,3}{}
      \caption{Plot of the permutation \(\perm{2,4,1,5,3}\) with an occurrence
      of \(\perm{2,3,1}\) indicated}
    \end{figure}
  \end{example}
\end{frame}

\subsection{Mesh Patterns}
\label{sub:Mesh Patterns}
\begin{frame}{Mesh Patterns}
  \emph{Mesh patterns} are a natural extension of classical permutation patterns.
  \begin{definition}
    A \emph{mesh pattern} is a pair
    \begin{equation*}
      p = (\tau,R)\text{ with } \tau \in \mathfrak{S}_k \text{ and } R \subseteq
      [0,k]\times [0,k].
    \end{equation*}
  \end{definition}
  We say that \(\tau\) is the \emph{underlying classical pattern} of \(p\).
\end{frame}

\begin{frame}
  \begin{example}
      The pattern \(p=\mperm{2,1,3}{\{(0,1),(0,2),(0,3),(1,0),(1,1),(2,1),(2,2)\}}=
      \textpattern{}{ 2, 1, 3 }{ 0/1, 0/2,0/3, 1/0, 1/1, 2/1, 2/2 }\) is contained in
      \(\pi = \perm{3,4,2,1,5}\).
      \begin{figure}[htb]
        \centering
      \pattern{1,3,5}{3,4,2,1,5}{0/2,0/3,0/4,0/5,
                                 1/0,1/1,1/2,
                                 2/0,2/1,2/2,
                                 3/2,3/3,3/4,
                                 4/2,4/3,4/4}
      \caption{ An occurrence of \(p\) in \(\pi\) }
      \end{figure}
  \end{example}
\end{frame}

\begin{frame}
  \begin{example}
      The pattern \(q=\mperm{2,1}{\{(0,1),(0,2),(1,0),(1,1)\}}=
      \textpattern{}{ 2, 1 }{ 0/1, 0/2, 1/0, 1/1}\) is contained in
      \(p=\mperm{2,1,3}{\{(0,1),(0,2),(0,3),(1,0),(1,1),(2,1),(2,2)\}}=
      \textpattern{}{ 2, 1, 3 }{ 0/1, 0/2,0/3, 1/0, 1/1, 2/1, 2/2 }\) as a
      subpattern.
      \begin{figure}[htb]
        \centering
      \pattern{1,2}{ 2, 1, 3 }{ 0/1, 0/2,0/3, 1/0, 1/1, 2/1, 2/2 }
      \caption{ An occurrence of \(q\) in \(p\) }
      \end{figure}
  \end{example}
\end{frame}

### 2016-05-17 writing/latex/coincidence.tex
More Presentation

@@
     \(\pi \preceq \adda{\sigma}{(a,b)}\)\\(\(\pi \preceq \addd{\sigma}{(a,b)}\))
-    then in any occurrence of \(m\) in a permutation \(\varrho\) the region
+    ,then in any occurrence of \(m\) in a permutation \(\varrho\), the region
     corresponding to the box \((a,b)\) can only contain an increasing

### 2016-05-17 writing/presentation/coincidences.tex
More Presentation

--- 2016-05-17 writing/latex/coincidence.tex
@@
-\chapter[Coincidence Classification]{Coincidences amongst families of mesh patterns and classical patterns}\label{chap:coincs}
-One interesting question to ask about permutation patterns is to consider when a
-pattern may be avoided by, or contained in, arbitrary permutations. Two
-patterns \(\pi\) and \(\sigma\) are said to be \emph{coincident} if the set
-of permutations that avoid \(\pi\) is the same as the set of permutations that
-avoid \(\sigma\), \ie \(\av{\pi} = \av{\sigma}\). This extends to sets of patterns as well as single patterns.
+\section{Coincidence Classes}
+\label{sec:Coincidence Classes}
+\subsection{Coincidence}
+\label{sub:Coincidence}
+\begin{frame}{Coincidence}
+  \begin{definition}
+    Two mesh patterns are said to be \emph{coincident} if they avoid the
+    same set of patterns at every length.
+  \end{definition}
+  \begin{block}{}
+    Classical patterns can never be coincident.
 
-We consider the avoidance sets, \(\av{\{\pi,q\}}\) where \(\pi\) is a classical pattern
-of length \(3\) and \(q\) is a mesh pattern of length \(2\) in order to
-establish some rules about when these two sets give the same avoidance set.
-We fix \(\pi\) in order to define the  coincidence and say that \(\pi\) is the
-\emph{dominating pattern}.
+    Aim to establish rules that classify coincidences when we
+    have one mesh pattern and one classical pattern.
+  \end{block}
+\end{frame}
 
-It is useful to be able to modify a mesh pattern by adding points to an
-already existing mesh pattern. First adding a single point into a pattern.
-\begin{definition}
-Given a mesh pattern \(p = (\pi,R)\) \(\addp[D]{p}{(a,b)}\) gives a mesh
-pattern \(p^\prime = (\pi^\prime,R^\prime)\) with length \(\setsize{\pi}+1\) defined by
-\begin{equation*}
-    \pi^\prime(i) = \begin{cases}
-    \pi(i) & \text{if } i\neq a+1 \text{ and } \pi(i)<b\\
-    \pi(i)+1 & \text{if } i\neq a+1 \text{ and } \pi(i)>b\\
-    b+1 & \text{if } i = a+1
-\end{cases}
-\end{equation*}
-and
-\begin{equation*}
-        R^\prime = \bigcup_{(i,j)\in R}r((i,j)),
-\end{equation*}
-where \(r((i,j))\) is defined by
-\begin{equation*}
-    r((i,j)) =
-    \begin{cases}
-        \{(i,j)\} & \text{if } i<a, j<b\\
-        \{(i,j),(i,j+1)\} & \text{if } i<a, j=b\\
-        \{(i,j+1)\} & \text{if } i<a, j>b\\
-        \{(i,j),(i+1,j)\} & \text{if } i=a, j<b\\
-        \{(i,j+1),(i+1,j+1)\} & \text{if } i=a, j>b\\
-        \{(i+1,j)\} & \text{if } i>a, j<b\\
-        \{(i+1,j),(i+1,j+1)\} & \text{if } i>a, j=b\\
-        \{(i+1,j+1)\} &\text{if } i>a, j>b\\
-    \end{cases}
-\end{equation*}
-
-If the shading set \(D\) is non-empty we can modify the definition of the directions
-slightly
-\begin{equation*}
-    \begin{aligned}
-        N &= \{(a,b+1),(a+1,b+1)\} \\
-        E &= \{(a+1,b),(a+1,b+1)\} \\
-        S &= \{(a,b),(a+1,b)\} \\
-        W &= \{(a,b),(a,b+1)\}
-    \end{aligned}
-\end{equation*}
-And we add the union of the sets in \(D\) into the mesh set \(R^\prime\).
-\end{definition}
-
-Given a mesh pattern \(p\), \(\addp[D]{p}{(a,b)}\) is the operation that returns
-a mesh pattern equivalent to placing a point in the center of box \((a,b)\),
-where \((a,b)\) is not in the mesh set of \(p\) , with shading defined by
-\(D\subseteq\{N,E,S,W\}\).
-The set \(D\) defines the shading by indicating that the boxes in the cardinal
-directions in \(D\) next to the point are shaded in the resulting pattern.
-Since there is no ambiguity we let \(\addpe[D]\) be equivalent to \(\addp[D]{\varepsilon}{(0,0)}\).
-
-\begin{example}
-    The result of adding a single point to the empty permutation for each cardinal direction.
- \begin{align*}
-     \addpe[\{N\}] = \textpattern{}{1}{0/1,1/1} &\qquad\addpe[\{E\}] =\textpattern{}{1}{1/1,1/0} \qquad \\
-     \addpe[\{S\}] = \textpattern{}{1}{0/0,1/0} &\qquad\addpe[\{W\}] =\textpattern{}{1}{0/0,0/1} \qquad
-\end{align*}
-    A more complex example for \texttt{add\_point} could be
-\begin{equation*}
-\addp[\{N,E\}]{\pattern{}{2,1,3}{0/3}}{(2,3)} = \pattern{}{2,1,4,3}{0/3,0/4,2/4,3/4,3/3}
-\end{equation*}
-\end{example}
-
-It is also useful to think about adding an ascent, or descent, into a pattern
-
-\begin{definition}
-Considering
-only adding the ascent, as adding a descent is very similar.
-Given a mesh pattern \(p = (\pi, R)\), \(\adda{p}{(a,b)}\) gives a mesh
-pattern \(p^\prime = (\pi^\prime,R^\prime)\) with length \(\setsize{\pi}+2\) defined by
-\begin{equation*}
-    \pi^\prime(i) = \begin{cases}
-    \pi(i) & \text{if } i\neq a+1,a+2 \text{ and } \pi(i)<b\\
-    \pi(i)+2 & \text{if } i\neq a+1,a+2 \text{ and } \pi(i)>b\\
-    b+j, b+j & \text{if } i = a+j, j\in\{1,2\}
-\end{cases}
-\end{equation*}
-and
-\begin{equation*}
-        R^\prime = \{(a+1,b),(a,b),(a+1,b),(a+2,b),(a+1,b+2)\} \cup \bigcup_{(i,j)\in R}r((i,j)),
-\end{equation*}
-where \(r((i,j))\) is defined by
-\begin{equation*}
-    r((i,j)) =
-    \begin{cases}
-        \{(i,j)\} & \text{if } i<a, j<b\\
-        \{(i,j),(i,j+1),(i,j+2)\} & \text{if } i<a, j=b\\
-        \{(i,j+2)\} & \text{if } i<a, j>b\\
-        \{(i,j),(i+1,j),(i+2,j)\} & \text{if } i=a, j<b\\
-        \{(i,j+2),(i+1,j+2),(i+2,j+2)\} & \text{if } i=a, j>b\\
-        \{(i+2,j)\} & \text{if } i>a, j<b\\
-        \{(i+2,j),(i+2,j+1),(i+2,j+2)\} & \text{if } i>a, j=b\\
-        \{(i+2,j+2)\} &\text{if } i>a, j>b\\
-    \end{cases}
-\end{equation*}
-\end{definition}
-
-Given a pattern \(p\), \(\addd{p}{(a,b)}\), and \\ \(\adda{p}{(a,b)}\),
-are the operations that return a mesh pattern equivalent to placing a
-decrease, or increase, in the center of box \((a,b)\), where \((a,b)\) is not
-in the mesh set of \(p\), in \(p\).
-
-\begin{example}
-    \begin{align*}
-        \addae = \pattern{}{1,2}{0/1,1/0,1/1,1/2,2/1}\\
-        \addde = \pattern{}{2,1}{0/1,1/0,1/1,1/2,2/1}
-    \end{align*}
-
-    \noindent A more complex example is
-\begin{equation*}
-\adda{\pattern{}{2,3,1}{0/0,1/0,1/3,
-                        3/0,3/1,3/3}}{(1,1)}
-= \pattern{}{4,2,3,5,1}{0/0,
-                        1/0,1/2,1/5,
-                        2/0,2/1,2/2,2/3,2/5,
-                        3/0,3/2,3/5,
-                        5/0,5/1,5/2,5/3,5/5}
-\end{equation*}
-\end{example}
-
-We now attempt to fully classify coincidences in families characterised by avoidance
-of a classical pattern of length \(3\) and a mesh pattern of length \(2\), that is
-finding and explaining all coincidences where \(\av{\{p,m\}} = \av{\{p,m^\prime\}}\).
-
-It can be easily seen that in order to classify coincidences one need only
-consider coincidences within the family of mesh patterns with the same underlying
-classical pattern, this is due to the fact that \(\perm{2,1}\in\av{\mperm{1,2}{R}}\)
-and \(\perm{1,2}\in\av{\mperm{2,1}{R}}\) for all mesh-sets \(R\).
-
-We know that there are a total of \(512\) mesh-sets for each underlying classical
-pattern. By use of the previous results of \textcite{DBLP:journals/corr/ClaessonTU14}\footnote{
-The authors use the Simultaneous Shading Lemma, a closure result and one worked out special case.
-} the number of coincidence classes can be reduced to \(220\).
-% the Shading
-% Lemma\cite[Lemma~3.11]{DBLP:journals/combinatorics/HilmarssonJSVU15},
-% Simultaneous Shading Lemma\textcite[Lemma~7.6]{DBLP:journals/corr/ClaessonTU14}, and
-% one special case, we can reduce the number of equivalence classes to \(220\).
-
-By discussion of a number of rules we will show that the number of coincidence
-classes follows the values shown in  \TableRef{tab:domclasses}. The experimental
-data in the last row of the table is calculated on permutations up to length 11.
-
+\subsection{Summary of Experimental Results}
+\label{sub:Summary of Experimental Results}
+\begin{frame}{Experimental Results}
+  \begin{block}{}
   \begin{table}[htb]
@@
 \end{center}
-    \caption{Coincidence class number reduction by application of Dominating Rules}
+    \caption{Coincidence class number reduction by application of Dominating rules}
     \label{tab:domclasses}
 \end{table}
+\end{block}
+\end{frame}
 
-From the table it can be seen that the rules established capture almost all
-coincidences. However, there are still some coincidences that are not able to
-be explained by the rules. This shows that complete coincidence classification
-of mesh patterns is a very difficult task, even when we have additional tools
-available.
-\section{Coincidence classes of Av(\{321, (21, \textit{R})\}).}
-Through experimentation, up to permutations of length 11, we discover that there
-are at least 29 coincidence classes of mesh patterns with underlying classical
-pattern \(\perm{2,1}\).
+\subsection{Dominating Pattern Rules}
+\label{sub:Dominating pattern Rules}
 
-\begin{proposition}[First Dominating Pattern Rule]
+\begin{frame}{First Dominating rule}
+  \begin{block}{Proposition: First Dominating rule}
       \label{prop:dom1}
+      \emph{
       Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
@@
           \item \(\pi \preceq \addp{\sigma}{(a,b)}\)\label{prop:dom1:cont}
-    \end{enumerate}
-\end{proposition}
-In order to prove this proposition we must first make the following note.
+      \end{enumerate}}
+  \end{block}
+\end{frame}
 
-\begin{note}
-    \label{not:downcmesh}
-    Let \(R^\prime \subseteq R\). Then any occurrence of \((\tau, R)\) in a permutation
-    is an occurrence of \((\tau,R^\prime)\).
-\end{note}
-% \begin{proof}
-%     Assume without meaningful loss of generality that \(R^\prime\) is a proper
-%     subset of \(R\).
-%
-%     Consider an occurence of \((\tau, R)\) in a permutation \(\sigma\), obviously
-%     this corresponds to an occurrence of \(\tau\) in \(\sigma\). Now consider
-%     the mesh sets \(R\) and \(R^\prime\), since \(R^\prime \subseteq R\) then
-%     there are more restrictions on where points are in an occurrence of \((\tau,R)\).
-%     Namely, for every shaded box in \(R\) the corresponding region in \(\sigma\)
-%     must contain no points, since \(R^\prime\) has less shading than \(R\) there
-%     exists a region in the occurence of \((\tau,R)\) in \(\sigma\) that is
-%     now devoid of restrictions. However, by removing restrictions we cannot
-%     make an occurence become not an occurence, and therefore the same occurrence
-%     of \(\tau \) in \(\sigma\) is now an occurence of \((\tau, R^\prime)\).
-% \end{proof}
-\begin{proof}[Proof of \PropositionRef{prop:dom1}]
-    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).
-
-    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
-    % Consider a permutation \(\omega\) that contains an occurrence of \(m_2\)
-    % by \NoteRef{not:downcmesh} any of these occurrences is also an occurence
-    % of \(m_1\). This proves that every occurence of \(m_2\) is also an
-    % occurence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).
-    Since \(R_1\) is a subset of \(R_2\), \NoteRef{not:downcmesh} states that
-    \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\)
-
-    Now we consider a permutation \(\omega^\prime\in \av{\pi}\),
-    containing an occurrence of \(m_1\). Consider placing a point in the
-    region corresponding to the box \((a,b)\), regardless of where in this
-    region we place the point by condition~\ref{prop:dom1:cont} of the
-    Proposition we create an occurrence of \(\pi\), therefore there can be no
-    points in this region, which could have been represented in the mesh set
-    \(R_1\) by adding the box \((a,b)\). Hence every occurrence of
-    \(m_1\) is in fact an occurrence of \(m_2\), and we have that
-    \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).
-
-    Taking both directions of the containment we can therefore draw the
-    conclusion that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).
-\end{proof}
-
-All coincidence classes of \(\av{\{\perm{3,2,1}, \mperm{2,1}{R}\}}\) can be
-explained by application of \PropositionRef{prop:dom1}. By experimentation
-we see that there are at least \(29\) coincidence classes, and all of these
-coincidences are explained by this Proposition.
-
-This rule can be understood very in graphical form. In the
-pattern in \FigureRef{fig:rule1} we can gain shading in three boxes since if
-there is a point in any of these boxes we would get an occurrence of the
-dominating pattern \(\perm{3,2,1}\).
-
-\begin{figure}[htb]
-    \begin{center}
+\begin{frame}
+  \begin{example}
+    The following two patterns are coincident in \(\av{\perm{3,2,1}}\)
+    \begin{equation*}
       \begin{tikzpicture}[scale=\picscale]
@@
     \end{tikzpicture}
-        \raisebox{2ex}{\(\mapsto\)}
+        \raisebox{2ex}{\qquad}
     \begin{tikzpicture}[scale=\picscale]
-        \filldraw (2.5,0.5) circle (3pt);
+        % \filldraw (2.5,0.5) circle (3pt);
         \modpattern{}{2,1}{1/0,1/1,1/2,2/2,2/0}
     \end{tikzpicture}
+    \end{equation*}
+  \end{example}
+  \begin{corollary}
+    All coincidences of classes the form \(\av{\{\perm{3,2,1},\mperm{2,1}{R}}\}\)
+    are fully explained by the First Dominating rule.
+  \end{corollary}
+  \begin{block}{}
+    There are \(29\) coincidences of mesh patterns of the form \(\av{\{\perm{3,2,1},\mperm{2,1}{R}}\}\)
+  \end{block}
+\end{frame}
 
-        \caption{Visual depiction of first dominating pattern rule.}
-        \label{fig:rule1}
-    \end{center}
-\end{figure}
-
-Note that there are two natural extensions of this rule. We can replace \(\pi\)
-with a set of classical patterns, or we can consider \(\pi\) to be a mesh pattern.
-
-\section{Equivalence classes of Av(\{231, (21, \textit{R})\}).}
-By application of \PropositionRef{prop:dom1} we obtain \(43\) coincidence
-classes. Experimentation shows that there are in fact at least \(39\) coincidence
-classes, for example the following two patterns are coincident in
-\(\av{\perm{2,3,1}}\) but this is not explained by \PropositionRef{prop:dom1}.
+\begin{frame}{Second Dominating rule}
+  \begin{block}{}
+    The patterns
     \begin{equation*}
@@
     \end{equation*}
+    are coincident in \(\av{\perm{2,3,1}}\).
+  \end{block}
+\end{frame}
 
-Consider an occurrence of \(m_1\) in a permutation in \(\av{\perm{2,3,1}}\), consisting
-of elements \(x\) and \(y\). If the region corresponding
-to the box \((1,1)\) is empty we have an occurrence of \(m_2\). Otherwise, if there is any
-increase in this box then we would have an occurrence of \(\perm{2,3,1}\),
-however, since we are in \(\av{\perm{2,3,1}}\) this is not possible. This box
-must therefore contain a (non-empty) decreasing subsequence.
- This gives rise to the following lemma:
-
+\begin{frame}
   \begin{lemma}
-    \label{lem:incdec}
     Given a mesh pattern \(m =(\sigma, R)\), where the box \((a,b)\) is not
@@
   \end{lemma}
-The proof is analogous to the proof of \PropositionRef{prop:dom1}.
+\end{frame}
 
-Going back to our example mesh patterns
+\begin{frame}
+  \begin{example}
+  Considering \(m_1\) again
   \begin{equation*}
     \begin{tikzpicture}[scale=\picscale]
+        {\only<3-> \fill[pattern=north east lines, pattern color=black!75] (1,1) rectangle +(1,1);}
         \modpattern{}{2,1}{0/0,0/2,1/0,2/1}
-        \filldraw (1.66,1.33) circle (3pt);
-        \filldraw (1.33,1.66) circle (3pt);
+        {\only<2> \filldraw (1.66,1.33) circle (3pt);}
+        {\only<2> \filldraw (1.33,1.66) circle (3pt);}
     \end{tikzpicture}
   \end{equation*}
-We know that the region corresponding to the box \((1,1)\) contains a decreasing subsequence.
-If we let \(z\) be the topmost point in this decreasing subsequence, then \(xz\)
-is an occurrence of \(m_2\). This shows that our two example patterns are coincident.
+  {\only<1-3> \vphantom{This is \(m_2\).}}
+  {\only<4-> {}}
+  \end{example}
+\end{frame}
 
-This result generalises into the following rule for categorising coincidences
-of mesh patterns in cases where there is a dominating classical pattern.
-
-\begin{proposition}[Second Dominating Pattern Rule]
+\begin{frame}
+  \begin{block}{Proposition: Second Dominating rule}
     \label{prop:dom2}
+    \emph{
     Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
@@
                                   \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b-1,b\)).
-                            \item \((a,b+1) \in \sigma\) and \((a-1,b+1)\notin R\) and \\
-                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
-                                  \((a-1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
+                            \item \(\dots\)
                         \end{enumerate}
-                    \item \(\pi \preceq \addd{\sigma}{(a,b)}\) and
-                        \begin{enumerate}
-                            \item \((a+1,b+1) \in \sigma\) and \((a+1,b+1)\notin R\) and \\
-                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a,a+1\)) and\\
-                                  \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
-                            \item \((a,b) \in \sigma\) and \((a-1,b-1)\notin R\) and \\
-                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
-                                  \((a-1,y)\in R \implies (a,y) \in R\)  (where \(y\neq b-1,b\)).
+                    \item \(\dots\)\begin{enumerate}
+                            \item \(\dots\)
+                            \item \(\dots\)
                           \end{enumerate}
@@
     \end{enumerate}
-\end{proposition}
-\begin{proof}
-    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).
+    }
+  \end{block}
+\end{frame}
 
-    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
-
-    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\).
-    By \NoteRef{not:downcmesh} any of these occurrences is also an occurrence
-    of \(m_1\). This proves that every occurrence of \(m_2\) is also an
-    occurrence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).
-
-    We will consider taking the first branch of every choice.
-    Now consider a permutation \(\omega^\prime \in \av{\pi}\). Suppose
-    \(\omega^\prime\) contains \(m_1\) and consider the region corresponding to
-    \((a,b)\) in \(R_1\).
-
-    If the region is empty, the occurrence of \(m_1\) is trivially an occurrence
-    of \(m_2\).
-
-    Now consider if the region is non-empty, by \LemmaRef{lem:incdec} and
-    condition~\ref{prop:dom2:condc} of the proposition this region must contain
-    a decreasing subsequence.
-    We can choose the topmost point in the region to replace the corresponding
-    point in the mesh pattern and the points from the subsequence are now in
-    the box southeast of the point. The other conditions allow this to be done
-    without points being present in regions that were shaded. Hence there are no
-    points in the region corresponding to the box \((a,b)\) in the mesh
-    pattern, and therefore we can shade this region. This implies that every
-    occurrence of \(m_1\) in \(\av{\pi}\) is in fact an occurrence of \(m_2\)
-    so \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).
-
-    Similar arguments satisfy the remainder of the branches.
-\end{proof}
+\begin{frame}
+  \begin{example}
     \begin{figure}[ht]
@@
 \end{tikzpicture}
-    \caption{If the conditions of \PropositionRef{prop:dom2} are satisfied the box
+    \caption{If the conditions of The Second Dominating rule are satisfied the box
     \((a-1,b)\) can be shaded.}
 \end{figure}
+  \end{example}
+\end{frame}
 
-This proposition essentially states that we slide all of the points in the box we desire
-to shade diagonally, and chose the topmost/bottommost point to replace the original
-point in the mesh pattern.
+\begin{frame}
+  \begin{corollary}
+    All coincidences of classes the form \(\av{\{\perm{2,3,1},\mperm{2,1}{R}}\}\)
+    are fully explained by applying the First Dominating rule, then applying the
+    Second Dominating rule.
+  \end{corollary}
+  \begin{block}{}
+    There are \(39\) coincidences of mesh patterns of the form
+    \(\av{\{\perm{2,3,1},\mperm{2,1}{R}}\}\)
+  \end{block}
+\end{frame}
 
-By taking the First Dominating Pattern Rule and the Second Dominating Pattern
-Rule together coincidences of classes of the form \(\av{\{\perm{2,3,1},\mperm{2,1}{R}\}}\)
-are fully explained, obtaining 39 coincidence classes of mesh patterns.
-
-\section{Equivalence classes of Av(\{231, (12, \textit{R})\}).}
-When considering the coincidence classes of \(\av{\perm{2,3,1},\mperm{1,2}{R}}\)
-we first apply the two Dominating Pattern rules previously established.
-Starting from \(220\) classes, application of the first Dominating Pattern rule
-gives \(85\) classes. Following this with the second Dominating Pattern rule
-reduces the number of classes to \(59\). However we know that there are patterns
-where the coincidences are not explained by the rules given above.
-
-For example the patterns
+\begin{frame}{Third Dominating rule}
+  \begin{block}{}
+    The patterns
     \begin{equation*}
@@
     \end{equation*}
-are experimentally coincident. This coincidence is not explained by our rules,
-it is necessary to attempt to capture these coincidences by establishing more
-rules.
+    are coincident in \(\av{\perm{2,3,1}}\).
+    Neither of the previous two rules explain this.
+  \end{block}
+\end{frame}
 
-Consider an occurrence of \(m_1\) in a permutation, if the region corresponding
-to the box \((1,0)\) is empty then we have an occurrence of \(m_2\). Now look at
-the case when this region is not empty. Consider choosing the rightmost point in
-region.
-% In order to rigorously establish this coincidence we need to consider what would
-% happen if we were to choose a point in the box \((1,0)\) that we would like to shade.
-% In order to get a chance of getting the shading we want consider choosing the
-% rightmost point in the box.
-This gives us an occurrence of the following mesh pattern.
+\begin{frame}
+  \begin{example}
     \begin{equation*}
-    \pattern{}{2,1,3}{0/3,2/0,2/1,3/0,3/1,3/2}
+    \only<1>{\vphantom{\pattern{2,3}{2,1,3}{0/3,1/3,2/0,2/1,3/0,3/1,3/2}}\pattern{}{1,2}{0/2,2/0,2/1}}
+    \only<2>{\pattern{}{2,1,3}{0/3,2/0,2/1,3/0,3/1,3/2}}
+    \only<3>{\pattern{}{2,1,3}{0/3,1/3,2/0,2/1,3/0,3/1,3/2}}
+    \only<4>{\pattern{2,3}{2,1,3}{0/3,1/3,2/0,2/1,3/0,3/1,3/2}}
+    \only<5->{\vphantom{\pattern{2,3}{2,1,3}{0/3,1/3,2/0,2/1,3/0,3/1,3/2}}\pattern{}{1,2}{0/2,1/0,2/0,2/1}}
   \end{equation*}
-By application of \PropositionRef{prop:dom1} we then achieve the following
-mesh pattern
-\begin{equation*}
-    \pattern{2,3}{2,1,3}{0/3,1/3,2/0,2/1,3/0,3/1,3/2}
-\end{equation*}
-If we look at the highlighted points we see that the subpattern is an occurrence
-of the mesh pattern that we originally desired. This gives rise to the following
-rule:
+  \end{example}
+\end{frame}
 
-\begin{proposition}[Third Dominating Pattern Rule]
-    \label{prop:dom3}
+\begin{frame}{}
+  \begin{block}{Proposition: Third Dominating rule}
+      \emph{
       Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
@@
       \end{enumerate}
-\end{proposition}
-\begin{proof}
-    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).
+      }
+  \end{block}
+\end{frame}
 
-    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
-
-    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\).
-    By \NoteRef{not:downcmesh},
-    %  any of these occurrences is also an occurrence
-    % of \(m_1\). This proves that every occurrence of \(m_2\) is also an
-    % occurrence of \(m_1\) and therefore
-     \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.
-
-    Now consider a permutation \(\varrho\) in \(\av{\pi}\) that contains an
-    occurrence of \(m_1\). If the region corresponding to the box \((a,b)\)
-    is empty then we have an occurrence of \(m_2\). If the region is non-empty
-    then by condition~\ref{prop:dom3:condocc}  of the proposition there exists
-    a direction such that there exists an occurrence of a mesh pattern of
-    length one longer than \(m_1\) in this position. This mesh pattern is
-    coincident with another mesh pattern that contains an
-    occurrence of \(m_2\). Hence, every occurrence of \(m_1\) leads to an occurrence
-    of \(m_2\).  Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the
-    two patterns are coincident.
-\end{proof}
-
-By application of this rule we can reduce the number of classes in
-\(\av{\{\perm{2,3,1},(\perm{1,2}, R)\}}\) to \(56\).
-
-\section{Equivalence classes of Av(\{321, (12, \textit{R})\}).}
-When considering coincidences of mesh patterns with underlying classical
-pattern \(\perm{1,2}\) in \(\av{\perm{3,2,1}}\) application of the previously
-established rules give no coincidences. Through experimentation we discover
-that there are \(7\) non-trivial coincidence classes (all others are singletons)
-which can be explained through the
-use of two different lines of reasoning. Since the number of coincidences is
-so small we will reason for these coincidences without attempting to generalise
-into concrete rules.
-
-Intuitively it is easy to see why our previous rules have no
-power here. There is nowhere that it is possible to add a single point to
-gain an occurrence of \(\pi = \perm{3,2,1}\). It is also impossible to have
-a position where addition of an increase, or decrease, provides extra
-shading power.
-
-The patterns
-\begin{equation*}
-    m_1 = \pattern{}{1,2}{0/0, 0/2, 1/1, 1/2, 2/1} \text { and } m_2 \pattern{}{1,2}{0/0, 0/1, 0/2, 1/1, 1/2, 2/1}
-\end{equation*}
-are equivalent in \(\av{\perm{3,2,1}}\). (There are 3 symmetries of these
-patterns that are also equivalent to each other by the same reasoning.)
-
-Consider the region corresponding to the box \((0,1)\) in any occurrence of \(m_1\),
-in a permutation. By \LemmaRef{lem:incdec} it must contain an increasing subsequence.
-If the region is empty then we have an occurrence of \(m_2\).
-If there is only one point in the region we can choose this to replace the \(1\) in
-the mesh pattern to get the required shading. If there is more than one point then
-choosing the two leftmost points gives us the following mesh pattern.
-\begin{equation*}
-    \pattern{3,4}{2,3,1,4}{0/0, 0/1, 0/2, 0/3, 0/4,
-                            1/0, 1/1, 1/2, 1/3, 1/4,
-                            2/0, 2/1,2/2, 2/4,
-                            3/1, 3/2, 3/3, 3/4,
-                            4/1, 4/2, 4/3}
-\end{equation*}
-Where the two highlighted points are the original two points.
-Now if we take the other two points as the points in our
-mesh pattern then we get an occurrence of the pattern
-we originally desired, and hence the two patterns are coincident.
-
-The other reasoning applies to the patterns
-\begin{equation*}
-    m_1 = \pattern{}{1,2}{0/1,0/2,2/0,2/1} \text{ and } m_2 = \pattern{}{1,2}{0/1,0/2,1/1,2/0,2/1}
-\end{equation*}
-which are coincident by experimentation.
-
-In order to prove this coincidence we will proceed by mathematical
-induction on the number of points in region corresponding to the middle box.
-We call this number \(n\).
-\begin{description}
-    \item [Base Case \((n=0)\):] The base case holds since we can freely shade the
-        box if it contains no points.
-    \item [Inductive Hypothesis \((n=k)\):] Suppose that we can find an occurrence
-        of the second pattern if we have an occurrence of the first with \(k\) points
-        in the middle box.
-    \item [Inductive Step \((n=k+1)\)] Suppose that we have \((k+1)\) points in
-        the middle box. Choose the bottom most point in the middle box, this gives the
-        mesh pattern
-     \begin{equation*}
-        \begin{tikzpicture}[scale=0.5]
-            \modpattern[5]{}{1,2,3}{0/1,0/2,0/3,1/1,2/1,3/0,3/1,3/2}
-            \draw (2.5,0.5) node {\(X\)};
-        \end{tikzpicture}
-    \end{equation*}
-        Now we need to consider the box labelled \(X\). If this box is empty then
-        we have an occurrence of \(m_2\) and are done. If this box contains
-        any points then we gain some extra shading on the mesh pattern
-        due to the dominating pattern
-     \begin{equation*}
-        \begin{tikzpicture}[scale=0.5]
-            \modpattern[5]{2,3}{1,2,3}{0/1,0/2,0/3,1/1,1/2,1/3,2/1,3/0,3/1,3/2}
-            \filldraw (2.5,0.5) circle (4 pt);
-        \end{tikzpicture}
-    \end{equation*}
-        The two highlighted points form an occurrence of \(m_1\) with \(k\)
-        points in the middle box, and thus by the Inductive Hypothesis we
-        are done.
-\end{description}
-By induction we have that every occurrence of \(m_1\) leads to an occurrence of \(m_2\)
-and by \NoteRef{not:downcmesh} every occurrence of \(m_2\) is an occurrence of
-\(m_1\) so the two patterns are coincident.
-This argument applies to another two pairs of classes. Therefore in
-total in \(\av{\perm{3,2,1}, \mperm{1,2}{R}}\) there are \(213\) coincidence
-classes.
+\begin{frame}
+  \begin{corollary}
+    All coincidences of classes the form \(\av{\{\perm{2,3,1},\mperm{1,2}{R}}\}\)
+    are fully explained by applying the First Dominating rule, the
+    Second Dominating rule, and then the Third Dominating rule.
+  \end{corollary}
+  \begin{block}{}
+    There are \(56\) coincidences of mesh patterns of the form
+    \(\av{\{\perm{2,3,1},\mperm{2,1}{R}}\}\)
+  \end{block}
+\end{frame}

### 2016-05-18 writing/latex/confut.tex
Anders chap3/4

@@
 Therefore the containers of \(p\) in \(\av{\perm{2,3,1}}\) have generating function
-\(\lastvar(x) = \frac{x(C(x)-1)}{1-x}\).
-Now, if we are \(\av{\perm{3,2,1}}\) then any points in the region corresponding
+\(\lastvar(x) = x(C(x)-1)/(1-x)\).
+Now, if we are in \(\av{\perm{3,2,1}}\) then any points in the region corresponding
 to the box \(b_1\) must be to the left of the points in the region corresponding to
@@
 Hence, the containers of \(p\) in \(\av{\perm{3,2,1}}\) have generating function
-\(\var(x) = \frac{x(C(x)-1)}{1-x}\) and both of these classes have the same
+\(\var(x) = x(C(x)-1)/(1-x)\) and both of these classes have the same
 enumeration.

### 2016-05-18 writing/latex/introduction.tex
Introduction for presentation
Anders chap3/4

@@
 occurring on the set \(\nrange{n} = \setrange{1}{n}\), therefore a permutation
-is a \emph{bijection} \(\pi:\nrange{n} \mapsto \nrange{n}\). A
-permutation \(\pi\) can be written in two-line notation
-\begin{equation*}
-\pi = \twoline{1,2,\dotso,n}{\pi(1),\pi(2),\dotso,\pi(n)}
-\end{equation*}
-However,
+is a \emph{bijection} \(\pi:\nrange{n} \mapsto \nrange{n}\). %A
+% permutation \(\pi\) can be written in two-line notation
+% \begin{equation*}
+% \pi = \twoline{1,2,\dotso,n}{\pi(1),\pi(2),\dotso,\pi(n)}
+% \end{equation*}
+% However,
 This
@@
 \begin{equation*}
-    p = (\pi,R)\text{ with } \tau \in \mathfrak{S}_k \text{ and } R \subseteq
+    p = (\tau,R)\text{ with } \tau \in \mathfrak{S}_k \text{ and } R \subseteq
     [0,k]\times [0,k].
@@
 
+Looking at the classical pattern \(\perm{2,3,1}\), any occurrences can be arbitrarily
+placed. However it may be interesting to consider occurrences where the elements
+corresponding to the \(2\) and \(3\) are adjacent in the permutation. In the past
+these sort of questions have lead to a variety of definitions.  \textcite{babstein2000}
+considered \emph{vincular} patterns (also known as \emph{generalised} or \emph{dashed} patterns),
+those where two adjacent entries in the pattern must be adjacent in the permutation,
+\ie \(R\) is a union of vertical strips.
+\textcite{MR2652101} look at classes of pattern where both columns and rows
+can be shaded, these are called \emph{bivincular} patterns.
+\emph{Bruhat-restricted} patterns were studied by \textcite{MR2264071} in order
+to establish necessary conditions for a Schubert variety to be Gorenstein.
+Mesh patterns also encompass a subset of \emph{barred} patterns introduced by
+\textcite{MR2716312}, those with only one barred letter.
+All of these definitions subsumed under the definition of mesh patterns.
+The example of an occurrence of \(\perm{2,3,1}\) with adjacent elements \(31\)
+can be represented as a mesh pattern by
+\begin{equation*}
+  \pattern{}{2,3,1}{1/0,1/1,1/2}
+\end{equation*}
+
 We define containment (denoted \(p\preceq \pi\)), and avoidance, of the pattern \(p\) in the permutation
@@
 no elements of \(\tau\) are allowed in the regions of the plot corresponding
-to shaded boxes in the mesh. Formally defined by \textcite{journals/combinatorics/BrandenC11},
+to shaded boxes in the mesh.
+
+Formally defined by \textcite{journals/combinatorics/BrandenC11},
 an \emph{occurrence} of \(p\) in \(\tau\) is a subset \(\omega\) of the plot of
@@
 
-In the past people have studied different classes
-of permutations that can be described by mesh patterns. \textcite{babstein2000}
-considered \emph{vincular} patterns (also known as \emph{generalised} or \emph{dashed} patterns),
-those where two adjacent entries in the pattern must be adjacent in the permutation,
-\ie \(R\) is a union of vertical strips.
-\textcite{MR2652101} look at classes of pattern where both columns and rows
-can be shaded, these are called \emph{bivincular} patterns.
-\emph{Bruhat-restricted} patterns were studied by \textcite{MR2264071} in order
-to establish necessary conditions for a Schubert variety to be Gorenstein.
-Mesh patterns also encompass a subset of \emph{barred} patterns introduced by
-\textcite{MR2716312}, those with only one barred letter.
+% In the past people have studied different classes
+% of permutations that can be described by mesh patterns.
+% \textcite{babstein2000}
+% considered \emph{vincular} patterns (also known as \emph{generalised} or \emph{dashed} patterns),
+% those where two adjacent entries in the pattern must be adjacent in the permutation,
+% \ie \(R\) is a union of vertical strips.
+% \textcite{MR2652101} look at classes of pattern where both columns and rows
+% can be shaded, these are called \emph{bivincular} patterns.
+% \emph{Bruhat-restricted} patterns were studied by \textcite{MR2264071} in order
+% to establish necessary conditions for a Schubert variety to be Gorenstein.
+% Mesh patterns also encompass a subset of \emph{barred} patterns introduced by
+% \textcite{MR2716312}, those with only one barred letter.
 

### 2016-05-18 writing/latex/wilf.tex
Anders chap3/4

@@
 
-Another question often asked in the field of permutation patterns is that
+One question often asked in the field of permutation patterns is that
 of Wilf-equivalence. Two patterns \(\pi\) and \(\sigma\) are said to be
@@
 There are a number of symmetries we can use when examining Wilf-equivalences
-to reduce the amount of work, it can be easily seen that the reverse, complement
+to reduce the amount of work. It can be easily seen that the reverse, complement
 and inverse operations (see \FigureRef{fig:symm}) preserve enumeration, and therefore these classes are trivially
@@
 of order \(8\), the group of symmetries of a square. \emph{Reverse-inverse} and
-\emph{reverse} correspond can be taken as generators of the dihedral group.
+\emph{reverse} correspond to generators of the dihedral group.
 
@@
 the Wilf-equivalence classes of the mesh patterns of length \(1\) inside
-\(\av{\perm{2,3,1}}\), this means that we are considering the set \(\av{\{231,p\}}\)
+\(\av{\perm{2,3,1}}\), this means that we are considering the set \(\av{231,p}\)
 where \(p\) is a mesh-pattern of length 1.
 
-The patterns in the following set are set equivalent,
+The patterns in the following set are coincident,
 \begin{equation*}
@@
 Alternatively we can enumerate
-avoiders by decomposing a non-empty avoider of \(p_1\) around the maximum element
+avoiders of \(p\) by decomposing a non-empty avoider of \(p\) around the maximum element
 in order to give the following structural decomposition.
@@
 \end{equation*}
-Since if the upper-right section was empty the maximum would create an
+If the upper-right section was empty the maximum would create an
 occurrence of the pattern, however no points in this section can create
 an occurrence since the maximum lies in a region corresponding to the
-shading in \(p\). The lower-left section however can create occurrences
+shading in \(p\), so we can use any avoider of \(\perm{2,3,1}\). The lower-left section however can create occurrences
 of \(p\) and therefore must also avoid \(p\), as well as \(\perm{2,3,1}\).
@@
     \shortintertext{Solving for \(\var\) gives}
-    \var(x) &= \frac{1}{1+x-xC(x)}
+    \var(x) &= \frac{1}{1+x-xC(x)}\\
+    \var(x) &= \frac{C(x)}{1+xC(x)}
 \end{align*}
 
-Evaluation of this generating function gives the Fine numbers
-(\href{https://oeis.org/A000957}{OEIS: A000957}).
+Calculating coefficients given by this generating function gives the Fine numbers.
 \begin{equation*}
-    1, 0, 1, 2, 6, 18, 57, 186, 622, 2120, 7338,\dotsc
+    1, 0, 1, 2, 6, 18, 57, 186, 622, 2120, 7338,\dotsc \tag{\href{https://oeis.org/A000957}{OEIS: A000957}}
 \end{equation*}
@@
 All of these classes have the same generating function, namely
+\begin{equation}
+    \var(x) = 1 + xC(x)(C(x)-1). \label{eqn:maxlgen}
+\end{equation}
+The coefficients of this generating function are
 \begin{equation*}
-    \var(x) = 1 + xC(x)(C(x)-1).
-\end{equation*}
-Enumeration of this generating function gives
-\begin{equation*}
-    1, 0, 1, 3, 9, 28, 90, 297, 1001, 3432, 11934,\dotsc \tag{\href{https://oeis.org/A000245}{OEIS: A000245} with offset}
+    1, 0, 1, 3, 9, 28, 90, 297, 1001, 3432, 11934,\dotsc \tag{\href{https://oeis.org/A000245}{OEIS: A000245} with offset \(1\)}
 \end{equation*}
@@
 of a \(\perm{2,3,1}\) avoider around the maximum we can see that an avoider
-of \(r\) must fit into precisely on of the following two forms.
+of \(r\) must fit into precisely one of the following two forms.
 \begin{equation*}
@@
 \end{equation*}
-This generating function is enumerated to give
+Computing coefficients of this generating function gives
 \begin{equation*}
@@
 
-The classes containing following patterns are experimentally Wilf-equivalent up to length 10 in
+The classes containing the following patterns are experimentally Wilf-equivalent up to length 10 in
 \(\av{\perm{2,3,1}}\)
@@
 \begin{equation*}
-    1, 1, 1, 4, 12, 37, 118, 387, 1298, 4433, 15366,\dotsc \tag{\(C_n\) - \href{https://oeis.org/A001453}{A001453} offset 2}
+    1, 1, 1, 4, 12, 37, 118, 387, 1298, 4433, 15366,\dotsc \tag{\(C_n - \)\href{https://oeis.org/A001453}{A001453} offset 2}
 \end{equation*}
@@
 Once again we consider the regions corresponding to the unshaded regions, For each
-pattern
-except the first the two regions are independent, and one contains
+pattern,
+except the first the two, these regions are independent, and one contains
 any avoider of \(\perm{2,3,1}\) and the other must contain a
@@
 The containers
-have generating function \(\frac{x^2C(x)}{(1-x)}\). Enumerating avoiders
+have generating function \(x^2C(x)/(1-x)\). Enumerating avoiders
 therefore gives us
@@
 Therefore we can have any number of points from \(\setrange{0}{k-2}\)
-points in the bottom box. Therefore there are \(k-1\) containers of
+in the bottom box. Therefore there are \(k-1\) containers of
 length \(k\). These permutations correspond to set partitions of \(k\)
-points into exactly two non-overlapping parts partitioned by the first
-element and the minimum.
+points into exactly two non-overlapping parts, such that the first part
+is the decreasing sequence from the first element to the minumum, and
+the second part is the elements in the sequence from the maximum to the last
+element.
+%partitioned by the first element and the minimum.
 
-Now consider the containers of \(m_2\), we know that the unshaded region
+Now consider the containers of \(m_2\). We know that the unshaded region
 must contain a decreasing subsequence, with the point corresponding to
 the \(1\) in the mesh pattern. This decreasing subsequence has \(k-1\)
-points, we can put the point corresponding to the \(2\) above any of these
+points. We can put the point corresponding to the \(2\) above any of these
 points and therefore there are \(k-1\) containers of length \(k\).
@@
 avoiders of length \(k\) for all \(k\) and are Wilf-equivalent.
-The avoiders have general form
+The number of avoiders of length \(k\) is given by
 \begin{equation*}
@@
 \(\av{\perm{2,3,1}}\)
-\begin{align}
-    m_1 = \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/1} &\text{ and }
-    m_2 = \pattern{}{2,1}{0/0,0/1,1/1,1/2,2/0,2/1} \label{eq:singleud}\\
-    \text{ and }
-    m_3 = \pattern{}{2,1}{0/1,0/2,1/0,1/1,1/2,2/1} &\text{ and }
+\begin{gather}
+    m_1 = \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/1},
+    m_2 = \pattern{}{2,1}{0/0,0/1,1/1,1/2,2/0,2/1}, \label{eq:singleud}\\
+    m_3 = \pattern{}{2,1}{0/1,0/2,1/0,1/1,1/2,2/1}, \text{ and }
     m_4 = \pattern{}{2,1}{0/1,0/2,1/0,1/1,1/2,2/0} \label{eq:other}
-\end{align}
+\end{gather}
 
@@
 \end{equation*}
-Since only the first point in the top right region can
+Only the first point in the top right region can
 create an occurrence of \(m_3\) if and only if it is the
-maximum in this region we must avoid starting with the maximum.
+element with largest value in this region, therefore the partial permutation
+in this region must avoid starting with the maximum.
 
@@
 \end{equation*}
-Since an occurrence of \(m_4\) can never occur in the
-top right region, and could only occur between the first point
-in the region and the maximum, if and only if this first point is the minimum.
+Any occurrence of \(m_4\) can never occur in the
+top right region. It could only occur between the maximum and the first point
+in the region, if and only if this first point is the lowest valued element in
+this region.
 Since both \(\mathcal{\varmaxl}_1\) and \(\mathcal{\varmaxl}_3\) have the same
@@
 consider the decomposition around the leftmost point of a permutation in
-\(\av{\perm{2,3,1},m_1}\) we gain the following.
+\(\av{\perm{2,3,1},m_1}\). We have the following
 \begin{equation*}
@@
 \end{equation*}
+Where \(\varmaxl(x)\) is the generating function given in \EquationRef{eqn:maxlgen}.
 This can be enumerated to give the sequence
@@
 \end{equation*}
-Where \(\mathcal{\varfine} = \av{(231,\textpattern{}{1}{0/1,1/0})}\).
+Where \(\mathcal{\varfine} = \av{231,\textpattern{}{1}{0/1,1/0}}\).
 Now consider the decomposition around the maximum of a permutation in
@@
 \(\av{\perm{2,3,1}}\)
-\begin{equation*}
-    \begin{aligned}
-        m_1 = \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/2} &\text{ and }
-        m_2 = \pattern{}{1,2}{0/0,0/1,1/0,2/0,2/1,2/2} \\
-        \text{ and }
-        m_3 = \pattern{}{1,2}{0/2,1/0,1/1,2/0,2/1,2/2} &\text{ and }
+    \begin{gather*}
+        m_1 = \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/2},
+        m_2 = \pattern{}{1,2}{0/0,0/1,1/0,2/0,2/1,2/2}, \\
+        m_3 = \pattern{}{1,2}{0/2,1/0,1/1,2/0,2/1,2/2}, \text{ and }
         m_4 = \pattern{}{1,2}{0/1,1/0,1/1,2/0,2/1,2/2}
-    \end{aligned}
-\end{equation*}
+    \end{gather*}
 
@@
 \end{equation*}
-Where \(\mathcal{\varmaxl}_1 = \av{(231,\textpattern{}{1}{0/0,0/1,1/1})}\).
+Where \(\mathcal{\varmaxl}_1 = \av{231,\textpattern{}{1}{0/0,0/1,1/1}}\).
 Now we decompose the avoiders of \(m_2\) around the leftmost point, we have similar
@@
 
-Consider an occurrence of of \(m_3\) in \(\av{\perm{2,3,1}}\)
+Consider an occurrence of \(m_3\) in \(\av{\perm{2,3,1}}\)
 \begin{equation*}
@@
 the last element is higher than the penultimate element. This can only occur
-if the last element is in a single block in the set partition context, In order
+if the last element is in a single block in the set partition context. In order
 to construct an avoider of length \(n\) we can take any avoider of
@@
         &= \frac{1}{1-x\varfine(x)} && \text{Solving for \(\var_3(x)\)}\\
-        &= \frac{1}{1 - \frac{x}{1+x-xC(x)}} &&\text{Substituting for \(\varfine(x)\)}\\
-        &= \frac{1-xC(x) +x}{1-xC(x)}\\
+        &= \frac{1}{1 - \frac{xC}{1+xC(x)}} &&\text{Substituting for \(\varfine(x)\)}\\
+        &= \frac{1+xC(x)}{1+xC(x)-xC(x)}\\
         &= 1+xC(x)
@@
 \(\av{\perm{2,3,1}}\)
-\begin{equation*}
-    \begin{aligned}
-        m_1 = \pattern{}{1,2}{0/0,0/1,1/0,1/2,2/1} \text{ and }
-        m_2 = \pattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0}\\
-        \text{ and }
-        m_3 = \pattern{}{1,2}{0/1,0/2,1/0,1/2,2/0,2/2} \text{ and }
+\begin{gather*}
+        m_1 = \pattern{}{1,2}{0/0,0/1,1/0,1/2,2/1},
+        m_2 = \pattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0},\\
+        m_3 = \pattern{}{1,2}{0/1,0/2,1/0,1/2,2/0,2/2}, \text{ and }
         m_4 = \pattern{}{2,1}{1/0,1/1,2/1,2/2}\\
-    \end{aligned}
-\end{equation*}
+\end{gather*}
  First we consider an occurrence of \(m_2\) in a
@@
 now contain avoiders of \(\perm{2,3,1}\), all points in the region corresponding
-to \(b_2\) must be to the right of those in the region corresponding to \(b_3\)
-\(\perm{2,3,1}\). Since we already have an occurrence of
+to \(b_2\) must be to the left of those in the region corresponding to \(b_3\) since
+we are in
+\(\av{\perm{2,3,1}}\). Since we already have an occurrence of
 \(m_2\) we do not need to care about creating more occurrences so there are no
@@
         \lastvar_1(x)&= 1+x(\lastvar_1(x)-1)C(x)+x\varfine(x)\\
-        &= 1+x\lastvar_1(x)C(x)-xC(x)+\frac{x}{1+x-xC(x)} &&\text{Substitution of \(\varfine(x)\)}\\
-        \lastvar_1(x)(1-xC(x))&= \frac{x^2C^2(x) - (x^2+2x)C(x) + 2x+1}{1+x-xC(x)}\\
-        \lastvar_1(x) &= \frac{x^2C^2(x) - (x^2+2x)C(x) + x+1+x}{x^2C^2(x) - (x^2+2x)C(x) +x+1}\\
-        &= 1 +\frac{x}{x^2C^2(x) - (x^2+2x)C(x) +x+1}\\
-        &= 1 + \frac{x}{1-x^2C(x)-xC(x)} && xC^2(x) = C(x)-1\\
-        &= 1 + \frac{xC^2(x)}{C^2(x) - xC^3(x)(x+1)}\\
-        &= 1 + \frac{xC^2(x)}{C(x) - xC^2(x) + xC(x)} && xC^2(x) = C(x)-1\\
-        \lastvar_1(x)&= 1 + \frac{xC^2(x)}{1+xC(x)} && C(x) = 1+xC^2(x)\\
+        &= 1+x\lastvar_1(x)C(x)-xC(x)+x\varfine(x)\\
+        \lastvar_1(x)&= 1+xC(x)\varfine(x)\\
+        \lastvar_1(x) &= 1 + \frac{xC^2(x)}{1+xC(x)} &&\text{Substitution of \(\varfine(x)\)}
+        % &= 1 +\frac{x}{x^2C^2(x) - (x^2+2x)C(x) +x+1}\\
+        % &= 1 + \frac{x}{1-x^2C(x)-xC(x)} && xC^2(x) = C(x)-1\\
+        % &= 1 + \frac{xC^2(x)}{C^2(x) - xC^3(x)(x+1)}\\
+        % &= 1 + \frac{xC^2(x)}{C(x) - xC^2(x) + xC(x)} && xC^2(x) = C(x)-1\\
+        % \lastvar_1(x)&= 1 + \frac{xC^2(x)}{1+xC(x)} && C(x) = 1+xC^2(x)\\
     \end{aligned}
@@
 This is a complete decomposition of avoiders of \(m_1\). Now we
-look at an avoider of \(m_2\), this time decomposition is around the maximum
+look at an avoider of \(m_2\), this time decomposition is around the leftmost point
 \begin{equation*}
@@
 \begin{equation*}
-    1, 1, 1, 2, 4, 9, 21, 51, 127, 323, 835,\dotsc
+    1, 1, 1, 2, 4, 9, 21, 51, 127, 323, 835,\dotsc \tag{\href{https://oeis.org/A001006}{OEIS: A001006 with offset 1}}
 \end{equation*}
-This is an offset of the Motzkin numbers \href{https://oeis.org/A001006}{(OEIS: A001006)}.
+This is an offset of the Motzkin numbers.
 
@@
 \end{equation*}
-Where \(\mathcal{\varmaxl}_2 = \av{(231,\textpattern{}{1}{0/1,1/0,1/1})}\).
+Where \(\mathcal{\varmaxl}_2 = \av{231,\textpattern{}{1}{0/1,1/0,1/1}}\).
 This gives us the generating function
@@
 \end{equation*}
-Where \(\mathcal{\varmaxl}_3 = \av{(231,\textpattern{}{1}{0/0,0/1,1/0})}\).
+Where \(\mathcal{\varmaxl}_3 = \av{231,\textpattern{}{1}{0/0,0/1,1/0}}\).
 This gives generating function \(\var_3(x)\) satisfying

### 2016-05-18 writing/presentation/coincidences.tex
Anders chap3/4

@@
 \hline
-\multicolumn{1}{|r|}{First Dominating rule}&\(85\)&\(43\)&\(220\)&\(29\)\\
+\multicolumn{1}{|r|}{\hspace{0pt}\invisible<2->{First Dominating rule}}&\hspace{0pt}\invisible<2->{\(85\)}&\hspace{0pt}\invisible<2->{\(43\)}&\hspace{0pt}\invisible<2->{\(220\)}&\hspace{0pt}\invisible<2->{\(29\)}\\
 \hline
-\multicolumn{1}{|r|}{Second Dominating rule}&\(59\)&\(39\)&\(220\)&\(29\)\\
+\multicolumn{1}{|r|}{\hspace{0pt}\invisible<2->{Second Dominating rule}}&\hspace{0pt}\invisible<2->{\(59\)}&\hspace{0pt}\invisible<2->{\(39\)}&\hspace{0pt}\invisible<2->{\(220\)}&\hspace{0pt}\invisible<2->{\(29\)}\\
 \hline
-\multicolumn{1}{|r|}{Third Dominating rule}&\(56\)&\(39\)&\(220\)&\(29\)\\
+\multicolumn{1}{|r|}{\hspace{0pt}\invisible<2->{Third Dominating rule}}&\hspace{0pt}\invisible<2->{\(56\)}&\hspace{0pt}\hspace{0pt}\invisible<2->{\(39\)}&\hspace{0pt}\invisible<2->{\(220\)}&\hspace{0pt}\hspace{0pt}\invisible<2->{\(29\)}\\
 \hline
@@
     There are \(29\) coincidences of mesh patterns of the form \(\av{\{\perm{3,2,1},\mperm{2,1}{R}}\}\)
+  \end{block}
+\end{frame}
+
+\begin{frame}{Experimental Results}
+  \begin{block}{}
+  \begin{table}[htb]
+\begin{center}
+\begin{tabularx}{\textwidth}{c|Y|Y|Y|Y|}
+\cline{2-5}
+& \multicolumn{4}{|c|}{Dominating Pattern}\\
+\cline{2-5}
+& \multicolumn{2}{|c|}{\(\perm{2,3,1}\)}&\multicolumn{2}{|c|}{\(\perm{3,2,1}\)}\\
+\cline{2-5}
+& \(\perm{1,2}\)&\(\perm{2,1}\)&\(\perm{1,2}\)&\(\perm{2,1}\)\\
+\hline
+\multicolumn{1}{|r|}{No Dominating rule}&\(220\)&\(220\)&\(220\)&\(220\)\\
+\hline
+\multicolumn{1}{|r|}{First Dominating rule}&\(85\)&\(43\)&\(220\)&\(29\)\\
+\hline
+\multicolumn{1}{|r|}{\hspace{0pt}\invisible{Second Dominating rule}}&\hspace{0pt}\invisible{\(59\)}&\hspace{0pt}\invisible{\(39\)}&\hspace{0pt}\invisible{\(220\)}&\hspace{0pt}\invisible{\(29\)}\\
+\hline
+\multicolumn{1}{|r|}{\hspace{0pt}\invisible{Third Dominating rule}}&\hspace{0pt}\invisible{\(56\)}&\hspace{0pt}\hspace{0pt}\invisible{\(39\)}&\hspace{0pt}\invisible{\(220\)}&\hspace{0pt}\hspace{0pt}\invisible{\(29\)}\\
+\hline
+\multicolumn{1}{|r|}{Experimental class size}&\(56\)&\(39\)&\(213\)&\(29\)\\
+\hline
+\end{tabularx}
+\end{center}
+    \caption{Coincidence class number reduction by application of Dominating rules}
+\end{table}
 \end{block}
@@
 
+\begin{frame}{Experimental Results}
+  \begin{block}{}
+  \begin{table}[htb]
+\begin{center}
+\begin{tabularx}{\textwidth}{c|Y|Y|Y|Y|}
+\cline{2-5}
+& \multicolumn{4}{|c|}{Dominating Pattern}\\
+\cline{2-5}
+& \multicolumn{2}{|c|}{\(\perm{2,3,1}\)}&\multicolumn{2}{|c|}{\(\perm{3,2,1}\)}\\
+\cline{2-5}
+& \(\perm{1,2}\)&\(\perm{2,1}\)&\(\perm{1,2}\)&\(\perm{2,1}\)\\
+\hline
+\multicolumn{1}{|r|}{No Dominating rule}&\(220\)&\(220\)&\(220\)&\(220\)\\
+\hline
+\multicolumn{1}{|r|}{First Dominating rule}&\(85\)&\(43\)&\(220\)&\(29\)\\
+\hline
+\multicolumn{1}{|r|}{Second Dominating rule}&\(59\)&\(39\)&\(220\)&\(29\)\\
+\hline
+\multicolumn{1}{|r|}{\hspace{0pt}\invisible{Third Dominating rule}}&\hspace{0pt}\invisible{\(56\)}&\hspace{0pt}\invisible{\(39\)}&\hspace{0pt}\invisible{\(220\)}&\hspace{0pt}\invisible{\(29\)}\\
+\hline
+\multicolumn{1}{|r|}{Experimental class size}&\(56\)&\(39\)&\(213\)&\(29\)\\
+\hline
+\end{tabularx}
+\end{center}
+    \caption{Coincidence class number reduction by application of Dominating rules}
+\end{table}
+\end{block}
+\end{frame}
+
 \begin{frame}{Third Dominating rule}
@@
 \end{frame}
+
+\begin{frame}{Experimental Results}
+  \begin{block}{}
+  \begin{table}[htb]
+\begin{center}
+\begin{tabularx}{\textwidth}{c|Y|Y|Y|Y|}
+\cline{2-5}
+& \multicolumn{4}{|c|}{Dominating Pattern}\\
+\cline{2-5}
+& \multicolumn{2}{|c|}{\(\perm{2,3,1}\)}&\multicolumn{2}{|c|}{\(\perm{3,2,1}\)}\\
+\cline{2-5}
+& \(\perm{1,2}\)&\(\perm{2,1}\)&\(\perm{1,2}\)&\(\perm{2,1}\)\\
+\hline
+\multicolumn{1}{|r|}{No Dominating rule}&\(220\)&\(220\)&\(220\)&\(220\)\\
+\hline
+\multicolumn{1}{|r|}{First Dominating rule}&\(85\)&\(43\)&\(220\)&\(29\)\\
+\hline
+\multicolumn{1}{|r|}{Second Dominating rule}&\(59\)&\(39\)&\(220\)&\(29\)\\
+\hline
+\multicolumn{1}{|r|}{Third Dominating rule}&\(56\)&\(39\)&\(220\)&\(29\)\\
+\hline
+\multicolumn{1}{|r|}{Experimental class size}&\(56\)&\(39\)&\(213\)&\(29\)\\
+\hline
+\end{tabularx}
+\end{center}
+    \caption{Coincidence class number reduction by application of Dominating rules}
+\end{table}
+\end{block}
+\end{frame}

### 2016-05-19 writing/latex/code.tex
Add changes to final section

@@
 and is split into the following files.
+
+\subsubsection*{\texttt{classes.py}}
+Contains the 220 equivalence classes of mesh patterns of length 2 shown in
+% \ChapterRef{cha:equivs}.
+\subsubsection*{\texttt{helper\_funcs.py}}
+Contains functions to convert between different classes used in experimentation
+as well as a class that functions as an accumulator.
+\subsubsection*{\texttt{p\_rule.py}}
+Contains functions that compute equivalences by the first and second dominating
+rules, and functions that return coincidences not explained by the rules.
+\subsubsection*{\texttt{wilf.py}}
+Contains a class that functions as a counter for Wilf-equivalences based on
+a dictionary, functions to calculate trivial Wilf classes and functions that
+calculate Wilf-equivalences experimentally
+\subsubsection*{\texttt{main.py}}
+The main file to import that automatically runs all of the experimental setup
+and calculations for coincidences and Wilf-equivalences.
+\subsubsection*{\texttt{scratchpad.py}}
+A collection of anonymous functions and utilities that do not fit into other
+files.
 

### 2016-05-19 writing/latex/confut.tex
Add changes to final section

@@
 \nextvar
-From \ChapterRef{chap:coincs} it can be seen that automatically classifying
-coincidences of mesh patterns is a difficult task, establishing rules for longer
-dominating patterns requires many more cases to be taken. It would however
-be interesting to consider the application of the third Dominating Rule to mesh
-patterns without any dominating pattern in order to try to capture some of the
-coincidences described in\textcite{DBLP:journals/combinatorics/HilmarssonJSVU15}
+% From \ChapterRef{chap:coincs} it can be seen that automatically classifying
+% coincidences of mesh patterns is a difficult task, establishing rules for longer
+% dominating patterns requires many more cases to be taken.
+If we consider a similar system for dominating patterns of length \(4\) and
+mesh patterns of length \(2\), it can be seen that the number of cases required
+to establish rules increases to a number that is infeasible to compute manually.
+For an extension of the First Dominating rule alone, we would have to consider
+placement of points in any pair of unshaded regions. The fact that the rules
+established do not completely cover the coincidences with a dominating pattern
+of length \(3\) (see \TableRef{tab:domclasses}) shows that this is a difficult
+task.
+
+It is interesting to consider the application of the Third Dominating rule,
+as well as the simple extension of allowing a sequence of \texttt{add\_point}
+operations, to meshpatterns without any dominating pattern in order to try to
+capture some of the coincidences described in
+\textcite{DBLP:journals/combinatorics/HilmarssonJSVU15}
  and \textcite{DBLP:journals/corr/ClaessonTU14}.
+ \begin{example}
+   We can establish the coincidence between the patterns
+   \begin{equation*}
+     m_1 = \pattern{}{1,2}{0/1,0/2,1/1,1/2,2/0}, \text{ and } m_2 = \pattern{}{1,2}{0/2,1/0,1/1,2/0,2/1}
+   \end{equation*}
+   That is not explained by the methods presented by \textcite{DBLP:journals/corr/ClaessonTU14}.
+
+   Consider a permutation containing \(m_1\),
+   \begin{equation*}
+     \begin{tikzpicture}[scale=0.5]
+         \modpattern[5]{}{1,2}{0/1,0/2,1/1,1/2,2/0}
+         \draw (1.5,0.5) node {\(Y\)};
+         \draw (2.5,1.5) node {\(X\)};
+     \end{tikzpicture}
+   \end{equation*}
+   If the regions corresponding to both \(X\) and \(Y\) are empty then we have
+   an occurrence of \(m_2\).
+   Consider if the region corresponding to \(X\) is non-empty, we can then choose
+   the lowest valued point in this region
+   \begin{equation*}
+     \begin{tikzpicture}[scale=0.5]
+         \modpattern[5]{1,3}{1,3,2}{0/1,0/2,0/3,1/1,1/2,1/3,2/0,2/1,3/0,3/1}
+         \draw (1.5,0.5) node {\(Y\)};
+     \end{tikzpicture}
+   \end{equation*}
+   If the region corresponding to \(Y\) is empty then we have an occurrence of
+   \(m_2\) with the indicated points.
+   Now if the region corresponding to \(Y\) is non-empty, we can choose the
+   rightmost point in this region.
+   \begin{equation*}
+     \begin{tikzpicture}[scale=0.5]
+         \modpattern[5]{2,4}{2,1,4,3}{0/2,0/3,0/4,1/2,1/3,1/4,2/0,2/1,2/2,2/3,2/4,3/0,3/1,3/2,4/0,4/1,4/2}
+     \end{tikzpicture}
+   \end{equation*}
+   And now the two indicated points form an occurrence of \(m_2\).
+   We have therefore shown that any occurrence of \(m_1\) is an occurrence of
+   \(m_2\) and we can easily show the converse by the same reasoning, so \(m_1\)
+   and \(m_2\) are coincident.
+   This is captured by an extension of the Third Dominating rule where we allow
+   multiple steps of adding points before we check for subpattern containment.
+ \end{example}
 It is not possible to apply the first and second Dominating Rules to the pattern
@@
 It would be interesting to consider a systematic explanation of Wilf-equivalences
-amongst classes where \(\perm{3,2,1}\) is the dominating pattern using the
+amongst classes where \(\perm{3,2,1}\) is the dominating pattern, possible using the
 construction presented in \cite[Sec.~12]{2015arXiv151203226B}, in order to directly
@@
 between classes with different dominating patterns. For example, it is possible to show that
-the sets \(\scriptvar = \av{\left\{\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1},\perm{2,3,1}\right\}}\)\nextvar and
-\(\scriptvar = \av{\left\{\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1},\perm{3,2,1}\right\}}\), are Wilf-equivalent.
+the sets \(\scriptvar = \av{\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1},\perm{2,3,1}}\)\nextvar and
+\(\scriptvar = \av{\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1},\perm{3,2,1}}\), are Wilf-equivalent.
 This can be seen by considering an occurrence of the mesh pattern,

### 2016-05-20 writing/latex/abstracteng.tex
Slides and fixes t bibliography

@@
 have the same number of avoiders at every length, we completely
-Wilf-classifies mesh patterns of length 2 when avoiding the classical pattern 231.
-% Finally we attempt to show some non-trivial Wilf-equivalences between avoiders
-% of sets of the form \(231,m_1\) and \(231,m_2\)
+Wilf-classify mesh patterns of length 2 when avoiding the classical pattern 231.
+Finally we attempt to show some non-trivial Wilf-equivalences between avoiders
+of sets of the form \(231,m_1\) and \(321,m_2\), as well as discussing possible
+future work that could be derived from this work.

### 2016-05-20 writing/latex/introduction.tex
Slides and fixes t bibliography

--- 2016-05-09 writing/latex/introduction.tex
@@
 \section{Mesh Patterns}
-Mesh patterns were introduced by \textcite{journals/combinatorics/BrandenC11} to
+When looking at a classical pattern, for example \(\perm{2,3,1}\), any occurrences can be arbitrarily
+placed. However it may be interesting to consider occurrences where the elements
+corresponding to the \(2\) and \(3\) are adjacent in the permutation. In the past
+these sort of questions have lead to a variety of definitions.  \textcite{babstein2000}
+considered \emph{vincular} patterns (also known as \emph{generalised} or \emph{dashed} patterns),
+those where two adjacent entries in the pattern must be adjacent in the permutation,
+\ie \(R\) is a union of vertical strips.
+\textcite{MR2652101} look at classes of pattern where both columns and rows
+can be shaded, these are called \emph{bivincular} patterns.
+\emph{Bruhat-restricted} patterns were studied by \textcite{MR2264071} in order
+to establish necessary conditions for a Schubert variety to be Gorenstein.
+Mesh patterns also encompass a subset of \emph{barred} patterns introduced by
+\textcite{MR2716312}, those with only one barred letter.
+
+All of these definitions are subsumed under the definition of \emph{mesh patterns},
+introduced by \textcite{journals/combinatorics/BrandenC11} to
 capture explicit expansions for certain permutation statistics. They are a
@@
 
+The example of an occurrence of \(\perm{2,3,1}\) with adjacent elements \(31\)
+can be represented as a mesh pattern by
+\begin{equation*}
+  \pattern{}{2,3,1}{1/0,1/1,1/2,1/3}
+\end{equation*}
+
 We define containment (denoted \(p\preceq \pi\)), and avoidance, of the pattern \(p\) in the permutation
@@
 
-In the past people have studied different classes
-of permutations that can be described by mesh patterns. \textcite{babstein2000}
-considered \emph{vincular} patterns (also known as \emph{generalised} or \emph{dashed} patterns),
-those where two adjacent entries in the pattern must be adjacent in the permutation,
-\ie \(R\) is a union of vertical strips.
-\textcite{MR2652101} look at classes of pattern where both columns and rows
-can be shaded, these are called \emph{bivincular} patterns.
-\emph{Bruhat-restricted} patterns were studied by \textcite{MR2264071} in order
-to establish necessary conditions for a Schubert variety to be Gorenstein.
-Mesh patterns also encompass a subset of \emph{barred} patterns introduced by
-\textcite{MR2716312}, those with only one barred letter.
-
 Avoiding pairs of patterns of the same length with certain properties has

### 2016-05-20 writing/latex/wilf.tex
Slides and fixes t bibliography

@@
 due to the fact that every permutation, except the empty
-permutation must contain an occurrence of all of these patterns.
+permutation, must contain an occurrence of all of these patterns.
 
@@
     Given the permutation \(\pi = \perm{5,4,2,1,3,9,6,8,7}\) this corresponds to the
-    partition \(\{\{5,4,2,1\},\{3\},\{9,6\},\{7,8\}\}\).
+    partition \(\{\{5,4,2,1\},\{3\},\{9,6\},\{8,7\}\}\).
 \end{example}
 
-We call the least element in each block the \emph{block bottom}
+We call the least element in each block the \emph{block bottom}.
 

### 2016-05-20 writing/presentation/coincidences.tex
Slides and fixes t bibliography

@@
     Two mesh patterns are said to be \emph{coincident} if they avoid the
-    same set of patterns at every length.
+    same set of permutations at every length.
   \end{definition}
@@
     \caption{If the conditions of The Second Dominating rule are satisfied the box
-    \((a-1,b)\) can be shaded.}
+    \((a,b)\) can be shaded.}
 \end{figure}
@@
 \end{frame}
+
+\subsection{Special Cases}
+\label{sub:Special Cases}
+\begin{frame}{Special Cases}
+  There are \(7\) coincidences of the form \(\av{\perm{3,2,1},m}\) that are not
+  explained by the rules.
+  \begin{example}
+    \begin{equation*}
+      m_1 = \pattern{}{1,2}{0/1,0/2,2/0,2/1} \text{ and } m_2 = \pattern{}{1,2}{0/1,0/2,1/1,2/0,2/1}
+    \end{equation*}
+    This coincidence is explained by mathematical induction on the number of
+    points in region corresponding to the middle box.
+  We call this number \(n\).
+  \end{example}
+\end{frame}
+\begin{frame}
+  \begin{example}
+    \begin{description}
+    \only<1-2>{\item [Base Case \((n=0)\):] The base case holds since we can freely shade the
+        box if it contains no points.}
+    \only<2>{\item [Inductive Hypothesis \((n=k)\):] Suppose that we can find an occurrence
+        of the second pattern if we have an occurrence of the first with \(k\) points
+        in the middle box.}
+    \only<3>{\item [Inductive Step \((n=k+1)\)] Suppose that we have \((k+1)\) points in
+        the middle box. Choose the bottom most point in the middle box, this gives the
+        mesh pattern
+     \begin{equation*}
+        \begin{tikzpicture}[scale=0.5]
+            \modpattern[5]{}{1,2,3}{0/1,0/2,0/3,1/1,2/1,3/0,3/1,3/2}
+            \draw (2.5,0.5) node {\(X\)};
+        \end{tikzpicture}
+    \end{equation*}
+        Now we need to consider the box labelled \(X\). If this box is empty then
+        we have an occurrence of \(m_2\) and are done.}
+        \only<4>{\item [Inductive Step \((n=k+1)\) (cont.)] If this box contains
+        any points then we gain some extra shading on the mesh pattern
+        due to the dominating pattern
+     \begin{equation*}
+        \begin{tikzpicture}[scale=0.5]
+            \modpattern[5]{2,3}{1,2,3}{0/1,0/2,0/3,1/1,1/2,1/3,2/1,3/0,3/1,3/2}
+            \filldraw (2.5,0.5) circle (4 pt);
+        \end{tikzpicture}
+    \end{equation*}
+        The two highlighted points form an occurrence of \(m_1\) with \(k\)
+        points in the middle box, and thus by the Inductive Hypothesis we
+        are done.}
+\end{description}
+  \end{example}
+\end{frame}

### 2016-05-20 writing/presentation/confut.tex
Slides and fixes t bibliography

\section{Conclusions and Future Work}
\label{sec:Conclusions and Future Work}
\subsection{Conclusions}
\label{sub:Conclusions}
\begin{frame}{Conclusions}
  \begin{itemize}
  \item Automatic coincidence classification is difficult.
  \begin{itemize}
    \item Completely classified for length 2 mesh patterns and length 3 classical patterns.
    \item Gets harder with larger dominating patterns.
  \end{itemize}
  \item There are a number of Wilf-classes that give interesting
  enumerations.
  \begin{itemize}
    \item Would be interesting to try and connect these sets to different objects.
  \end{itemize}
  \end{itemize}
\end{frame}
\subsection{Future Work}
\label{sub:Future Work}
\subsubsection{Extensions of rules}
\label{subs:Extensions of rules}
\begin{frame}{Future Work: Extensions of rules}
    One can consider application of the Third Dominating rule, as well as a
    simple extension, without any dominating pattern. This can capture the special
    case described by Claesson, Tenner, and Ulfarsson in \cite{ABH}
    \begin{equation*}
     m_1 = \pattern{}{1,2}{0/1,0/2,1/1,1/2,2/0}, \text{ and } m_2 = \pattern{}{1,2}{0/2,1/0,1/1,2/0,2/1}
   \end{equation*}

   We cannot apply the first, or second, rule without a dominating pattern.

   We can also consider taking sets of mesh patterns, or sets of dominating patterns.
\end{frame}

\subsubsection{Equivalences with different dominating patterns}
\label{subs:Equivalences with different dominating patterns}
\begin{frame}{Future Work: Equivalences with different dominating patterns}
  It would be interesting to consider Wilf-equivalences
amongst classes where \(\perm{3,2,1}\) is the dominating pattern.

It is also interesting to consider Wilf-equivalence when we have different dominating
patterns
We can show that the sets \(\mathcal{T} = \av{\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1},\perm{2,3,1}}\) and
\(\mathcal{U} = \av{\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1},\perm{3,2,1}}\)
are Wilf-equivalent.
\pause
\begin{equation*}
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
        \draw (0.5,1.5) node {\(b_1\)};
        \draw (0.5,2.5) node {\( b_2\)};
        \draw (2.5,2.5) node {\( b_3\)};
        \modpattern[2]{2}{1,2}{0/0,1/0,1/1,1/2,2/0,2/1}
    \end{tikzpicture}
    }
\end{equation*}
\end{frame}

### 2016-05-20 writing/presentation/introduction.tex
More Presentation
Slides and fixes t bibliography

@@
 
-  More intuitively \textquote[Knuth]{A \emph{permutation of \(n\) objects} is
+  More intuitively \textquote[Knuth~\cite{Knuth}]{A \emph{permutation of \(n\) objects} is
   an arrangement of \(n\) distinct objects in a row}.
@@
   We write permutations in \emph{one-line notation}, writing the entries of
-  the entries of the permutation in order \(\pi\)
+  the entries of the permutation in order
   \begin{equation*}

### 2016-05-20 writing/presentation/wilf.tex
Slides and fixes t bibliography

\section{Wilf-equivalences}
\label{sec:Wilf-equivalences}
\subsection{Wilf-equivalence}
\label{sub:Wilf-equivalence}
\begin{frame}{Wilf-equivalence}
  \begin{definition}
  Two patterns are said to be \emph{Wilf-equivalent} if they avoid (or contain)
  the same number of patterns at every length.
\end{definition}
 All coincident pattern classes are Wilf-equivalent.

 The permutations in \(\av{\perm{2,3,1}}\) of length \(n\) are counted by the
 Catalan numbers.
 \begin{equation}
   C_0 = 1, \text{ and } C_{n+1} =\sum_{i=0}^{n}C_iC_{n-i}
 \end{equation}
 This gives the sequence
 \begin{equation*}
   1,1,2,5,14,42,132,1430,4862,\dotsc
 \end{equation*}
\end{frame}
\subsection{Set Partitions}
\label{sub:Set Partitions}
\begin{frame}{Set Partitions}
  \begin{block}{Note}
    The avoiders of the pattern \(q = \mperm{2,3,1}{\{(1,0),(1,1),(1,2),(1,3)\}}\),
    \textpattern{}{2,3,1}{1/0,1/1,1/2,1/3},
    in \(\mathfrak{S}_n\) are in one-to-one correspondence with partitions of
    \(\nrange{n}\). (Claesson)
\end{block}
\begin{example}
    Given the permutation \(\pi = \perm{5,4,2,1,3,9,6,8,7}\) this corresponds to the
    partition \(\{\{5,4,2,1\},\{3\},\{9,6\},\{7,8\}\}\).
\end{example}
All permutations in \(\av{\perm{2,3,1}}\) are also in \(\av{q}\).
\end{frame}
\begin{frame}
  Set Partitions can be used to explain certain Wilf-equivalences.
  \begin{example}
    The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
  \begin{equation*}
      m_1 = \pattern{}{1,2}{0/0,0/2,
                      1/0,1/1,
                      2/0,2/1,2/2}
      \text{ and }
      m_2 = \pattern{}{1,2}{0/0,
                      1/0,1/1,1/2,
                      2/0,2/1,2/2}
  \end{equation*}
  Consider containers of the patterns in \(\av{\perm{2,3,1}}\).

  \only<2>{There can only ever be a single occurrence of \(m_1\). The shading
  shows us that this is a partition of \(k\) elements into two non-overlapping
  parts by the first element and the last element.}
  \only<3>{There can only ever be a single occurrence of \(m_2\). The shading
  shows us that this is a partition of \(k\) elements into two non-overlapping
  parts, where the split in parts is determined by the value of the last element
  in the permutation.}
  \only<4->{Therefore for each of these patterns a container of length \(k\)
  consists of a decreasing sequence of length \(k-2\) split into two parts.
  There are \(k-1\) ways to perform such a split. So the number of avoiders of
  length \(k\) is
  \begin{equation*}
    K_k = C_k - (k-1), K_0 = 1
  \end{equation*}}
  \end{example}

\end{frame}
\subsection{Generating Functions}
\label{sub:Generating Functions}
\begin{frame}{Generating {\huge FUN}ctions}
  \begin{block}{}
    \blockquote[Herbert Wilf, \emph{generatingfunctionology}~\cite{Wilf}]{
    A generating function is a clothesline on which we hang up a sequence of numbers for display.}
    % \flushright -Herbert Wilf, \emph{generatingfunctionology}
  \end{block}
  \begin{definition}
    The \emph{ordinary generating function (OGF)} of a sequence \((A_n)\) is the
    formal power series
    \begin{equation}
      A(x) = \sum_{n=0}^{\infty} A_nx^n
    \end{equation}
    The right hand side is the Taylor expansion of the left hand side at \(0\).
  \end{definition}
\end{frame}
\begin{frame}
  \begin{example}
    The generating function for the constant sequence \(\{1,1,1,1,\dotsc\}\) is
    \begin{equation*}
      f(x) = \frac{1}{1-x}
    \end{equation*}
  \end{example}
\end{frame}
\begin{frame}
  \begin{example}
  The generating function for the avoiders of the permutation \(\perm{2,3,1}\)
  satisfies
  \begin{equation}
    C(x) = 1 +xC^2(x)
  \end{equation}
  This can be seen structurally (\(\mathcal{C}\) is the set of \(\perm{2,3,1}\) avoiders)
  \begin{figure}[!ht]
    \centering
    \(\mathcal{C} = \varepsilon \mathrel{\sqcup}\) \decompmax{0/1,1/0}{\only<1>A\only<2->{\mathcal{C}}}{}{}{\only<1>B\only<2->{\mathcal{C}}}
\end{figure}
\only<3->{
Solving for \(C(x)\) and taking the Taylor expansion at \(0\) gives
\begin{equation*}
  1+x+2x^2+5x^3+14x^4+42x^5+132x^6+1430x^7+4862x^8+\dotsb
\end{equation*}
}

\end{example}
\end{frame}
\begin{frame}
  Now we consider one of the Wilf-equivalences between mesh patterns in \(\av{\perm{2,3,1}}\).
  \begin{equation*}
    m_1 = \pattern{}{1,2}{1/0,1/1,1/2,2/0,2/1} \text{ and }
    m_2 = \pattern{}{2,1}{0/1,1/0,1/1,1/2,2/1}
  \end{equation*}
\end{frame}

\begin{frame}
  First we consider the avoiders \(\mathcal{M}_1\) of \(m_1\) in \(\av{\perm{2,3,1}}\).
  \begin{equation*}
    \mathcal{M}_1 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{M}_1}{}{}{\mathcal{M}^\prime_1}
\end{equation*}
Where \(\mathcal{M}^\prime_1\) is a permutation avoiding \(\perm{2,3,1}, m_1\)
and \(\textpattern{}{1}{0/0,0/1,1/0}\).\pause Now considering a permutation in \(\mathcal{M}^\prime_1\)
\begin{equation*}
    \mathcal{M}^\prime_1 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\scriptstyle \mathcal{M}_1\setminus\varepsilon}{}{}{\mathcal{M}^\prime_1}
\end{equation*}
\end{frame}

\begin{frame}
  Next we consider the avoiders \(\mathcal{M}_2\) of \(m_2\) in \(\av{\perm{2,3,1}}\).
  \begin{equation*}
    \mathcal{M}_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{M}^\prime_2}{}{}{\mathcal{M}_2}
\end{equation*}
Where \(\mathcal{M}^\prime_2\) is a permutation avoiding \(\perm{2,3,1}, m_2\)
and \(\textpattern{}{1}{0/0,0/1,1/1}\).\pause Now considering a permutation in \(\mathcal{M}^\prime_1\)
\begin{equation*}
    \mathcal{M}^\prime_2 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{M}^\prime_2}{}{}{\scriptstyle \mathcal{M}_2\setminus\varepsilon}
\end{equation*}
\end{frame}
\begin{frame}
  Both of these sets therefore have generating function \(M(x)\) satisfying
  \begin{align}
    M(x) &= 1 + xM(x)M^\prime(x) \label{eqn:Pgen}\\
    M^\prime(x) &= 1 + x(M(x)-1)M^\prime(x)\label{eqn:PprimeGen}
 \end{align}\pause
 Solving Equation~\ref{eqn:PprimeGen} for \(M^\prime(x)\) and substituting into
Equation~\ref{eqn:Pgen} gives us that the generating function for
\(M(x)\) satisfies
\begin{equation}
    M(x) = xM^2(x) - x(M(x) - 1) + 1
\end{equation}\pause
Evaluating coefficients of \(M(x)\) then gives the sequence
\begin{equation*}
    1, 1, 1, 2, 4, 9, 21, 51, 127, 323, 835,\dotsc
\end{equation*}
This is an offset of the Motzkin numbers.
\end{frame}

### 2016-05-21 writing/latex/coincidence.tex
Add changes to final section
Slides and fixes t bibliography
Changes to presentation from Henning's notes

--- 2016-05-09 writing/latex/coincidence.tex
@@
     \(\pi \preceq \adda{\sigma}{(a,b)}\)\\(\(\pi \preceq \addd{\sigma}{(a,b)}\))
-    then in any occurrence of \(m\) in a permutation \(\varrho\) the region
-    corresponding to the box \((a,b)\) can only contain an increasing
-    (decreasing) subsequence of \(\varrho\).
+    ,then in any occurrence of \(m\) in a permutation \(\varrho\), the region
+    corresponding to the box \((a,b)\) can only contain an decreasing
+    (increasing) subsequence of \(\varrho\).
 \end{lemma}
@@
     \caption{If the conditions of \PropositionRef{prop:dom2} are satisfied the box
-    \((a-1,b)\) can be shaded.}
+    \((a,b)\) can be shaded.}
 \end{figure}
@@
 we originally desired, and hence the two patterns are coincident.
+It is also possible to calculate this coincidence by an extension of the Third
+Dominating rule, where we allow a sequence of \texttt{add\_point} operations, this
+is discussed further in the future work section.
 

### 2016-05-21 writing/presentation/coincidences.tex
Changes to presentation from Henning's notes

@@
   \begin{block}{}
-    Classical patterns can never be coincident.
+    Distinct classical patterns can never be coincident.
 
@@
 in \(R\), and a dominating classical pattern \(\pi = (\pi,\emptyset)\) if
-\(\pi \preceq \adda{\sigma}{(a,b)}\)\\(\(\pi \preceq \addd{\sigma}{(a,b)}\))
+\(\pi \preceq \adda{\sigma}{(a,b)}\)\\
 ,then in any occurrence of \(m\) in a permutation \(\varrho\), the region
-corresponding to the box \((a,b)\) can only contain an increasing
-(decreasing) subsequence of \(\varrho\).
+corresponding to the box \((a,b)\) can only contain an decreasing
+subsequence of \(\varrho\).
   \end{lemma}
@@
     This coincidence is explained by mathematical induction on the number of
-    points in region corresponding to the middle box.
+    points in the region corresponding to the middle box.
   We call this number \(n\).

### 2016-05-21 writing/presentation/introduction.tex
Changes to presentation from Henning's notes

@@
 
-  We write permutations in \emph{one-line notation}, writing the entries of
+  We write permutations in \emph{one-line notation}, writing
   the entries of the permutation in order
@@
 
-  We the set of permutations of length \(n\) avoiding a pattern \(\sigma\) is
+  The set of permutations of length \(n\) avoiding a pattern \(\sigma\) is
   denoted as \(\Av_n(\sigma)\) and
-  \(\av{\sigma} = \bigcup_{i=0}^{\infty}\Av_i(\sigma)\).
+  \begin{equation*}
+    \av{\sigma} = \bigcup_{i=0}^{\infty}\Av_i(\sigma)
+  \end{equation*}
 \end{frame}

### 2016-05-21 writing/presentation/wilf.tex
Changes to presentation from Henning's notes

@@
   \begin{definition}
-  Two patterns are said to be \emph{Wilf-equivalent} if they avoid (or contain)
-  the same number of patterns at every length.
+  Two patterns are said to be \emph{Wilf-equivalent} if the set of avoiders
+  of the patterns is the same size at every length.
 \end{definition}
@@
     in \(\mathfrak{S}_n\) are in one-to-one correspondence with partitions of
-    \(\nrange{n}\). (Claesson)
+    \(\nrange{n}\). (Claesson \cite{A})
 \end{block}
@@
  \end{align}\pause
- Solving Equation~\ref{eqn:PprimeGen} for \(M^\prime(x)\) and substituting into
-Equation~\ref{eqn:Pgen} gives us that the generating function for
+ Solving \ref{eqn:PprimeGen} for \(M^\prime(x)\) and substituting into
+\ref{eqn:Pgen} gives us that the generating function for
 \(M(x)\) satisfies

### 2016-05-23 writing/latex/abstracteng.tex
Henning pres & IS abstract

@@
 A permutation is an arrangement of \(n\) objects.
-Permutation classes of classical permutation patterns
-capture many interesting properties, such as
+Sets of permutations classified by the avoidance of classical permutation
+patterns capture many interesting properties, such as
 stack-sortability, and have links to many different combinatorial objects.
@@
 coincidence among mesh patterns, whilst also avoiding a longer classical
-pattern. These conditions, along with two special cases are used to completely
-classify coincidence amongst families containing a mesh pattern of length 2
-and a classical pattern of length 3. Two patterns are Wilf-equivalent if they
+pattern. These conditions, along with two special cases, are used to completely
+classify coincidence amongst families containing a mesh pattern of length \(2\)
+and a classical pattern of length \(3\).
+
+Two patterns are Wilf-equivalent if they
 have the same number of avoiders at every length, we completely
-Wilf-classify mesh patterns of length 2 when avoiding the classical pattern 231.
+Wilf-classify mesh patterns of length \(2\) when avoiding the classical pattern \(231\).
 Finally we attempt to show some non-trivial Wilf-equivalences between avoiders

### 2016-05-23 writing/presentation/coincidences.tex
Henning pres & IS abstract

@@
   \begin{definition}
-    Two mesh patterns are said to be \emph{coincident} if they avoid the
+    Two mesh patterns are said to be \emph{coincident} if they are avoided by the
     same set of permutations at every length.
@@
 
+    Coincidences of mesh patterns of length \(2\) are classified.
+
     Aim to establish rules that classify coincidences when we
     have one mesh pattern and one classical pattern.
+
+    We call the classical pattern a \emph{dominating pattern}.
   \end{block}
@@
     \begin{tikzpicture}[scale=\picscale]
-        {\only<3-> \fill[pattern=north east lines, pattern color=black!75] (1,1) rectangle +(1,1);}
+        {\only<3-4> \fill[pattern=north east lines, pattern color=black!75] (1,1) rectangle +(1,1);}
         \modpattern{}{2,1}{0/0,0/2,1/0,2/1}
@@
         {\only<2> \filldraw (1.33,1.66) circle (3pt);}
+        {\only<3> \filldraw (2.66,0.33) circle (3pt);}
+        {\only<3> \filldraw (2.33,0.66) circle (3pt);}
     \end{tikzpicture}

### 2016-05-23 writing/presentation/wilf.tex
Henning pres & IS abstract

@@
     Given the permutation \(\pi = \perm{5,4,2,1,3,9,6,8,7}\) this corresponds to the
-    partition \(\{\{5,4,2,1\},\{3\},\{9,6\},\{7,8\}\}\).
+    partition \(\{\{5,4,2,1\},\{3\},\{9,6\},\{8,7\}\}\).
 \end{example}
@@
 Where \(\mathcal{M}^\prime_2\) is a permutation avoiding \(\perm{2,3,1}, m_2\)
-and \(\textpattern{}{1}{0/0,0/1,1/1}\).\pause Now considering a permutation in \(\mathcal{M}^\prime_1\)
+and \(\textpattern{}{1}{0/0,0/1,1/1}\).\pause Now considering a permutation in \(\mathcal{M}^\prime_2\)
 \begin{equation*}

### 2016-05-26 writing/latex/introduction.tex
Make changes suggested by Michael to thesis. Bjarki's pedantry in presentation.

@@
 considered \emph{vincular} patterns (also known as \emph{generalised} or \emph{dashed} patterns),
-those where two adjacent entries in the pattern must be adjacent in the permutation,
-\ie \(R\) is a union of vertical strips.
+those where two adjacent entries in the pattern must be adjacent in the permutation.
 \textcite{MR2652101} look at classes of pattern where both columns and rows
@@
 
-The example of an occurrence of \(\perm{2,3,1}\) with adjacent elements \(31\)
+The example of an occurrence of \(\perm{2,3,1}\) with adjacent elements \(23\)
 can be represented as a mesh pattern by
@@
 
-Two patterns are said to be \emph{coincident} if they avoid the same set of
-permutations and \emph{Wilf-equivalent} if they avoid the same number of permutations
+Two patterns are said to be \emph{coincident} if they are avoided by the same set of
+permutations and \emph{Wilf-equivalent} if they are avoided by the same number of permutations
 at every length.

### 2016-05-26 writing/latex/wilf.tex
Make changes suggested by Michael to thesis. Bjarki's pedantry in presentation.

@@
 Computing avoiders up to length \(10\) suggest that there are at least \(23\)
-Wilf-classes, of which \(13\) are non-trivial.
+Wilf-classes, of which \(13\) are non-trivial, this means that there are
+Wilf-equivalences between patterns that are not explained by coincidences.
 
@@
 \end{equation*}
-Consider containers of these patterns.
+Consider containers of these patterns in \(\av{\perm{2,3,1}}\).
 For each of these patterns there is precisely one occurrence

### 2016-05-26 writing/presentation/coincidences.tex
Make changes suggested by Michael to thesis. Bjarki's pedantry in presentation.

@@
     \begin{equation*}
-    m_1 = \pattern{}{2,1}{0/0,0/2,1/0,2/1} \text{ and } m_2 = \pattern{}{2,1}{0/0,0/2,1/0,1/1,2/1}
+    m_1 = \pattern{}{2,1}{0/0,0/1,0/2,1/0,2/1} \text{ and } m_2 = \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,2/1}
     \end{equation*}
@@
         {\only<3-4> \fill[pattern=north east lines, pattern color=black!75] (1,1) rectangle +(1,1);}
-        \modpattern{}{2,1}{0/0,0/2,1/0,2/1}
+        \modpattern{}{2,1}{0/0,0/1,0/2,1/0,2/1}
         {\only<2> \filldraw (1.66,1.33) circle (3pt);}

### 2016-05-26 writing/presentation/confut.tex
Changes to presentation from Henning's notes
Make changes suggested by Michael to thesis. Bjarki's pedantry in presentation.

@@
     simple extension, without any dominating pattern. This can capture the special
-    case described by Claesson, Tenner, and Ulfarsson in \cite{ABH}
+    case described by Claesson, Tenner, and Ulfarsson in \cite{ABH}, that was
+    unexplained by the general results proved in that paper.
     \begin{equation*}
@@
 It is also interesting to consider Wilf-equivalence when we have different dominating
-patterns
+patterns.
 We can show that the sets \(\mathcal{T} = \av{\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1},\perm{2,3,1}}\) and

### 2016-05-26 writing/presentation/wilf.tex
Make changes suggested by Michael to thesis. Bjarki's pedantry in presentation.

@@
   \begin{block}{}
-    \blockquote[Herbert Wilf, \emph{generatingfunctionology}~\cite{Wilf}]{
+    \blockquote[Herbert Wilf, \emph{generatingfunctionology}~\cite{Wilf}]{%
     A generating function is a clothesline on which we hang up a sequence of numbers for display.}
@@
 \begin{frame}
-  First we consider the avoiders \(\mathcal{M}_1\) of \(m_1\) in \(\av{\perm{2,3,1}}\).
+  First we consider the avoiders \(\mathcal{M}_1\) of \(m_1 = \pattern{}{1,2}{1/0,1/1,1/2,2/0,2/1} \) in \(\av{\perm{2,3,1}}\).
   \begin{equation*}
@@
 Where \(\mathcal{M}^\prime_1\) is a permutation avoiding \(\perm{2,3,1}, m_1\)
-and \(\textpattern{}{1}{0/0,0/1,1/0}\).\pause Now considering a permutation in \(\mathcal{M}^\prime_1\)
+and \(\textpattern{}{1}{0/0,0/1,1/0}\). \pause Now considering a permutation in \(\mathcal{M}^\prime_1\)
 \begin{equation*}
@@
 \begin{frame}
-  Next we consider the avoiders \(\mathcal{M}_2\) of \(m_2\) in \(\av{\perm{2,3,1}}\).
+  Next we consider the avoiders \(\mathcal{M}_2\) of \(m_2 = \pattern{}{2,1}{0/1,1/0,1/1,1/2,2/1}\) in \(\av{\perm{2,3,1}}\).
   \begin{equation*}
@@
 Where \(\mathcal{M}^\prime_2\) is a permutation avoiding \(\perm{2,3,1}, m_2\)
-and \(\textpattern{}{1}{0/0,0/1,1/1}\).\pause Now considering a permutation in \(\mathcal{M}^\prime_2\)
+and \(\textpattern{}{1}{0/0,0/1,1/1}\). \pause Now considering a permutation in \(\mathcal{M}^\prime_2\)
 \begin{equation*}

### 2016-05-30 writing/ppabstract/PP2016Abstract_Tannock.tex
PP Abstract

% ------------------------------------------------------
% Abstract for talk at Permutation Patterns 2016, Washington D.C. (Based on 2015
% abstract template)
% ------------------------------------------------------
\documentclass[11pt,a4paper]{article}

% 'European' formatting
\frenchspacing \setlength{\parskip}{9pt plus 3pt minus 1pt}
\setlength{\parindent}{0pt}

\setlength{\textwidth}{6.27in}
\setlength{\textheight}{9.69in}
\setlength{\topmargin}{0pt}
\setlength{\headsep}{0pt}
\setlength{\headheight}{0pt}
\setlength{\oddsidemargin}{0pt}
\setlength{\evensidemargin}{0pt}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{tikz}

% ------- extra packages
\usepackage{xinttools}
\usepackage{tabularx}
\usepackage{hyperref}
% -------

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{question}[theorem]{Question}
\newtheorem{example}[theorem]{Example}

\newcommand{\talkTitle}[1]{\hrule\section*{\textsf{\textmd{\LARGE #1}}}}
\newcommand{\talkPresenter}[1]{\textbf{\large #1}}
\newcommand{\talkAffiliation}[1]{(#1)}
\newcommand{\talkJointWork}[1]{This talk is based on joint work with #1.}

\renewcommand{\refname}{\normalsize References}
\newenvironment{talkBibliog}{\begin{thebibliography}{9}\small}{\end{thebibliography}}

% ----------------PERSONAL PREAMBLE---------------------
\DeclareMathOperator{\Av}{Av}
\newcolumntype{Y}{>{\centering\arraybackslash}X}
% ------------------------------------------------------
\begin{document}
\begingroup
% ------------------------------------------------------
% Please do not modify above this line.
% Personal macros can be placed here.
% -----------------------------------------------------------------------------
\newcommand{\ie}{\textit{i.e.}}

\newcommand{\av}[2][]{\Av_{#1}\!\left(#2\right)}
\newcommand{\setsize}[1]{%
    \left|#1\right|}

\newcommand{\addp}[3][\emptyset]{\mathtt{add\_point}\left(#2, #3, #1\right)}
\newcommand{\addpe}[1][\emptyset]{\mathtt{add\_point}\left(\varepsilon, #1\right)}

\newcommand{\adda}[2]{\mathtt{add\_ascent}\left(#1, #2\right)}
\newcommand{\addae}[0]{\mathtt{add\_ascent}\left(\varepsilon\right)}
\newcommand{\addd}[2]{\mathtt{add\_descent}\left(#1, #2\right)}
\newcommand{\addde}[0]{\mathtt{add\_descent}\left(\varepsilon\right)}

\newcommand{\perm}[1]{%
    \xintFor ##1 in {#1}\do {##1}
}
\newcommand{\mperm}[2]{%
\left(
\perm{#1},#2\right)
}

% --------------TIKZ MACROS---------------------------
% For shading boxes
\usetikzlibrary{patterns}
\usetikzlibrary{decorations.markings}

\tikzset{->-/.style={decoration={
  markings,
  mark=at position #1 with {\arrow{>}}},postaction={decorate}}}

\providecommand{\picscale}{0.35}
\providecommand{\textscale}{0.175}

\newcommand{\drawthegrid}[1]{%
\draw (0.01,0.01) grid (#1+0.99,#1+0.99);
}

\newcommand{\drawthepoints}[2]{%
\foreach[count=\x] \y in {#1}
\filldraw (\x,\y) circle (#2 pt);
}

\newcommand{\shadeboxes}[1]{
    \foreach \x/\y in {#1}
    \fill[pattern=north east lines, pattern color=black!75] (\x,\y) rectangle +(1,1);

}

\newcommand{\clpattern}[3][5]{%
    \pgfmathsetmacro\circlesize{#1+4}
    \drawthegrid{\xintNthElt{0}{\xintCSVtoList {#3}}}
    \drawthepoints{#3}{#1}
    \foreach \x in {#2}
    \draw (\x,\xintNthElt{\x}{\xintCSVtoList {#3}}) circle (\circlesize pt);
}

\newcommand{\pattern}[4][]{%
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=\picscale, baseline=(current bounding box.center), #1]
        \shadeboxes{#4}
        \clpattern[5]{#2}{#3}
    \end{tikzpicture}}\;
}

\newcommand{\textpattern}[4][]{%
    \raisebox{0.6ex}{
    \begin{tikzpicture}[scale=\textscale, baseline=(current bounding box.center), #1]
        \shadeboxes{#4}
        \clpattern[8]{#2}{#3}
    \end{tikzpicture}}\;
}

\newcommand{\modpattern}[4][5]{%
    \shadeboxes{#4}
    \clpattern[#1]{#2}{#3}}

    \newcommand{\decompleft}[5]{%
        \begin{tikzpicture}[scale=1.0, baseline=(current bounding box.center)]
            \fill[pattern=north east lines, pattern color=black!75] (0,-1) rectangle +(-0.5,2);
            \foreach \x/\y in {#1}
                \fill[pattern=north east lines, pattern color=black!75] (\x,\y-1) rectangle +(1,1);
            \draw (0,1) -- (0,-1)
                  (-0.5,0) -- (2,0)
                  (1,1) -- (1,-1);
            \filldraw (0,0) circle (3pt);
            \draw (1.5, 0.5) node {\(#5\)};
            \draw (1.5, -0.5) node {\(#3\)};
            \draw (0.5, 0.5) node {\(#4\)};
            \draw (0.5, -0.5) node {\(#2\)};
        \end{tikzpicture}
    }

% ------------------------------------------------------
% Talk title and presenter info goes here.

\talkTitle{Equivalence classes of mesh patterns with a dominating pattern}
\talkPresenter{Murray Tannock}
\talkAffiliation{Reykjavik University}

% ------------------------------------------------------
% The abstract goes here; please restrict to no more than 2 pages.
% Graphics can be included using TikZ (only).

We establish sufficient conditions for coincidence between mesh patterns when
considered inside a set of avoiders of a classical permutation pattern, and
establish general rules that can be used to construct such coincidences. These
rules along with two special cases allow us to completely classify coincidence
between all sets of avoiders of a mesh pattern of length \(2\) and a classical
pattern of length \(3\). We then go on to consider Wilf-equivalence amongst
mesh patterns of length \(1\) and \(2\) whilst also avoiding the classical pattern
\(231\).

When we consider avoidance of a pattern, \(p\), inside a permutation class we gain
additional restrictions on the occurrences of \(p\).
We will restrict the set of permutations that form the basis
for our permutation class to a single classical pattern, \(\pi\), which we shall
refer to as the \emph{dominating pattern}. In particular, we consider the case
where \(p\) is mesh pattern of length \(2\) and \(\pi\) is either of the classical
permutation patterns \(231\) or \(321\).

By use of previous results of Claesson, Tenner and Ulfarsson\cite{ABH}
we can begin to reduce the number of potential coincidences by only looking for
coincidences between patterns that are not coincident outwith the setting of
having a dominating pattern. This gives us an initial upper bound of \(220\)
coincidence classes in each of the four cases we consider.

By examining experimental coincidences of non-classical permutation classes,
considered up to length \(11\) we can find rules that allow us to establish
coincidences, and thus establish the number of coincidence classes as follows.

\begin{table}[htb]
\begin{center}
\begin{tabularx}{\textwidth}{c|Y|Y|Y|Y|}
\cline{2-5}
& \multicolumn{4}{|c|}{Dominating Pattern}\\
\cline{2-5}
& \multicolumn{2}{|c|}{\(\av{\perm{2,3,1}}\)}&\multicolumn{2}{|c|}{\(\av{\perm{3,2,1}}\)}\\
\hline
\multicolumn{1}{|r|}{\emph{Underlying Classical pattern}}& \(\perm{1,2}\)&\(\perm{2,1}\)&\(\perm{1,2}\)&\(\perm{2,1}\)\\
\hline
\hline
\multicolumn{1}{|r|}{No Dominating rule}&\(220\)&\(220\)&\(220\)&\(220\)\\
\hline
\multicolumn{1}{|r|}{First Dominating rule}&\(85\)&\(43\)&\(220\)&\(29\)\\
\hline
\multicolumn{1}{|r|}{Second Dominating rule}&\(59\)&\(39\)&\(220\)&\(29\)\\
\hline
\multicolumn{1}{|r|}{Third Dominating rule}&\(56\)&\(39\)&\(220\)&\(29\)\\
\hline
\multicolumn{1}{|r|}{Experimental class size}&\(56\)&\(39\)&\(213\)&\(29\)\\
\hline
\end{tabularx}
\end{center}
    \caption{Coincidence class number reduction by application of Dominating Rules}
    \label{tab:domclasses}
\end{table}

By considering the classical permutation classes \(\av{231}\) and \(\av{321}\)
along with a mesh pattern with underlying classical pattern \(21\), we gain the
following two rules that completely classify the coincidence in theses cases.

\begin{proposition}[First Dominating Pattern Rule]
    \label{prop:dom1}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\) such that
    \(\setsize{\pi} \le \setsize{\sigma} + 1\), the sets \(\av{\{\pi,m_1\}}\) and
    \(\av{\{\pi,m_2\}}\) are coincident if

    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item \(\pi \) is contained in \( \addp{\sigma}{(a,b)}\)\label{prop:dom1:cont}
    \end{enumerate}
\end{proposition}

\begin{figure}[hbt]
    \begin{center}
    \begin{tikzpicture}[scale=\picscale]
        \modpattern{}{2,1}{1/0,1/1,1/2,2/2}
    \end{tikzpicture}
        \raisebox{2ex}{\(\mapsto\)}
    \begin{tikzpicture}[scale=\picscale]
        \filldraw (2.5,0.5) circle (3pt);
        \modpattern{}{2,1}{1/0,1/1,1/2,2/2,2/0}
    \end{tikzpicture}

        \caption{Visual depiction of first dominating pattern rule.}
        \label{fig:rule1}
    \end{center}
\end{figure}

\begin{proposition}[Second Dominating Pattern Rule]
    \label{prop:dom2}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\) such that
    \(\setsize{\pi} \le \setsize{\sigma} + 2\), the sets \(\av{\{\pi,m_1\}}\) and
    \(\av{\{\pi,m_2\}}\) are coincident if

    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item   \begin{enumerate}
                \item\label{prop:dom2:condc} \(\pi \) is contained in
                \( \adda{\sigma}{(a,b)}\) and some additional conditions on the
                mesh are satisfied that allow selection of different points in
                the occurrence
                \item \(\pi \) is contained in \( \addd{\sigma}{(a,b)}\) and
                some additional conditions on the
                mesh are satisfied that allow selection of different points in
                the occurrence
                        % \begin{enumerate}
                        %     \item \((a+1,b) \in \sigma\) and \((a+1,b-1)\notin R\) and \\
                        %         \((x,b-1)\in R \implies (x,b) \in R \) (where \(x\neq a,a+1\)) and\\
                        %           \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b-1,b\)).
                        %     \item \((a,b+1) \in \sigma\) and \((a-1,b+1)\notin R\) and \\
                        %           \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
                        %           \((a-1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
                        % \end{enumerate}
                    % \item \(\pi \) is contained in \( \addd{\sigma}{(a,b)}\) and
                    %     \begin{enumerate}
                    %         \item \((a+1,b+1) \in \sigma\) and \((a+1,b+1)\notin R\) and \\
                    %               \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a,a+1\)) and\\
                    %               \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
                    %         \item \((a,b) \in \sigma\) and \((a-1,b-1)\notin R\) and \\
                    %               \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
                    %               \((a-1,y)\in R \implies (a,y) \in R\)  (where \(y\neq b-1,b\)).
                    %     \end{enumerate}
                \end{enumerate}
    \end{enumerate}
\end{proposition}

\begin{figure}[ht]
\centering
\begin{tikzpicture}[scale = 0.75]

     % Make the grid
     \draw[step=1cm,gray!50] (2/3,2/3) grid (5+1/3,5+1/3);

     % Make black cross in the middle
     \draw[-,very thick] (3,2/3) -- (3,5+1/3);
     \draw[-,very thick] (2/3,3) -- (5+1/3,3);
     \draw[-,very thick,gray] (4,2) -- (2,2) -- (2,4) -- (4,4) -- (4,2);

     \fill[black] (3,3) circle (0.1);
     \fill[red] (2,3) circle (0.1);

     % Draw the dots on the side
     \foreach \x in {1,2,4,5}{
         \draw[dotted, gray!50] (\x, 0) -- (\x, 2/3);
         \draw[dotted, gray!50] (\x, 5+1/3) -- (\x, 6);
         \draw[dotted, gray!50] (0,\x) -- (2/3,\x);
         \draw[dotted, gray!50] (5+1/3,\x) -- (6,\x);
        }
     % Draw black dots next to the cross
     \draw[dotted, black, very thick] (3, 0) -- (3, 2/3);
     \draw[dotted, black, very thick] (3, 5+1/3) -- (3, 6);
     \draw[dotted, black, very thick] (0, 3) -- (2/3, 3);
     \draw[dotted, black, very thick] (5+1/3, 3) -- (6, 3);

     \draw[->-=.5,thick] (3,3) to[bend right=15](2,4);

     \draw[->, thick] (5,3) -- (5,4);
     \draw[->, thick] (1,3) -- (1,4);
     \draw[->, thick] (3,5) -- (2,5);
     \draw[->, thick] (3,1) -- (2,1);

     \node [black,right] at (6,3) {$b$};
     \node [black,below] at (2,0) {$a$};

\end{tikzpicture}
    \caption{If the conditions of Proposition~\ref{prop:dom2} are satisfied the box
    \((a,b)\) can be shaded.}
\end{figure}

We then define the notion of containment of a mesh pattern, \(p\), within
another mesh pattern, \(q\), we call this notion \emph{subpattern} containment.
This is defined as an extension of regular mesh pattern containment, with the
additional condition that if \((i,j)\in R\text{ then } R_{ij}\) is contained in
the mesh set of \(q\). This notion is then used to establish the following rule
that fully classifies the coincidences between non-classical permutation classes
where we avoid the classical pattern \(\perm{2,3,1}\) and a mesh pattern with
underlying classical pattern \(\perm{1,2}\).
\begin{proposition}[Third Dominating Pattern Rule]
    \label{prop:dom3}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\), the sets
    \(\av{\{\pi,m_1\}}\) and \(\av{\{\pi,m_2\}}\) are coincident if
    \begin{enumerate}
        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
        \item\label{prop:dom3:condocc} \(\addp[D]{\mperm{\sigma}{R_1}}{(a,b)}\)
            where \(D\in\{N,E,S,W\}\)
            is coincident with a mesh pattern containing an occurrence of
            \(\mperm{\sigma}{R_2}\) as a subpattern.
    \end{enumerate}
\end{proposition}

These three rules, along with two special cases, give a complete coincidence
classification of the non-classical permutation classes \(\av{\pi,p}\) where
\(\pi\) is a fixed classical permutation pattern of length \(3\) and \(p\) is a
mesh pattern of length \(2\).

We then go on to consider Wilf-equivalences between non-classical permutation
classes where we avoid the classical permutation pattern \(\perm{2,3,1}\) and
a mesh pattern of length \(2\). I order to establish these Wilf-equivalences
we use the results of the prior section as well as some observations about
symmetries in order to reduce the initial upper bound on Wilf-equivalence classes.
Using the bijection between avoiders of the vincular pattern
\textpattern{}{2,3,1}{1/0,1/1,1/2,1/3} and set partitions presented by
Claesson\cite{A}, and structural decomposition of avoiders to obtain generating
functions we fully classify the Wilf-equivalences. This results in \(13\)
non-trivial (\ie{} not containing only a single coincidence class) Wilf-equivalence
classes. When using generating function methods we also provide enumeration of
the non-classical permutation class.

\begin{example}
  The following patterns are experimentally Wilf-equivalent up to length 10 in
\(\av{\perm{2,3,1}}\)
\begin{equation*}
    m_1 = \pattern{}{1,2}{1/0,1/1,1/2,2/0,2/1} \text{ and }
    m_2 = \pattern{}{2,1}{0/1,1/0,1/1,1/2,2/1}
\end{equation*}
Considering structural decomposition of an avoider of \(m_1\) in \(\av{2,3,1}\)
around the leftmost point, we find that the set of avoiders has form
\begin{equation*}
    \mathcal{M}_1 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\mathcal{M}_1}{}{}{\mathcal{M}^\prime_1}
\end{equation*}
Where \(\mathcal{M}^\prime_1\) is a permutation avoiding \(\perm{2,3,1}, m_1\)
and \(\textpattern{}{1}{0/0,0/1,1/0}\), these permutations have form
\(\mathcal{M}^\prime_1\)
\begin{equation*}
    \mathcal{M}^\prime_1 = \varepsilon \sqcup
    \decompleft{0/1,1/0}{\scriptstyle \mathcal{M}_1\setminus\varepsilon}{}{}{\mathcal{M}^\prime_1}
\end{equation*}
The avoiders of \(m_2\) in \(\av{2,3,1}\) can be split in a similar manner
around the leftmost point. This gives us that both of these sets have generating
function \(M(x)\) satisfying
\begin{align}
    M(x) &= 1 + xM(x)M^\prime(x) \label{eqn:Pgen}\\
    M^\prime(x) &= 1 + x(M(x)-1)M^\prime(x)\label{eqn:PprimeGen}
 \end{align}

 Solving \ref{eqn:PprimeGen} for \(M^\prime(x)\) and substituting into
 \ref{eqn:Pgen} gives us that the generating function for
 \(M(x)\) satisfies
 \begin{equation}
     M(x) = xM^2(x) - x(M(x) - 1) + 1
 \end{equation}
 Solving \(M(x)\) and evaluating the coefficients gives the sequence
 \begin{equation*}
    1, 1, 1, 2, 4, 9, 21, 51, 127, 323, 835,\dotsc \tag{\href{https://oeis.org/A001006}{OEIS: A001006 with offset 1}}
\end{equation*}
This is an offset of the Motzkin numbers.
\end{example}

% ------------------------------------------------------
% Please comment out if not joint work.

% \talkJointWork{Christian Bean and Henning Ulfarsson}

% ------------------------------------------------------
% Please comment out if no references.

\begin{talkBibliog}
 % \bibitem{comtet} L.~Comtet, Advanced combinatorics, D. Reidel Publishing Co., Dordrecht (1974).
 \bibitem{ABH}
  A.~Claesson, B.~E.~Tenner, and H.~Ulfarsson,
  Coincidence among families of mesh patterns,
  \emph{CoRR}, 2014.
  \url{http://arxiv.org/abs/1412.0703}
 \bibitem{A}
 A.~Claesson
 \newblock Generalized pattern avoidance
 \emph{Eur. J. Comb}, 2001.
 \url{http://dx.doi. org/10.1006/eujc.2001.0515}
 % \bibitem{flanoy} P.~Flajolet and M.~Noy, Analytic combinatorics of non-crossing configurations, Discrete Math., 204 (1999), 203--229
\end{talkBibliog}

% ------------------------------------------------------
\endgroup
\end{document}
% ------------------------------------------------------

### 2016-06-25 writing/ppabstract/PP2016Abstract_Tannock.tex
PP presentation

@@
 \newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{definition}[theorem]{Definition}
 \newtheorem{conjecture}[theorem]{Conjecture}
 \newtheorem{question}[theorem]{Question}
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
 \newtheorem{example}[theorem]{Example}
@@
 refer to as the \emph{dominating pattern}. In particular, we consider the case
-where \(p\) is mesh pattern of length \(2\) and \(\pi\) is either of the classical
+where \(p\) a is mesh pattern of length \(2\) and \(\pi\) is either of the classical
 permutation patterns \(231\) or \(321\).
 
-By use of previous results of Claesson, Tenner and Ulfarsson\cite{ABH}
+By use of previous results of Claesson, Tenner and Ulfarsson~\cite{ABH}
 we can begin to reduce the number of potential coincidences by only looking for
-coincidences between patterns that are not coincident outwith the setting of
-having a dominating pattern. This gives us an initial upper bound of \(220\)
-coincidence classes in each of the four cases we consider.
+coincidences between patterns that are not coincident before considering the
+setting of having a dominating pattern. This gives us an initial upper bound of
+\(220\) coincidence classes in each of the four cases we consider, in Table~\ref{tab:domclasses}.
 
@@
 We then go on to consider Wilf-equivalences between non-classical permutation
-classes where we avoid the classical permutation pattern \(\perm{2,3,1}\) and
-a mesh pattern of length \(2\). I order to establish these Wilf-equivalences
+classes where we avoid the permutation pattern \(\perm{2,3,1}\) and
+a mesh pattern of length \(2\). In order to establish these Wilf-equivalences
 we use the results of the prior section as well as some observations about
@@
 \textpattern{}{2,3,1}{1/0,1/1,1/2,1/3} and set partitions presented by
-Claesson\cite{A}, and structural decomposition of avoiders to obtain generating
+Claesson~\cite{A}, and structural decomposition of avoiders to obtain generating
 functions we fully classify the Wilf-equivalences. This results in \(13\)
@@
 classes. When using generating function methods we also provide enumeration of
-the non-classical permutation class.
+the non-classical permutation classes.
 
@@
 \end{equation*}
-Considering structural decomposition of an avoider of \(m_1\) in \(\av{2,3,1}\)
+    Considering structural decomposition of an avoider of \(m_1\) in \(\av{\perm{2,3,1}}\)
 around the leftmost point, we find that the set of avoiders has form
@@
 \end{equation*}
-The avoiders of \(m_2\) in \(\av{2,3,1}\) can be split in a similar manner
+    The avoiders of \(m_2\) in \(\av{\perm{2,3,1}}\) can be split in a similar manner
 around the leftmost point. This gives us that both of these sets have generating
@@
 
-% \talkJointWork{Christian Bean and Henning Ulfarsson}
+\talkJointWork{Henning Ulfarsson}
 

### 2016-06-25 writing/pppresentation/coincidences.tex
PP presentation

--- 2016-05-26 writing/presentation/coincidences.tex
@@
 \begin{frame}{Coincidence}
-  \begin{definition}
-    Two mesh patterns are said to be \emph{coincident} if they are avoided by the
-    same set of permutations at every length.
-  \end{definition}
+  % \begin{definition}
+  %   Two mesh patterns are said to be \emph{coincident} if they are avoided by the
+  %   same set of permutations at every length.
+  % \end{definition}
   \begin{block}{}
-    Distinct classical patterns can never be coincident.
-
+    % Distinct classical patterns can never be coincident.
+    %
     Coincidences of mesh patterns of length \(2\) are classified.
@@
 \cline{2-5}
-& \multicolumn{2}{|c|}{\(\perm{2,3,1}\)}&\multicolumn{2}{|c|}{\(\perm{3,2,1}\)}\\
+    & \multicolumn{2}{|c|}{\(\av{\perm{2,3,1}}\)}&\multicolumn{2}{|c|}{\(\av{\perm{3,2,1}}\)}\\
 \cline{2-5}
@@
 in \(R\), and a dominating classical pattern \(\pi = (\pi,\emptyset)\) if
-\(\pi \preceq \adda{\sigma}{(a,b)}\)\\
-,then in any occurrence of \(m\) in a permutation \(\varrho\), the region
+\(\pi \preceq \adda{\sigma}{(a,b)}\),
+then in any occurrence of \(m\) in a permutation \(\varrho\), the region
 corresponding to the box \((a,b)\) can only contain an decreasing
@@
   \begin{example}
+    \begin{equation*}
+      m_1 = \pattern{}{1,2}{0/1,0/2,2/0,2/1}
+    \end{equation*}
     \begin{description}

### 2016-06-25 writing/pppresentation/wilf.tex
PP presentation

--- 2016-05-26 writing/presentation/wilf.tex
@@
 \begin{frame}{Wilf-equivalence}
-  \begin{definition}
-  Two patterns are said to be \emph{Wilf-equivalent} if the set of avoiders
-  of the patterns is the same size at every length.
-\end{definition}
+%   \begin{definition}
+%   Two patterns are said to be \emph{Wilf-equivalent} if the set of avoiders
+%   of the patterns is the same size at every length.
+% \end{definition}
  All coincident pattern classes are Wilf-equivalent.
 
- The permutations in \(\av{\perm{2,3,1}}\) of length \(n\) are counted by the
- Catalan numbers.
- \begin{equation}
-   C_0 = 1, \text{ and } C_{n+1} =\sum_{i=0}^{n}C_iC_{n-i}
- \end{equation}
- This gives the sequence
- \begin{equation*}
-   1,1,2,5,14,42,132,1430,4862,\dotsc
- \end{equation*}
+ We Wilf-classify classes of the form \(\av{231,m}\) where \(m\)
+ is a mesh pattern.
+
+ % The permutations in \(\av{\perm{2,3,1}}\) of length \(n\) are counted by the
+ % Catalan numbers.
+ % \begin{equation}
+ %   C_0 = 1, \text{ and } C_{n+1} =\sum_{i=0}^{n}C_iC_{n-i}
+ % \end{equation}
+ % This gives the sequence
+ % \begin{equation*}
+ %   1,1,2,5,14,42,132,1430,4862,\dotsc
+ % \end{equation*}
 \end{frame}
@@
 \label{sub:Generating Functions}
-\begin{frame}{Generating {\huge FUN}ctions}
-  \begin{block}{}
-    \blockquote[Herbert Wilf, \emph{generatingfunctionology}~\cite{Wilf}]{%
-    A generating function is a clothesline on which we hang up a sequence of numbers for display.}
-    % \flushright -Herbert Wilf, \emph{generatingfunctionology}
-  \end{block}
-  \begin{definition}
-    The \emph{ordinary generating function (OGF)} of a sequence \((A_n)\) is the
-    formal power series
-    \begin{equation}
-      A(x) = \sum_{n=0}^{\infty} A_nx^n
-    \end{equation}
-    The right hand side is the Taylor expansion of the left hand side at \(0\).
-  \end{definition}
-\end{frame}
-\begin{frame}
-  \begin{example}
-    The generating function for the constant sequence \(\{1,1,1,1,\dotsc\}\) is
-    \begin{equation*}
-      f(x) = \frac{1}{1-x}
-    \end{equation*}
-  \end{example}
-\end{frame}
-\begin{frame}
+% \begin{frame}{Generating {\huge FUN}ctions}
+%   \begin{block}{}
+%     \blockquote[Herbert Wilf, \emph{generatingfunctionology}~\cite{Wilf}]{%
+%     A generating function is a clothesline on which we hang up a sequence of numbers for display.}
+%     % \flushright -Herbert Wilf, \emph{generatingfunctionology}
+%   \end{block}
+%   \begin{definition}
+%     The \emph{ordinary generating function (OGF)} of a sequence \((A_n)\) is the
+%     formal power series
+%     \begin{equation}
+%       A(x) = \sum_{n=0}^{\infty} A_nx^n
+%     \end{equation}
+%     The right hand side is the Taylor expansion of the left hand side at \(0\).
+%   \end{definition}
+% \end{frame}
+% \begin{frame}
+%   \begin{example}
+%     The generating function for the constant sequence \(\{1,1,1,1,\dotsc\}\) is
+%     \begin{equation*}
+%       f(x) = \frac{1}{1-x}
+%     \end{equation*}
+%   \end{example}
+% \end{frame}
+\begin{frame}{Generating Functions by Structure}
   \begin{example}
@@
     \centering
-    \(\mathcal{C} = \varepsilon \mathrel{\sqcup}\) \decompmax{0/1,1/0}{\only<1>A\only<2->{\mathcal{C}}}{}{}{\only<1>B\only<2->{\mathcal{C}}}
+    \(\mathcal{C} = \varepsilon \mathrel{\sqcup}\) \decompleft{0/1,1/0}{{\mathcal{C}}}{}{}{{\mathcal{C}}}
 \end{figure}
-\only<3->{
-Solving for \(C(x)\) and taking the Taylor expansion at \(0\) gives
-\begin{equation*}
-  1+x+2x^2+5x^3+14x^4+42x^5+132x^6+1430x^7+4862x^8+\dotsb
-\end{equation*}
-}
-
+% \only<2->{
+% Solving for \(C(x)\) and taking the Taylor expansion at \(0\) gives
+% \begin{equation*}
+%   1+x+2x^2+5x^3+14x^4+42x^5+132x^6+1430x^7+4862x^8+\dotsb
+% \end{equation*}
+% }
+%
 \end{example}

### 2016-10-24 writing/thesis/abstractis.tex
Rename writing/latex to writing/thesis

Umröðun er endurröðun á \(n\) hlutum. Mengi umraðana sem forðast klassísk
umraðanamynstur geta lýst mörgum áhugaverðum eiginleikum, s.s. staflaraðanlegum
umröðunum, og tengjast mörgum fléttufræðilegum hugtökum. Möskvamynstur eru
útvíkkun á klassískum mynsturum sem leyfa auka skorður á tilvik mynstursins. Tvö
möskvamynstur eru samtilfallandi ef sömu umraðanir forðast hvert um sig. Við
finnum nægjanleg skilyrði fyrir því að tvö möskvamynstur séu samtilfallandi,
þegar umraðanir forðast einnig lengra klassískt mynstur. Þessi skilyrði, ásamt
tveimur sértilfellum, eru notuð til að gera greiningu á hvaða pör af
möskvamynstri af lengd \(2\) og klassísku mynstri af lengd \(3\) eru
samtilfallandi.

Tvö mynstur eru Wilf-jafngild ef jafn margar umraðanir af hverri lengd forðast
hvert mynstur um sig. Við Wilf-flokkum öll pör sem innihalda möskvamynstur af
lengd \(2\) og klassíska mynstrið \(231\). Að lokum finnum við ófáfengileg
Wilf-jafngildi milli pars \(231,m_1\) og \(321,m_2\), ásamt því fjalla um vinnu
sem má byggja á þessari ritgerð.

### 2016-10-24 writing/thesis/coincidence.tex
Rename writing/latex to writing/thesis

--- 2016-05-21 writing/latex/coincidence.tex
@@
 This rule can be understood very in graphical form. In the
-pattern in \FigureRef{fig:rule1} we can gain shading in three boxes since if
+pattern in \FigureRef{fig:rule1} we can gain shading in two boxes since if
 there is a point in any of these boxes we would get an occurrence of the
@@
 
-Note that there are two natural extensions of this rule. We can replace \(\pi\)
+There are two natural extensions of this rule. We can replace \(\pi\)
 with a set of classical patterns, or we can consider \(\pi\) to be a mesh pattern.
@@
 \begin{equation*}
-    m_1 = \pattern{}{2,1}{0/0,0/2,1/0,2/1} \text{ and } m_2 = \pattern{}{2,1}{0/0,0/2,1/0,1/1,2/1}
+    m_1 = \pattern{}{2,1}{0/0,0/1,0/2,1/0,2/1} \text{ and } m_2 = \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,2/1}
 \end{equation*}

### 2016-10-24 writing/thesis/confut.tex
Rename writing/latex to writing/thesis

--- 2016-05-19 writing/latex/confut.tex
@@
 as well as the simple extension of allowing a sequence of \texttt{add\_point}
-operations, to meshpatterns without any dominating pattern in order to try to
+operations, to mesh patterns without any dominating pattern in order to try to
 capture some of the coincidences described in

### 2016-10-24 writing/thesis/wilf.tex
Rename writing/latex to writing/thesis

--- 2016-05-26 writing/latex/wilf.tex
@@
 use \(\mathcal{C}\) to denote \(\av{\perm{2,3,1}}\) and \(C(x)\) will
-be the usual Catalan generating function satisfying \(C(x) = 1 + C(x)^2\).
+be the usual Catalan generating function satisfying \(C(x) = 1 + xC(x)^2\).
 This is easy to see by structural decomposition around the maximum, as
@@
 \end{note}
-\begin{proof}
+
+The idea of the bijection is as follows.
     Let \(\pi\) be a permutation in \(\av[n]{q}\)
@@
     with the elements written in each block in descending order.
-\end{proof}
 

### 2016-10-24 writing/paper/abstract.tex
Add author and abstract to paper.
Add latexmk files to ignore

\begin{abstract}
    Two mesh patterns are coincident if they are avoided by the same set of permutations, and are
    Wilf-equivalent if they have the same number of avoiders at each length. We provide
    sufficient conditions for coincidence among mesh patterns, whilst also avoiding a longer
    classical pattern. Using theses conditions we completely classify coincidences between families
    containing a mesh pattern of length 2 and a classical pattern of length 3. Furthermore, we
    completely Wilf-classify equivalences of mesh patterns of length 2 whilst also avoiding the
    classical pattern 231.
\end{abstract}

### 2016-10-24 writing/paper/abstract.tex
Introduction writing.

@@
     sufficient conditions for coincidence among mesh patterns, whilst also avoiding a longer
-    classical pattern. Using theses conditions we completely classify coincidences between families
+    classical pattern. Using these conditions we completely classify coincidences between families
     containing a mesh pattern of length 2 and a classical pattern of length 3. Furthermore, we
@@
     classical pattern 231.
+
+    \noindent \\
+    \emph{Keywords: } permutation, pattern, mesh pattern, pattern coincidence
 \end{abstract}

### 2016-10-24 writing/paper/introduction.tex
Introduction writing.

The study of permutation patterns began as a result of Knuth's statements on stack sorting in
\emph{The Art of Computer Programming}\cite[p.~243, Ex.~5,6]{Knuth:1997:ACP:260999}. This original
concept---a subsequence of symbols having a particular relative order, now known as classical
patterns---has been expanded to a variety of definitions. Mesh patterns \textcite{babstein2000} 
considered \emph{vincular} patterns---also known as \emph{generalised} or \emph{dashed} 
patterns---where two adjacent entries in the pattern must also be adjacent in the permutation. 
\textcite{MR2652101} look at classes of pattern where both columns and rows can be shaded, these are
called \emph{bivincular} patterns. \emph{Bruhat-restricted} patterns were studied by 
\textcite{MR2264071} in order to establish necessary conditions for a Schubert variety to be
Gorenstein. All of these definitions are subsumed under the definition of \emph{mesh patterns},
introduced by \textcite{journals/combinatorics/BrandenC11} to capture explicit expansions for 
certain permutation statistics.

When considering permutation patterns some of the main questions posed relate to how and when a
pattern is avoided by, or contained in, a arbitrary set of permutations. Two patterns \(\pi\) and
\(\sigma\) are \emph{Wilf-equivalent} if the number of permutations that avoid \(\pi\) of length 
\(n\) is equal to the number of permutations that avoid \(\sigma\) of length \(n\). A stronger 
equivalence condition is that of \emph{coincidence}, where the set of permutations avoiding \(\pi\)
is exactly equal to the set of permutations avoiding \(\sigma\). Avoiding pairs of patterns of the
same length with certain properties has been studied previously, \textcite{MR2178749} considered 
avoiding a pair of vincular patterns of length 3. \textcite{2015arXiv151203226B} study avoiding a 
vincular and a covincular pattern simultaneously in order to achieve some interesting counting 
results. However, very little work has been done on avoiding a mesh pattern and a classical pattern 
simultaneously.

In this work we aim to establish some ground in this field by computing coincidences
and Wilf-classes and calculating some of the enumerations of avoiders of a mesh pattern of length
2 and a classical pattern of length 3. We begin by establishing coincidences between mesh patterns
of length 2 while avoiding a classical pattern of length 3, this is used to establish sufficient
conditions for coincidence. We then establish Wilf-equivalence classes of these coincidence classes
who avoid the classical pattern 231.

### 2016-10-25 writing/paper/sections/introduction.tex
mesh patterns, minor changes to introduction

--- 2016-10-24 writing/paper/introduction.tex
@@
 patterns---has been expanded to a variety of definitions. Mesh patterns \textcite{babstein2000} 
-considered \emph{vincular} patterns---also known as \emph{generalised} or \emph{dashed} 
-patterns---where two adjacent entries in the pattern must also be adjacent in the permutation. 
+considered \emph{vincular} patterns (also known as \emph{generalised} or \emph{dashed} 
+patterns) where two adjacent entries in the pattern must also be adjacent in the permutation. 
 \textcite{MR2652101} look at classes of pattern where both columns and rows can be shaded, these are

### 2016-10-25 writing/paper/sections/meshpatterns.tex
mesh patterns, minor changes to introduction

--- 2016-05-26 writing/latex/introduction.tex
@@
-\chapter{Introduction\label{cha:introduction}}
-\section{What is a Permutation?}
-In \emph{The Art of Computer Programming} \cite[p.~45]{Knuth:1997:ACP:260999} Knuth states,
-\enquote{A \emph{permutation of \(n\) objects} is an arrangement of \(n\) distinct
-objects in a row}.
-When considering permutations we can consider them as
-occurring on the set \(\nrange{n} = \setrange{1}{n}\), therefore a permutation
-is a \emph{bijection} \(\pi:\nrange{n} \mapsto \nrange{n}\). %A
-% permutation \(\pi\) can be written in two-line notation
-% \begin{equation*}
-% \pi = \twoline{1,2,\dotso,n}{\pi(1),\pi(2),\dotso,\pi(n)}
-% \end{equation*}
-% However,
-This
-% most frequent
-notation
-% used in computer science
-is called
-\emph{one-line notation}.
-% in this form we drop the top line of the two line
-% notation and are left with the following.
-In this form we write the entries of the permutation in order, and get
-\begin{equation*}
-\pi = \perm{\pi(1),\pi(2),\dotso,\pi(n)}
-\end{equation*}
-
-\begin{example}
-The 6 permutations on \(\nrange{3}\), in one-line notation, are
-\begin{equation*}
-\perm{1,2,3},\perm{1,3,2},\perm{2,1,3},\perm{2,3,1}, \perm{3,1,2}, \perm{3,2,1}
-\end{equation*}
-\end{example}
-
-We can display a permutation in a \emph{plot} in order to give a graphical
-representation of the permutation. In such a plot we display the points
-\((i,\pi(i))\) in a Cartesian coordinate system. The plot of the permutation
-\(\pi = \perm{2,3,1}\) is shown below
-\begin{equation*}
-    \pattern{}{2,3,1}{}
-\end{equation*}
-It is convenient to call the elements of the permutation \emph{points} when
-referring to these plots.
-
-The set of all permutations of length \(n\) is \(\mathfrak{S}_n\) and
-has size \(n!\). The set of all permutations is
-\(\mathfrak{S}=\bigcup_{i=0}^{\infty}\mathfrak{S}_i\). Note that \(\mathfrak{S}_0\)
-has exactly one element, the empty permutation \(\varepsilon\). As a function
-this is equivalent to the unique bijection \(\emptyset \mapsto \emptyset\), and
-it's one-line representation is the empty string.
-
-\section{Classical Permutation Patterns}
-Classical permutation patterns began to be studied as a result of Knuth's
-statements about stack-sorting in \emph{The Art of Computer Programming}
-\cite[p.~243, Ex.~5,6]{Knuth:1997:ACP:260999}.
+A \emph{permutation} is a bijection from the set \(\nrange{n} = \setrange{1}{n}\) to itself. The
+set of all such bijections on this set if denoted \(\mathfrak{S}_n\) and has size \(n!\). We can
+denote an individual permutation \(\pi\in\mathfrak{S}_n\) in \emph{one-line notation} by writing the
+entries of the permutation in order, therefore \(\pi = \perm{\pi(1),\pi(2),\dotso,\pi(n)}\). The
+set \(\mathfrak{S}_n\) has exactly one element, the empty permutation \(\varepsilon\).
 
@@
 
-A permutation \(\pi\) is said to \emph{contain} the permutation \(\sigma\) of
-length \(k\) as a \emph{pattern} (denoted \(\sigma \preceq \pi\)) if there is some
-subsequence \(i_1i_2\dotsm{}i_k\) such that the sequence
-\(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is order isomorphic to
-\(\sigma(1)\sigma(2)\dotsm\sigma(k)\). If \(\pi\) does not contain
-\(\sigma\), we say that \(\pi\) \emph{avoids} \(\sigma\).
+The definition of order isomorphism allows us to give the meaning of containment for classical
+permutation patterns.
+\begin{definition}
+    A permutation \(\pi\in\mathfrak{S}_n\) contains the pattern \(\sigma\in\mathfrak{S}_k\) (\mbox{denoted} 
+    \(\sigma \preceq \pi\)) if there is some subsequence of indices of \(\pi, i_1i_2\dotsm{}i_k\)
+    such that the sequence \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is order isomorphic to
+    \(\sigma(1)\sigma(2)\dotsm\sigma(k)\). If \(\pi\) does not contain \(\sigma\), we say that 
+    \(\pi\) \emph{avoids} \(\sigma\).
+\end{definition}
+\begin{example}
+\label{ex:contexmpl}
+The permutation \(\pi = \perm{2,4,1,5,3}\) contains the pattern 
+\(\sigma = \perm{2,3,1}\), since the second, fourth and fifth elements (\(453\)) are order 
+isomorphic to \(\perm{231}\), it also contains the occurrence \(241\). The permutation 
+\(\perm{2,4,1,5,3}\) avoids the pattern \(\perm{3,2,1}\).
+\end{example}
 
-For example the permutation \(\pi = \perm{2,4,1,5,3}\) contains the pattern
-\(\sigma = \perm{2,3,1}\), since the second, fourth and fifth elements
-(\(453\)) are order isomorphic to \(\perm{231}\).
-This can be seen graphically below, the subsequence order isomorphic to \(\sigma\)
-is highlighted.
-\begin{equation*}
-    \pattern{2,4,5}{2,4,1,5,3}{}
-\end{equation*}
+We denote the set of permutations of length \(n\) avoiding a pattern \(\sigma\) as \(\Av_n(\sigma)\)
+and \(\av{\sigma}=\bigcup_{i=0}^{\infty}\Av_i(\sigma)\).
 
-We denote the set of permutations of length \(n\) avoiding a pattern \(\sigma\)
-as \(\Av_n(\sigma)\) and \(\av{\sigma} = \bigcup_{i=0}^{\infty}\Av_i(\sigma)\).
+We can display a permutation graphically in a \emph{plot}, in such a plot we display the points
+\(G(\pi) = \{(i,\pi(i))\mid i\in[1,n]\}\) in a Cartesian coordinate system. The plots of the permutations \(\pi = 
+\perm{2,4,1,5,3}\) and \(\sigma = \perm{2,3,1}\) can be seen in \FigureRef{fig:plots}.
+\FigureRef{fig:containment} shows the containment of \(\sigma\) in \(\pi\) as in \ExampleRef{ex:contexmpl}.
 
-Knuth's statements were exercises in showing that the permutations avoiding the
-pattern \(\perm{2,3,1}\) are precisely the permutations that are sortable to
-the identity permutation using a single stack, and that permutations avoiding the
-pattern \(\perm{3,2,1}\) are precisely the permutations that are sortable to
-the identity permutation using a single queue with bypass.
+\begin{figure}[htb]
+    \begin{center}
+    \raisebox{6ex}{\(G(\pi)=\)}
+    \begin{tikzpicture}[scale=\picscale]
+        \modpattern{}{2,4,1,5,3}{}
+    \end{tikzpicture}
+    \raisebox{6ex}{\(\quad{}G(\sigma)=\)}
+    \raisebox{0.35cm}{
+    \begin{tikzpicture}[scale=\picscale]
+        \modpattern{}{2,3,1}{}
+    \end{tikzpicture}}
 
-\section{Mesh Patterns}
-When looking at a classical pattern, for example \(\perm{2,3,1}\), any occurrences can be arbitrarily
-placed. However it may be interesting to consider occurrences where the elements
-corresponding to the \(2\) and \(3\) are adjacent in the permutation. In the past
-these sort of questions have lead to a variety of definitions.  \textcite{babstein2000}
-considered \emph{vincular} patterns (also known as \emph{generalised} or \emph{dashed} patterns),
-those where two adjacent entries in the pattern must be adjacent in the permutation.
-\textcite{MR2652101} look at classes of pattern where both columns and rows
-can be shaded, these are called \emph{bivincular} patterns.
-\emph{Bruhat-restricted} patterns were studied by \textcite{MR2264071} in order
-to establish necessary conditions for a Schubert variety to be Gorenstein.
-Mesh patterns also encompass a subset of \emph{barred} patterns introduced by
-\textcite{MR2716312}, those with only one barred letter.
+        \caption{The plots of the permutations \(\pi\) and \(\sigma\).}
+        \label{fig:plots}
+    \end{center}
+\end{figure}
 
-All of these definitions are subsumed under the definition of \emph{mesh patterns},
-introduced by \textcite{journals/combinatorics/BrandenC11} to
-capture explicit expansions for certain permutation statistics. They are a
-natural extension of classical permutation patterns.
+\begin{figure}[htb]
+    \begin{center}
+    \(\pattern{2,4,5}{2,4,1,5,3}{}\)
+        \caption{The occurrence of 231 in 24153 corresponding to 453.}
+        \label{fig:containment}
+    \end{center}
+\end{figure}
+
+\begin{definition}
 A \emph{mesh pattern} is a pair
 \begin{equation*}
-    p = (\tau,R)\text{ with } \tau \in \mathfrak{S}_k \text{ and } R \subseteq
-    [0,k]\times [0,k].
+    p = (\tau,R)\text{ with } \tau \in \mathfrak{S}_k \text{ and } R \subseteq [0,k]\times [0,k].
 \end{equation*}
-
-The set \(R\) is called the \emph{mesh set} of the mesh pattern \(p\).
-The plot for a mesh pattern looks similar to that of a classical pattern with the
-addition that we shade the unit square with bottom left corner \((i,j)\) for each \((i,j) \in R\):
-\begin{equation*}
-    \pattern{}{ 2, 1, 3 }{ 0/1, 0/2, 1/0, 1/1, 2/1, 2/2 }
-\end{equation*}
-
-The example of an occurrence of \(\perm{2,3,1}\) with adjacent elements \(23\)
-can be represented as a mesh pattern by
-\begin{equation*}
-  \pattern{}{2,3,1}{1/0,1/1,1/2,1/3}
-\end{equation*}
-
+\end{definition}
 We define containment (denoted \(p\preceq \pi\)), and avoidance, of the pattern \(p\) in the permutation
@@
 regions corresponding to the boxes \((0,1)\) and \((0,2)\), which are shaded in
-\(p\). This is shown in the figure below.
+\(p\). This is shown in the plot below.
 
@@
 classical pattern} of \(p\).
-\begin{note}
-    \label{not:classmesh}
-    Classical patterns can be thought of as a subset of mesh patterns: the
-    classical pattern \(\pi\) can be represented by a mesh pattern as
-    \((\pi,\emptyset)\).
-\end{note}
-
-Two patterns are said to be \emph{coincident} if they are avoided by the same set of
-permutations and \emph{Wilf-equivalent} if they are avoided by the same number of permutations
-at every length.
-
-Avoiding pairs of patterns of the same length with certain properties has
-also been studied in the past, \textcite{MR2178749} considered avoiding a
-pair of vincular patterns of length 3. \textcite{2015arXiv151203226B} study avoiding
-a vincular and a covincular pattern simultaneously in order to achieve some
-interesting counting results. However, very little work has been done on avoiding
-a mesh pattern and a classical pattern simultaneously. In this work we aim to
-establish some ground in this field by computing coincidences and Wilf-classes
-and calculating some of the enumerations of avoiders of a mesh pattern of length
-2 and a classical pattern of length 3.
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: "msc-tannock-2016"
-%%% End:

### 2016-10-28 writing/paper/sections/coincidences.tex
Start coincidences section.
Fix errors in mesh patterns from variable name changes

Coincidences among small mesh patterns have previously been considered by 
\textcite{DBLP:journals/corr/ClaessonTU14}, in which the authors use the Simultaneous Shading Lemma,
a closure result and one worked out special case to fully classify coincidences among mesh patterns
of length 2.

Two patterns \(\lambda\) and \(\gamma\) are considered \emph{coincident} if the set of permutations
that avoid \(\lambda\) is the same as the set of permutations that avoid \(\gamma\), \ie \(\av{\lambda} = \av{\gamma}\).
Equivalently we can say that they have the same set of \emph{containers},\ie \(\cont{\lambda} = \cont{\gamma}\).

We will consider the avoidance sets \(\av{\pi,p}\) where \(\pi\) is a classical pattern of length 3
and \(p\) is a mesh pattern of length 2 in order to establish sufficient conditions for two such
sets to be coincident. We will fix \(\pi\) in order to define these coincidences and say that \(\pi\)
is the \emph{dominating pattern}.

In order to describe the rules it is useful to have a notion for inserting points, ascents, and descents
into a mesh pattern.
\begin{definition}
Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\). We define
a mesh pattern \(p^\boks{i,j} = (\tau^\prime,R^\prime)\) of length \(n+1\) as the pattern where a 
point is \emph{inserted} into the box \(\boks{i,j}\) in \(G(p)\). Formally the new underlying classical
pattern is defined by 
\begin{equation*}
\tau^\prime(k) = \begin{cases}
    j+1 & \text{if } k = i+1\\
    \pi(k) & \text{if } \pi(k)\le j\\
    \pi(k)+1 & \text{otherwise.}
\end{cases} 
\end{equation*}
\end{definition}

\(p^\boks{0,2},
p^\boksright{0,2},
p^\boksrightup{0,2},
p^\boksrightdown{0,2},
p^\boksleft{0,2},
p^\boksleftup{0,2},
p^\boksleftdown{0,2},
p^\boksup{0,2},
p^\boksdown{0,2}\)

### 2016-10-28 writing/paper/sections/meshpatterns.tex
Start coincidences section.
Fix errors in mesh patterns from variable name changes

@@
 We define containment (denoted \(p\preceq \pi\)), and avoidance, of the pattern \(p\) in the permutation
-\(\tau\) on mesh patterns analogously to classical containment, and avoidance,
-of \(\pi\) in \(\tau\) with the additional restrictions on the relative
-position of the occurrence of \(\pi\) in \(\tau\). These restrictions say that
-no elements of \(\tau\) are allowed in the regions of the plot corresponding
-to shaded boxes in the mesh.
+\(\pi\) on mesh patterns analogously to classical containment, and avoidance,
+of \(\tau\) in \(\pi\) with the additional restrictions on the relative
+position of the occurrence of \(\tau\) in \(\pi\). These restrictions say that
+no elements of \(\pi\) are allowed in the regions of the plot corresponding
+to shaded boxes in the mesh. These boxes are denoted by \(\boks{i,j}\), where the point \((i,j)\) is
+the lower left corner of the box.
 
 Formally defined by \textcite{journals/combinatorics/BrandenC11},
-an \emph{occurrence} of \(p\) in \(\tau\) is a subset \(\omega\) of the plot of
-\(\tau, G(\tau) = \{(i,\tau(i)\mid i\in[1,n]\}\) such that there are order-preserving
+an \emph{occurrence} of \(p\) in \(\pi\) is a subset \(\omega\) of the plot of
+\(\pi, G(\pi) = \{(i,\pi(i)\mid i\in[1,n]\}\) such that there are order-preserving
 injections \(\alpha,\beta:[1,k]\mapsto[1,n]\) satisfying the following two conditions.
@@
   \begin{enumerate}[i.]
-    \item \(\omega = \{(\alpha(i),\beta(j)):(i,j)\in G(\pi)\}\)
+    \item \(\omega = \{(\alpha(i),\beta(j)):(i,j)\in G(\tau)\}\)
   \end{enumerate}\saveenum
@@
   \begin{enumerate}[i.]\resetenum
-  \item if \((i,j) \in R \text{ then } R_{ij} \cap G(\tau) = \emptyset\)
+  \item if \(\boks{i,j} \in R \text{ then } R_{ij} \cap G(\pi) = \emptyset\)
 \end{enumerate}
-We call \(R_{ij}\) the region corresponding to \((i,j)\).
-We define containment of a mesh pattern \(p\) in another mesh \(\kappa\) as above,
-with the additional condition that if \((i,j)\in R\text{ then } R_{ij}\) is
-contained in the mesh set of \(\kappa\), in this case we call \(p\) a \emph{subpattern}
-of \(\kappa\).
+We call \(R_{ij}\) the region corresponding to \(\boks{i,j}\).
+We define containment of a mesh pattern \(p\) in another mesh pattern \(q\) as above,
+with the additional condition that if \(\boks{i,j}\in R\text{ then } R_{ij}\) is
+contained in the mesh set of \(q\), in this case we call \(p\) a \emph{subpattern}
+of \(q\).
+\begin{definition}
+A mesh pattern \(q=(\kappa,T)\) \emph{contains} a mesh pattern \(p=(\tau,R)\) as a \emph{subpattern}
+if \(\kappa\) contains \(\tau\) and \(\left(\bigcup_{\boks{i,j}\in R} R_{ij}\right)\subseteq T\).
+\end{definition}
 
@@
 not the mesh pattern since the points with values \(4\) and \(2\) are in the
-regions corresponding to the boxes \((0,1)\) and \((0,2)\), which are shaded in
+regions corresponding to the boxes \(\boks{0,1}\) and \(\boks{0,2}\), which are shaded in
 \(p\). This is shown in the plot below.

### 2016-11-08 writing/paper/sections/coincidences.tex
Add preliminaries for equivalences

@@
 \begin{definition}
+\label{def:ap}
 Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\). We define
@@
     j+1 & \text{if } k = i+1\\
-    \pi(k) & \text{if } \pi(k)\le j\\
-    \pi(k)+1 & \text{otherwise.}
+    \tau(k) & \text{if } \tau(k)\le j \text{ and }k\le i\\
+    \tau(k)+1 & \text{if } \tau(k)> j \text{ and }k\le i\\
+    \tau(k-1) & \text{if } \tau(k)\le j \text{ and }k> i+1\\
+    \tau(k-1)+1 & \text{if } \tau(k)> j \text{ and }k> i+1\\
 \end{cases} 
 \end{equation*}
+While the mesh becomes
+\begin{equation*}
+\begin{aligned}
+R^\prime=&\{\boks{k,\ell}\mid k\le i, \ell\le j, \boks{k,\ell}\in R\} \cup\\
+&\{\boks{k,\ell}\mid k\le i, \ell > j, \boks{k,\ell-1}\in R\}\cup\\
+&\{\boks{k,\ell}\mid k > i, \ell \le j, \boks{k-1,\ell}\in R\}\cup\\
+&\{\boks{k,\ell}\mid k > i, \ell > j, \boks{k-1,\ell-1}\in R\}\\
+\end{aligned}
+\end{equation*}
 \end{definition}
+In addition, we give the following definitions:
+\begin{definition}
+Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\) and 
+\(p^\boks{i,j} = (\tau^\prime,R^\prime)\) to be as defined in \DefinitionRef{def:ap}. We define
+the following five modifications of the mesh patterns of the same length as \(p^\boks{i,j}\). 
+\begin{align*}
+p^\boksup{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j+1},\boks{i+1,j+1}\})\\
+p^\boksright{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i+1,j},\boks{i+1,j+1}\})\\
+p^\boksdown{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j},\boks{i+1,j}\})\\
+p^\boksleft{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j},\boks{i,j+1}\})\\
+\end{align*}
+\end{definition}
+Informally, these are considering the topmost, rightmost, leftmost, or bottommost point in \(\boks{i,j}\).
+We allow composition of these modifications, and collect the resulting mesh patterns in a set
+\begin{equation*}
+p^\boksall{i,j}=\left\{
+p^\boks{i,j},
+p^\boksright{i,j},
+p^\boksrightup{i,j},
+p^\boksrightdown{i,j},
+p^\boksleft{i,j},
+p^\boksleftup{i,j},
+p^\boksleftdown{i,j},
+p^\boksup{i,j},
+p^\boksdown{i,j}\right\}
+\end{equation*}
+See \FigureRef{fig:addp} for an example of adding a point into a mesh pattern.
 
-\(p^\boks{0,2},
-p^\boksright{0,2},
-p^\boksrightup{0,2},
-p^\boksrightdown{0,2},
-p^\boksleft{0,2},
-p^\boksleftup{0,2},
-p^\boksleftdown{0,2},
-p^\boksup{0,2},
-p^\boksdown{0,2}\)
+\begin{figure}
+\begin{equation*}
+%\adda{\pattern{}{2,3,1}{0/0,1/0,1/3,
+%                        3/0,3/1,3/3}}{(1,1)}
+p = \pattern{}{1,2}{0/1,2/2} \quad
+p^\boksrightup{2,1}
+= \pattern{}{1,3,2}{0/1,0/2,2/2,2/3,3/1,3/2,3/3}
+\end{equation*}
+\caption{The result of inserting a point into \(p=(12,\{(0,1),(2,2)\})\)}
+\label{fig:addp}
+\end{figure}
+
+\begin{definition}
+Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\). We define
+a mesh pattern \(p^{\boks{i,j}_a} = (\tau^\prime,R^\prime)\) (\(p^{\boks{i,j}_d}\)) of length \(n+2\)
+as the pattern where an ascent (descent) is \emph{inserted} into the box \(\boks{i,j}\) in \(G(p)\). Formally the new underlying classical
+pattern is defined by
+\begin{equation*}
+\tau^\prime(k) = \begin{cases}
+    j+t & \text{if } k = i+t,t\in\{1,2\}\\
+    \tau(k) & \text{if } \tau(k)\le j \text{ and }k\le i\\
+    \tau(k)+2 & \text{if } \tau(k)> j \text{ and }k\le i\\
+    \tau(k-2) & \text{if } \tau(k)\le j \text{ and }k> i+2\\
+    \tau(k-2)+2 & \text{if } \tau(k)> j \text{ and }k> i+2\\
+\end{cases} 
+\end{equation*}
+The ordering of the top branch determines whether an ascent(or descent) is added.
+The mesh becomes
+\begin{equation*}
+\begin{aligned}
+R^\prime=&\{\boks{k,\ell}\mid k\le i, \ell\le j, \boks{k,\ell}\in R\} \cup\\
+&\{\boks{k,\ell}\mid k\le i, \ell > j, \boks{k,\ell-2}\in R\}\cup\\
+&\{\boks{k,\ell}\mid k > i, \ell \le j, \boks{k-2,\ell}\in R\}\cup\\
+&\{\boks{k,\ell}\mid k > i, \ell > j, \boks{k-2,\ell-2}\in R\}\cup\\
+&\{\boks{i,j+1},\boks{i+1,j},\boks{i+1,j+1},\boks{i+1,j+2},\boks{i+2,j+1}\}
+\end{aligned}
+\end{equation*}
+\end{definition}
+An example of adding an ascent to a mesh pattern can be seen in \FigureRef{fig:adda}.
+\begin{figure}
+\begin{equation*}
+p = \pattern{}{2,3,1}{0/0,1/0,1/3,
+                      3/0,3/1,3/3} \quad
+p^{\boks{1,1}_a}= \pattern{}{4,2,3,5,1}{0/0,
+                        1/0,1/2,1/5,
+                        2/0,2/1,2/2,2/3,2/5,
+                        3/0,3/2,3/5,
+                        5/0,5/1,5/2,5/3,5/5}
+\end{equation*}
+\caption{The result of inserting an ascent into \(p=(231,\{(0,0),(1,0),(1,3),(3,0),(3,1),(3,3)\})\)}
+\label{fig:adda}
+\end{figure}
+
+\subsection{Coincidence classes of Av(\{321, (21, \textit{R})\})}

### 2016-11-08 writing/paper/sections/coincidences.tex
First dominating rule in paper

@@
 
+We now attempt to fully classify coincidences in families characterised by avoidance
+of a classical pattern of length \(3\) and a mesh pattern of length \(2\), that is
+finding and explaining all coincidences where \(\av{\{p,m\}} = \av{\{p,m^\prime\}}\).
+
+It can be easily seen that in order to classify coincidences one need only
+consider coincidences within the family of mesh patterns with the same underlying
+classical pattern, this is due to the fact that \(\perm{2,1}\in\av{\mperm{1,2}{R}}\)
+and \(\perm{1,2}\in\av{\mperm{2,1}{R}}\) for all mesh-sets \(R\).
+
+We know that there are a total of \(512\) mesh-sets for each underlying classical
+pattern. By use of the previous results of \textcite{DBLP:journals/corr/ClaessonTU14}\footnote{
+The authors use the Simultaneous Shading Lemma, a closure result and one worked out special case.
+} the number of coincidence classes can be reduced to \(220\).
+
 \subsection{Coincidence classes of Av(\{321, (21, \textit{R})\})}
+Through experimentation, considering avoidance of permutations of up to length \(11\), we discover
+that there are at least \(29\) coincidence classes of mesh patterns with underlying classical
+pattern \(21\).
+
+\begin{proposition}[First Dominating Pattern Rule]
+    \label{prop:dom1}
+    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
+    and a dominating classical pattern \(\pi = (\pi,\emptyset)\) such that
+    \(\setsize{\pi} \le \setsize{\sigma} + 1\), the sets \(\av{\{\pi,m_1\}}\) and
+    \(\av{\{\pi,m_2\}}\) are coincident if
+
+    \begin{enumerate}
+        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
+        \item \(\pi \preceq \sigma^\boks{a,b}\)\label{prop:dom1:cont}
+    \end{enumerate}
+\end{proposition}
+In order to prove this proposition we must first make the following note.
+
+\begin{note}
+    \label{not:downcmesh}
+    Let \(R^\prime \subseteq R\). Then any occurrence of \((\tau, R)\) in a permutation
+    is an occurrence of \((\tau,R^\prime)\).
+\end{note}
+
+\begin{proof}[Proof of \PropositionRef{prop:dom1}]
+    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).
+
+    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
+    Since \(R_1\) is a subset of \(R_2\), \NoteRef{not:downcmesh} states that
+    \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\)
+
+    Now we consider a permutation \(\omega^\prime\in \av{\pi}\),
+    containing an occurrence of \(m_1\). Consider placing a point in the
+    region corresponding to the box \((a,b)\), regardless of where in this
+    region we place the point by condition~\ref{prop:dom1:cont} of the
+    Proposition we create an occurrence of \(\pi\), therefore there can be no
+    points in this region, which could have been represented in the mesh set
+    \(R_1\) by adding the box \((a,b)\). Hence every occurrence of
+    \(m_1\) is in fact an occurrence of \(m_2\), and we have that
+    \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).
+
+    Taking both directions of the containment we can therefore draw the
+    conclusion that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).
+\end{proof}

### 2017-02-20 writing/paper/sections/coincidences.tex
Finish writing paper v1

@@
-Coincidences among small mesh patterns have previously been considered by 
+Coincidences among small mesh patterns have been considered by
 \textcite{DBLP:journals/corr/ClaessonTU14}, in which the authors use the Simultaneous Shading Lemma,
@@
 Two patterns \(\lambda\) and \(\gamma\) are considered \emph{coincident} if the set of permutations
-that avoid \(\lambda\) is the same as the set of permutations that avoid \(\gamma\), \ie \(\av{\lambda} = \av{\gamma}\).
-Equivalently we can say that they have the same set of \emph{containers},\ie \(\cont{\lambda} = \cont{\gamma}\).
+that avoid \(\lambda\) is the same as the set of permutations that avoid \(\gamma\),
+\ie \(\av{\lambda} = \av{\gamma}\). Equivalently we can say that they have the same set of
+\emph{containers},\ie \(\cont{\lambda} = \cont{\gamma}\).
 
@@
 and \(p\) is a mesh pattern of length 2 in order to establish sufficient conditions for two such
-sets to be coincident. We will fix \(\pi\) in order to define these coincidences and say that \(\pi\)
-is the \emph{dominating pattern}.
+sets to be coincident. We will fix \(\pi\) in order to define these coincidences and say that
+\(\pi\)is the \emph{dominating pattern}.
 
-In order to describe the rules it is useful to have a notion for inserting points, ascents, and descents
-into a mesh pattern.
+In order to describe the rules it is useful to have a notion for inserting points, ascents, and
+descents into a mesh pattern.
 \begin{definition}
@@
 a mesh pattern \(p^\boks{i,j} = (\tau^\prime,R^\prime)\) of length \(n+1\) as the pattern where a
-point is \emph{inserted} into the box \(\boks{i,j}\) in \(G(p)\). Formally the new underlying classical
-pattern is defined by 
+point is \emph{inserted} into the box \(\boks{i,j}\) in \(G(p)\). Formally the new underlying
+classical pattern is defined by
 \begin{equation*}
@@
 \end{definition}
-Informally, these are considering the topmost, rightmost, leftmost, or bottommost point in \(\boks{i,j}\).
-We allow composition of these modifications, and collect the resulting mesh patterns in a set
+Informally, these are considering the topmost, rightmost, leftmost, or bottommost point in
+\(\boks{i,j}\). We allow composition of these modifications, and collect the resulting mesh patterns
+in a set
 \begin{equation*}
@@
 p^\boksright{i,j},
-p^\boksrightup{i,j},
-p^\boksrightdown{i,j},
+% p^\boksrightup{i,j},
+% p^\boksrightdown{i,j},
 p^\boksleft{i,j},
-p^\boksleftup{i,j},
-p^\boksleftdown{i,j},
+% p^\boksleftup{i,j},
+% p^\boksleftdown{i,j},
 p^\boksup{i,j},
@@
 a mesh pattern \(p^{\boks{i,j}_a} = (\tau^\prime,R^\prime)\) (\(p^{\boks{i,j}_d}\)) of length \(n+2\)
-as the pattern where an ascent (descent) is \emph{inserted} into the box \(\boks{i,j}\) in \(G(p)\). Formally the new underlying classical
-pattern is defined by
+as the pattern where an ascent (descent) is \emph{inserted} into the box \(\boks{i,j}\) in \(G(p)\).
+Formally the new underlying classical pattern is defined by
 \begin{equation*}

### 2017-02-20 writing/paper/sections/conclfut.tex
Finish writing paper v1

--- 2016-05-19 writing/latex/confut.tex
@@
-\chapter{Conclusions and Future work}
-\nextvar
-% From \ChapterRef{chap:coincs} it can be seen that automatically classifying
-% coincidences of mesh patterns is a difficult task, establishing rules for longer
-% dominating patterns requires many more cases to be taken.
 If we consider a similar system for dominating patterns of length \(4\) and
@@
 established do not completely cover the coincidences with a dominating pattern
-of length \(3\) (see \TableRef{tab:domclasses}) shows that this is a difficult
-task.
+of length \(3\) shows that this is a difficult task.
 
-It is interesting to consider the application of the Third Dominating rule,
-as well as the simple extension of allowing a sequence of \texttt{add\_point}
-operations, to meshpatterns without any dominating pattern in order to try to
-capture some of the coincidences described in
-\textcite{DBLP:journals/combinatorics/HilmarssonJSVU15}
- and \textcite{DBLP:journals/corr/ClaessonTU14}.
+It is interesting to consider the application of the Third Dominating rule, as well as the simple
+extension of allowing a sequence of point insertions, to mesh patterns without any dominating
+pattern in order to try to capture some of the coincidences described in
+\textcite{DBLP:journals/combinatorics/HilmarssonJSVU15} and
+\textcite{DBLP:journals/corr/ClaessonTU14}.
  \begin{example}
@@
  \end{example}
-It is not possible to apply the first and second Dominating Rules to the pattern
-itself, since when applying the rules we consider containers of the pattern inside
-the avoiders of the dominating pattern. For example if we were to attempt to apply
-the first rule to the pattern \(\perm{1,2}\) then we would have to consider
-containers of \(\perm{1,2}\) inside \(\av{\perm{1,2}}\), and obviously this can
-never occur.
 
-It is also possible to take sets of mesh patterns instead of a single mesh pattern
-when considering dominating rules, and expressing coincidence between these sets.
-Doing this may give nice results. Coincidences between sets with
-multiple dominating patterns can also be considered, as this provides even more
-power to the rules discussed, these methods may be useful for classifying coincidence
-in large sets of pattern. It is also possible to apply these rules to sets where the
-dominating pattern is a mesh pattern.
-
-It would be interesting to consider a systematic explanation of Wilf-equivalences
-amongst classes where \(\perm{3,2,1}\) is the dominating pattern, possible using the
-construction presented in \cite[Sec.~12]{2015arXiv151203226B}, in order to directly
-reach enumeration and hopefully establish some of the non-trivial Wilf-equivalences
-between classes with different dominating patterns. For example, it is possible to show that
-the sets \(\scriptvar = \av{\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1},\perm{2,3,1}}\)\nextvar and
-\(\scriptvar = \av{\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1},\perm{3,2,1}}\), are Wilf-equivalent.
-This can be seen by considering an occurrence of the mesh pattern,
-\(p=\textpattern{}{1,2}{1/2,2/0,0/0,1/0,1/1,2/1}\) in a permutation
-\begin{equation*}
-    \raisebox{0.6ex}{
-    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
-        \draw (0.5,1.5) node {\(b_1\)};
-        \draw (0.5,2.5) node {\( b_2\)};
-        \draw (2.5,2.5) node {\( b_3\)};
-        \modpattern[2]{2}{1,2}{0/0,1/0,1/1,1/2,2/0,2/1}
-    \end{tikzpicture}
-    }
-\end{equation*}
-
-If we are inside \(\av{\perm{2,3,1}}\) then any points in the region corresponding
-to the box \(b_1\) must be to the right of the points in the region corresponding to
-box \(b_2\), and must also form a decreasing subsequence by \LemmaRef{lem:incdec}.
-Furthermore, the points in the regions corresponding to \(b_2\) and
-\(b_3\) must form an avoider of \(\perm{2,3,1}\) with the indicated point.
-Therefore the containers of \(p\) in \(\av{\perm{2,3,1}}\) have generating function
-\(\lastvar(x) = x(C(x)-1)/(1-x)\).
-Now, if we are in \(\av{\perm{3,2,1}}\) then any points in the region corresponding
-to the box \(b_1\) must be to the left of the points in the region corresponding to
-box \(b_2\), and must also form a increasing subsequence by \LemmaRef{lem:incdec}.
-Furthermore, the points in the regions corresponding to \(b_2\) and
-\(b_3\) must form an avoider of \(\perm{3,2,1}\) with the indicated point.
-Hence, the containers of \(p\) in \(\av{\perm{3,2,1}}\) have generating function
-\(\var(x) = x(C(x)-1)/(1-x)\) and both of these classes have the same
-enumeration.
+It would be interesting to consider a systematic explanation of Wilf-equivalences among classes
+where \(\perm{2,3,1}\) is the dominating pattern, possibly using the  construction presented in
+\cite[Sec.~12]{2015arXiv151203226B}, in order to directly reach enumeration and hopefully establish
+some of the non-trivial Wilf-equivalences between classes with different dominating patterns.

### 2017-02-20 writing/paper/sections/introduction.tex
Finish writing paper v1

@@
-The study of permutation patterns began as a result of Knuth's statements on stack sorting in
-\emph{The Art of Computer Programming}\cite[p.~243, Ex.~5,6]{Knuth:1997:ACP:260999}. This original
-concept---a subsequence of symbols having a particular relative order, now known as classical
-patterns---has been expanded to a variety of definitions. Mesh patterns \textcite{babstein2000} 
-considered \emph{vincular} patterns (also known as \emph{generalised} or \emph{dashed} 
-patterns) where two adjacent entries in the pattern must also be adjacent in the permutation. 
-\textcite{MR2652101} look at classes of pattern where both columns and rows can be shaded, these are
-called \emph{bivincular} patterns. \emph{Bruhat-restricted} patterns were studied by 
-\textcite{MR2264071} in order to establish necessary conditions for a Schubert variety to be
-Gorenstein. All of these definitions are subsumed under the definition of \emph{mesh patterns},
-introduced by \textcite{journals/combinatorics/BrandenC11} to capture explicit expansions for 
-certain permutation statistics.
+The study of permutation patterns began as a result of Knuth's statements on
+stack sorting in \emph{The Art of Computer Programming}\cite[p.~243,
+Ex.~5,6]{Knuth:1997:ACP:260999}. This original concept---a subsequence of
+symbols having a particular relative order, now known as classical
+patterns---has been expanded to a variety of definitions.
+\textcite{babstein2000} considered \emph{vincular} patterns (also known as
+\emph{generalised} or \emph{dashed} patterns) where two adjacent entries in the
+pattern must also be adjacent in the permutation. \textcite{MR2652101} look at
+classes of pattern where both columns and rows can be shaded, these are called
+\emph{bivincular} patterns. \emph{Bruhat-restricted} patterns were studied by
+\textcite{MR2264071} to establish necessary conditions for a Schubert variety to
+be Gorenstein. These definitions are subsumed under the definition of \emph{mesh
+patterns}, introduced by \textcite{journals/combinatorics/BrandenC11} to capture
+explicit expansions for certain permutation statistics.
 
-When considering permutation patterns some of the main questions posed relate to how and when a
-pattern is avoided by, or contained in, a arbitrary set of permutations. Two patterns \(\pi\) and
-\(\sigma\) are \emph{Wilf-equivalent} if the number of permutations that avoid \(\pi\) of length 
-\(n\) is equal to the number of permutations that avoid \(\sigma\) of length \(n\). A stronger 
-equivalence condition is that of \emph{coincidence}, where the set of permutations avoiding \(\pi\)
-is exactly equal to the set of permutations avoiding \(\sigma\). Avoiding pairs of patterns of the
-same length with certain properties has been studied previously, \textcite{MR2178749} considered 
-avoiding a pair of vincular patterns of length 3. \textcite{2015arXiv151203226B} study avoiding a 
-vincular and a covincular pattern simultaneously in order to achieve some interesting counting 
-results. However, very little work has been done on avoiding a mesh pattern and a classical pattern 
-simultaneously.
+When considering permutation patterns some of the main questions posed relate to
+how and when a pattern is avoided by, or contained in, a arbitrary set of
+permutations. Two patterns \(\pi\) and \(\sigma\) are \emph{Wilf-equivalent} if
+the number of permutations that avoid \(\pi\) of length \(n\) is equal to the
+number of permutations that avoid \(\sigma\) of length \(n\). A stronger
+equivalence condition is that of \emph{coincidence}, where the set of
+permutations avoiding \(\pi\) is exactly equal to the set of permutations
+avoiding \(\sigma\). Avoiding pairs of patterns of the same length with certain
+properties has been studied, \textcite{MR2178749} considered avoiding
+a pair of vincular patterns of length 3. \textcite{2015arXiv151203226B} study
+avoiding a vincular and a covincular pattern simultaneously in order to achieve
+some interesting counting results. However, little work has been done on
+avoiding a mesh pattern and a classical pattern simultaneously.
 
-In this work we aim to establish some ground in this field by computing coincidences
-and Wilf-classes and calculating some of the enumerations of avoiders of a mesh pattern of length
-2 and a classical pattern of length 3. We begin by establishing coincidences between mesh patterns
-of length 2 while avoiding a classical pattern of length 3, this is used to establish sufficient
-conditions for coincidence. We then establish Wilf-equivalence classes of these coincidence classes
-who avoid the classical pattern 231.
+In this work we aim to establish some ground in this field by computing
+coincidences and Wilf-classes and calculating some of the enumerations of
+avoiders of a mesh pattern of length 2 and a classical pattern of length 3. We
+begin by establishing coincidences between mesh patterns of length 2 while
+avoiding a classical pattern of length 3, this is used to establish sufficient
+conditions for coincidence. We then establish Wilf-equivalence classes of these
+coincidence classes who avoid the classical pattern 231.

### 2017-02-20 writing/paper/sections/wilf.tex
Finish writing paper v1

--- 2016-10-24 writing/thesis/wilf.tex
@@
-\chapter[Wilf-Classification]{Wilf-equivalences under dominating  patterns}
+Wilf-equivalence is an important aspect to study in the field of permutation
+patterns. If two patterns \(\pi\) and \(\sigma\) are said to be Wilf-equivalent
+if their avoidance sets have the same size at each length. More formally:
 
-One question often asked in the field of permutation patterns is that
-of Wilf-equivalence. Two patterns \(\pi\) and \(\sigma\) are said to be
-Wilf-equivalent if their avoidance sets have the same size at each
-length. More formally:
-
-\begin{definition}[Wilf-equivalence]
+\begin{definition}{(Wilf Equivalence.)}
   Two patterns \(\pi\) and \(\sigma\) are said to be \emph{Wilf-equivalent}
-    if for all \(k_{} \ge 0\), \(\size{\av[k]{\pi}} = \size{\av[k]{\sigma}}\).
-    Two sets of permutation patterns \(R\) and \(S\) are are
-    \emph{Wilf-equivalent} if for all \(k_{} \ge 0\),
-    \(\size{\av[k]{R}} = \size{\av[k]{S}}\).
+  if for all \(k\ge0, \size{\av[k]{\pi}} = \size{\av[k]{\sigma}}\). Two sets of
+  permutation patterns \(R\) and \(S\) are \emph{Wilf-equivalent} if for all
+  \(k\ge0, \size{\av[k]{R}} = \size{\av[k]{S}}\).
 \end{definition}
 
-Wilf-equivalence is of interest since if two permutation classes are enumerated
-in the same way then there should exist a bijection between them, and therefore
-any other combinatorial object that they represent.
+Wilf-equivalence is of interest due to the fact that if two permutation classes
+are enumerated in the same way then there should exist a bijection between
+them, and therefore any combinatorial object they represent.
 
-Coincident pattern classes are also Wilf-equivalent. This is due to the fact that
-if \(\av[k]{S} = \av[k]{R}\) then obviously \(\size{\av[k]{R}} = \size{\av[k]{S}}\).
-Coincidence is therefore a stronger equivalence condition than Wilf-equivalence.
+Coincident patterns are trivially Wilf-equivalent, if \(\av[k]{R} = \av[k]{S}\)
+then trivially \(\size{\av[k]{R}} = \size{\av[k]{S}}\). Coincidence is
+therefore a stronger equivalence condition than Wilf-equivalence.
 
-There are a number of symmetries we can use when examining Wilf-equivalences
-to reduce the amount of work. It can be easily seen that the reverse, complement
-and inverse operations (see \FigureRef{fig:symm}) preserve enumeration, and therefore these classes are trivially
+When examining Wilf-equivalences we can use a number of symmetries to reduce
+the amount of work required. It can be seen that the reverse, complement
+and inverse operations (see \FigureRef{fig:symm}) preserve enumeration, and
+therefore these classes are trivially
 Wilf-equivalent.
@@
 
-The group of symmetries on permutations is isomorphic to the dihedral group
-of order \(8\), the group of symmetries of a square. \emph{Reverse-inverse} and
-\emph{reverse} correspond to generators of the dihedral group.
-
 Since we are always considering Wilf-equivalences in the set \(\av{S}\) we
 must only use these symmetries when they preserve the dominating pattern, if we
-were to allow other symmetries, then the equivalences calculated in the previous
-section do not necessarily hold.
+were to allow other symmetries, then the equivalences calculated in the
+previous section do not necessarily hold.
 
@@
 The elements to the left of the maximum, \(A\), have the structure of a \(231\)
-avoiding permutation, and the elements to the right of the maximum, \(B\), have the
-structure of a \(\perm{2,3,1}\) avoiding permutation. Furthermore, all the
+avoiding permutation, and the elements to the right of the maximum, \(B\), have
+the structure of a \(\perm{2,3,1}\) avoiding permutation. Furthermore, all the
 elements in \(A\) lie below all of the elements in \(B\). We call \(A\) the
@@
 
-We can also decompose a permutation avoiding \(\perm{2,3,1}\) around the leftmost
-point, giving a similar figure.
+We can also decompose a permutation avoiding \(\perm{2,3,1}\) around the
+leftmostpoint, giving a similar figure.
 
-\section{Wilf-classes with patterns of length 1.}
+\subsection{Wilf-classes with mesh patterns of length 1.}
 When considering the mesh patterns of length \(2\) it will be useful to know
 the Wilf-equivalence classes of the mesh patterns of length \(1\) inside
-\(\av{\perm{2,3,1}}\), this means that we are considering the set \(\av{231,p}\)
+\(\av{\perm{2,3,1}}\), this means that we are considering the set
+\(\av{231,p}\)
 where \(p\) is a mesh-pattern of length 1.
@@
 Alternatively we can enumerate
-avoiders of \(p\) by decomposing a non-empty avoider of \(p\) around the maximum element
+avoiders of \(p\) by decomposing a non-empty avoider of \(p\) around the
+maximum element
 in order to give the following structural decomposition.
+
 \begin{equation*}
-    \scriptvar = \varepsilon \sqcup \decompmax{0/1,1/0}{\scriptvar}{}{}{\mathcal{C}\setminus\varepsilon}
+    \scriptvar = \varepsilon \sqcup
+\decompmax{0/1,1/0}{\scriptvar}{}{}{\mathcal{C}\setminus\varepsilon}
 \end{equation*}
@@
 an occurrence since the maximum lies in a region corresponding to the
-shading in \(p\), so we can use any avoider of \(\perm{2,3,1}\). The lower-left section however can create occurrences
+shading in \(p\), so we can use any avoider of \(\perm{2,3,1}\). The lower-left
+section however can create occurrences
 of \(p\) and therefore must also avoid \(p\), as well as \(\perm{2,3,1}\).
@@
 
-Calculating coefficients given by this generating function gives the Fine numbers.
+Calculating coefficients given by this generating function gives the Fine
+numbers.
 \begin{equation*}
-    1, 0, 1, 2, 6, 18, 57, 186, 622, 2120, 7338,\dotsc \tag{\href{https://oeis.org/A000957}{OEIS: A000957}}
+    1, 0, 1, 2, 6, 18, 57, 186, 622, 2120, 7338,\dotsc
+\tag{\href{https://oeis.org/A000957}{OEIS: A000957}}
 \end{equation*}
@@
 \begin{equation*}
-    \scriptvar_1 =\varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
+    \scriptvar_1 =\varepsilon \sqcup
+\decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
 \end{equation*}
@@
 \(q_1\) starts with it's maximum, by not allowing the lower-left section of the
-\(\perm{2,3,1}\) avoider to be empty we prevent an occurrence from ever happening.
+\(\perm{2,3,1}\) avoider to be empty we prevent an occurrence from ever
+happening.
 
 Consider \(q_2 = \textpattern{}{1}{0/1,1/0,1/1}\), avoiding this pattern
-means that a permutation does not end with it's maximum. We can perform a similar
+means that a permutation does not end with it's maximum. We can perform a
+similar
 decomposition as before to get
 \begin{equation*}
-    \mathcal{\var}_2 =\varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\mathcal{C}\setminus\varepsilon}
+    \mathcal{\var}_2 =\varepsilon \sqcup
+\decompmax{0/1,1/0}{\mathcal{C}}{}{}{\mathcal{C}\setminus\varepsilon}
 \end{equation*}
@@
 \begin{equation*}
-    \mathcal{\var}_3 =\varepsilon \sqcup \decompleft{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
+    \mathcal{\var}_3 =\varepsilon \sqcup
+\decompleft{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
 \end{equation*}
@@
 \begin{equation*}
-    1, 0, 1, 3, 9, 28, 90, 297, 1001, 3432, 11934,\dotsc \tag{\href{https://oeis.org/A000245}{OEIS: A000245} with offset \(1\)}
+    1, 0, 1, 3, 9, 28, 90, 297, 1001, 3432, 11934,\dotsc
+\tag{\href{https://oeis.org/A000245}{OEIS: A000245} with offset \(1\)}
 \end{equation*}
@@
     \scriptvar = \varepsilon \sqcup
-    \underbrace{\decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}}_\text{
-        Minimum comes before the maximum.
-    } \sqcup
+    \underbrace{\decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}}_\text{Minimum comes before the maximum.} \sqcup
 \underbrace{\decompmax{0/0,0/1,1/0}{}{}{}{\scriptstyle{\scriptvar\setminus\varepsilon}}}_\text{
@@
 \begin{equation*}
-    	1, 0, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc \tag{\href{https://oeis.org/A141364}{OEIS: A141364}}
+    	1, 0, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc
+\tag{\href{https://oeis.org/A141364}{OEIS: A141364}}
 \end{equation*}
@@
 \section{Wilf-classes with patterns of length 2}
-By use of the set equivalences from \ChapterRef{chap:coincs} we know there are
-at most \(95\) Wilf-equivalence classes.
+By the use of set equivalences established in
+\SectionRef{sec:coincs} we know that there are at most 95 Wilf-equivalence classes.
 
-In order to consider symmetries we must only take the symmetries that preserve
-the pattern \(\perm{2,3,1}\). The only symmetry that preserves the pattern
-\(\perm{2,3,1}\) is \emph{reverse-complement-inverse}. Using this
-symmetry to reduces the number of Wilf-classes gives us \(61\) classes of trivial
-Wilf-equivalences, these Wilf-equivalences are explained by patterns being either
-coincident, or being the reverse-complement-inverse of a pattern.
+The only symmetry that we are able to consider is \emph{reverse-complement-inverse} as this is the only symmetry that preserves the
+\(\perm{2,3,1}\) pattern. Using this symmetry we can find 61 classes of trivial
+Wilf-equivalence, these equivalences being explained by either the patterns being
+coincident in \(\av{\perm{2,3,1}}\), or by one pattern being the reverse-complement-inverse of some other pattern.
 
-Computing avoiders up to length \(10\) suggest that there are at least \(23\)
-Wilf-classes, of which \(13\) are non-trivial, this means that there are
-Wilf-equivalences between patterns that are not explained by coincidences.
+Computing avoiders up to length 10 suggests that there are at least 23 Wilf-classes,
+of which 13 are non-trivial, therefore there are Wilf-equivalences that are not
+explained by coincidences or symmetry.
 
-When considering explanations of Wilf-equivalences we consider how the
-permutations correspond to set-partitions.
+When considering the Wilf-equivalences we consider how permutations correspond
+to set-partitions.
 
 \begin{note}
-    \label{not:setequiv}
-    The avoiders of the pattern \(q = \mperm{2,3,1}{\{(1,0),(1,1),(1,2),(1,3)\}}\),
-    \textpattern{}{2,3,1}{1/0,1/1,1/2,1/3},
-    in \(\mathfrak{S}_n\) are in one-to-one correspondence with partitions of
-    \(\nrange{n}\). (\textcite[Prop.~2]{DBLP:journals/ejc/Claesson01})
+  The avoiders of the pattern \(q = \mperm{\perm{2,3,1}}{(1,0),(1,1),(1,2),(1,3)},
+\textpattern{}{2,3,1}{1/0,1/1,1/2,1/3}\) in \(\mathfrak{S}_n\) are in one-to-one correspondence with partitions
+of \(\nrange{n}\) (\textcite[Prop.~2]{DBLP:journals/ejc/Claesson01})
 \end{note}
+We will call the least element in each block the \emph{block-bottom}, and note that all permutations
+in \(\av{\perm{2,3,1}}\) also avoid this mesh pattern.
 
-The idea of the bijection is as follows.
-    Let \(\pi\) be a permutation in \(\av[n]{q}\)
-    in one-line notation and insert a dash between each ascent in \(\pi\). This
-    corresponds to set partitions where the blocks are the elements between the
-    dashes, the blocks are listed in increasing order of their least element,
-    with the elements written in each block in descending order.
-
-\begin{example}
-    Given the permutation \(\pi = \perm{5,4,2,1,3,9,6,8,7}\) this corresponds to the
-    partition \(\{\{5,4,2,1\},\{3\},\{9,6\},\{8,7\}\}\).
-\end{example}
-
-We call the least element in each block the \emph{block bottom}.
-
-We are looking at permutations in \(\av{\perm{2,3,1}}\), all of these permutations
-also avoid the mesh pattern in \NoteRef{not:setequiv}, \ie
-\(\av{\perm{2,3,1}} \subset \av{q}\).
-
-The classes containing the following patterns are experimentally Wilf-equivalent up to length 10 in
-\(\av{\perm{2,3,1}}\)
-\begin{equation*}
-    \pattern{}{1,2}{} \text{ and }
-    \pattern{}{2,1}{}
-\end{equation*}
-This is true since the only avoiders of these patterns are the decreasing
-sequence and the increasing sequence respectively, and both of these
-avoid \(\perm{2,3,1}\) in all cases. There is therefore \(1\) avoider at
-every length.
-
-\nextvar[\varminl]
-The following patterns are experimentally Wilf-equivalent up to length 10 in
-\(\av{\perm{2,3,1}}\)
-\begin{equation*}
-    m_1 = \pattern{}{1,2}{0/0,0/1,0/2,
-                    1/0,1/1,1/2,
-                    2/0,2/1,2/2}
-    \text{ and }
-    m_2 = \pattern{}{2,1}{0/0,0/1,0/2,
-                    1/0,1/1,1/2,
-                    2/0,2/1,2/2}
-\end{equation*}
-It is obvious that these two are Wilf-equivalent since the only
-permutations that contain these patterns are \(\perm{1,2}\) and
-\(\perm{2,1}\) respectively, therefore the avoiders of these patterns
-are counted by the Catalan numbers at all lengths except for length
-\(2\) where there is precisely \(1\) avoider.
-Therefore the generating function is
-\begin{equation*}
-  \var(x) = C(x)-x^2
-\end{equation*}
-
-\subsection{}
-\nextvar
-The following patterns are experimentally Wilf-equivalent up to length 10 in
-\(\av{\perm{2,3,1}}\)
-\begin{equation*}
-    \pattern{}{1,2}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
-    \pattern{}{1,2}{1/1,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
-    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,1/1,2/1,2/2},
-    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,1/1,2/2},
-    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,1/1},
-    \pattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0,2/1,2/2}
-\end{equation*}
-Consider containers of these patterns in \(\av{\perm{2,3,1}}\).
-For each of these patterns there is precisely one occurrence
-in any permutation containing the pattern. Now consider
-the points in the region corresponding to the unshaded box in each case. Each
-must contain an avoider of \(\perm{2,3,1}\) that is of length
-\(n-2\). Therefore these classes are all Wilf-equivalent
-and the number of length \(n\) avoiders is
-\begin{equation*}
-    \var_n = C_n - C_{n-2}
-\end{equation*}
-for \(n \ge 2\) where \(C_n\) is the \(n\)th Catalan number, the number of
-\(\perm{2,3,1}\) avoiders of length \(n\). This gives the sequence
-\begin{equation*}
-    1, 1, 1, 4, 12, 37, 118, 387, 1298, 4433, 15366,\dotsc \tag{\(C_n - \)\href{https://oeis.org/A001453}{A001453} offset 2}
-\end{equation*}
-
-\subsection{}
-The following patterns are experimentally Wilf-equivalent up to length 10 in
-\(\av{\perm{2,3,1}}\)
-\begin{equation*}
-    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1,2/2},
-    \pattern{}{1,2}{0/0,1/1,0/2,1/0,1/2,2/0,2/1,2/2},
-    \pattern{}{1,2}{0/0,0/1,1/1,1/0,1/2,2/0,2/1,2/2}
-\end{equation*}
-Consider containers of these patterns.
-Each of these patterns again occurs precisely once in
-any containing permutation. However this time when considering
-the region corresponding to the unshaded box we need to take into consideration
-\LemmaRef{lem:incdec}
-and so the empty box can only contain a decreasing subsequence. There
-is precisely one decreasing subsequence at every length, and so there
-is exactly one container of each pattern at each length. The three
-patterns are Wilf-equivalent and have \(C_n - 1\) avoiders of
-length \(n\) for all \(n \ge 2\).
-This gives the sequence
-\begin{equation*}
-    1, 1, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc\tag{\href{https://oeis.org/A001453}{OEIS: A001453} offset 2}
-\end{equation*}
-
-\subsection{}
-The following patterns are experimentally Wilf-equivalent up to length 10 in
-\(\av{\perm{2,3,1}}\)
-\begin{equation*}
-    \pattern{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
-    \pattern{}{2,1}{0/0,0/1,1/0,1/1,1/2,2/0,2/2}
-    \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/1,2/2}
-    \pattern{}{1,2}{0/1,0/2,1/0,1/1,2/0,2/1,2/2}
-    \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/2,2/0,2/1}
-    \pattern{}{1,2}{0/0,0/1,1/0,1/2,2/0,2/1,2/2}
-\end{equation*}
-The containers of the patterns
-have exactly one occurrence.
-Once again we consider the regions corresponding to the unshaded regions, For each
-pattern,
-except the first the two, these regions are independent, and one contains
-any avoider of \(\perm{2,3,1}\) and the other must contain a
-decreasing sequence by \LemmaRef{lem:incdec}.
-Let us consider the first pattern separately. In order to
-avoid \(\perm{2,3,1}\) across the regions corresponding to the unshaded boxes
-we can add some additional restrictions, \ie all elements in the top region must
-be to the right of all elements in the bottom region.
-\begin{equation}
-    \raisebox{0.6ex}{
-    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
-        \fill[pattern=north east lines, pattern color=black!75] (1.5, 0) rectangle +(0.5,1);
-        \fill[pattern=north east lines, pattern color=black!75] (1.0, 2) rectangle +(0.5,1);
-        \draw (1.5,0) -- (1.5,1)
-              (1.5,2) -- (1.5,3);
-        \modpattern[2]{}{1,2}{0/0,0/1,0/2,1/1,2/0,2/1,2/2}
-    \end{tikzpicture}
-    }
-    \label{eqn:splitboxes}
-\end{equation}
-Now we can see that the region corresponding to the top free box must contain a decreasing sequence,
-and the bottom must contain an avoider of \(\perm{2,3,1}\) and these
-two do not interact in any manner. The containers of this pattern are counted
-the same as the other patterns, and due to this they are Wilf-equivalent in \(\av{231}\).
-The containers
-have generating function \(x^2C(x)/(1-x)\). Enumerating avoiders
-therefore gives us
-\begin{equation*}
-    1, 1, 1, 3, 10, 33, 109, 364, 1233, 4236, 14740,\dotsc \tag{\(C_n -\) \href{https://oeis.org/A014137}{A014137} offset 2}
-\end{equation*}
-
-\subsection{}
-\nextvar
-The following patterns are experimentally Wilf-equivalent up to length 10 in
-\(\av{\perm{2,3,1}}\)
-\begin{equation*}
-    m_1 = \pattern{}{1,2}{0/0,0/2,
-                    1/0,1/1,
-                    2/0,2/1,2/2}
-    \text{ and }
-    m_2 = \pattern{}{1,2}{0/0,
-                    1/0,1/1,1/2,
-                    2/0,2/1,2/2}
-\end{equation*}
-In this case it is better to consider the containers of the patterns
-instead of the avoiders due to the amount of shadings in the mesh.
-
-We look at the containers of the pattern \(m_1\), there can only ever
-be one occurrence of this pattern in a permutation corresponding to
-the last point in the permutation and the minimum. Consider an occurrence of \(m_1\),
-the points in the two regions corresponding to the the two boxes must form decreasing
-subsequences.
-For a permutation of length \(k\) if we fix the number of points in
-one of the boxes the number of points in the other box is determined.
-Therefore we can have any number of points from \(\setrange{0}{k-2}\)
-in the bottom box. Therefore there are \(k-1\) containers of
-length \(k\). These permutations correspond to set partitions of \(k\)
-points into exactly two non-overlapping parts, such that the first part
-is the decreasing sequence from the first element to the minumum, and
-the second part is the elements in the sequence from the maximum to the last
-element.
-%partitioned by the first element and the minimum.
-
-Now consider the containers of \(m_2\). We know that the unshaded region
-must contain a decreasing subsequence, with the point corresponding to
-the \(1\) in the mesh pattern. This decreasing subsequence has \(k-1\)
-points. We can put the point corresponding to the \(2\) above any of these
-points and therefore there are \(k-1\) containers of length \(k\).
-
-Therefore these two patterns have been shown to have the same number of
-avoiders of length \(k\) for all \(k\) and are Wilf-equivalent.
-The number of avoiders of length \(k\) is given by
-\begin{equation*}
-  \var_k = C_k - (k-1), \var_0 = 1
-\end{equation*}
-and have enumeration
-\begin{equation*}
-    1, 1, 1, 3, 11, 38, 127, 423, 1423, 4854, 16787,\dotsc \tag{\(C_n -\) \href{https://oeis.org/A000027}{A000027} offset 2}
-\end{equation*}
+We will use two main methods of establishing Wilf-equivalence between mesh patterns of length \(2\)
+in \(\av{\perm{2,3,1}}\): the structural decomposition of avoiders, via generating functions; or the
+structure of the set-partition induced by the pattern, looking at a particular occurrence of the
+pattern in a permutation avoiding \(\perm{2,3,1}\). Sometimes it will be necessary to use both of
+these methods to consolidate a single Wilf-class.
 
@@
 
-\subsection{}
-\nextvar
-The following patterns are experimentally Wilf-equivalent up to length 10 in
-\(\av{\perm{2,3,1}}\)
-\begin{equation*}
-    m_1 = \pattern{}{2,1}{0/0,0/1,0/2,
-                    1/1,1/2,
-                    2/0,2/2}
-    \text{ and }
-    m_2 = \pattern{}{1,2}{0/1,0/2,
-                    1/0,1/2,
-                    2/0,2/1,2/2}
-\end{equation*}
-
-First consider the structure of an avoider of \(m_1\) in \(\av{\perm{2,3,1}}\).
-We can perform the usual structural decomposition of an avoider
-of \(\perm{2,3,1}\) where we consider decomposition around the maximum.
-% Any permutation in \(\av{\perm{2,3,1},m_1}\) is in the following set
-% precisely once.
-If \(\scriptvar_1\) is the set \(\av{\perm{2,3,1},m_1}\) then any permutation in
-\(\scriptvar_1\) either starts with a maximum or does not, giving us the
-decomposition
-\begin{equation*}
-    \scriptvar_1 = \varepsilon \sqcup
-    \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}} \sqcup
-    \decompmax{0/1,1/0,0/0}{}{}{}{\mathcal{\varfine}}
-\end{equation*}
-Where \(\mathcal{\varfine} = \av{231,\textpattern{}{1}{0/1,1/0}}\).
-Now consider the decomposition around the maximum of a permutation in
-\(\scriptvar_2 = \av{\perm{2,3,1},m_2}\), the permutation either ends with
-the maximum, or it does not, so we get
-\begin{equation*}
-    \scriptvar_2 = \varepsilon \sqcup
-    \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\mathcal{C}\setminus\varepsilon} \sqcup
-    \decompmax{0/1,1/0,1/1}{\mathcal{\varfine}}{}{}{}
-\end{equation*}
-Therefore both of these sets of avoiders are enumerated in the same manner
-having generating function satisfying
-\begin{equation*}
-    \var(x) = 1 + xC(x)(C(x)-1) + x\varfine(x)
-\end{equation*}
-This generating function gives
-\begin{equation*}
-    1, 1, 1, 4, 11, 34, 108, 354, 1187, 4054, 14054,\dotsc \tag{\(C_n - \)\href{https://oeis.org/A000958}{A000958} offset 2}
-\end{equation*}
-
-\subsection{}
-\nextvar
-The following patterns are experimentally Wilf-equivalent up to length 10 in
-\(\av{\perm{2,3,1}}\)
-    \begin{gather*}
-        m_1 = \pattern{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/2},
-        m_2 = \pattern{}{1,2}{0/0,0/1,1/0,2/0,2/1,2/2}, \\
-        m_3 = \pattern{}{1,2}{0/2,1/0,1/1,2/0,2/1,2/2}, \text{ and }
-        m_4 = \pattern{}{1,2}{0/1,1/0,1/1,2/0,2/1,2/2}
-    \end{gather*}
-
-First consider the decomposition of avoiders of \(m_1\) in \(\av{\perm{2,3,1}}\)
-around the maximum. We have different conditions if we start with the maximum or not.
-\begin{equation*}
-    \scriptvar_1 =
-    \varepsilon \sqcup
-    \decompmax{0/1,1/0}{\scriptstyle \mathcal{\varmaxl}_1\setminus\varepsilon}{}{}{\mathcal{C}}
-    \sqcup
-    \decompmax{0/0,0/1,1/0}{}{}{}{\scriptvar_1}
-\end{equation*}
-Where \(\mathcal{\varmaxl}_1 = \av{231,\textpattern{}{1}{0/0,0/1,1/1}}\).
-Now we decompose the avoiders of \(m_2\) around the leftmost point, we have similar
-conditions for starting with the maximum
-\begin{equation*}
-    \scriptvar_2 =
-    \varepsilon \sqcup
-    \decompleft{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}\setminus\varepsilon}
-    \sqcup
-    \decompleft{0/1,1/0,1/1}{\scriptvar_2}{}{}{}
-\end{equation*}
-This gives us two generating functions satisfying the following pair of equations
-\begin{align}
-    \var_1(x) &= 1 + xC(x)(\varmaxl(x)-1) +x\var_1(x) \label{eqn:genL1}\\
-    \text{ and }\var_2(x) &= 1 + x(C(x)-1)^2 + x\var_2(x) \label{eqn:genL2}
-\end{align}
-In order for these two functions to give the same value it is necessary to show that
-\eqref{eqn:genL1} and \eqref{eqn:genL2} are equal, this occurs if
-\(C(x)(\varmaxl(x)-1) = (C(x)-1)^2\).
-\begin{align*}
-    &&(C(x)-1)^2 &= C(x)(\varmaxl(x)-1)\\
-        &\Leftrightarrow& x^2C^4(x) &=  xC(x)(C(x)-1)C(x) &&\text{By definition of \(\varmaxl\) and \(C\)}\\
-        &\Leftrightarrow& x^2C^4(x)&=xC^3(x)-xC^2(x)\\
-        &\Leftrightarrow& xC^2(x) &= C(x) -1 &&\text{Divide by \(xC^2(x)\)}\\
-        &\Leftrightarrow& C(x) &= 1 +xC^2(x)
-\end{align*}
-The final line is always satisfied since it is the form of \(C(x)\), and
-therefore the two generating functions are equal.
-
-Now we look at the other patterns. In particular note that any container of these
-patterns can contain the pattern precisely once, \(m_2\) specifies the
-minimum and last point, \(m_3\) and \(m_4\) both use the last point and the
-previous block bottom (in the set partition context).
-
-Consider an occurrence of \(m_3\) in \(\av{\perm{2,3,1}}\)
-\begin{equation*}
-    \raisebox{0.6ex}{
-        \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
-            \modpattern[2]{}{1,2}{0/2,1/0,1/1,2/0,2/1,2/2}
-            \draw (0.5,0.5) node {\(b_1\)};
-            \draw (0.5,1.5) node {\(b_2\)};
-            \draw (1.5,2.5) node {\(b_3\)};
-        \end{tikzpicture}
-    }
-\end{equation*}
-The regions \(b_2\) and \(b_3\) must contain a decreasing sequence by
-\LemmaRef{lem:incdec}. The box labelled \(b_1\) must contain an avoider
-of \(\perm{2,3,1}\). However note that the points in this box can have
-interaction with any points in box \(b_2\). If there is a point
-in \(b_2\) then any points in \(b_1\) to the left of this point must
-be lower than any points to the right of this point. By extension,
-if \(b_2\) contains a decreasing sequence with \(k\) points, there
-are \(k+1\) non-interacting avoiders of \(\perm{2,3,1}\) in \(b_1\).
-
-Now in \(m_2\) and \(m_4\) containers we can use the same method as in
-\eqref{eqn:splitboxes} to separate the two decreasing sequences in the
-free regions in the top row, and the mixing happens in the same manner
-as in a container of \(m_3\).
-We now have that \(m_2,m_3\) and \(m_4\) have the same number of containers
-so are Wilf-equivalent, and that \(m_1\) and \(m_2\) have the same generating
-function so all four classes are Wilf-equivalent.
-
-Evaluating the generating function \(\var(x)\) gives us the enumeration
-\begin{equation*}
-    1, 1, 1, 2, 6, 20, 68, 233, 805, 2807, 9879,\dotsc\tag{\(C_n - \) \href{https://oeis.org/A014138}{A014138} offset 1}
-\end{equation*}
-
-\subsection{}
-\nextvar
-The following patterns are experimentally Wilf-equivalent up to length 10 in
-\(\av{\perm{2,3,1}}\)
-\begin{equation*}
-    \pattern{}{1,2}{0/0,0/2,1/0,1/2,2/0,2/2},
-    \pattern{}{1,2}{1/1,2/1,1/0,1/2,2/0,2/2},
-    \pattern{}{1,2}{0/1,1/0,1/2,2/0,2/2},
-    \pattern{}{1,2}{0/1,0/2,1/1,1/2,2/0}, \text{ and }
-    \pattern{}{1,2}{0/0,0/1,0/2,1/0,1/1,1/2}
-\end{equation*}
-
-If \(\scriptvar_1\) is the set of avoiders of
-\textpattern{}{1,2}{0/0,0/2,1/0,1/2,2/0,2/2}, then by the structural
-decomposition around the maximum we have
-\begin{equation*}
-    \scriptvar_1 = \varepsilon \sqcup
-    \decompmax{0/0,0/1,1/0}{}{}{}{\mathcal{C}}
-\end{equation*}
-The lower-left section is empty because the minimum must occur after the maximum.
-These are counted by \(\var(x) = 1+xC(x)\)
-% This is because \(\scriptvar_1\) is the set of permutations
-% who have their minimum occur after their maximum.
-
-% If \(\mathcal{M}_2\) is the set of avoiders of
-% \textpattern{}{1,2}{1/1,2/1,1/0,1/2,2/0,2/2}, then by structural
-% decomposition around the maximum we have
-% \begin{equation*}
-%     \mathcal{M}_2 = \varepsilon \sqcup
-%     \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\scriptstyle\mathcal{M}_1\setminus\varepsilon}
-%     \sqcup \tikz{\filldraw (0,0) circle (3pt);}
-% \end{equation*}
-% Alternatively,
-The pattern \textpattern{}{1,2}{1/1,2/1,1/0,1/2,2/0,2/2} occurs if
-the last element is higher than the penultimate element. This can only occur
-if the last element is in a single block in the set partition context. In order
-to construct an avoider of length \(n\) we can take any avoider of
-\(\perm{2,3,1}\) of length \(n-1\) and insert the new maximum into the last
-block.  This ensures that the last block is never a singleton. This means
-that these permutations are also counted by \(\var(x)\).
-
-Considering the last pattern, the only way we can construct an
-avoider is to take any \(\perm{2,3,1}\) avoider and add a new minimum
-at the start of the permutation. Adding a new leftmost point with any other
-value would either create an occurrence of \(\perm{2,3,1}\) or the mesh pattern.
-Therefore these permutations are also counted by \(\var(x) = 1+xC(x)\).
-
-The avoiders of the third pattern can be decomposed by the maximum
-to give
-\begin{equation*}
-    \scriptvar_3 = \varepsilon \sqcup
-    \decompmax{0/1,1/0}{\mathcal{\varfine}}{}{}{\scriptvar_3}
-\end{equation*}
-Where \(\mathcal{\varfine} = \av{(231,\textpattern{}{1}{0/1,1/0})}\).
-The generating function derived satisfies \(\var_3(x) = 1 + x\varfine(x)\var_3(x)\).
-The fourth pattern can be decomposed around the maximum in a similar
-manner.
-\begin{equation*}
-    \scriptvar_4 = \varepsilon \sqcup
-    \decompmax{0/1,1/0}{\scriptvar_4}{}{}{\mathcal{\varfine}}
-\end{equation*}
-
-So clearly \(\var_4(x) = \var_3(x)\).
-We need to show that the generating function \(\var_3(x)\) is the same as \(\var(x)\)
-\begin{equation*}
-    \begin{aligned}
-        \var_3(x) &= 1 + x\varfine(x)\var_3(x) \\
-        &= \frac{1}{1-x\varfine(x)} && \text{Solving for \(\var_3(x)\)}\\
-        &= \frac{1}{1 - \frac{xC}{1+xC(x)}} &&\text{Substituting for \(\varfine(x)\)}\\
-        &= \frac{1+xC(x)}{1+xC(x)-xC(x)}\\
-        &= 1+xC(x)
-    \end{aligned}
-\end{equation*}
-We have that \(\var_3(x) = 1+xC(x)= \var(x) \) so all four patterns are Wilf-equivalent
- and have enumeration sequence
- \begin{equation*}
-     1, 1, 1, 2, 5, 14, 42, 132, 429, 1430, 4862,\dotsc \tag{\href{https://oeis.org/A000108}{OEIS: A000108} offset 1}
- \end{equation*}
- This is an offset of the Catalan numbers.
-
-\subsection{}
-\nextvar
-The following patterns are experimentally Wilf-equivalent up to length 10 in
-\(\av{\perm{2,3,1}}\)
-\begin{gather*}
-        m_1 = \pattern{}{1,2}{0/0,0/1,1/0,1/2,2/1},
-        m_2 = \pattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0},\\
-        m_3 = \pattern{}{1,2}{0/1,0/2,1/0,1/2,2/0,2/2}, \text{ and }
-        m_4 = \pattern{}{2,1}{1/0,1/1,2/1,2/2}\\
-\end{gather*}
- First we consider an occurrence of \(m_2\) in a
-permutation in \(\av{\perm{2,3,1}}\)
-\begin{equation*}
-    \raisebox{0.6ex}{
-    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
-        \fill[pattern=north east lines, pattern color=black!75] (2.5, 1) rectangle +(0.5,1);
-        \fill[pattern=north east lines, pattern color=black!75] (2.0, 2) rectangle +(0.5,1);
-        \draw (2.5,1) -- (2.5,3);
-        \draw (1.5,0.5) node {\(b_1\)};
-        \draw (2.25,1.5) node {\(\scriptstyle b_2\)};
-        \draw (2.75,2.5) node {\(\scriptstyle b_3\)};
-        \modpattern[2]{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0}
-    \end{tikzpicture}
-    }
-\end{equation*}
-We can choose the lowest occurrence of \(m_2\) in the sense that the region
-corresponding to \(b_1\) must avoid the pattern \(\textpattern{}{1}{0/1,1/0}\)
-as well as \(\perm{2,3,1}\). The regions corresponding to \(b_2\) and \(b_3\) must
-now contain avoiders of \(\perm{2,3,1}\), all points in the region corresponding
-to \(b_2\) must be to the left of those in the region corresponding to \(b_3\) since
-we are in
-\(\av{\perm{2,3,1}}\). Since we already have an occurrence of
-\(m_2\) we do not need to care about creating more occurrences so there are no
-other conditions on these boxes.
-
-Now looking at an occurrence of \(m_3\) in \(\pi \in \av{\perm{2,3,1}}\)
-\begin{equation*}
-    \raisebox{0.6ex}{
-    \begin{tikzpicture}[scale=1, baseline=(current bounding box.center)]
-        \fill[pattern=north east lines, pattern color=black!75] (1, 1.5) rectangle +(1,0.5);
-        \fill[pattern=north east lines, pattern color=black!75] (2, 1) rectangle +(1,0.5);
-        \draw (1,1.5) -- (3,1.5);
-        \draw (0.5,0.5) node {\(b_1\)};
-        \draw (1.5,1.25) node {\(\scriptstyle b_2\)};
-        \draw (2.5,1.75) node {\(\scriptstyle b_3\)};
-        \modpattern[2]{}{1,2}{0/1,0/2,1/0,1/2,2/0,2/2}
-    \end{tikzpicture}
-    }
-\end{equation*}
-We consider the leftmost occurrence of \(m_3\) in the sense that the region
-corresponding
-to \(b_1\) must avoid the pattern \(\textpattern{}{1}{0/1,1/0}\) as well as
-\(\perm{2,3,1}\) once more. The regions corresponding to \(b_2\) and \(b_3\)
-must avoid \(\perm{2,3,1}\) and as in a container of \(m_2\) the points in the
-region corresponding to the box containing \(b_2\) must be lower in value than
-all of those in the region corresponding to the box containing \(b_3\), as doing
-so would lead to an occurrence of \(\perm{2,3,1}\).
-Therefore both of these sets of containers are enumerated in the same way.
-
-Now we find a structural decomposition for an avoider of \(m_2\).
-Decomposing around the maximum we see the set of avoiders of \(m_2\)
-have the form
-\begin{equation*}
-    \scriptvar_2 = \varepsilon \sqcup
-    \decompmax{0/1,1/0}{\scriptstyle\scriptvar_2\setminus\varepsilon}{}{}{\mathcal{C}}
-    \sqcup
-    \decompmax{0/0,0/1,1/0}{}{}{}{\mathcal{\varfine}}
-\end{equation*}
-We can decompose an avoider of \(m_1\) in \(\av{\perm{2,3,1}}\) around the leftmost point
-in a similar manner:
-\begin{equation*}
-    \scriptvar_1 = \varepsilon \sqcup
-    \decompleft{0/1,1/0}{\scriptstyle\scriptvar_1\setminus\varepsilon}{}{}{\mathcal{C}}
-    \sqcup
-    \decompleft{0/0,0/1,1/0}{}{}{}{\mathcal{\varfine}}
-\end{equation*}
-These two decompositions tell us that these two patterns are Wilf-equivalent
-and have generating function
-\begin{equation}
-    \var_1 = 1 + x(\var_1(x)-1)C(x) +x\varfine(x)
-    \label{eqn:ngen}
-\end{equation}
-Now consider an avoider of \(m_4\) decomposed around the maximum\nextvar %subvar
-\begin{equation*}
-    \mathcal{\lastvar}_4 = \varepsilon \sqcup
-    \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\scriptvar}
-\end{equation*}
-Here \(\scriptvar\) is the permutations avoiding \(\perm{2,3,1},m_4\) and
-\(p = \textpattern{}{1}{0/0,0/1,1/1}\), since if the subsequence in this box
-were to start with the maximum then this point and the maximum would create
-an occurrence of \(m_4\).
-Now consider decomposition of a permutation in \(\scriptvar\) around its
-leftmost point.
-\begin{equation*}
-    \scriptvar = \varepsilon \sqcup
-    \decompleft{0/1,1/0}{\mathcal{C}}{}{}{\scriptstyle\mathcal{\lastvar}_4\setminus\varepsilon}
-\end{equation*}
-This gives us the generating function
-\begin{equation*}
-    \var(x) = 1 + xC(x)(\lastvar_4(x)-1)
-\end{equation*}
-Now we get the following for \(\lastvar_4\)
-\begin{equation}
-    \lastvar_4(x) = 1 +xC(x)(xC(x)(\lastvar_4(x)-1)+1)
-    \label{eqn:ngen2}
-\end{equation}
-
-All that remains to show Wilf-equivalence is to show that \EquationRef{eqn:ngen} and
-\EquationRef{eqn:ngen2} are the same generating function.
-First solve \EquationRef{eqn:ngen2} for \(\lastvar_4(x)\)
-\begin{equation}
-    \begin{aligned}
-        \lastvar_4(x) &= 1+xC(x)(xC(x)(\lastvar_4(x)-1)+1)\\
-        &= 1 + x^2C^2(x)\lastvar_4(x) -x^2C^2(x) +xC(x)\\
-        &= 1 + \frac{xC(x)}{1 - x^2C^2(x)}\\
-        &= 1 + \frac{xC(x)}{(1-xC(x))(1+xC(x))} && \text{Difference of squares}\\
-        \lastvar_4(x)&= 1 + \frac{xC^2(x)}{1+xC(x)} && C(x) = \frac{1}{1-xC(x)}
-    \end{aligned}
-\end{equation}
-
-Now we solve \EquationRef{eqn:ngen} for \(\lastvar_1(x)\)
-\begin{equation}
-    \begin{aligned}
-        \lastvar_1(x)&= 1+x(\lastvar_1(x)-1)C(x)+x\varfine(x)\\
-        &= 1+x\lastvar_1(x)C(x)-xC(x)+x\varfine(x)\\
-        \lastvar_1(x)&= 1+xC(x)\varfine(x)\\
-        \lastvar_1(x) &= 1 + \frac{xC^2(x)}{1+xC(x)} &&\text{Substitution of \(\varfine(x)\)}
-        % &= 1 +\frac{x}{x^2C^2(x) - (x^2+2x)C(x) +x+1}\\
-        % &= 1 + \frac{x}{1-x^2C(x)-xC(x)} && xC^2(x) = C(x)-1\\
-        % &= 1 + \frac{xC^2(x)}{C^2(x) - xC^3(x)(x+1)}\\
-        % &= 1 + \frac{xC^2(x)}{C(x) - xC^2(x) + xC(x)} && xC^2(x) = C(x)-1\\
-        % \lastvar_1(x)&= 1 + \frac{xC^2(x)}{1+xC(x)} && C(x) = 1+xC^2(x)\\
-    \end{aligned}
-\end{equation}
-We have shown that \(\lastvar_1\) and \(\lastvar_4\) are indeed the same generating function,
-and we have that the classes containing these four patterns are Wilf-equivalent.
-Evaluating the generating function \(\lastvar(x)\) gives
-\begin{equation*}
-    1, 1, 1, 3, 8, 24, 75, 243, 808, 2742, 9458,\dotsc \tag{\href{https://oeis.org/A001453}{OEIS: A001453}}
-\end{equation*}
+This particular example is interesting as it shows that both methods can be used in tandem to establish a coincidence class.
 
@@
 
-\subsection{}
-\nextvar
-The following patterns are experimentally Wilf-equivalent up to length 10 in
-\(\av{\perm{2,3,1}}\)
-\begin{gather*}
-        \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,1/2,2/2},
-        \pattern{}{1,2}{0/1,0/2,1/0,1/1,1/2,2/0,2/2},
-        \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,1/2,2/1},
-        \pattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0,2/1},
-        \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,1/2,2/0},\\
-        \pattern{}{1,2}{0/0,0/1,0/2,1/2,2/0,2/1,2/2},
-        \pattern{}{1,2}{0/0,0/1,0/2,1/1,1/2,2/0,2/2},
-        \pattern{}{1,2}{0/0,0/1,1/0,1/1,1/2,2/0,2/1},
-        \pattern{}{1,2}{0/0,0/1,1/0,1/1,1/2,2/0,2/1}
-\end{gather*}
-
-In order to gain enumeration, consider decomposition of avoiders
-of first pattern, \(\textpattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,1/2,2/2}\), around
-the maximum.
-\begin{equation*}
-    \scriptvar_1 = \varepsilon \sqcup
-    \tikz{\filldraw (0,0) circle (3pt);}
-    \sqcup
-    \decompmax{0/1,1/0}{\mathcal{C\setminus\varepsilon}}{}{}{\mathcal{C}}
-\end{equation*}
-
-This gives us the following generating function
-\begin{equation}
-    \var(x) = 1 + x + xC(x)(C(x)-1) \label{eq:qgen}
-\end{equation}
-
-Now we consider decomposition of an avoider of the second pattern,
-\textpattern{}{1,2}{0/1,0/2,1/0,1/1,1/2,2/0,2/2},  in \(\av{231}\) around the maximum.
-This avoider has form
-\begin{equation*}
-    \scriptvar_2 = \varepsilon \sqcup
-    \decompmax{0/1,1/0}{\mathcal{\varmaxl}_2}{}{}{\mathcal{C}}
-\end{equation*}
-Where \(\mathcal{\varmaxl}_2 = \av{231,\textpattern{}{1}{0/1,1/0,1/1}}\).
-This gives us the generating function
-\begin{equation*}
-    \begin{aligned}
-        \var_2(x) &= 1+xC(x)\varmaxl(x)\\
-        &= 1+xC(x)(1+xC(x)(C(x)-1))\\
-        &= 1 +xC(x)(C(x)-xC(x)) && C(x) = 1+xC^2(x)\\
-        &= 1 +x + xC^2(x) - xC(x) \\
-        &= 1 +x + xC(x)(C(x)-1)
-    \end{aligned}
-\end{equation*}
-Therefore this generating function is the same as \EquationRef{eq:qgen}.
-We can decompose
-\textpattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,1/2,2/1},
-\textpattern{}{2,1}{0/0,0/1,0/2,1/1,1/2,2/0,2/1},
-\textpattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,1/2,2/0},
-\textpattern{}{1,2}{0/0,0/1,0/2,1/2,2/0,2/1,2/2},
-and \textpattern{}{1,2}{0/0,0/1,0/2,1/1,1/2,2/0,2/2} around the leftmost point
-into an avoider of one of the patterns with generating function \(\varmaxl(x)\) and
-an avoider of \(\perm{2,3,1}\).
-
-Now decompose an avoider of \textpattern{}{1,2}{0/0,0/1,1/0,1/1,1/2,2/0,2/1}
-around the leftmost point.
-\begin{equation*}
-    \scriptvar_3 = \varepsilon \sqcup
-    \decompleft{0/1,1/0}{\scriptstyle \scriptvar_3\setminus\varepsilon}{}{}{\mathcal{C}}
-    \sqcup \decompleft{0/0,0/1,1/0}{}{}{}{\varmaxl_3}
-\end{equation*}
-Where \(\mathcal{\varmaxl}_3 = \av{231,\textpattern{}{1}{0/0,0/1,1/0}}\).
-This gives generating function \(\var_3(x)\) satisfying
-\begin{equation*}
-    \begin{aligned}
-        \var_3(x) &= 1 +xC(x)(\var_3(x)-1) + x\varmaxl(x)\\
-        &= C(x) - xC^2(x) + xC(x) +x^2C^3(x) - x^2C^2 && \text{Solving for \(\var_3(x)\)}\\
-        &= 1 + xC(x) +x^2C^3(x)-xC^2(x) && C = 1+xC^2(x)\\
-        &= 1 + x + x^2C^3(x) && xC^2(x) = C(x)-1\\
-        &= 1 + x + xC(x)(C(x)-1)
-    \end{aligned}
-\end{equation*}
-This is equivalent to \EquationRef{eq:qgen}, and therefore these patterns are Wilf-equivalent.
-The classes have enumeration
-\begin{equation*}
-    1, 1, 1, 3, 9, 28, 90, 297, 1001, 3432, 11934,\dotsc \tag{\href{https://oeis.org/A071724}{OEIS: A071724 with offset}}
-\end{equation*}
+In order to establish the remainder of the Wilf-equivalences of the form \(\av{231,p}\) where \(p\)
+is a mesh pattern we can use similar methods to allow us to consolidate experimental classes into
+actual classes.

### 2017-03-07 writing/paper/sections/coincidences.tex
reflow and add henning as author

@@
 \end{proof}
+
+All coincidence classes of \(\av{\{\perm{3,2,1}, \mperm{2,1}{R}\}}\) can be
+explained by application of \PropositionRef{prop:dom1}. By experimentation
+we see that there are at least \(29\) coincidence classes, and all of these
+coincidences are explained by this Proposition.
+
+This rule can be understood very in graphical form. In the
+pattern in \FigureRef{fig:rule1} we can gain shading in two boxes since if
+there is a point in any of these boxes we would get an occurrence of the
+dominating pattern \(\perm{3,2,1}\).
+
+\begin{figure}[htb]
+    \begin{center}
+    \begin{tikzpicture}[scale=\picscale]
+        \modpattern{}{2,1}{1/0,1/1,1/2,2/2}
+    \end{tikzpicture}
+        \raisebox{2ex}{\(\mapsto\)}
+    \begin{tikzpicture}[scale=\picscale]
+        \filldraw (2.5,0.5) circle (3pt);
+        \modpattern{}{2,1}{1/0,1/1,1/2,2/2,2/0}
+    \end{tikzpicture}
+
+        \caption{Visual depiction of first dominating pattern rule.}
+        \label{fig:rule1}
+    \end{center}
+\end{figure}
+
+\subsection{Equivalence classes of Av(\{231, (21, \textit{R})\}).}
+By application of \PropositionRef{prop:dom1} we obtain \(43\) coincidence
+classes. Experimentation shows that there are in fact at least \(39\) coincidence
+classes, for example the following two patterns are coincident in
+\(\av{\perm{2,3,1}}\) but this is not explained by \PropositionRef{prop:dom1}.
+\begin{equation*}
+    m_1 = \pattern{}{2,1}{0/0,0/1,0/2,1/0,2/1} \text{ and } m_2 = \pattern{}{2,1}{0/0,0/1,0/2,1/0,1/1,2/1}
+\end{equation*}
+
+Consider an occurrence of \(m_1\) in a permutation in \(\av{\perm{2,3,1}}\), consisting
+of elements \(x\) and \(y\). If the region corresponding
+to the box \((1,1)\) is empty we have an occurrence of \(m_2\). Otherwise, if there is any
+increase in this box then we would have an occurrence of \(\perm{2,3,1}\),
+however, since we are in \(\av{\perm{2,3,1}}\) this is not possible. This box
+must therefore contain a (non-empty) decreasing subsequence.
+ This gives rise to the following lemma:
+
+\begin{lemma}
+    \label{lem:incdec}
+    Given a mesh pattern \(m =(\sigma, R)\), where the box \((a,b)\) is not
+    in \(R\), and a dominating classical pattern \(\pi = (\pi,\emptyset)\) if
+    \(\pi \preceq p^{\boks{i,j}_a}\)\\(\(\pi \preceq p^{\boks{i,j}_d}\))
+    ,then in any occurrence of \(m\) in a permutation \(\varrho\), the region
+    corresponding to the box \((a,b)\) can only contain an decreasing
+    (increasing) subsequence of \(\varrho\).
+\end{lemma}
+The proof is analogous to the proof of \PropositionRef{prop:dom1}.
+
+Going back to our example mesh patterns
+\begin{equation*}
+    \begin{tikzpicture}[scale=\picscale]
+        \modpattern{}{2,1}{0/0,0/2,1/0,2/1}
+        \filldraw (1.66,1.33) circle (3pt);
+        \filldraw (1.33,1.66) circle (3pt);
+    \end{tikzpicture}
+\end{equation*}
+We know that the region corresponding to the box \((1,1)\) contains a decreasing subsequence.
+If we let \(z\) be the topmost point in this decreasing subsequence, then \(xz\)
+is an occurrence of \(m_2\). This shows that our two example patterns are coincident.
+
+This result generalises into the following rule for categorising coincidences
+of mesh patterns in cases where there is a dominating classical pattern.
+
+\begin{proposition}[Second Dominating Pattern Rule]
+    \label{prop:dom2}
+    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
+    and a dominating classical pattern \(\pi = (\pi,\emptyset)\) such that
+    \(\setsize{\pi} \le \setsize{\sigma} + 2\), the sets \(\av{\{\pi,m_1\}}\) and
+    \(\av{\{\pi,m_2\}}\) are coincident if
+
+    \begin{enumerate}
+        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
+        \item   \begin{enumerate}
+                \item\label{prop:dom2:condc} \(\pi \preceq p^{\boks{i,j}_a}\) and
+                        \begin{enumerate}
+                            \item \((a+1,b) \in \sigma\) and \((a+1,b-1)\notin R\) and \\
+                                \((x,b-1)\in R \implies (x,b) \in R \) (where \(x\neq a,a+1\)) and\\
+                                  \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b-1,b\)).
+                            \item \((a,b+1) \in \sigma\) and \((a-1,b+1)\notin R\) and \\
+                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
+                                  \((a-1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
+                        \end{enumerate}
+                    \item \(\pi \preceq p^{\boks{i,j}_d}\) and
+                        \begin{enumerate}
+                            \item \((a+1,b+1) \in \sigma\) and \((a+1,b+1)\notin R\) and \\
+                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a,a+1\)) and\\
+                                  \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
+                            \item \((a,b) \in \sigma\) and \((a-1,b-1)\notin R\) and \\
+                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
+                                  \((a-1,y)\in R \implies (a,y) \in R\)  (where \(y\neq b-1,b\)).
+                        \end{enumerate}
+                \end{enumerate}
+    \end{enumerate}
+\end{proposition}
+\begin{proof}
+    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).
+
+    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
+
+    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\).
+    By \NoteRef{not:downcmesh} any of these occurrences is also an occurrence
+    of \(m_1\). This proves that every occurrence of \(m_2\) is also an
+    occurrence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).
+
+    We will consider taking the first branch of every choice.
+    Now consider a permutation \(\omega^\prime \in \av{\pi}\). Suppose
+    \(\omega^\prime\) contains \(m_1\) and consider the region corresponding to
+    \((a,b)\) in \(R_1\).
+
+    If the region is empty, the occurrence of \(m_1\) is trivially an occurrence
+    of \(m_2\).
+
+    Now consider if the region is non-empty, by \LemmaRef{lem:incdec} and
+    condition~\ref{prop:dom2:condc} of the proposition this region must contain
+    a decreasing subsequence.
+    We can choose the topmost point in the region to replace the corresponding
+    point in the mesh pattern and the points from the subsequence are now in
+    the box southeast of the point. The other conditions allow this to be done
+    without points being present in regions that were shaded. Hence there are no
+    points in the region corresponding to the box \((a,b)\) in the mesh
+    pattern, and therefore we can shade this region. This implies that every
+    occurrence of \(m_1\) in \(\av{\pi}\) is in fact an occurrence of \(m_2\)
+    so \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).
+
+    Similar arguments satisfy the remainder of the branches.
+\end{proof}
+
+This proposition essentially states that we slide all of the points in the box we desire
+to shade diagonally, and chose the topmost/bottommost point to replace the original
+point in the mesh pattern.
+
+By taking the First Dominating Pattern Rule and the Second Dominating Pattern
+Rule together coincidences of classes of the form \(\av{\{\perm{2,3,1},\mperm{2,1}{R}\}}\)
+are fully explained, obtaining 39 coincidence classes of mesh patterns.
+
+\subsection{Equivalence classes of Av(\{231, (12, \textit{R})\}).}
+When considering the coincidence classes of \(\av{\perm{2,3,1},\mperm{1,2}{R}}\)
+we first apply the two Dominating Pattern rules previously established.
+Starting from \(220\) classes, application of the first Dominating Pattern rule
+gives \(85\) classes. Following this with the second Dominating Pattern rule
+reduces the number of classes to \(59\). However we know that there are patterns
+where the coincidences are not explained by the rules given above.
+
+For example the patterns
+\begin{equation*}
+    m_1 = \pattern{}{1,2}{0/2,2/0,2/1} \text{ and } m_2 = \pattern{}{1,2}{0/2,1/0,2/0,2/1}
+\end{equation*}
+are experimentally coincident. This coincidence is not explained by our rules,
+it is necessary to attempt to capture these coincidences by establishing more
+rules.
+
+Consider an occurrence of \(m_1\) in a permutation, if the region corresponding
+to the box \((1,0)\) is empty then we have an occurrence of \(m_2\). Now look at
+the case when this region is not empty. Consider choosing the rightmost point in
+region.
+
+This gives us an occurrence of the following mesh pattern.
+\begin{equation*}
+    \pattern{}{2,1,3}{0/3,2/0,2/1,3/0,3/1,3/2}
+\end{equation*}
+By application of \PropositionRef{prop:dom1} we then achieve the following
+mesh pattern
+\begin{equation*}
+    \pattern{2,3}{2,1,3}{0/3,1/3,2/0,2/1,3/0,3/1,3/2}
+\end{equation*}
+If we look at the highlighted points we see that the subpattern is an occurrence
+of the mesh pattern that we originally desired. This gives rise to the following
+rule:
+
+\begin{proposition}[Third Dominating Pattern Rule]
+    \label{prop:dom3}
+    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
+    and a dominating classical pattern \(\pi = (\pi,\emptyset)\), the sets
+    \(\av{\{\pi,m_1\}}\) and \(\av{\{\pi,m_2\}}\) are coincident if
+    \begin{enumerate}
+        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
+        \item\label{prop:dom3:condocc} \(\addp[D]{\mperm{\sigma}{R_1}}{(a,b)}\)
+            where \(D\in\{N,E,S,W\}\)
+            is coincident with a mesh pattern containing an occurrence of
+            \(\mperm{\sigma}{R_2}\) as a subpattern.
+    \end{enumerate}
+\end{proposition}
+\begin{proof}
+    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).
+
+    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
+
+    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\).
+    By \NoteRef{not:downcmesh},
+     \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.
+
+    Now consider a permutation \(\varrho\) in \(\av{\pi}\) that contains an
+    occurrence of \(m_1\). If the region corresponding to the box \((a,b)\)
+    is empty then we have an occurrence of \(m_2\). If the region is non-empty
+    then by condition~\ref{prop:dom3:condocc}  of the proposition there exists
+    a direction such that there exists an occurrence of a mesh pattern of
+    length one longer than \(m_1\) in this position. This mesh pattern is
+    coincident with another mesh pattern that contains an
+    occurrence of \(m_2\). Hence, every occurrence of \(m_1\) leads to an occurrence
+    of \(m_2\).  Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the
+    two patterns are coincident.
+\end{proof}
+
+By application of this rule we can reduce the number of classes in
+\(\av{\{\perm{2,3,1},(\perm{1,2}, R)\}}\) to \(56\).
+
+\subsection{Equivalence classes of Av(\{321, (12, \textit{R})\}).}
+When considering coincidences of mesh patterns with underlying classical
+pattern \(\perm{1,2}\) in \(\av{\perm{3,2,1}}\) application of the previously
+established rules give no coincidences. Through experimentation we discover
+that there are \(7\) non-trivial coincidence classes (all others are singletons)
+which can be explained through the
+use of two different lines of reasoning. Since the number of coincidences is
+so small we will reason for these coincidences without attempting to generalise
+into concrete rules.
+
+Intuitively it is easy to see why our previous rules have no
+power here. There is nowhere that it is possible to add a single point to
+gain an occurrence of \(\pi = \perm{3,2,1}\). It is also impossible to have
+a position where addition of an increase, or decrease, provides extra
+shading power.
+
+The patterns
+\begin{equation*}
+    m_1 = \pattern{}{1,2}{0/0, 0/2, 1/1, 1/2, 2/1} \text { and } m_2 \pattern{}{1,2}{0/0, 0/1, 0/2, 1/1, 1/2, 2/1}
+\end{equation*}
+are equivalent in \(\av{\perm{3,2,1}}\). (There are 3 symmetries of these
+patterns that are also equivalent to each other by the same reasoning.)
+
+Consider the region corresponding to the box \((0,1)\) in any occurrence of \(m_1\),
+in a permutation. By \LemmaRef{lem:incdec} it must contain an increasing subsequence.
+If the region is empty then we have an occurrence of \(m_2\).
+If there is only one point in the region we can choose this to replace the \(1\) in
+the mesh pattern to get the required shading. If there is more than one point then
+choosing the two leftmost points gives us the following mesh pattern.
+\begin{equation*}
+    \pattern{3,4}{2,3,1,4}{0/0, 0/1, 0/2, 0/3, 0/4,
+                            1/0, 1/1, 1/2, 1/3, 1/4,
+                            2/0, 2/1,2/2, 2/4,
+                            3/1, 3/2, 3/3, 3/4,
+                            4/1, 4/2, 4/3}
+\end{equation*}
+Where the two highlighted points are the original two points.
+Now if we take the other two points as the points in our
+mesh pattern then we get an occurrence of the pattern
+we originally desired, and hence the two patterns are coincident.
+It is also possible to calculate this coincidence by an extension of the Third
+Dominating rule, where we allow a sequence of \texttt{add\_point} operations, this
+is discussed further in the future work section.
+
+The other reasoning applies to the patterns
+\begin{equation*}
+    m_1 = \pattern{}{1,2}{0/1,0/2,2/0,2/1} \text{ and } m_2 = \pattern{}{1,2}{0/1,0/2,1/1,2/0,2/1}
+\end{equation*}
+which are coincident by experimentation.
+
+In order to prove this coincidence we will proceed by mathematical
+induction on the number of points in region corresponding to the middle box.
+We call this number \(n\).
+\begin{description}
+    \item [Base Case \((n=0)\)] The base case holds since we can freely shade the
+        box if it contains no points.
+    \item [Inductive Hypothesis \((n=k)\)] Suppose that we can find an occurrence
+        of the second pattern if we have an occurrence of the first with \(k\) points
+        in the middle box.
+    \item [Inductive Step \((n=k+1)\)] Suppose that we have \((k+1)\) points in
+        the middle box. Choose the bottom most point in the middle box, this gives the
+        mesh pattern
+     \begin{equation*}
+        \begin{tikzpicture}[scale=0.5]
+            \modpattern[5]{}{1,2,3}{0/1,0/2,0/3,1/1,2/1,3/0,3/1,3/2}
+            \draw (2.5,0.5) node {\(X\)};
+        \end{tikzpicture}
+    \end{equation*}
+        Now we need to consider the box labelled \(X\). If this box is empty then
+        we have an occurrence of \(m_2\) and are done. If this box contains
+        any points then we gain some extra shading on the mesh pattern
+        due to the dominating pattern
+     \begin{equation*}
+        \begin{tikzpicture}[scale=0.5]
+            \modpattern[5]{2,3}{1,2,3}{0/1,0/2,0/3,1/1,1/2,1/3,2/1,3/0,3/1,3/2}
+            \filldraw (2.5,0.5) circle (4 pt);
+        \end{tikzpicture}
+    \end{equation*}
+        The two highlighted points form an occurrence of \(m_1\) with \(k\)
+        points in the middle box, and thus by the Inductive Hypothesis we
+        are done.
+\end{description}
+By induction we have that every occurrence of \(m_1\) leads to an occurrence of \(m_2\)
+and by \NoteRef{not:downcmesh} every occurrence of \(m_2\) is an occurrence of
+\(m_1\) so the two patterns are coincident.
+This argument applies to another two pairs of classes. Therefore in
+total in \(\av{\perm{3,2,1}, \mperm{1,2}{R}}\) there are \(213\) coincidence
+classes.

### 2017-03-07 writing/paper/sections/coincidences.tex
fix minor errors with math and spacing

@@
 sets to be coincident. We will fix \(\pi\) in order to define these coincidences and say that
-\(\pi\)is the \emph{dominating pattern}.
+\(\pi\) is the \emph{dominating pattern}.
 
@@
         \item \(R_1 \triangle R_2 = \{(a,b)\}\)
-        \item\label{prop:dom3:condocc} \(\addp[D]{\mperm{\sigma}{R_1}}{(a,b)}\)
-            where \(D\in\{N,E,S,W\}\)
+        \item\label{prop:dom3:condocc} One of \((\sigma,R_1)^\boksall{a,b}\)
             is coincident with a mesh pattern containing an occurrence of
@@
 It is also possible to calculate this coincidence by an extension of the Third
-Dominating rule, where we allow a sequence of \texttt{add\_point} operations, this
+Dominating rule, where we allow a sequence of point addition operations, this
 is discussed further in the future work section.

### 2017-03-07 writing/paper/sections/coincidences.tex
Make Changes suggest by Michael

@@
 p = \pattern{}{1,2}{0/1,2/2} \quad
-p^\boksrightup{2,1}
-= \pattern{}{1,3,2}{0/1,0/2,2/2,2/3,3/1,3/2,3/3}
+p^\boks{2,1}= \pattern{}{1,3,2}{0/1,0/2,2/3,3/3}
+\quad
+p^\boksup{2,1}
+= \pattern{}{1,3,2}{0/1,0/2,2/2,2/3,3/2,3/3}
 \end{equation*}
@@
 &\{\boks{k,\ell}\mid k > i, \ell > j, \boks{k-2,\ell-2}\in R\}\cup\\
-&\{\boks{i,j+1},\boks{i+1,j},\boks{i+1,j+1},\boks{i+1,j+2},\boks{i+2,j+1}\}
+&\{\boks{i+1,j},\boks{i+1,j+1},\boks{i+1,j+2}\}
 \end{aligned}
@@
 p^{\boks{1,1}_a}= \pattern{}{4,2,3,5,1}{0/0,
-                        1/0,1/2,1/5,
+                        1/0,1/5,
                         2/0,2/1,2/2,2/3,2/5,
-                        3/0,3/2,3/5,
+                        3/0,3/5,
                         5/0,5/1,5/2,5/3,5/5}
@@
 All coincidence classes of \(\av{\{\perm{3,2,1}, \mperm{2,1}{R}\}}\) can be
-explained by application of \PropositionRef{prop:dom1}. By experimentation
-we see that there are at least \(29\) coincidence classes, and all of these
-coincidences are explained by this Proposition.
+explained by application of \PropositionRef{prop:dom1}. Thus explaining the
+29 experimental coincidence classes opserved.
 
@@
 classes. Experimentation shows that there are in fact at least \(39\) coincidence
-classes, for example the following two patterns are coincident in
+classes, for example the following two patterns seem to be coincident in
 \(\av{\perm{2,3,1}}\) but this is not explained by \PropositionRef{prop:dom1}.
@@
 
-    Now consider if the region is non-empty, by \LemmaRef{lem:incdec} and
-    condition~\ref{prop:dom2:condc} of the proposition this region must contain
+    If the region is non-empty, then by \LemmaRef{lem:incdec} and
+    condition~\ref{prop:dom2:condc} of the proposition it must contain
     a decreasing subsequence.
@@
 
-    Similar arguments satisfy the remainder of the branches.
+    Similar arguments cover the remainder of the branches.
 \end{proof}
@@
 Rule together coincidences of classes of the form \(\av{\{\perm{2,3,1},\mperm{2,1}{R}\}}\)
-are fully explained, obtaining 39 coincidence classes of mesh patterns.
+are fully explained, obtaining 39 coincidence classes of mesh patterns, confirming experimental
+observations.
 
@@
 By application of this rule we can reduce the number of classes in
-\(\av{\{\perm{2,3,1},(\perm{1,2}, R)\}}\) to \(56\).
+\(\av{\{\perm{2,3,1},(\perm{1,2}, R)\}}\) to \(56\), which coincides with the 
+56 classes observed through experimentation
 
@@
         any points then we gain some extra shading on the mesh pattern
-        due to the dominating pattern
+        as any points in these boxes would create an occurrence of the dominating pattern \(\perm{2,3,1}\)
      \begin{equation*}

### 2017-03-07 writing/paper/sections/meshpatterns.tex
Make Changes suggest by Michael

@@
     A permutation \(\pi\in\mathfrak{S}_n\) contains the pattern \(\sigma\in\mathfrak{S}_k\) (\mbox{denoted} 
-    \(\sigma \preceq \pi\)) if there is some subsequence of indices of \(\pi, i_1i_2\dotsm{}i_k\)
-    such that the sequence \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is order isomorphic to
+    \(\sigma \preceq \pi\)) if there is some sequence\( i_1,i_2,\dotsm{},i_k\)
+    such that \(1\le i_1<i_2<\dotsm<i_k\le n\) and the sequence \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is order isomorphic to
     \(\sigma(1)\sigma(2)\dotsm\sigma(k)\). If \(\pi\) does not contain \(\sigma\), we say that 
@@
 \(\sigma = \perm{2,3,1}\), since the second, fourth and fifth elements (\(453\)) are order 
-isomorphic to \(\perm{231}\), it also contains the occurrence \(241\). The permutation 
+isomorphic to \(\perm{231}\), it also contains the occurrence \(241\) of the sam pattern. The permutation 
 \(\perm{2,4,1,5,3}\) avoids the pattern \(\perm{3,2,1}\).
@@
 \FigureRef{fig:containment} shows the containment of \(\sigma\) in \(\pi\) as in \ExampleRef{ex:contexmpl}.
+
+The boxes in the plot of a permutation are denoted by \(\boks{i,j}\), where the point \((i,j)\) is
+the lower left corner of the box.
 
@@
 \end{definition}
-We define containment (denoted \(p\preceq \pi\)), and avoidance, of the pattern \(p\) in the permutation
-\(\pi\) on mesh patterns analogously to classical containment, and avoidance,
-of \(\tau\) in \(\pi\) with the additional restrictions on the relative
-position of the occurrence of \(\tau\) in \(\pi\). These restrictions say that
-no elements of \(\pi\) are allowed in the regions of the plot corresponding
-to shaded boxes in the mesh. These boxes are denoted by \(\boks{i,j}\), where the point \((i,j)\) is
-the lower left corner of the box.
 
@@
 
-\noindent Firstly, \(\omega\) is an occurrence of \(\pi\) in the classical sense
+\noindent Firstly, \(\omega\) is an occurrence of \(\tau\) in the classical sense
   \begin{enumerate}[i.]

### 2017-03-07 writing/paper/sections/wilf.tex
Make Changes suggest by Michael

@@
 Wilf-equivalence is an important aspect to study in the field of permutation
-patterns. If two patterns \(\pi\) and \(\sigma\) are said to be Wilf-equivalent
+patterns. Two patterns \(\pi\) and \(\sigma\) are said to be Wilf-equivalent
 if their avoidance sets have the same size at each length. More formally:
@@
 
-Wilf-equivalence is of interest due to the fact that if two permutation classes
-are enumerated in the same way then there should exist a bijection between
-them, and therefore any combinatorial object they represent.
-
 Coincident patterns are trivially Wilf-equivalent, if \(\av[k]{R} = \av[k]{S}\)
@@
 and inverse operations (see \FigureRef{fig:symm}) preserve enumeration, and
-therefore these classes are trivially
+therefore classes related by these symmetries are trivially
 Wilf-equivalent.
@@
 be the usual Catalan generating function satisfying \(C(x) = 1 + xC(x)^2\).
-This is easy to see by structural decomposition around the maximum, as
+The fact that \(C(x)\) is teh generating function for \(\mathcal{C}\)
+can be seen by structural decomposition around the maximum, as
 shown in \FigureRef{fig:decompmax}.
@@
 This can be explained succinctly by the fact that a permutation containing
-\(q_1\) starts with it's maximum, by not allowing the lower-left section of the
+\(q_1\) starts with its maximum, by not allowing the lower-left section of the
 \(\perm{2,3,1}\) avoider to be empty we prevent an occurrence from ever
@@
 Consider \(q_2 = \textpattern{}{1}{0/1,1/0,1/1}\), avoiding this pattern
-means that a permutation does not end with it's maximum. We can perform a
+means that a permutation does not end with its maximum. We can perform a
 similar
@@
 \(r = \textpattern{}{1}{0/0,1/0,1/1}\) is avoided by all permutations
-that do not end in their minimum. Considering the standard decomposition
-of a \(\perm{2,3,1}\) avoider around the maximum we can see that an avoider
-of \(r\) must fit into precisely one of the following two forms.
+that do not end in their minimum. Any avoider of \(\perm{2,3,1}\) that
+ends in its minimum must be a decreasing sequence
+Therefore this particular class has equation
 \begin{equation*}
-    \scriptvar = \varepsilon \sqcup
-    \underbrace{\decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}}_\text{Minimum comes before the maximum.} \sqcup
-\underbrace{\decompmax{0/0,0/1,1/0}{}{}{}{\scriptstyle{\scriptvar\setminus\varepsilon}}}_\text{
-        Minimum is after the maximum, minimum cannot be last.
-    }
+    \var(0) = 1, \var(x) = C(x)-1
 \end{equation*}
-Therefore this particular class has generating function \(\var(x)\) satisfying
-\begin{equation*}
-    \var(x) = 1 + xC(x)(C(x)-1) + x(\var(x)-1)
-\end{equation*}
-Computing coefficients of this generating function gives
+Computing these values gives
 \begin{equation*}
@@
 
+\begin{definition} A \emph{Set Partition} of \(\nrange{n}\) is a set of non-empty subsets of \(\nrange{n}\)
+such that \(\nrange{n}\) is a \emph{disjoint} union of these subsets.
+\end{definition}
+
 \begin{note}
@@
 \end{note}
+\begin{example}
+    The permutation \(\pi = \perm{5,4,2,1,3,9,6,8,7}\) corresponds to the
+    partition \(\{\{5,4,2,1\},\{3\},\{9,6\},\{8,7\}\}\).
+\end{example}
+
 We will call the least element in each block the \emph{block-bottom}, and note that all permutations
@@
 is a mesh pattern we can use similar methods to allow us to consolidate experimental classes into
-actual classes.
+actual classes, these methods allow us to explain all 23 of the observed Wilf-classes seen in experimentation.

### 2017-04-05 writing/paper/sections/abstract.tex
Finish writing paper v1
Comments from Henning up to section 4

--- 2016-10-24 writing/paper/abstract.tex
@@
-\begin{abstract}
-    Two mesh patterns are coincident if they are avoided by the same set of permutations, and are
-    Wilf-equivalent if they have the same number of avoiders at each length. We provide
-    sufficient conditions for coincidence among mesh patterns, whilst also avoiding a longer
-    classical pattern. Using these conditions we completely classify coincidences between families
-    containing a mesh pattern of length 2 and a classical pattern of length 3. Furthermore, we
-    completely Wilf-classify equivalences of mesh patterns of length 2 whilst also avoiding the
-    classical pattern 231.
+\begin{abstract} Two mesh patterns are coincident if they are avoided by the
+same set of permutations, and are Wilf-equivalent if they have the same number
+of avoiders of each length. We provide sufficient conditions for coincidence
+among mesh patterns, whilst also avoiding a longer classical pattern. Using
+these conditions we completely classify coincidences between families containing
+a mesh pattern of length 2 and a classical pattern of length 3. Furthermore, we
+completely Wilf-classify equivalences of mesh patterns of length 2 whilst also
+avoiding the classical pattern 231.
 
-    \noindent \\
-    \emph{Keywords: } permutation, pattern, mesh pattern, pattern coincidence
-\end{abstract}
+    \noindent \\ \emph{Keywords: } permutation, pattern, mesh pattern, pattern
+    coincidence \end{abstract}

### 2017-04-05 writing/paper/sections/coincidences.tex
Comments from Henning up to section 4

@@
 
-Two patterns \(\lambda\) and \(\gamma\) are considered \emph{coincident} if the set of permutations
+Recall that two patterns \(\lambda\) and \(\gamma\) are considered \emph{coincident} if the set of permutations
 that avoid \(\lambda\) is the same as the set of permutations that avoid \(\gamma\),
 \ie \(\av{\lambda} = \av{\gamma}\). Equivalently we can say that they have the same set of
-\emph{containers},\ie \(\cont{\lambda} = \cont{\gamma}\).
+\emph{containers}, \ie \(\cont{\lambda} = \cont{\gamma}\).
 
 We will consider the avoidance sets \(\av{\pi,p}\) where \(\pi\) is a classical pattern of length 3
-and \(p\) is a mesh pattern of length 2 in order to establish sufficient conditions for two such
-sets to be coincident. We will fix \(\pi\) in order to define these coincidences and say that
-\(\pi\) is the \emph{dominating pattern}.
+and \(p\) is a mesh pattern of length 2 and establish sufficient conditions for two such
+sets to be coincident. The classical pattern \(\pi\) will be fixed and called
+the \emph{dominating pattern}.
 
@@
 In addition, we give the following definitions:
+
 \begin{definition}
 Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\) and
-\(p^\boks{i,j} = (\tau^\prime,R^\prime)\) to be as defined in \DefinitionRef{def:ap}. We define
-the following five modifications of the mesh patterns of the same length as \(p^\boks{i,j}\).
+\(p^\boks{i,j} = (\tau^\prime,R^\prime)\) is as defined in \DefinitionRef{def:ap}. We define
+the following four modifications of \(p^\boks{i,j}\).
 \begin{align*}
@@
 Informally, these are considering the topmost, rightmost, leftmost, or bottommost point in
-\(\boks{i,j}\). We allow composition of these modifications, and collect the resulting mesh patterns
-in a set
+\(\boks{i,j}\). We collect the resulting mesh patterns in a set
 \begin{equation*}
@@
 \end{equation*}
+We allow composition of these modifications. % TODO Are they ever used?
 See \FigureRef{fig:addp} for an example of adding a point into a mesh pattern.
@@
 finding and explaining all coincidences where \(\av{\{p,m\}} = \av{\{p,m^\prime\}}\).
+% TODO There is some inconsistency in when you put the set brackets into Av
 
@@
 We know that there are a total of \(512\) mesh-sets for each underlying classical
-pattern. By use of the previous results of \textcite{DBLP:journals/corr/ClaessonTU14}\footnote{
-The authors use the Simultaneous Shading Lemma, a closure result and one worked out special case.
-} the number of coincidence classes can be reduced to \(220\).
+pattern. By use of the previous results of \textcite{DBLP:journals/corr/ClaessonTU14}
+the number of coincidence classes can be reduced to \(220\).
 
@@
     \begin{enumerate}
-        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
+        \item The symmetric difference, \(R_1 \triangle R_2\), is \(\{(a,b)\}\)
         \item \(\pi \preceq \sigma^\boks{a,b}\)\label{prop:dom1:cont}
@@
 \end{proposition}
-In order to prove this proposition we must first make the following note.
-
-\begin{note}
-    \label{not:downcmesh}
-    Let \(R^\prime \subseteq R\). Then any occurrence of \((\tau, R)\) in a permutation
-    is an occurrence of \((\tau,R^\prime)\).
-\end{note}
-
-\begin{proof}[Proof of \PropositionRef{prop:dom1}]
-    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).
-
-    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
-    Since \(R_1\) is a subset of \(R_2\), \NoteRef{not:downcmesh} states that
-    \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\)
-
-    Now we consider a permutation \(\omega^\prime\in \av{\pi}\),
-    containing an occurrence of \(m_1\). Consider placing a point in the
-    region corresponding to the box \((a,b)\), regardless of where in this
-    region we place the point by condition~\ref{prop:dom1:cont} of the
-    Proposition we create an occurrence of \(\pi\), therefore there can be no
-    points in this region, which could have been represented in the mesh set
-    \(R_1\) by adding the box \((a,b)\). Hence every occurrence of
-    \(m_1\) is in fact an occurrence of \(m_2\), and we have that
-    \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).
-
-    Taking both directions of the containment we can therefore draw the
-    conclusion that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).
-\end{proof}
-
-All coincidence classes of \(\av{\{\perm{3,2,1}, \mperm{2,1}{R}\}}\) can be
-explained by application of \PropositionRef{prop:dom1}. Thus explaining the
-29 experimental coincidence classes opserved.
 
@@
 
+In order to prove the proposition we must first make the following note.
+
+\begin{note}
+    \label{not:downcmesh}
+    Let \(R_1 \subseteq R_2\). Then any occurrence of \((\tau, R_2)\) in a permutation
+    is an occurrence of \((\tau,R_1)\).
+\end{note}
+
+\begin{proof}[Proof of \PropositionRef{prop:dom1}]
+    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).
+
+    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
+    Since \(R_1\) is a subset of \(R_2\), \NoteRef{not:downcmesh} states that
+    \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\)
+
+    Now we consider a permutation \(\omega \in \av{\pi}\),
+    containing an occurrence of \(m_1\). Consider placing a point in the
+    region corresponding to the box \((a,b)\), regardless of where in this
+    region we place the point by condition~\ref{prop:dom1:cont} of the
+    proposition we create an occurrence of \(\pi\). Therefore there can be no
+    points in this region, which implies this is an occurrence of $m_2$. Hence every occurrence of
+    \(m_1\) is in fact an occurrence of \(m_2\), and we have that
+    \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).
+
+    Taking both directions of the containment we can therefore draw the
+    conclusion that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).
+\end{proof}
+
+All coincidence classes of \(\av{\{\perm{3,2,1}, \mperm{2,1}{R}\}}\) can be
+explained by application of \PropositionRef{prop:dom1}. Thus explaining the
+29 experimental coincidence classes opserved.
+
 \subsection{Equivalence classes of Av(\{231, (21, \textit{R})\}).}
@@
     \label{lem:incdec}
-    Given a mesh pattern \(m =(\sigma, R)\), where the box \((a,b)\) is not
-    in \(R\), and a dominating classical pattern \(\pi = (\pi,\emptyset)\) if
-    \(\pi \preceq p^{\boks{i,j}_a}\)\\(\(\pi \preceq p^{\boks{i,j}_d}\))
-    ,then in any occurrence of \(m\) in a permutation \(\varrho\), the region
+    Let \(m =(\sigma, R)\) be a mesh pattern, where the box \((a,b)\) is not
+    in \(R\), and \(\pi = (\pi,\emptyset)\) be a dominating classical pattern. If
+    \(\pi \preceq p^{\boks{i,j}_a}\)\\(\(\pi \preceq p^{\boks{i,j}_d}\)),
+    then in any occurrence of \(m\) in a permutation \(\varrho\), the region
     corresponding to the box \((a,b)\) can only contain an decreasing
@@
 \end{equation*}
-We know that the region corresponding to the box \((1,1)\) contains a decreasing subsequence.
+we know that the region corresponding to the box \((1,1)\) contains a decreasing subsequence.
 If we let \(z\) be the topmost point in this decreasing subsequence, then \(xz\)
@@
 \begin{proof}
-    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).
-
-    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
-
-    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\).
-    By \NoteRef{not:downcmesh} any of these occurrences is also an occurrence
-    of \(m_1\). This proves that every occurrence of \(m_2\) is also an
-    occurrence of \(m_1\) and therefore \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).
-
-    We will consider taking the first branch of every choice.
-    Now consider a permutation \(\omega^\prime \in \av{\pi}\). Suppose
-    \(\omega^\prime\) contains \(m_1\) and consider the region corresponding to
-    \((a,b)\) in \(R_1\).
-
-    If the region is empty, the occurrence of \(m_1\) is trivially an occurrence
+    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
+    By \NoteRef{not:downcmesh} we only need to show that an occurrence of \(m_1\)
+    implies and occurrence of \(m_2\).
+    We consider taking the first branch of every choice.
+    Now consider a permutation \(\omega \in \av{\pi}\). Suppose
+    \(\omega\) contains \(m_1\) and consider the region corresponding to
+    \((a,b)\) in \(R_1\).If the region is empty, the occurrence of \(m_1\) is trivially an occurrence
     of \(m_2\).
-
     If the region is non-empty, then by \LemmaRef{lem:incdec} and
@@
 
-By taking the First Dominating Pattern Rule and the Second Dominating Pattern
-Rule together coincidences of classes of the form \(\av{\{\perm{2,3,1},\mperm{2,1}{R}\}}\)
-are fully explained, obtaining 39 coincidence classes of mesh patterns, confirming experimental
-observations.
+By applying both the First Dominating Pattern Rule and the Second Dominating
+Pattern Rule fully explains coincidences of classes of the form
+\(\av{\{\perm{2,3,1},\mperm{2,1}{R}\}}\), obtaining 39 coincidence classes of
+mesh patterns, confirming experimental observations.
 
@@
 \end{equation*}
-are experimentally coincident. This coincidence is not explained by our rules,
-it is necessary to attempt to capture these coincidences by establishing more
-rules.
+are experimentally coincident.
 
@@
 to the box \((1,0)\) is empty then we have an occurrence of \(m_2\). Now look at
-the case when this region is not empty. Consider choosing the rightmost point in
-region.
+the case when this region is not empty, adn consider choosing the rightmost
+point in the region.
 
@@
         \item \(R_1 \triangle R_2 = \{(a,b)\}\)
-        \item\label{prop:dom3:condocc} One of \((\sigma,R_1)^\boksall{a,b}\)
+        \item\label{prop:dom3:condocc} One of the patterns in \((\sigma,R_1)^\boksall{a,b}\)
             is coincident with a mesh pattern containing an occurrence of
@@
 \begin{proof}
-    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).
+    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
+    \NoteRef{not:downcmesh}, implies \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.
 
-    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
-
-    Consider a permutation \(\omega\) that contains an occurrence of \(m_2\).
-    By \NoteRef{not:downcmesh},
-     \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.
-
-    Now consider a permutation \(\varrho\) in \(\av{\pi}\) that contains an
+    Now consider a permutation \(\sigma\) in \(\av{\pi}\) that contains an
     occurrence of \(m_1\). If the region corresponding to the box \((a,b)\)
@@
     then by condition~\ref{prop:dom3:condocc}  of the proposition there exists
-    a direction such that there exists an occurrence of a mesh pattern of
+    a pattern in \((\sigma,R_1)^\boksall{a,b}\) such that there exists an occurrence of a mesh pattern of
     length one longer than \(m_1\) in this position. This mesh pattern is
@@
 \(\av{\{\perm{2,3,1},(\perm{1,2}, R)\}}\) to \(56\), which coincides with the
-56 classes observed through experimentation
+56 classes observed through experimentation.
 
@@
     \item [Inductive Step \((n=k+1)\)] Suppose that we have \((k+1)\) points in
-        the middle box. Choose the bottom most point in the middle box, this gives the
+        the middle box. Choose the bottom most point in the middle box, which gives the
         mesh pattern

### 2017-04-05 writing/paper/sections/introduction.tex
Make Changes suggest by Michael
Comments from Henning up to section 4

@@
 The study of permutation patterns began as a result of Knuth's statements on
-stack sorting in \emph{The Art of Computer Programming}\cite[p.~243,
+stack sorting in \emph{The Art of Computer Programming}~\cite[p.~243,
 Ex.~5,6]{Knuth:1997:ACP:260999}. This original concept---a subsequence of
@@
 \emph{generalised} or \emph{dashed} patterns) where two adjacent entries in the
-pattern must also be adjacent in the permutation. \textcite{MR2652101} look at
-classes of pattern where both columns and rows can be shaded, these are called
-\emph{bivincular} patterns. \emph{Bruhat-restricted} patterns were studied by
-\textcite{MR2264071} to establish necessary conditions for a Schubert variety to
-be Gorenstein. These definitions are subsumed under the definition of \emph{mesh
-patterns}, introduced by \textcite{journals/combinatorics/BrandenC11} to capture
-explicit expansions for certain permutation statistics.
+pattern can be required to be adjacent in the permutation. \textcite{MR2652101}
+look at classes of pattern where entries can be required to be consecutive in
+value, these are called \emph{bivincular} patterns. \emph{Bruhat-restricted}
+patterns were studied by \textcite{MR2264071} to establish necessary conditions
+for a Schubert variety to be Gorenstein. These definitions are subsumed under
+the definition of \emph{mesh patterns}, introduced by
+\textcite{journals/combinatorics/BrandenC11} to capture explicit expansions for
+certain permutation statistics.
 
@@
 begin by establishing coincidences between mesh patterns of length 2 while
-avoiding a classical pattern of length 3, this is used to establish sufficient
+avoiding a classical pattern of length 3; this is used to establish sufficient
 conditions for coincidence. We then establish Wilf-equivalence classes of these
-coincidence classes who avoid the classical pattern 231.
+coincidence classes where the classical pattern 231.

### 2017-04-05 writing/paper/sections/meshpatterns.tex
Comments from Henning up to section 4

@@
 A \emph{permutation} is a bijection from the set \(\nrange{n} = \setrange{1}{n}\) to itself. The
-set of all such bijections on this set if denoted \(\mathfrak{S}_n\) and has size \(n!\). We can
+set of all such bijections is denoted \(\mathfrak{S}_n\) and has \(n!\) elements. We can
 denote an individual permutation \(\pi\in\mathfrak{S}_n\) in \emph{one-line notation} by writing the
 entries of the permutation in order, therefore \(\pi = \perm{\pi(1),\pi(2),\dotso,\pi(n)}\). The
-set \(\mathfrak{S}_n\) has exactly one element, the empty permutation \(\varepsilon\).
+set \(\mathfrak{S}_0\) has exactly one element, the empty permutation \(\varepsilon\).
 
 \begin{definition}{(Order isomorphism.)}
-    Two substrings \(\alpha_1\alpha_2\dotsm\alpha_n\) and
+    Two strings \(\alpha_1\alpha_2\dotsm\alpha_n\) and
     \(\beta_1\beta_2\dotsm\beta_n\) are said to be \emph{order isomorphic}
@@
 \begin{definition}
-    A permutation \(\pi\in\mathfrak{S}_n\) contains the pattern \(\sigma\in\mathfrak{S}_k\) (\mbox{denoted} 
-    \(\sigma \preceq \pi\)) if there is some sequence\( i_1,i_2,\dotsm{},i_k\)
+    A permutation \(\pi\in\mathfrak{S}_n\) \emph{contains} the permutation \(\sigma\in\mathfrak{S}_k\) (\mbox{denoted}
+    \(\sigma \preceq \pi\)) if there is some sequence \( i_1,i_2,\dotsm{},i_k\)
     such that \(1\le i_1<i_2<\dotsm<i_k\le n\) and the sequence \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is order isomorphic to
     \(\sigma(1)\sigma(2)\dotsm\sigma(k)\). If \(\pi\) does not contain \(\sigma\), we say that
-    \(\pi\) \emph{avoids} \(\sigma\).
+    \(\pi\) \emph{avoids} \(\sigma\). If this is the case the sequence
+    \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is called an \emph{occurrence} of \(\sigma\)
+    in \(\pi\). In this context \(\sigma\) is called a (\emph{classical})
+    \emph{permutation pattern}.
 \end{definition}
+
 \begin{example}
@@
 \(\sigma = \perm{2,3,1}\), since the second, fourth and fifth elements (\(453\)) are order
-isomorphic to \(\perm{231}\), it also contains the occurrence \(241\) of the sam pattern. The permutation 
+isomorphic to \(\perm{231}\). The permutation also contains the occurrence \(241\) of the same pattern. The permutation
 \(\perm{2,4,1,5,3}\) avoids the pattern \(\perm{3,2,1}\).
@@
 We denote the set of permutations of length \(n\) avoiding a pattern \(\sigma\) as \(\Av_n(\sigma)\)
-and \(\av{\sigma}=\bigcup_{i=0}^{\infty}\Av_i(\sigma)\).
+and let \(\av{\sigma}=\bigcup_{i=0}^{\infty}\Av_i(\sigma)\).
 
-We can display a permutation graphically in a \emph{plot}, in such a plot we display the points
-\(G(\pi) = \{(i,\pi(i))\mid i\in[1,n]\}\) in a Cartesian coordinate system. The plots of the permutations \(\pi = 
+We can display a permutation graphically in a \emph{plot}, where we display the points
+\(G(\pi) = \{(i,\pi(i))\mid i \in \nrange{n} \}\) in a Cartesian coordinate system. The plots of the permutations \(\pi =
 \perm{2,4,1,5,3}\) and \(\sigma = \perm{2,3,1}\) can be seen in \FigureRef{fig:plots}.
@@
 an \emph{occurrence} of \(p\) in \(\pi\) is a subset \(\omega\) of the plot of
-\(\pi, G(\pi) = \{(i,\pi(i)\mid i\in[1,n]\}\) such that there are order-preserving
-injections \(\alpha,\beta:[1,k]\mapsto[1,n]\) satisfying the following two conditions.
+\(\pi, G(\pi) = \{(i,\pi(i)\mid i\in\nrange{n}\}\) such that there are order-preserving
+injections \(\alpha,\beta:\nrange{k}\mapsto\nrange{n}\) satisfying the following two conditions.
 
@@
   \begin{enumerate}[i.]
-    \item \(\omega = \{(\alpha(i),\beta(j)):(i,j)\in G(\tau)\}\)
+    \item \(\omega = \{(\alpha(i),\beta(j)):(i,j)\in G(\tau)\}\).
   \end{enumerate}\saveenum
@@
   \begin{enumerate}[i.]\resetenum
-  \item if \(\boks{i,j} \in R \text{ then } R_{ij} \cap G(\pi) = \emptyset\)
+  \item if \(\boks{i,j} \in R \text{ then } R_{ij} \cap G(\pi) = \emptyset\).
 \end{enumerate}
-We call \(R_{ij}\) the region corresponding to \(\boks{i,j}\).
-We define containment of a mesh pattern \(p\) in another mesh pattern \(q\) as above,
-with the additional condition that if \(\boks{i,j}\in R\text{ then } R_{ij}\) is
-contained in the mesh set of \(q\), in this case we call \(p\) a \emph{subpattern}
-of \(q\).
-\begin{definition}
-A mesh pattern \(q=(\kappa,T)\) \emph{contains} a mesh pattern \(p=(\tau,R)\) as a \emph{subpattern}
-if \(\kappa\) contains \(\tau\) and \(\left(\bigcup_{\boks{i,j}\in R} R_{ij}\right)\subseteq T\).
-\end{definition}
+We call \(R_{ij}\) the \emph{region corresponding to} \(\boks{i,j}\).
 
@@
 \end{equation*}
-The subsequence \(325\) is therefore an occurrence of the pattern \(p\) in \(\pi\) and
-\(\pi\) contains \(p\).
+The subsequence \(325\) is therefore an occurrence of the pattern \(p\) in \(\pi\).
+
+We define containment of a mesh pattern \(p\) in another mesh pattern \(q\) as above,
+with the additional condition that if \(\boks{i,j}\in R\text{ then } R_{ij}\) is
+contained in the mesh set of \(q\).
+
+\begin{definition}
+A mesh pattern \(q=(\kappa,T)\) \emph{contains} a mesh pattern \(p=(\tau,R)\) as a \emph{subpattern}
+if \(\kappa\) contains \(p\) and \(\left(\bigcup_{\boks{i,j}\in R} R_{ij}\right)\subseteq T\).
+\end{definition}
+
+TODO: I changed a symbol in the above definition. See the comment.
+% if \(\kappa\) contains \(\tau\)
+% to
+% if \(\kappa\) contains \(p\)
+% that is what we want right?
+
+TODO: Change the following example to an example of a mesh pattern in a mesh
+pattern.
 

### 2017-04-05 writing/paper/sections/conclfut.tex
Make Changes suggest by Michael
All comments from Henning. See TODOs

@@
-If we consider a similar system for dominating patterns of length \(4\) and
+If we consider a similar approach to dominating patterns of length \(4\) and
 mesh patterns of length \(2\), it can be seen that the number of cases required
@@
    an occurrence of \(m_2\).
-   Consider if the region corresponding to \(X\) is non-empty, we can then choose
+   If the region corresponding to \(X\) is non-empty, we can then choose
    the lowest valued point in this region
@@
    And now the two indicated points form an occurrence of \(m_2\).
-   We have therefore shown that any occurrence of \(m_1\) is an occurrence of
+   We have therefore shown that any occurrence of \(m_1\) leads to an occurrence of
    \(m_2\) and we can easily show the converse by the same reasoning, so \(m_1\)
@@
 It would be interesting to consider a systematic explanation of Wilf-equivalences among classes
-where \(\perm{2,3,1}\) is the dominating pattern, possibly using the  construction presented in
-\cite[Sec.~12]{2015arXiv151203226B}, in order to directly reach enumeration and hopefully establish
+where \(\perm{3,2,1}\) is the dominating pattern, possibly using the  construction presented in
+Bean, Claesson and Ulfarsson~\cite[Sec.~12]{2015arXiv151203226B}, in order to directly reach enumeration and hopefully establish
 some of the non-trivial Wilf-equivalences between classes with different dominating patterns.

### 2017-04-05 writing/paper/sections/wilf.tex
All comments from Henning. See TODOs

@@
 Wilf-equivalence is an important aspect to study in the field of permutation
-patterns. Two patterns \(\pi\) and \(\sigma\) are said to be Wilf-equivalent
-if their avoidance sets have the same size at each length. More formally:
+patterns.
 
@@
 
-Coincident patterns are trivially Wilf-equivalent, if \(\av[k]{R} = \av[k]{S}\)
+Coincident patterns are trivially Wilf-equivalent: if \(\av[k]{R} = \av[k]{S}\)
 then trivially \(\size{\av[k]{R}} = \size{\av[k]{S}}\). Coincidence is
@@
 
-Since we are always considering Wilf-equivalences in the set \(\av{S}\) we
+Since we are always considering Wilf-equivalences in a set \(\av{S}\) we
 must only use these symmetries when they preserve the dominating pattern, if we
@@
 be the usual Catalan generating function satisfying \(C(x) = 1 + xC(x)^2\).
-The fact that \(C(x)\) is teh generating function for \(\mathcal{C}\)
+The fact that \(C(x)\) is the generating function for \(\mathcal{C}\)
 can be seen by structural decomposition around the maximum, as
@@
 This gives the generating function of avoiders to be the function \(\var(x)\)
-satisfying.
+satisfying
 \begin{align*}
@@
     \var(x) &= \frac{C(x)}{1+xC(x)}
-\end{align*}
+\end{align*} % TODO Do we need both representations?
 
+\noindent
 Calculating coefficients given by this generating function gives the Fine
@@
 This can be explained succinctly by the fact that a permutation containing
-\(q_1\) starts with its maximum, by not allowing the lower-left section of the
+\(q_1\) starts with its maximum, and by not allowing the lower-left section of the
 \(\perm{2,3,1}\) avoider to be empty we prevent an occurrence from ever
@@
 
-Consider \(q_2 = \textpattern{}{1}{0/1,1/0,1/1}\), avoiding this pattern
+Consider \(q_2 = \textpattern{}{1}{0/1,1/0,1/1}\). Avoiding this pattern
 means that a permutation does not end with its maximum. We can perform a
@@
 
-Now consider \(q_3 = \textpattern{}{1}{0/0,0/1,1/0}\), the avoiders of this
+% TODO Isn't q_3 a symmetry of q_2? So we don't need to repeat the work
+Now consider \(q_3 = \textpattern{}{1}{0/0,0/1,1/0}\). The avoiders of this
 pattern are permutations that do not start with their minimum. In this case we
@@
 that do not end in their minimum. Any avoider of \(\perm{2,3,1}\) that
-ends in its minimum must be a decreasing sequence
+ends in its minimum must be a decreasing sequence.
 Therefore this particular class has equation
@@
 \section{Wilf-classes with patterns of length 2}
-By the use of set equivalences established in
+By the use of coincidence classes established in
 \SectionRef{sec:coincs} we know that there are at most 95 Wilf-equivalence classes.
@@
 Computing avoiders up to length 10 suggests that there are at least 23 Wilf-classes,
-of which 13 are non-trivial, therefore there are Wilf-equivalences that are not
+of which 13 are non-trivial. Therefore there are Wilf-equivalences that are not
 explained by coincidences or symmetry.
@@
 
-\begin{definition} A \emph{Set Partition} of \(\nrange{n}\) is a set of non-empty subsets of \(\nrange{n}\)
+\begin{definition} A \emph{set partition} of \(\nrange{n}\) is a set of non-empty subsets of \(\nrange{n}\)
 such that \(\nrange{n}\) is a \emph{disjoint} union of these subsets.
@@
 We will call the least element in each block the \emph{block-bottom}, and note that all permutations
-in \(\av{\perm{2,3,1}}\) also avoid this mesh pattern.
+in \(\av{\perm{2,3,1}}\) also avoid the mesh pattern \(q\).
 
@@
 First we prove the Wilf-equivalence between \(m_1\) and \(m_2\) shown in
-\eqref{eq:singleud}. The easiest way to
-show that these are equinumerous is to consider the containers as set partitions.
+\eqref{eq:singleud}, by considering the containers as set partitions consider
+the containers as set partitions.
 
@@
 \end{equation*}
-Any occurrence of \(m_4\) can never occur in the
+An occurrence of \(m_4\) can never occur in the
 top right region. It could only occur between the maximum and the first point
@@
 \end{equation*}
-Where \(\varmaxl(x)\) is the generating function given in \EquationRef{eqn:maxlgen}.
+where \(\varmaxl(x)\) is the generating function given in \EquationRef{eqn:maxlgen}.
 This can be enumerated to give the sequence
@@
 \end{equation*}
-Where \(\scriptvar^\prime_1\) is a permutation avoiding \(\perm{2,3,1}, m_1\)
+
+\noindent
+Here \(\scriptvar^\prime_1\) is a permutation avoiding \(\perm{2,3,1}, m_1\)
 and \(\textpattern{}{1}{0/0,0/1,1/0}\).
@@
 \end{equation*}
-This is an offset of the Motzkin numbers.
+Which is an offset of the Motzkin numbers.
 

### 2017-04-05 writing/paper/sections/coincidences.tex
Missing shading in a pattern
Fixed adn
Do most of Henning's changes
- Change makefile to output to pdf
- Never need compositions of boksall, so removed.
- fixed consistency with av brackets (I hope)
- fixed Henning  using dollar signs in his edits
- changed Henning's long-form cite to nice automatic textcite
- removed duplicate line about considering as set partitions

@@
 
-We will consider the avoidance sets \(\av{\pi,p}\) where \(\pi\) is a classical pattern of length 3
-and \(p\) is a mesh pattern of length 2 and establish sufficient conditions for two such
+We will consider the avoidance sets \(\av{\{\pi,p\}}\) where \(\pi\) is a classical pattern of length 3
+and \(p\) is a mesh pattern of length 2 to establish sufficient conditions for two such
 sets to be coincident. The classical pattern \(\pi\) will be fixed and called
@@
 p^\boksright{i,j},
-% p^\boksrightup{i,j},
-% p^\boksrightdown{i,j},
 p^\boksleft{i,j},
-% p^\boksleftup{i,j},
-% p^\boksleftdown{i,j},
 p^\boksup{i,j},
@@
 \end{equation*}
-We allow composition of these modifications. % TODO Are they ever used?
 See \FigureRef{fig:addp} for an example of adding a point into a mesh pattern.
@@
 \begin{equation*}
-%\adda{\pattern{}{2,3,1}{0/0,1/0,1/3,
-%                        3/0,3/1,3/3}}{(1,1)}
 p = \pattern{}{1,2}{0/1,2/2} \quad
@@
 finding and explaining all coincidences where \(\av{\{p,m\}} = \av{\{p,m^\prime\}}\).
-% TODO There is some inconsistency in when you put the set brackets into Av
 
@@
 classical pattern, this is due to the fact that \(\perm{2,1}\in\av{\mperm{1,2}{R}}\)
-and \(\perm{1,2}\in\av{\mperm{2,1}{R}}\) for all mesh-sets \(R\).
+and \(\perm{1,2}\in\av{\{\mperm{2,1}{R}\}}\) for all mesh-sets \(R\).
 
@@
     proposition we create an occurrence of \(\pi\). Therefore there can be no
-    points in this region, which implies this is an occurrence of $m_2$. Hence every occurrence of
+    points in this region, which implies this is an occurrence of \(m_2\). Hence every occurrence of
     \(m_1\) is in fact an occurrence of \(m_2\), and we have that
@@
     \begin{tikzpicture}[scale=\picscale]
-        \modpattern{}{2,1}{0/0,0/2,1/0,2/1}
+        \modpattern{}{2,1}{0/0,0/1,0/2,1/0,2/1}
         \filldraw (1.66,1.33) circle (3pt);
@@
 \subsection{Equivalence classes of Av(\{231, (12, \textit{R})\}).}
-When considering the coincidence classes of \(\av{\perm{2,3,1},\mperm{1,2}{R}}\)
+When considering the coincidence classes of \(\av{\{\perm{2,3,1},\mperm{1,2}{R}\}}\)
 we first apply the two Dominating Pattern rules previously established.
@@
 to the box \((1,0)\) is empty then we have an occurrence of \(m_2\). Now look at
-the case when this region is not empty, adn consider choosing the rightmost
+the case when this region is not empty, and consider choosing the rightmost
 point in the region.
@@
     \item [Inductive Step \((n=k+1)\)] Suppose that we have \((k+1)\) points in
-        the middle box. Choose the bottom most point in the middle box, which gives the
+        the middle box. Choose the bottom most point in the middle box, giving the
         mesh pattern
@@
 This argument applies to another two pairs of classes. Therefore in
-total in \(\av{\perm{3,2,1}, \mperm{1,2}{R}}\) there are \(213\) coincidence
+total in \(\av{\{\perm{3,2,1}, \mperm{1,2}{R}\}}\) there are \(213\) coincidence
 classes.

### 2017-04-05 writing/paper/sections/conclfut.tex
Do most of Henning's changes
- Change makefile to output to pdf
- Never need compositions of boksall, so removed.
- fixed consistency with av brackets (I hope)
- fixed Henning  using dollar signs in his edits
- changed Henning's long-form cite to nice automatic textcite
- removed duplicate line about considering as set partitions

@@
 where \(\perm{3,2,1}\) is the dominating pattern, possibly using the  construction presented in
-Bean, Claesson and Ulfarsson~\cite[Sec.~12]{2015arXiv151203226B}, in order to directly reach enumeration and hopefully establish
+\textcite[Sec.~12]{2015arXiv151203226B}, in order to directly reach enumeration and hopefully establish
 some of the non-trivial Wilf-equivalences between classes with different dominating patterns.

### 2017-04-05 writing/paper/sections/wilf.tex
Do most of Henning's changes
- Change makefile to output to pdf
- Never need compositions of boksall, so removed.
- fixed consistency with av brackets (I hope)
- fixed Henning  using dollar signs in his edits
- changed Henning's long-form cite to nice automatic textcite
- removed duplicate line about considering as set partitions

@@
 \(\av{\perm{2,3,1}}\), this means that we are considering the set
-\(\av{231,p}\)
+\(\av{\{\perm{2,3,1},p}\}\)
 where \(p\) is a mesh-pattern of length 1.
@@
 First we prove the Wilf-equivalence between \(m_1\) and \(m_2\) shown in
-\eqref{eq:singleud}, by considering the containers as set partitions consider
-the containers as set partitions.
+\eqref{eq:singleud}, by considering the containers as set partitions.
 
@@
 consider the decomposition around the leftmost point of a permutation in
-\(\av{\perm{2,3,1},m_1}\). We have the following
+\(\av{\{\perm{2,3,1},m_1\}}\). We have the following
 \begin{equation*}
@@
 
-In order to establish the remainder of the Wilf-equivalences of the form \(\av{231,p}\) where \(p\)
+In order to establish the remainder of the Wilf-equivalences of the form \(\av{\{\perm{2,3,1},p\}}\) where \(p\)
 is a mesh pattern we can use similar methods to allow us to consolidate experimental classes into

### 2017-04-06 writing/paper/sections/wilf.tex
Symmetry use in Wilf
- Refer to symmetry instead of doing the structural work.

@@
 \end{equation*}
-
-% TODO Isn't q_3 a symmetry of q_2? So we don't need to repeat the work
-Now consider \(q_3 = \textpattern{}{1}{0/0,0/1,1/0}\). The avoiders of this
-pattern are permutations that do not start with their minimum. In this case we
-perform the decomposition around the leftmost element
-\begin{equation*}
-    \mathcal{\var}_3 =\varepsilon \sqcup
-\decompleft{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
-\end{equation*}
+The pattern \(q_3 = \textpattern{}{1}{0/0,0/1,1/0}\) is the reverse-complement-inverse
+of \(q_2\) and hence the avoiders of \(q_2\) and \(q_3\) are equinumerous.
 All of these classes have the same generating function, namely

### 2017-04-06 writing/paper/sections/meshpatterns.tex
Do most of Henning's changes
- Change makefile to output to pdf
- Never need compositions of boksall, so removed.
- fixed consistency with av brackets (I hope)
- fixed Henning  using dollar signs in his edits
- changed Henning's long-form cite to nice automatic textcite
- removed duplicate line about considering as set partitions
Results of Discussions with Henning
- Changed example to be about subpatterns, may need more words
- Removed second description of one example

@@
     \textpattern{}{ 2, 1, 3 }{ 0/1, 0/2, 1/0, 1/1, 2/1, 2/2 }\) is contained in
-    \(\pi = \perm{3,4,2,1,5}\) but is not contained in \(\sigma = \perm{4,2,3,1,5}\).
+    \(\pi = \perm{3,4,2,1,5}\).
 \end{example}
@@
 
-TODO: I changed a symbol in the above definition. See the comment.
-% if \(\kappa\) contains \(\tau\)
-% to
-% if \(\kappa\) contains \(p\)
-% that is what we want right?
+\begin{example}
+    The pattern \(p=\mperm{2,1,3}{\{(0,1),(0,2),(1,0),(2,2)\}}=
+    \textpattern{}{ 2, 1, 3 }{ 0/1, 0/2, 1/0, 2/2 }\) is contained in the pattern
+    \(r = \textpattern{}{4,2,3,1,5}{0/1,0/2,0/3,0/4,
+                               1/0,1/1,1/2,1/3,1/4,
+                               2/1,2/2,2/3,2/4,
+                               3/2,3/4,
+                               4/2,4/3,4/4} \) as a subpattern.
+\end{example}
 
-TODO: Change the following example to an example of a mesh pattern in a mesh
-pattern.
-
-Now we consider the plot for the permutation \(\sigma\). This permutation
-avoids \(p\) since for every occurrence of the classical pattern \(\perm{2,1,3}\)
-there is at least one point in one of the shaded boxes. For example, consider the
-subsequence \(315\) in \(\sigma\), this is an occurrence of \(\perm{2,1,3}\) but
-not the mesh pattern since the points with values \(4\) and \(2\) are in the
-regions corresponding to the boxes \(\boks{0,1}\) and \(\boks{0,2}\), which are shaded in
-\(p\). This is shown in the plot below.
-
+The highlighted points form an occurrence of \(p\) in \(r\)
 \begin{equation*}
-    \pattern{3,4,5}{4,2,3,1,5}{0/1,0/2,0/3,0/4,
-                               1/1,1/2,1/3,1/4,
+    \pattern{1,2,5}{4,2,3,1,5}{0/1,0/2,0/3,0/4,
+                               1/0,1/1,1/2,1/3,1/4,
                                2/1,2/2,2/3,2/4,
-                               3/0,3/1,3/2,
-                               4/1,4/2,4/3,4/4}
+                               3/2,3/4,
+                               4/2,4/3,4/4}
 \end{equation*}
-This is true for all occurrences of \(\perm{2,1,3}\) in \(\sigma\) and
-therefore \(\sigma\) avoids \(p\).
+The permutation \(\perm{4,2,3,1,5}\) also contains \(p\) in the usual sense.
 

### 2017-04-20 writing/paper/sections/abstract.tex
Final comments from Henning

@@
-\begin{abstract} Two mesh patterns are coincident if they are avoided by the
-same set of permutations, and are Wilf-equivalent if they have the same number
-of avoiders of each length. We provide sufficient conditions for coincidence
-among mesh patterns, whilst also avoiding a longer classical pattern. Using
-these conditions we completely classify coincidences between families containing
-a mesh pattern of length 2 and a classical pattern of length 3. Furthermore, we
-completely Wilf-classify equivalences of mesh patterns of length 2 whilst also
-avoiding the classical pattern 231.
+\begin{abstract}
+    Two mesh patterns are coincident if they are avoided by the same set of
+    permutations, and are Wilf-equivalent if they have the same number of
+    avoiders of each length. We provide sufficient conditions for coincidence of
+    mesh patterns, when only permutations also avoiding a longer classical
+    pattern are considered. Using these conditions we completely classify
+    coincidences between families containing a mesh pattern of length 2 and a
+    classical pattern of length 3. Furthermore, we completely Wilf-classify mesh
+    patterns of length 2 inside the class of 231-avoiding permutations.
 

### 2017-04-20 writing/paper/sections/coincidences.tex
Final comments from Henning

@@
 that avoid \(\lambda\) is the same as the set of permutations that avoid \(\gamma\),
-\ie \(\av{\lambda} = \av{\gamma}\). Equivalently we can say that they have the same set of
-\emph{containers}, \ie \(\cont{\lambda} = \cont{\gamma}\).
+\ie, \(\av{\lambda} = \av{\gamma}\). Equivalently we can say that they have the same set of
+\emph{containers}, \ie, \(\cont{\lambda} = \cont{\gamma}\).
 
@@
 of a classical pattern of length \(3\) and a mesh pattern of length \(2\), that is
-finding and explaining all coincidences where \(\av{\{p,m\}} = \av{\{p,m^\prime\}}\).
+finding and explaining all coincidences where \(\av{\{\pi,m\}} = \av{\{\pi,m^\prime\}}\),
+where \(\pi\) is a classical pattern of length \(3\).
 
@@
 
-This rule can be understood very in graphical form. In the
-pattern in \FigureRef{fig:rule1} we can gain shading in two boxes since if
-there is a point in any of these boxes we would get an occurrence of the
-dominating pattern \(\perm{3,2,1}\).
+This rule can be understood in graphical form. In the pattern in
+\FigureRef{fig:rule1} we can gain shading in the boxes \((0,2)\), \((2,0)\)
+since if there is a point in either of these boxes there would be an occurrence of
+the dominating pattern \(\perm{3,2,1}\).
 
@@
     \begin{tikzpicture}[scale=\picscale]
+        \filldraw (2.5,0.5) circle (3pt);
         \modpattern{}{2,1}{1/0,1/1,1/2,2/2}
@@
     \begin{tikzpicture}[scale=\picscale]
-        \filldraw (2.5,0.5) circle (3pt);
         \modpattern{}{2,1}{1/0,1/1,1/2,2/2,2/0}
@@
     Since \(R_1\) is a subset of \(R_2\), \NoteRef{not:downcmesh} states that
-    \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\)
+    \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).
 
     Now we consider a permutation \(\omega \in \av{\pi}\),
-    containing an occurrence of \(m_1\). Consider placing a point in the
-    region corresponding to the box \((a,b)\), regardless of where in this
-    region we place the point by condition~\ref{prop:dom1:cont} of the
-    proposition we create an occurrence of \(\pi\). Therefore there can be no
-    points in this region, which implies this is an occurrence of \(m_2\). Hence every occurrence of
+    containing an occurrence of \(m_1\). If there is a point in the
+    region corresponding to the box \((a,b)\), then that point, along with the
+    points of the occurrence of \(m_1\), form an occurrence of \(\sigma^\boks{a,b}\).
+    Then condition~\eqref{prop:dom1:cont} of the
+    proposition implies an occurrence of \(\pi\). Therefore there can be no
+    points in this region, which implies the occurrence of \(m_1\) is an occurrence of \(m_2\). Hence every occurrence of
     \(m_1\) is in fact an occurrence of \(m_2\), and we have that
@@
         \item \(R_1 \triangle R_2 = \{(a,b)\}\)
-        \item   \begin{enumerate}
+        \item\label{prop:dom2:cond} Either one of the following hold \begin{enumerate}
                 \item\label{prop:dom2:condc} \(\pi \preceq p^{\boks{i,j}_a}\) and
@@
     By \NoteRef{not:downcmesh} we only need to show that an occurrence of \(m_1\)
-    implies and occurrence of \(m_2\).
-    We consider taking the first branch of every choice.
+    implies an occurrence of \(m_2\).
+    We consider taking the first branch of every choice in condition~\eqref{prop:dom2:cond}.
     Now consider a permutation \(\omega \in \av{\pi}\). Suppose
     \(\omega\) contains \(m_1\) and consider the region corresponding to
-    \((a,b)\) in \(R_1\).If the region is empty, the occurrence of \(m_1\) is trivially an occurrence
+    \((a,b)\) in \(R_1\). If the region is empty, the occurrence of \(m_1\) is trivially an occurrence
     of \(m_2\).
     If the region is non-empty, then by \LemmaRef{lem:incdec} and
-    condition~\ref{prop:dom2:condc} of the proposition it must contain
+    condition~\eqref{prop:dom2:condc} of the proposition it must contain
     a decreasing subsequence.
@@
 
-By applying both the First Dominating Pattern Rule and the Second Dominating
-Pattern Rule fully explains coincidences of classes of the form
+Together the First Dominating Pattern Rule and the Second Dominating
+Pattern Rule fully explain coincidences of classes of the form
 \(\av{\{\perm{2,3,1},\mperm{2,1}{R}\}}\), obtaining 39 coincidence classes of
@@
     \NoteRef{not:downcmesh}, implies \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.
-
     Now consider a permutation \(\sigma\) in \(\av{\pi}\) that contains an
@@
 
-By application of this rule we can reduce the number of classes in
+This rule reduces the number of classes in
 \(\av{\{\perm{2,3,1},(\perm{1,2}, R)\}}\) to \(56\), which coincides with the
@@
 Intuitively it is easy to see why our previous rules have no
-power here. There is nowhere that it is possible to add a single point to
-gain an occurrence of \(\pi = \perm{3,2,1}\). It is also impossible to have
+power here. It is impossible to add a single point to a mesh pattern \((12, R)\) and
+create an occurrence of \(\pi = \perm{3,2,1}\). It is also impossible to have
 a position where addition of an increase, or decrease, provides extra
@@
 \end{equation*}
-are equivalent in \(\av{\perm{3,2,1}}\). (There are 3 symmetries of these
-patterns that are also equivalent to each other by the same reasoning.)
+are coincident in \(\av{\perm{3,2,1}}\). (There are 3 symmetries of these
+patterns that are also coincident to each other by the same reasoning.)
 
@@
 Where the two highlighted points are the original two points.
-Now if we take the other two points as the points in our
-mesh pattern then we get an occurrence of the pattern
+The other two points are an occurrence of the pattern
 we originally desired, and hence the two patterns are coincident.

### 2017-04-20 writing/paper/sections/conclfut.tex
Final comments from Henning

@@
 where \(\perm{3,2,1}\) is the dominating pattern, possibly using the  construction presented in
-\textcite[Sec.~12]{2015arXiv151203226B}, in order to directly reach enumeration and hopefully establish
+\textcite[Sec.~11]{2015arXiv151203226B}, in order to directly reach enumeration and hopefully establish
 some of the non-trivial Wilf-equivalences between classes with different dominating patterns.

### 2017-04-20 writing/paper/sections/introduction.tex
Do most of Henning's changes
- Change makefile to output to pdf
- Never need compositions of boksall, so removed.
- fixed consistency with av brackets (I hope)
- fixed Henning  using dollar signs in his edits
- changed Henning's long-form cite to nice automatic textcite
- removed duplicate line about considering as set partitions
Final comments from Henning

@@
 pattern can be required to be adjacent in the permutation. \textcite{MR2652101}
-look at classes of pattern where entries can be required to be consecutive in
+look at classes of patterns where entries can also be required to be consecutive in
 value, these are called \emph{bivincular} patterns. \emph{Bruhat-restricted}
@@
 When considering permutation patterns some of the main questions posed relate to
-how and when a pattern is avoided by, or contained in, a arbitrary set of
+how and when a pattern is avoided by, or contained in, an arbitrary set of
 permutations. Two patterns \(\pi\) and \(\sigma\) are \emph{Wilf-equivalent} if
@@
 avoiding a vincular and a covincular pattern simultaneously in order to achieve
-some interesting counting results. However, little work has been done on
+some several counting results. However, little work has been done on
 avoiding a mesh pattern and a classical pattern simultaneously.
@@
 conditions for coincidence. We then establish Wilf-equivalence classes of these
-coincidence classes where the classical pattern 231.
+coincidence classes where the classical pattern is \perm{2,3,1}.

### 2017-04-20 writing/paper/sections/wilf.tex
Results of Discussions with Henning
- Changed example to be about subpatterns, may need more words
- Removed second description of one example
Final comments from Henning

@@
 Since we are always considering Wilf-equivalences in a set \(\av{S}\) we
-must only use these symmetries when they preserve the dominating pattern, if we
+must only use symmetries that preserve the dominating pattern(s), if we
 were to allow other symmetries, then the equivalences calculated in the
@@
     \shortintertext{Solving for \(\var\) gives}
-    \var(x) &= \frac{1}{1+x-xC(x)}\\
-    \var(x) &= \frac{C(x)}{1+xC(x)}
-\end{align*} % TODO Do we need both representations?
+    \var(x) &= \frac{1}{1-x(C(x)-1)}
+\end{align*}
 
@@
 \begin{note}
-  The avoiders of the pattern \(q = \mperm{\perm{2,3,1}}{(1,0),(1,1),(1,2),(1,3)},
+  The avoiders of the mesh pattern \(q = %= \mperm{\perm{2,3,1}}{(1,0),(1,1),(1,2),(1,3)}
 \textpattern{}{2,3,1}{1/0,1/1,1/2,1/3}\) in \(\mathfrak{S}_n\) are in one-to-one correspondence with partitions
-of \(\nrange{n}\) (\textcite[Prop.~2]{DBLP:journals/ejc/Claesson01})
+of \(\nrange{n}\) (\textcite[Prop.~2]{DBLP:journals/ejc/Claesson01}).
 \end{note}
@@
 
-This particular example is interesting as it shows that both methods can be used in tandem to establish a coincidence class.
+% This particular example is interesting as it shows that both methods can be used in tandem to establish a coincidence class.
 
@@
 This is a complete decomposition of avoiders of \(m_1\). Now we
-look at an avoider of \(m_2\), this time decomposition is around the leftmost point
+look at an avoider of \(m_2\), decomposed around the leftmost point
 \begin{equation*}

### 2017-08-21 writing/pgsymposium/abstract_murray_tannock.tex
Add PG symposium abstract

\section{Equivalence classes of Mesh patterns with a dominating
pattern}\label{equivalence-classes-of-mesh-patterns-with-a-dominating-pattern}

\subsection{Murray Tannock}\label{murray-tannock}

A permutation is an arrangement of \(n\) objects. We can define sets of
permutations by the avoidance of subsequences with certain properties.
One of these is \emph{classical pattern avoidance} where the values of
the entries in the subsequence must not have a particular ordering.
Classical patterns capture interesting properties of permutations such
as stack sortability, and as such have links to different combinatorial
objects. Mesh patterns are an extension of classical patterns that allow
more restrictions on the occurrence of the patterns. We say that two
sets of patterns are \emph{coincident} if they are avoided by the same
set of permutations. We establish sufficient conditions for coincidence
of mesh patterns, whilst also avoiding a longer classical pattern. These
conditions, along with two special cases, completely classify
coincidence between families containing a mesh pattern of length \(2\)
and a classical pattern of length \(3\).

(Supervised by Michael Albert)

### 2017-09-11 writing/paper/sections/coincidences.tex
Fix small fixes proposed by reviewer 2

@@
 All coincidence classes of \(\av{\{\perm{3,2,1}, \mperm{2,1}{R}\}}\) can be
-explained by application of \PropositionRef{prop:dom1}. Thus explaining the
-29 experimental coincidence classes opserved.
+explained by application of \PropositionRef{prop:dom1}, thus explaining the
+29 experimental coincidence classes observed.
 
-\subsection{Equivalence classes of Av(\{231, (21, \textit{R})\}).}
+\subsection{Coincidence classes of Av(\{231, (21, \textit{R})\}).}
 By application of \PropositionRef{prop:dom1} we obtain \(43\) coincidence
@@
         \item\label{prop:dom2:cond} Either one of the following hold \begin{enumerate}
-                \item\label{prop:dom2:condc} \(\pi \preceq p^{\boks{i,j}_a}\) and
+                \item\label{prop:dom2:condc} \(\pi \preceq p^{\boks{a,b}_a}\) and
                         \begin{enumerate}
@@
                         \end{enumerate}
-                    \item \(\pi \preceq p^{\boks{i,j}_d}\) and
+                    \item \(\pi \preceq p^{\boks{a,b}_d}\) and
                         \begin{enumerate}
@@
                             \item \((a,b) \in \sigma\) and \((a-1,b-1)\notin R\) and \\
-                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
+                                  \((x,b-1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
                                   \((a-1,y)\in R \implies (a,y) \in R\)  (where \(y\neq b-1,b\)).
@@
 
-\subsection{Equivalence classes of Av(\{231, (12, \textit{R})\}).}
+\subsection{Coincidence classes of Av(\{231, (12, \textit{R})\}).}
 When considering the coincidence classes of \(\av{\{\perm{2,3,1},\mperm{1,2}{R}\}}\)
@@
 
-\subsection{Equivalence classes of Av(\{321, (12, \textit{R})\}).}
+\subsection{Coincidence classes of Av(\{321, (12, \textit{R})\}).}
 When considering coincidences of mesh patterns with underlying classical
@@
         any points then we gain some extra shading on the mesh pattern
-        as any points in these boxes would create an occurrence of the dominating pattern \(\perm{2,3,1}\)
+        as any points in these boxes would create an occurrence of the dominating pattern \(\perm{3,2,1}\)
      \begin{equation*}

### 2017-09-14 writing/paper/sections/wilf.tex
Update on anders bijection, start thinking about bijection in first example

@@
 of \(\nrange{n}\) (\textcite[Prop.~2]{DBLP:journals/ejc/Claesson01}).
+I order to form this correspondence we read the permutation from left to right starting a
+new block in the partition whenever we see an ascent.
 \end{note}
@@
 We will call the least element in each block the \emph{block-bottom}, and note that all permutations
-in \(\av{\perm{2,3,1}}\) also avoid the mesh pattern \(q\).
+in \(\av{\perm{2,3,1}}\) also avoid the mesh pattern \(q\), this classical pattern case
+corresponds precisely to the non-crossing partitions.
 
@@
     If these regions are non-empty then the block containing the point
-    corresponding to the second point in both patterns contains only the point
+    corresponding to the second point in both patterns contains only that point
     (it is a singleton block).
 \end{itemize}
-This tells us that an occurrence of the patterns must happen when there
-is a singleton block occurring after the first block. The difference
-between the patterns is in the underlying classical pattern. This
-means that permutations containing \(m_1\) correspond to set partitions
+
+This means that permutations containing \(m_1\) correspond to set partitions
 with a singleton block with value one higher than some element in the
@@
 one lower than some element in the block containing \(1\) and if this
-block is not the block containing \(1\) then it is a singleton block. This
-proves that the containers of both of these patterns in \(\av{231}\) are
-equinumerous, and therefore so are their avoiders.
+block is not the block containing \(1\) then it is a singleton block.
+%%%%%%%%%%
+%%%%%%%%%%
+%%%%%%%%%%
+
+The classes defined by containment of \(m_1\) and \(m_2\) are
+equinumerous  and therefore so are their avoiders.
 

### 2017-09-15 writing/paper/sections/coincidences.tex
start on reviewer 1

@@
     \begin{enumerate}
-        \item The symmetric difference, \(R_1 \triangle R_2\), is \(\{(a,b)\}\)
+        \item The mesh set \(R_2 = R_1 \cup \{(a,b)\}\)
         \item \(\pi \preceq \sigma^\boks{a,b}\)\label{prop:dom1:cont}
@@
 to the box \((1,1)\) is empty we have an occurrence of \(m_2\). Otherwise, if there is any
-increase in this box then we would have an occurrence of \(\perm{2,3,1}\),
+ascent in this box then we would have an occurrence of \(\perm{2,3,1}\),
 however, since we are in \(\av{\perm{2,3,1}}\) this is not possible. This box
@@
     in \(R\), and \(\pi = (\pi,\emptyset)\) be a dominating classical pattern. If
-    \(\pi \preceq p^{\boks{i,j}_a}\)\\(\(\pi \preceq p^{\boks{i,j}_d}\)),
+    \(\pi \preceq m^{\boks{a,b}_a}\)\\(\(\pi \preceq m^{\boks{a,b}_d}\)),
     then in any occurrence of \(m\) in a permutation \(\varrho\), the region
@@
     \begin{enumerate}
-        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
-        \item\label{prop:dom2:cond} Either one of the following hold \begin{enumerate}
-                \item\label{prop:dom2:condc} \(\pi \preceq p^{\boks{a,b}_a}\) and
+        \item The mesh set \(R_2 = R_1 \cup \{(a,b)\}\)
+        \item\label{prop:dom2:cond} Any one of the following four conditions hold \begin{enumerate}
+                \item\label{prop:dom2:condc} \(\pi \preceq \sigma^{\boks{a,b}_a}\) and
                         \begin{enumerate}
-                            \item \((a+1,b) \in \sigma\) and \((a+1,b-1)\notin R\) and \\
-                                \((x,b-1)\in R \implies (x,b) \in R \) (where \(x\neq a,a+1\)) and\\
-                                  \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b-1,b\)).
-                            \item \((a,b+1) \in \sigma\) and \((a-1,b+1)\notin R\) and \\
-                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
-                                  \((a-1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
+                            \item \((a+1,b) \in \sigma\) and \((a+1,b-1)\notin R_1\) and \\
+                                \((x,b-1)\in R_1 \implies (x,b) \in R_1 \) (where \(x\neq a,a+1\)) and\\
+                                  \((a+1,y)\in R_1 \implies (a,y) \in R_1\) (where \(y\neq b-1,b\)).
+                            \item \((a,b+1) \in \sigma\) and \((a-1,b+1)\notin R_1\) and \\
+                                  \((x,b+1)\in R_1 \implies (x,b) \in R_1\) (where \(x\neq a-1,a\)) and\\
+                                  \((a-1,y)\in R_1 \implies (a,y) \in R_1\) (where \(y\neq b,b+1\)).
                         \end{enumerate}
-                    \item \(\pi \preceq p^{\boks{a,b}_d}\) and
+                    \item \(\pi \preceq \sigma^{\boks{a,b}_d}\) and
                         \begin{enumerate}
-                            \item \((a+1,b+1) \in \sigma\) and \((a+1,b+1)\notin R\) and \\
-                                  \((x,b+1)\in R \implies (x,b) \in R\) (where \(x\neq a,a+1\)) and\\
-                                  \((a+1,y)\in R \implies (a,y) \in R\) (where \(y\neq b,b+1\)).
-                            \item \((a,b) \in \sigma\) and \((a-1,b-1)\notin R\) and \\
-                                  \((x,b-1)\in R \implies (x,b) \in R\) (where \(x\neq a-1,a\)) and\\
-                                  \((a-1,y)\in R \implies (a,y) \in R\)  (where \(y\neq b-1,b\)).
+                            \item \((a+1,b+1) \in \sigma\) and \((a+1,b+1)\notin R_1\) and \\
+                                  \((x,b+1)\in R_1 \implies (x,b) \in R_1\) (where \(x\neq a,a+1\)) and\\
+                                  \((a+1,y)\in R_1 \implies (a,y) \in R_1\) (where \(y\neq b,b+1\)).
+                            \item \((a,b) \in \sigma\) and \((a-1,b-1)\notin R_1\) and \\
+                                  \((x,b-1)\in R_1 \implies (x,b) \in R_1\) (where \(x\neq a-1,a\)) and\\
+                                  \((a-1,y)\in R_1 \implies (a,y) \in R_1\)  (where \(y\neq b-1,b\)).
                         \end{enumerate}
@@
 \begin{proof}
-    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
     By \NoteRef{not:downcmesh} we only need to show that an occurrence of \(m_1\)
@@
     of \(m_2\).
+
     If the region is non-empty, then by \LemmaRef{lem:incdec} and
@@
     point in the mesh pattern and the points from the subsequence are now in
-    the box southeast of the point. The other conditions allow this to be done
-    without points being present in regions that were shaded. Hence there are no
+    the box southeast of the point.
+
+    The conditions on the mesh ensure that no elements of the permutation
+    that were inside a region corresponding to an unshaded box in the occurrence
+    of \(m_1\) would be in a region corresponding to a shaded box in an occurrence of
+    \(m_2\).
+
+    Hence there are no
     points in the region corresponding to the box \((a,b)\) in the mesh
@@
     \begin{enumerate}
-        \item \(R_1 \triangle R_2 = \{(a,b)\}\)
+        \item The mesh set \(R_2 = R_1 \cup \{(a,b)\}\)
         \item\label{prop:dom3:condocc} One of the patterns in \((\sigma,R_1)^\boksall{a,b}\)
@@
 create an occurrence of \(\pi = \perm{3,2,1}\). It is also impossible to have
-a position where addition of an increase, or decrease, provides extra
+a position where addition of an ascent, or descent, provides extra
 shading power.

### 2017-09-15 writing/paper/sections/wilf.tex
start on reviewer 1

@@
 \subsection{Wilf-classes with mesh patterns of length 1.}
+\label{wilf1}
 When considering the mesh patterns of length \(2\) it will be useful to know
@@
 \textpattern{}{1}{0/0,1/1} and \(q_1 = \textpattern{}{1}{0/0,0/1,1/1}\) are
-coincident. Consider the decomposition of a non-empty avoider of \(q_1\) in
+coincident inside \(\av{\perm{2,3,1}}\). Consider the decomposition of a non-empty avoider of \(q_1\) in
 \(\av{\perm{2,3,1}}\) around the maximum:
@@
 The pattern \(q_3 = \textpattern{}{1}{0/0,0/1,1/0}\) is the reverse-complement-inverse
-of \(q_2\) and hence the avoiders of \(q_2\) and \(q_3\) are equinumerous.
+of \(q_2\) and hence the avoiders of \(q_2\) and \(q_3\) (\(\scriptvar_2\) and \(\scriptvar_3\)) are equinumerous.
 All of these classes have the same generating function, namely
@@
     \(m_1\) is empty, then the point corresponding to the second point is
-    precisely the last block bottom. If the region corresponding to box \((0,1)\)
+    precisely the last block bottom. If the region corresponding to box \((1,0)\)
     in an occurrence of \(m_2\) is empty, then the point corresponding to the
@@
 
-The classes defined by containment of \(m_1\) and \(m_2\) are
+The classes defined by containment of \(m_1\) and \(m_2\) inside \(\av{\perm{2,3,1}}\)are
 equinumerous and therefore so are their avoiders.
 
-Consider an avoider of \(\perm{2,3,1}\) and \(m_3\). We can perform
-the decomposition around the maximum
+Consider the class, \(\scriptvar_3\), of permutations defined by avoiding \(\perm{2,3,1}\) and \(m_3\). We can decompose a member of this class around the maximum
 \begin{equation*}
-    \scriptvar_1 = \varepsilon \sqcup
-    \decompmax{1/0,0/1}{\scriptvar_1}{}{}{\mathcal{\varmaxl}_1}
+    \scriptvar_3 = \varepsilon \sqcup
+    \decompmax{1/0,0/1}{\scriptvar_3}{}{}{\mathcal{\varmaxl}_1}
 \end{equation*}
@@
 \begin{equation*}
-    \scriptvar_2 = \varepsilon \sqcup
-    \decompmax{1/0,0/1}{\scriptvar_2}{}{}{\mathcal{\varmaxl}_3}
+    \scriptvar_4 = \varepsilon \sqcup
+    \decompmax{1/0,0/1}{\scriptvar_4}{}{}{\mathcal{\varmaxl}_3}
 \end{equation*}
@@
 in the region, if and only if this first point is the lowest valued element in
-this region.
+this region, so this top right region must contain a sub-permutation that does not
+start with it's minimum, i.e., it is a member of \(\mathcal{\varmaxl}_3\) described
+in \SectionRef{wilf1}.
 Since both \(\mathcal{\varmaxl}_1\) and \(\mathcal{\varmaxl}_3\) have the same
-enumeration, \(\scriptvar_1\) and \(\scriptvar_2\) must also have
+enumeration, \(\scriptvar_3\) and \(\scriptvar_4\) must also have
 the same enumeration and are therefore Wilf-equivalent.
@@
 \begin{equation*}
-    \scriptvar_3 = \varepsilon \sqcup
-    \decompleft{1/0,0/1}{\scriptvar_3}{}{}{\mathcal{\varmaxl}_3}
+    \scriptvar_1 = \varepsilon \sqcup
+    \decompleft{1/0,0/1}{\scriptvar_1}{}{}{\mathcal{\varmaxl}_3}
 \end{equation*}
@@
 \end{equation*}
-
-% This particular example is interesting as it shows that both methods can be used in tandem to establish a coincidence class.
 

### 2017-10-06 writing/paper/sections/coincidences.tex
Some of reviewer 1s comments,change away from set partitions

@@
 
-We will consider the avoidance sets \(\av{\{\pi,p\}}\) where \(\pi\) is a classical pattern of length 3
+We will consider the avoidance sets \(\av{\pi,p}\) where \(\pi\) is a classical pattern of length 3
 and \(p\) is a mesh pattern of length 2 to establish sufficient conditions for two such
@@
 the \emph{dominating pattern}.
+Given such a dominating pattern \(\pi\) we will write \(q_1 \coinc{\pi} q_2\) if for the mesh patterns, \(q_1\) and \(q_2\),  the sets \(\av{p,q_1}\) and \(\av{p,q_2}\) are coincident.
+
+Our first step is to calculate whether \(\av{p,q_1}\) and \(\av{p,q_2}\) are equal up to permutations of length 11 computationally, if they are we write \(q_1 \coinc[comp]{\pi} q_2\).
+If \(q_1 \coinc{\pi} q_2\), it can be seen that \(q_1 \coinc[comp]{\pi} q_2\), and therefore the \(\coinc{\pi}\)-equivalence classes form partitions of the \(\coinc[comp]{\pi}\)-coincidence classes.
 
@@
 of a classical pattern of length \(3\) and a mesh pattern of length \(2\), that is
-finding and explaining all coincidences where \(\av{\{\pi,m\}} = \av{\{\pi,m^\prime\}}\),
+finding and explaining all coincidences between mesh patterns \(m\) and \(m^\prime\), \(m\coinc{\pi} m^\prime\),
 where \(\pi\) is a classical pattern of length \(3\).
@@
 Through experimentation, considering avoidance of permutations of up to length \(11\), we discover
-that there are at least \(29\) coincidence classes of mesh patterns with underlying classical
-pattern \(21\).
+that there are at least \(29\) \(\coinc[comp]{321}\)-coincidence classes where the underlying classical pattern of the mesh pattern is \(21\).
 

### 2017-10-06 writing/paper/sections/introduction.tex
Fix small fixes proposed by reviewer 2
Some of reviewer 1s comments,change away from set partitions

@@
 avoiding a vincular and a covincular pattern simultaneously in order to achieve
-some several counting results. However, little work has been done on
+several counting results. However, little work has been done on
 avoiding a mesh pattern and a classical pattern simultaneously.
@@
 begin by establishing coincidences between mesh patterns of length 2 while
-avoiding a classical pattern of length 3; this is used to establish sufficient
-conditions for coincidence. We then establish Wilf-equivalence classes of these
-coincidence classes where the classical pattern is \perm{2,3,1}.
+avoiding a classical pattern of length 3 by computational methods, which are then
+used to establish three ``Dominating Pattern Rules'' as well as some special cases
+that can be used to calculate coincidences. We then use these coincidence classes
+to calculate Wilf-equivalence classes showing some of the methods used.

### 2017-10-06 writing/paper/sections/wilf.tex
Some of reviewer 1s comments,change away from set partitions

@@
 
-When considering the Wilf-equivalences we consider how permutations correspond
-to set-partitions.
-
-\begin{definition} A \emph{set partition} of \(\nrange{n}\) is a set of non-empty subsets of \(\nrange{n}\)
-such that \(\nrange{n}\) is a \emph{disjoint} union of these subsets.
-\end{definition}
-
-\begin{note}
-  The avoiders of the mesh pattern \(q = %= \mperm{\perm{2,3,1}}{(1,0),(1,1),(1,2),(1,3)}
-\textpattern{}{2,3,1}{1/0,1/1,1/2,1/3}\) in \(\mathfrak{S}_n\) are in one-to-one correspondence with partitions
-of \(\nrange{n}\) (\textcite[Prop.~2]{DBLP:journals/ejc/Claesson01}).
-I order to form this correspondence we read the permutation from left to right starting a
-new block in the partition whenever we see an ascent.
-\end{note}
-\begin{example}
-    The permutation \(\pi = \perm{5,4,2,1,3,9,6,8,7}\) corresponds to the
-    partition \(\{\{5,4,2,1\},\{3\},\{9,6\},\{8,7\}\}\).
-\end{example}
-
-We will call the least element in each block the \emph{block-bottom}, and note that all permutations
-in \(\av{\perm{2,3,1}}\) also avoid the mesh pattern \(q\), this classical pattern case
-corresponds precisely to the non-crossing partitions.
-
 We will use two main methods of establishing Wilf-equivalence between mesh patterns of length \(2\)
 in \(\av{\perm{2,3,1}}\): the structural decomposition of avoiders, via generating functions; or the
-structure of the set-partition induced by the pattern, looking at a particular occurrence of the
+structure of a general permutation containing the pattern, looking at a particular occurrence of the
 pattern in a permutation avoiding \(\perm{2,3,1}\). Sometimes it will be necessary to use both of
@@
 First we prove the Wilf-equivalence between \(m_1\) and \(m_2\) shown in
-\eqref{eq:singleud}, by considering the containers as set partitions.
+\eqref{eq:singleud}, by considering the form of a general permutation containing
+either of the two patterns. First looking at a general occurrence of \(m_1\) in a permutation
+in \(\av{231}\)
+\begin{equation*}
+\begin{tikzpicture}[scale=1.0]
+\modpattern[4]{}{1,2}{0/0,0/1,1/1,1/2,2/0,2/1}
+\node at (0.5,2.5) {\Large \(A\)};
+\node at (2.5,2.5) {\Large \(B\)};
+\node at (1.5,0.5) {\Large \(C\)};
+\end{tikzpicture}	
+\end{equation*}
 
-Considering an occurrence of either of these patterns in a permutation  we know
-the following about the points corresponding to the points in the patterns.
-\begin{itemize}
-    \item The point corresponding to the first point in both patterns
-        must lie in the first block of the set partition (there are
-        no points southwest from it in the permutation).
-    \item The point corresponding to the second point in both patterns
-        is a block bottom (there are no points southeast of it in
-        the permutation).
-    \item If the region corresponding to box \((2,2)\) in an occurrence of
-    \(m_1\) is empty, then the point corresponding to the second point is
-    precisely the last block bottom. If the region corresponding to box \((1,0)\)
-    in an occurrence of \(m_2\) is empty, then the point corresponding to the
-    second point is precisely the first block bottom.
-    If these regions are non-empty then the block containing the point
-    corresponding to the second point in both patterns contains only that point
-    (it is a singleton block).
-\end{itemize}
+If there is an occurrence of \(m_1\) there must be a lowest occurrence of \(m_1\), \ie the occurrence has the lowest possible values for any occurrence. Now consider the top
+regions, labelled \(A \et B\), the subpermutation contained in these regions must avoid the permutation \(231\).
+Also, the subpermutation in the region \(A\) must be a decreasing subsequence, otherwise an occurrence of \(231\) will be created with either of the points in the occurrence.
+Now, consider the region labelled \(C\), since we specified that we were focused on the lowest possible occurrence of \(m_1\) this region cannot contain an occurrence of either \(m_1\) or \(231\).
+We now have a structural decomposition of a container of \(m_1\) inside \(\av{231}\).
 
-This means that permutations containing \(m_1\) correspond to set partitions
-with a singleton block with value one higher than some element in the
-block containing \(1\). The permutations containing \(m_2\) correspond
-to the set partitions containing a block with block bottom having value
-one lower than some element in the block containing \(1\) and if this
-block is not the block containing \(1\) then it is a singleton block.
-%%%%%%%%%%
-%%%%%%%%%%
-%%%%%%%%%%
+Now consider a general occurrence of \(m_2\) in a permutation
+in \(\av{231}\)
+\begin{equation*}
+\begin{tikzpicture}[scale=1.0]
+\modpattern[4]{}{2,1}{0/0,0/1,1/1,1/2,2/0,2/1}
+\node at (0.5,2.5) {\Large \(A^\prime\)};
+\node at (2.5,2.5) {\Large \(B^\prime\)};
+\node at (1.5,0.5) {\Large \(C^\prime\)};
+\end{tikzpicture}	
+\end{equation*}
 
-The classes defined by containment of \(m_1\) and \(m_2\) inside \(\av{\perm{2,3,1}}\)are
+Similarly to a container of \(m_1\) we will look at the lowest possible occurrence of \(m_2\), similar to before the regions \(A^\prime \et B^\prime\) together contain an avoider of \(231\) and \(A\) contains a decreasing subsequence.
+Since we are considering the lowest occurrence of \(m_2\) so the region \(C^\prime\) does not contain an occurrence of either \(m_2\) or \(231\).
+
+Since all the regions in both of these cases contain the same parts,  the classes defined by containment of \(m_1\) and \(m_2\) inside \(\av{\perm{2,3,1}}\) are
 equinumerous and therefore so are their avoiders.

### 2017-10-09 writing/paper/sections/coincidences.tex
Fixes for comments DMTCS

@@
 If \(q_1 \coinc{\pi} q_2\), it can be seen that \(q_1 \coinc[comp]{\pi} q_2\), and therefore the \(\coinc{\pi}\)-equivalence classes form partitions of the \(\coinc[comp]{\pi}\)-coincidence classes.
+
+We then form three Propositions called ``Dominating Pattern Rules'', if the First Dominating
+Pattern Rule shows that \(q_1 \coinc{\pi} q_2\) then we write \(q_1 \coinc[(1)]{\pi} q_2\).
+If a combination of the First and Second Dominating Pattern Rules show that \(\av{p,q_1}= \av{p,q_2}\) then we write \(q_1 \coinc[(2)]{\pi} q_2\); similarly we write \(q_1 \coinc[(3)]{\pi} q_2\) if a combination of all three rules shows coincidences.
 
@@
     and a dominating classical pattern \(\pi = (\pi,\emptyset)\) such that
-    \(\setsize{\pi} \le \setsize{\sigma} + 1\), the sets \(\av{\{\pi,m_1\}}\) and
-    \(\av{\{\pi,m_2\}}\) are coincident if
-
+    \(\setsize{\pi} \le \setsize{\sigma} + 1\), then \(m_1 \coinc{\pi} m_2 \) 
+    as a consequence of the first dominating pattern rule if
     \begin{enumerate}
@@
 \begin{proof}[Proof of \PropositionRef{prop:dom1}]
-    We need to prove that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).
+    We need to prove that \(m_1 \coinc{\pi} m_2\).
 
@@
     Taking both directions of the containment we can therefore draw the
-    conclusion that \(\av{\{\pi,m_1\}} = \av{\{\pi,m_2\}}\).
+    conclusion that \(m_1 \coinc{\pi} m_2\).
 \end{proof}
@@
 \subsection{Coincidence classes of Av(\{231, (21, \textit{R})\}).}
-By application of \PropositionRef{prop:dom1} we obtain \(43\) coincidence
-classes. Experimentation shows that there are in fact at least \(39\) coincidence
-classes, for example the following two patterns seem to be coincident in
+By application of \PropositionRef{prop:dom1} we obtain \(43\) \(\coinc[(1)]{231}\)-coincidence
+classes between mesh patterns with \(21\) as an underlying classical pattern.
+Experimentation shows that there are \(39 \coinc[comp]{231}\)-coincidence
+classes between mesh patterns with \(21\) as an underlying classical pattern, for example the following two patterns seem to be coincident in
 \(\av{\perm{2,3,1}}\) but this is not explained by \PropositionRef{prop:dom1}.
@@
     \(\setsize{\pi} \le \setsize{\sigma} + 2\), the sets \(\av{\{\pi,m_1\}}\) and
-    \(\av{\{\pi,m_2\}}\) are coincident if
+    \(\av{\{\pi,m_2\}}\) are coincident as a consequence of the Second Dominating Rule if
 
@@
 point in the mesh pattern.
+\FigureRef{fig:d2} show how the cases apply to a general container of \(m_1\) to transform it into a container of \(m_2\) , the circled points are the same points in the permutation.
+\begin{figure}
+\begin{tikzpicture}
+	\modpattern[3]{1,2}{2,1}{0/0,0/1,0/2,1/0,2/1}
+	\filldraw (1.33,1.66) circle (1.5pt);
+	\filldraw (1.66,1.33) circle (1.5pt);
+	\filldraw[pattern=crosshatch dots, pattern color=black!80] (2.1,0.9) rectangle (2.5,0.0);
+	\filldraw[pattern=crosshatch dots, pattern color=black!80] (2.5,3) rectangle (2.9 ,2.1);
+\end{tikzpicture}
+\raisebox{8ex}{\(\coinc{}\)}
+\begin{tikzpicture}
+	\modpattern[3]{1}{2,1}{0/0,0/1,0/2,1/0,1/1,2/1}
+	\filldraw (2.15,0.85) circle (1.5pt);
+	\filldraw (2.3,0.7) circle (1.5pt);
+	\draw (2.3,0.7) circle (3pt);
+	\filldraw[pattern=crosshatch dots, pattern color=black!80] (2.4,0.6) rectangle (2.65,0.0);
+	\filldraw[pattern=crosshatch dots, pattern color=black!80] (2.65,3) rectangle (2.9 ,2.1);
+\end{tikzpicture}
+\caption{A depiction of the second dominating pattern rule using our example patterns.}
+\label{fig:d2}
+\end{figure}
 
@@
 we first apply the two Dominating Pattern rules previously established.
-Starting from \(220\) classes, application of the first Dominating Pattern rule
-gives \(85\) classes. Following this with the second Dominating Pattern rule
-reduces the number of classes to \(59\). However we know that there are patterns
+Starting from \(220\) classes, we result in \(85 \coinc[(1)]{231}\)-coincidence classes 
+reducing to \(59 \coinc[(2)]{231}\)-coincidence classes, considering mesh patterns with
+\(12\) as the underlying classical patterns. However we know that there are patterns
 where the coincidences are not explained by the rules given above.

### 2017-10-09 writing/paper/sections/introduction.tex
Fixes for comments DMTCS

@@
 begin by establishing coincidences between mesh patterns of length 2 while
-avoiding a classical pattern of length 3 by computational methods, which are then
+avoiding a dominating pattern by computational methods, which are then
 used to establish three ``Dominating Pattern Rules'' as well as some special cases
-that can be used to calculate coincidences. We then use these coincidence classes
+that can be used to calculate coincidences.
+
+ We then use these coincidence classes
 to calculate Wilf-equivalence classes showing some of the methods used.

### 2017-10-12 writing/paper/sections/coincidences.tex
Some changes from Henning

@@
 \textcite{DBLP:journals/corr/ClaessonTU14}, in which the authors use the Simultaneous Shading Lemma,
-a closure result and one worked out special case to fully classify coincidences among mesh patterns
+a closure result and one special case to fully classify coincidences among mesh patterns
 of length 2.
@@
 that avoid \(\lambda\) is the same as the set of permutations that avoid \(\gamma\),
-\ie, \(\av{\lambda} = \av{\gamma}\). Equivalently we can say that they have the same set of
-\emph{containers}, \ie, \(\cont{\lambda} = \cont{\gamma}\).
+\ie, \(\av{\lambda} = \av{\gamma}\). This is equivalent to \(\lambad\) and \(\gamma\) being contained in
+the same set of permutations, \ie, \(\cont{\lambda} = \cont{\gamma}\).
 
-We will consider the avoidance sets \(\av{\pi,p}\) where \(\pi\) is a classical pattern of length 3
-and \(p\) is a mesh pattern of length 2 to establish sufficient conditions for two such
-sets to be coincident. The classical pattern \(\pi\) will be fixed and called
-the \emph{dominating pattern}.
-Given such a dominating pattern \(\pi\) we will write \(q_1 \coinc{\pi} q_2\) if for the mesh patterns, \(q_1\) and \(q_2\),  the sets \(\av{p,q_1}\) and \(\av{p,q_2}\) are coincident.
+We will consider the avoidance sets \(\av{\pi,p}\) where \(\pi\) is a fixed
+classical pattern of length 3 and \(p\) is a mesh pattern of length 2 that
+varies. The classical pattern \(\pi\) will be called the \emph{dominating
+pattern}. Given such a dominating pattern \(\pi\) we will write \(p_1
+\coinc{\pi} p_2\) if for the mesh patterns, \(p_1\) and \(p_2\),  the sets
+\(\av{\pi,p_1}\) and \(\av{\pi,p_2}\) are equal.
 
-Our first step is to calculate whether \(\av{p,q_1}\) and \(\av{p,q_2}\) are equal up to permutations of length 11 computationally, if they are we write \(q_1 \coinc[comp]{\pi} q_2\).
-If \(q_1 \coinc{\pi} q_2\), it can be seen that \(q_1 \coinc[comp]{\pi} q_2\), and therefore the \(\coinc{\pi}\)-equivalence classes form partitions of the \(\coinc[comp]{\pi}\)-coincidence classes.
+Our first step is to calculate whether \(\av{\pi,p_1}\) and \(\av{\pi,p_2}\) are
+equal up to permutations of length 11 computationally, if they are we write
+\(p_1 \coinc[comp]{\pi} p_2\). The equivalence relation \(\coinc{\pi}\) is a
+refinement of the equivalence relation \(\coinc[comp]{\pi}\), and therefore the
+\(\coinc{\pi}\)-equivalence classes form partitions of the
+\(\coinc[comp]{\pi}\)-coincidence classes.
 
-We then form three Propositions called ``Dominating Pattern Rules'', if the First Dominating
-Pattern Rule shows that \(q_1 \coinc{\pi} q_2\) then we write \(q_1 \coinc[(1)]{\pi} q_2\).
-If a combination of the First and Second Dominating Pattern Rules show that \(\av{p,q_1}= \av{p,q_2}\) then we write \(q_1 \coinc[(2)]{\pi} q_2\); similarly we write \(q_1 \coinc[(3)]{\pi} q_2\) if a combination of all three rules shows coincidences.
+If \(p_1 \coinc{\pi} p_2\), it can be seen that
+\(p_1 \coinc[comp]{\pi} p_2\), and therefore the \(\coinc{\pi}\)-equivalence
+classes form partitions of the \(\coinc[comp]{\pi}\)-coincidence classes.
+
+We then prove three propositions called ``Dominating Pattern Rules'', if the
+First Dominating Pattern Rule shows that \(p_1 \coinc{\pi} p_2\) then we write
+\(p_1 \coinc[(1)]{\pi} p_2\). If a combination of the First and Second
+Dominating Pattern Rules show that \(p_1 \coinc{\pi} p_2\) then we write
+\(p_1 \coinc[(2)]{\pi} p_2\); similarly we write \(p_1 \coinc[(3)]{\pi} p_2\) if
+a combination of all three rules shows coincidences. These three equivalence
+relations are successive refinements of \(\coinc{\pi}\). Therefore if we find
+that \eg, \(\coinc[(1)]{\pi} = \coinc[comp]{\pi}\) then \(\coinc[(1)]{\pi} = \coinc{\pi}\).
 

### 2017-10-12 writing/paper/sections/introduction.tex
Some changes from Henning

@@
 stack sorting in \emph{The Art of Computer Programming}~\cite[p.~243,
-Ex.~5,6]{Knuth:1997:ACP:260999}. This original concept---a subsequence of
+Ex.~5,6]{Knuth:1997:ACP:260999}. The original concept---a subsequence of
 symbols having a particular relative order, now known as classical

### 2017-10-12 writing/paper/sections/meshpatterns.tex
Final comments from Henning
Some changes from Henning

@@
 denote an individual permutation \(\pi\in\mathfrak{S}_n\) in \emph{one-line notation} by writing the
-entries of the permutation in order, therefore \(\pi = \perm{\pi(1),\pi(2),\dotso,\pi(n)}\). The
+entries of the permutation in order, therefore \(\pi = \perm{\pi(1),\pi(2),\dotsm,\pi(n)}\). The
 set \(\mathfrak{S}_0\) has exactly one element, the empty permutation \(\varepsilon\).
 
-\begin{definition}{(Order isomorphism.)}
-    Two strings \(\alpha_1\alpha_2\dotsm\alpha_n\) and
+\begin{definition}
+    Two strings of integers \(\alpha_1\alpha_2\dotsm\alpha_n\) and
     \(\beta_1\beta_2\dotsm\beta_n\) are said to be \emph{order isomorphic}
@@
     A permutation \(\pi\in\mathfrak{S}_n\) \emph{contains} the permutation \(\sigma\in\mathfrak{S}_k\) (\mbox{denoted}
-    \(\sigma \preceq \pi\)) if there is some sequence \( i_1,i_2,\dotsm{},i_k\)
+    \(\sigma \preceq \pi\)) if there is some sequence \( i_1,i_2,\dotsc,i_k\)
     such that \(1\le i_1<i_2<\dotsm<i_k\le n\) and the sequence \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is order isomorphic to
-    \(\sigma(1)\sigma(2)\dotsm\sigma(k)\). If \(\pi\) does not contain \(\sigma\), we say that
-    \(\pi\) \emph{avoids} \(\sigma\). If this is the case the sequence
+    \(\sigma(1)\sigma(2)\dotsm\sigma(k)\). If this is the case the sequence
     \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is called an \emph{occurrence} of \(\sigma\)
-    in \(\pi\). In this context \(\sigma\) is called a (\emph{classical})
+    in \(\pi\). If \(\pi\) does not contain \(\sigma\), we say that
+    \(\pi\) \emph{avoids} \(\sigma\). In this context \(\sigma\) is called a (\emph{classical})
     \emph{permutation pattern}.
@@
 injections \(\alpha,\beta:\nrange{k}\mapsto\nrange{n}\) satisfying the following two conditions.
-
-\noindent Firstly, \(\omega\) is an occurrence of \(\tau\) in the classical sense
+Firstly, \(\omega\) is an occurrence of \(\tau\) in the classical sense
   \begin{enumerate}[i.]
@@
 \begin{example}
-    The pattern \(p=\mperm{2,1,3}{\{(0,1),(0,2),(1,0),(1,1),(2,1),(2,2)\}}=
+    The pattern \(p=\mperm{2,1,3}{\{\boks{0,1},\boks{0,2},\boks{1,0},\boks{1,1},\boks{2,1},\boks{2,2}\}}=
     \textpattern{}{ 2, 1, 3 }{ 0/1, 0/2, 1/0, 1/1, 2/1, 2/2 }\) is contained in
     \(\pi = \perm{3,4,2,1,5}\).
-\end{example}
-
     Let us consider the plot for the permutation \(\pi\). The subsequence \(325\) is
@@
     The subsequence \(325\) is therefore an occurrence of the pattern \(p\) in \(\pi\).
+\end{example}
+
+We denote the avoidance sets for mesh patterns in the same way as for
+classical patterns. Given a mesh pattern \(p=(\sigma,R)\) we say that \(\sigma\) is the \emph{underlying
+classical pattern} of \(p\).
 
@@
 with the additional condition that if \(\boks{i,j}\in R\text{ then } R_{ij}\) is
-contained in the mesh set of \(q\).
+contained in the mesh set of \(q\). More formally:
 
@@
 \begin{example}
-    The pattern \(p=\mperm{2,1,3}{\{(0,1),(0,2),(1,0),(2,2)\}}=
+    The pattern \(p=\mperm{2,1,3}{\{\boks{0,1},\boks{0,2},\boks{1,0},\boks{2,2}\}}=
     \textpattern{}{ 2, 1, 3 }{ 0/1, 0/2, 1/0, 2/2 }\) is contained in the pattern
-    \(r = \textpattern{}{4,2,3,1,5}{0/1,0/2,0/3,0/4,
+    \(q = \textpattern{}{4,2,3,1,5}{0/1,0/2,0/3,0/4,
                                1/0,1/1,1/2,1/3,1/4,
@@
                                4/2,4/3,4/4} \) as a subpattern.
-\end{example}
-
-The highlighted points form an occurrence of \(p\) in \(r\)
+    The highlighted points form an occurrence of \(p\) in \(q\)
     \begin{equation*}
@@
     The permutation \(\perm{4,2,3,1,5}\) also contains \(p\) in the usual sense.
-
-We denote the avoidance sets for mesh patterns in the same way as for
-classical patterns. Given a mesh pattern \(p=(\sigma,R)\) we say that \(\sigma\) is the \emph{underlying
-classical pattern} of \(p\).
+\end{example}

### 2017-10-12 writing/paper/sections/wilf.tex
Some changes from Henning

@@
 
-\begin{definition}{(Wilf Equivalence.)}
+\begin{definition}
   Two patterns \(\pi\) and \(\sigma\) are said to be \emph{Wilf-equivalent}

### 2017-10-12 writing/paper/sections/coincidences.tex
More changes

@@
 that avoid \(\lambda\) is the same as the set of permutations that avoid \(\gamma\),
-\ie, \(\av{\lambda} = \av{\gamma}\). This is equivalent to \(\lambad\) and \(\gamma\) being contained in
+\ie, \(\av{\lambda} = \av{\gamma}\). This is equivalent to \(\lambda\) and \(\gamma\) being contained in
 the same set of permutations, \ie, \(\cont{\lambda} = \cont{\gamma}\).
@@
 \coinc{\pi} p_2\) if for the mesh patterns, \(p_1\) and \(p_2\),  the sets
-\(\av{\pi,p_1}\) and \(\av{\pi,p_2}\) are equal.
+\(\av{\pi,p_1}\) and \(\av{\pi,p_2}\) are equal. If this is the case we say
+the two mesh patterns are \emph{coincidenc under} \(\pi\).
 
@@
 \(\coinc{\pi}\)-equivalence classes form partitions of the
-\(\coinc[comp]{\pi}\)-coincidence classes.
-
-If \(p_1 \coinc{\pi} p_2\), it can be seen that
-\(p_1 \coinc[comp]{\pi} p_2\), and therefore the \(\coinc{\pi}\)-equivalence
-classes form partitions of the \(\coinc[comp]{\pi}\)-coincidence classes.
+\(\coinc[comp]{\pi}\)-equivalence classes.
 
@@
 \(p_1 \coinc[(2)]{\pi} p_2\); similarly we write \(p_1 \coinc[(3)]{\pi} p_2\) if
-a combination of all three rules shows coincidences. These three equivalence
-relations are successive refinements of \(\coinc{\pi}\). Therefore if we find
-that \eg, \(\coinc[(1)]{\pi} = \coinc[comp]{\pi}\) then \(\coinc[(1)]{\pi} = \coinc{\pi}\).
+a combination of all three rules shows coincidence. These three equivalence
+relations are successive refinements of \(\coinc{\pi}\). For future reference
+we record the following:
+
+\begin{note}
+\label{note:main}
+    If we find that \(\coinc[(i)]{\pi}\), for \(i = 1, 2\), or \(3\),
+    equals \(\coinc[comp]{\pi}\) then the equivalence relations \(\coinc[(i)]{\pi}\), \(\coinc{\pi}\),
+    and \(\coinc[comp]{\pi}\) are all equal.
+\end{note}
 
@@
 \end{equation*}
-\caption{The result of inserting a point into \(p=(12,\{(0,1),(2,2)\})\)}
+\caption{The result of inserting a point into \(p=(12,\{\boks{0,1},\boks{2,2}\})\)}
 \label{fig:addp}
@@
 \end{equation*}
-\caption{The result of inserting an ascent into \(p=(231,\{(0,0),(1,0),(1,3),(3,0),(3,1),(3,3)\})\)}
+\caption{The result of inserting an ascent into \(p=(231,\{\boks{0,0},\boks{1,0},\boks{1,3},\boks{3,0},\boks{3,1},\boks{3,3}\})\)}
 \label{fig:adda}
@@
 \subsection{Coincidence classes of Av(\{321, (21, \textit{R})\})}
-Through experimentation, considering avoidance of permutations of up to length \(11\), we discover
-that there are at least \(29\) \(\coinc[comp]{321}\)-coincidence classes where the underlying classical pattern of the mesh pattern is \(21\).
+Through experimentation, considering avoidance of permutations up to length \(11\), we discover
+that there are \(29\) \(\coinc[comp]{321}\)-equivalence classes where the underlying classical pattern of the mesh pattern is \(21\).
 
@@
     \(\setsize{\pi} \le \setsize{\sigma} + 1\), then \(m_1 \coinc{\pi} m_2 \)
-    as a consequence of the first dominating pattern rule if
+    if
     \begin{enumerate}
-        \item The mesh set \(R_2 = R_1 \cup \{(a,b)\}\)
+        \item The mesh set \(R_2 = R_1 \cup \{\boks{a,b}\}\)
         \item \(\pi \preceq \sigma^\boks{a,b}\)\label{prop:dom1:cont}
@@
 This rule can be understood in graphical form. In the pattern in
-\FigureRef{fig:rule1} we can gain shading in the boxes \((0,2)\), \((2,0)\)
+\FigureRef{fig:rule1} we can gain shading in the boxes \(\boks{0,2}\), \(\boks{2,0}\)
 since if there is a point in either of these boxes there would be an occurrence of
@@
 \begin{proof}[Proof of \PropositionRef{prop:dom1}]
-    We need to prove that \(m_1 \coinc{\pi} m_2\).
-
-    \noindent Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
-    Since \(R_1\) is a subset of \(R_2\), \NoteRef{not:downcmesh} states that
+    Since \(R_1\) is a subset of \(R_2\), \NoteRef{not:downcmesh} implies that
     \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\).
@@
     containing an occurrence of \(m_1\). If there is a point in the
-    region corresponding to the box \((a,b)\), then that point, along with the
+    region corresponding to the box \(\boks{a,b}\), then that point, along with the
     points of the occurrence of \(m_1\), form an occurrence of \(\sigma^\boks{a,b}\).
@@
 
-All coincidence classes of \(\av{\{\perm{3,2,1}, \mperm{2,1}{R}\}}\) can be
-explained by application of \PropositionRef{prop:dom1}, thus explaining the
-29 experimental coincidence classes observed.
+After implementing the First Dominating Pattern Rule we find that there are
+\(29\) \(\coinc[(1)]{321}\)-classes, of mesh patterns where the underlying
+classical pattern is \(21\). By Note~\NoteRef{note:main} there are
+therefore exactly \(29\) \(\coinc{321}\)-classes.
 
 \subsection{Coincidence classes of Av(\{231, (21, \textit{R})\}).}
-By application of \PropositionRef{prop:dom1} we obtain \(43\) \(\coinc[(1)]{231}\)-coincidence
+By application of First Dominating Pattern Rule we obtain \(43\) \(\coinc[(1)]{231}\)-coincidence
 classes between mesh patterns with \(21\) as an underlying classical pattern.
 Experimentation shows that there are \(39 \coinc[comp]{231}\)-coincidence
-classes between mesh patterns with \(21\) as an underlying classical pattern, for example the following two patterns seem to be coincident in
-\(\av{\perm{2,3,1}}\) but this is not explained by \PropositionRef{prop:dom1}.
+classes between mesh patterns with \(21\) as an underlying classical pattern, for example the following two patterns are in the same
+\(\coinc[comp]{231}\)-class, but this is not explained by the First Dominating Pattern Rule.
 \begin{equation*}

### 2017-10-12 writing/paper/sections/coincidences.tex
Such changes

@@
 After implementing the First Dominating Pattern Rule we find that there are
-\(29\) \(\coinc[(1)]{321}\)-classes, of mesh patterns where the underlying
-classical pattern is \(21\). By Note~\NoteRef{note:main} there are
-therefore exactly \(29\) \(\coinc{321}\)-classes.
+\(29\) \(\coinc[(1)]{321}\)-equivalence classes, of mesh patterns where the underlying
+classical pattern is \(21\). By \NoteRef{note:main} there are
+therefore exactly \(29\) \(\coinc{321}\)-equivalence classes.
 
 \subsection{Coincidence classes of Av(\{231, (21, \textit{R})\}).}
-By application of First Dominating Pattern Rule we obtain \(43\) \(\coinc[(1)]{231}\)-coincidence
+By application of First Dominating Pattern Rule we obtain \(43\) \(\coinc[(1)]{231}\)-equivalence
 classes between mesh patterns with \(21\) as an underlying classical pattern.
-Experimentation shows that there are \(39 \coinc[comp]{231}\)-coincidence
+Experimentation shows that there are \(39 \coinc[comp]{231}\)-equivalence
 classes between mesh patterns with \(21\) as an underlying classical pattern, for example the following two patterns are in the same
-\(\coinc[comp]{231}\)-class, but this is not explained by the First Dominating Pattern Rule.
+\(\coinc[comp]{231}\)-equivalence class, but this is not explained by the First Dominating Pattern Rule.
 \begin{equation*}
@@
 of elements \(x\) and \(y\). If the region corresponding
-to the box \((1,1)\) is empty we have an occurrence of \(m_2\). Otherwise, if there is any
+to the box \(\boks{1,1}\) is empty we have an occurrence of \(m_2\). Otherwise, if there is any
 ascent in this box then we would have an occurrence of \(\perm{2,3,1}\),
-however, since we are in \(\av{\perm{2,3,1}}\) this is not possible. This box
+however, since the permutation is in \(\av{\perm{2,3,1}}\) this is not possible. This box
 must therefore contain a (non-empty) decreasing subsequence.
@@
     \label{lem:incdec}
-    Let \(m =(\sigma, R)\) be a mesh pattern, where the box \((a,b)\) is not
-    in \(R\), and \(\pi = (\pi,\emptyset)\) be a dominating classical pattern. If
+    Let \(m =(\sigma, R)\) be a mesh pattern, where \(\boks{a,b} \notin R\),
+    and \(\pi = (\pi,\emptyset)\) be a dominating classical pattern. If
     \(\pi \preceq m^{\boks{a,b}_a}\)\\(\(\pi \preceq m^{\boks{a,b}_d}\)),
     then in any occurrence of \(m\) in a permutation \(\varrho\), the region
-    corresponding to the box \((a,b)\) can only contain an decreasing
+    corresponding to the box \(\boks{a,b}\) can only contain an decreasing
     (increasing) subsequence of \(\varrho\).
@@
 \end{equation*}
-we know that the region corresponding to the box \((1,1)\) contains a decreasing subsequence.
+we know that the region corresponding to the box \(\boks{1,1}\) contains a decreasing subsequence.
 If we let \(z\) be the topmost point in this decreasing subsequence, then \(xz\)
 is an occurrence of \(m_2\). This shows that our two example patterns are coincident.
-
-This result generalises into the following rule for categorising coincidences
-of mesh patterns in cases where there is a dominating classical pattern.
+This example generalises into the following rule.
 
@@
     and a dominating classical pattern \(\pi = (\pi,\emptyset)\) such that
-    \(\setsize{\pi} \le \setsize{\sigma} + 2\), the sets \(\av{\{\pi,m_1\}}\) and
-    \(\av{\{\pi,m_2\}}\) are coincident as a consequence of the Second Dominating Rule if
-
+    \(\setsize{\pi} \le \setsize{\sigma} + 2\), then \(m_1 \coinc{\pi} m_2 \)
+    if
     \begin{enumerate}
-        \item The mesh set \(R_2 = R_1 \cup \{(a,b)\}\)
+        \item The mesh set \(R_2 = R_1 \cup \{\boks{a,b}\}\)
         \item\label{prop:dom2:cond} Any one of the following four conditions hold \begin{enumerate}
@@
                         \begin{enumerate}
-                            \item \((a+1,b) \in \sigma\) and \((a+1,b-1)\notin R_1\) and \\
-                                \((x,b-1)\in R_1 \implies (x,b) \in R_1 \) (where \(x\neq a,a+1\)) and\\
-                                  \((a+1,y)\in R_1 \implies (a,y) \in R_1\) (where \(y\neq b-1,b\)).
-                            \item \((a,b+1) \in \sigma\) and \((a-1,b+1)\notin R_1\) and \\
-                                  \((x,b+1)\in R_1 \implies (x,b) \in R_1\) (where \(x\neq a-1,a\)) and\\
-                                  \((a-1,y)\in R_1 \implies (a,y) \in R_1\) (where \(y\neq b,b+1\)).
+                            \item \((a+1,b) \in \sigma\) and \(\boks{a+1,b-1}\notin R_1\) and \\
+                                \(\boks{x,b-1}\in R_1 \implies \boks{x,b} \in R_1 \) (where \(x\neq a,a+1\)) and\\
+                                  \(\boks{a+1,y}\in R_1 \implies \boks{a,y} \in R_1\) (where \(y\neq b-1,b\)).
+                            \item \((a,b+1) \in \sigma\) and \(\boks{a-1,b+1}\notin R_1\) and \\
+                                  \(\boks{x,b+1}\in R_1 \implies \boks{x,b} \in R_1\) (where \(x\neq a-1,a\)) and\\
+                                  \(\boks{a-1,y}\in R_1 \implies \boks{a,y} \in R_1\) (where \(y\neq b,b+1\)).
                         \end{enumerate}
@@
                         \begin{enumerate}
-                            \item \((a+1,b+1) \in \sigma\) and \((a+1,b+1)\notin R_1\) and \\
-                                  \((x,b+1)\in R_1 \implies (x,b) \in R_1\) (where \(x\neq a,a+1\)) and\\
-                                  \((a+1,y)\in R_1 \implies (a,y) \in R_1\) (where \(y\neq b,b+1\)).
-                            \item \((a,b) \in \sigma\) and \((a-1,b-1)\notin R_1\) and \\
-                                  \((x,b-1)\in R_1 \implies (x,b) \in R_1\) (where \(x\neq a-1,a\)) and\\
-                                  \((a-1,y)\in R_1 \implies (a,y) \in R_1\)  (where \(y\neq b-1,b\)).
+                            \item \((a+1,b+1) \in \sigma\) and \(\boks{a+1,b+1}\notin R_1\) and \\
+                                  \(\boks{x,b+1}\in R_1 \implies \boks{x,b} \in R_1\) (where \(x\neq a,a+1\)) and\\
+                                  \(\boks{a+1,y}\in R_1 \implies \boks{a,y} \in R_1\) (where \(y\neq b,b+1\)).
+                            \item \((a,b) \in \sigma\) and \(\boks{a-1,b-1}\notin R_1\) and \\
+                                  \(\boks{x,b-1}\in R_1 \implies \boks{x,b} \in R_1\) (where \(x\neq a-1,a\)) and\\
+                                  \(\boks{a-1,y}\in R_1 \implies \boks{a,y} \in R_1\)  (where \(y\neq b-1,b\)).
                         \end{enumerate}
@@
     We consider taking the first branch of every choice in condition~\eqref{prop:dom2:cond}.
-    Now consider a permutation \(\omega \in \av{\pi}\). Suppose
+    Consider a permutation \(\omega \in \av{\pi}\). Suppose
     \(\omega\) contains \(m_1\) and consider the region corresponding to
-    \((a,b)\) in \(R_1\). If the region is empty, the occurrence of \(m_1\) is trivially an occurrence
+    \(\boks{a,b}\) in \(R_1\). If the region is empty, the occurrence of \(m_1\) is trivially an occurrence
     of \(m_2\).
@@
     Hence there are no
-    points in the region corresponding to the box \((a,b)\) in the mesh
+    points in the region corresponding to the box \(\boks{a,b}\) in the mesh
     pattern, and therefore we can shade this region. This implies that every
-    occurrence of \(m_1\) in \(\av{\pi}\) is in fact an occurrence of \(m_2\)
+    occurrence of \(m_1\) in \(\av{\pi}\) can be modified into an occurrence of \(m_2\)
     so \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).
@@
 point in the mesh pattern.
-\FigureRef{fig:d2} show how the cases apply to a general container of \(m_1\) to transform it into a container of \(m_2\) , the circled points are the same points in the permutation.
+\FigureRef{fig:d2} show how the cases apply to a general container of \(m_1\) to transform it into a container of \(m_2\), the circled points are the same points in the permutation.
 \begin{figure}
@@
 
-Together the First Dominating Pattern Rule and the Second Dominating
-Pattern Rule fully explain coincidences of classes of the form
-\(\av{\{\perm{2,3,1},\mperm{2,1}{R}\}}\), obtaining 39 coincidence classes of
-mesh patterns, confirming experimental observations.
+After implementing the Second Dominating Pattern Rule we find that there are
+\(39\) \(\coinc[(2)]{231}\)-equivalence classes, of mesh patterns where the underlying
+classical pattern is \(21\). By \NoteRef{note:main} there are
+therefore exactly \(39\) \(\coinc{231}\)-equivalence classes.
 
@@
 we first apply the two Dominating Pattern rules previously established.
-Starting from \(220\) classes, we result in \(85 \coinc[(1)]{231}\)-coincidence classes
-reducing to \(59 \coinc[(2)]{231}\)-coincidence classes, considering mesh patterns with
-\(12\) as the underlying classical patterns. However we know that there are patterns
-where the coincidences are not explained by the rules given above.
+Starting from \(220\) classes, we result in \(85 \coinc[(1)]{231}\)-equivalence classes
+reducing to \(59 \coinc[(2)]{231}\)-equivalence classes, considering mesh patterns with
+\(12\) as the underlying classical patterns. However there are fewer \(\coinc[comp]{231}\)-equivalence classes.
 
@@
 \end{equation*}
-are experimentally coincident.
+are \(\coinc[comp]{231}\)-equivalent.
 
 Consider an occurrence of \(m_1\) in a permutation, if the region corresponding
-to the box \((1,0)\) is empty then we have an occurrence of \(m_2\). Now look at
+to the box \(\boks{1,0}\) is empty then we have an occurrence of \(m_2\). Now look at
 the case when this region is not empty, and consider choosing the rightmost
 point in the region.
-
 This gives us an occurrence of the following mesh pattern.
@@
 \end{equation*}
-By application of \PropositionRef{prop:dom1} we then achieve the following
+By application of First Dominating Patter Rule we obtain the following
 mesh pattern
@@
 \end{equation*}
-If we look at the highlighted points we see that the subpattern is an occurrence
-of the mesh pattern that we originally desired. This gives rise to the following
+The highlighted points are an occurrence
+of the mesh pattern \(m_2\). This gives rise to the following
 rule:
@@
     Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
-    and a dominating classical pattern \(\pi = (\pi,\emptyset)\), the sets
-    \(\av{\{\pi,m_1\}}\) and \(\av{\{\pi,m_2\}}\) are coincident if
+    and a dominating classical pattern \(\pi = (\pi,\emptyset)\), then
+    \(m_1 \coinc{\pi} m_2\) if
     \begin{enumerate}
-        \item The mesh set \(R_2 = R_1 \cup \{(a,b)\}\)
-        \item\label{prop:dom3:condocc} One of the patterns in \((\sigma,R_1)^\boksall{a,b}\)
+        \item The mesh set \(R_2 = R_1 \cup \{\boks{a,b}\}\)
+        \item\label{prop:dom3:condocc} One of the patterns in \(m_1^\boksall{a,b}\)
             is coincident with a mesh pattern containing an occurrence of
-            \(\mperm{\sigma}{R_2}\) as a subpattern.
+            \(m_2\) as a subpattern.
     \end{enumerate}
@@
 \begin{proof}
-    Assume without meaningful loss of generality that \(R_2 = R_1 \cup \{(a,b)\}\).
-    \NoteRef{not:downcmesh}, implies \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.
-    Now consider a permutation \(\sigma\) in \(\av{\pi}\) that contains an
-    occurrence of \(m_1\). If the region corresponding to the box \((a,b)\)
-    is empty then we have an occurrence of \(m_2\). If the region is non-empty
-    then by condition~\ref{prop:dom3:condocc}  of the proposition there exists
-    a pattern in \((\sigma,R_1)^\boksall{a,b}\) such that there exists an occurrence of a mesh pattern of
-    length one longer than \(m_1\) in this position. This mesh pattern is
-    coincident with another mesh pattern that contains an
-    occurrence of \(m_2\). Hence, every occurrence of \(m_1\) leads to an occurrence
-    of \(m_2\).  Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the
+    \NoteRef{not:downcmesh} implies \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.
+    Now consider a permutation \(\omega\) in \(\av{\pi}\) that contains an
+    occurrence of \(m_1\). If the region corresponding to the box \(\boks{a,b}\)
+    is empty then we have an occurrence of \(m_2\).
+
+    By condition~\ref{prop:dom3:condocc} of the proposition there exists
+    a pattern \(r\) in \(m_1^\boksall{a,b}\) such that \(m_2\) is a subpattern in some
+    pattern \(r'\) coincident with \(r\). If the region is non-empty then there exists
+    an occurrence of \(r\) in \(\omega\). Therefore there is an occurrence of \(r'\)
+    in \(\omega\), which implies that there is an occurrence of \(m_2\) in \(\omega\).
+    Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the
     two patterns are coincident.

### 2017-10-12 writing/paper/sections/coincidences.tex
Near final fixes

@@
     \NoteRef{not:downcmesh} implies \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.
-    Now consider a permutation \(\omega\) in \(\av{\pi}\) that contains an
+    Now consider a permutation \(\omega \in \av{\pi}\) that contains an
     occurrence of \(m_1\). If the region corresponding to the box \(\boks{a,b}\)
     is empty then we have an occurrence of \(m_2\).
-
     By condition~\ref{prop:dom3:condocc} of the proposition there exists
@@
 
-This rule reduces the number of classes in
-\(\av{\{\perm{2,3,1},(\perm{1,2}, R)\}}\) to \(56\), which coincides with the
-56 classes observed through experimentation.
+After implementing the Third Dominating Pattern Rule we find that there are
+\(56\) \(\coinc[(3)]{231}\)-equivalence classes, of mesh patterns where the underlying
+classical pattern is \(12\). By \NoteRef{note:main} there are
+therefore exactly \(56\) \(\coinc{231}\)-equivalence classes, since we have
+observed \(56\) \(\coinc[comp]{231}\)-equivalence classes through experimentation.
 
@@
 pattern \(\perm{1,2}\) in \(\av{\perm{3,2,1}}\) application of the previously
-established rules give no coincidences. Through experimentation we discover
-that there are \(7\) non-trivial coincidence classes (all others are singletons)
-which can be explained through the
-use of two different lines of reasoning. Since the number of coincidences is
+established rules gives no coincidences. Through experimentation we discover
+that there are \(7\) \(\coinc[comp]{321}\)-equivalence classes are unexplained.
+Since the number of coincidences is
 so small we will reason for these coincidences without attempting to generalise
@@
 
-Intuitively it is easy to see why our previous rules have no
+Intuitively, it is easy to see why our previous rules have no
 power here. It is impossible to add a single point to a mesh pattern \((12, R)\) and
@@
 \end{equation*}
-are coincident in \(\av{\perm{3,2,1}}\). (There are 3 symmetries of these
-patterns that are also coincident to each other by the same reasoning.)
+are \(\coinc[comp]{321}\)-equivalent. (There are 3 symmetries of these
+patterns that are also equivalent to each other by the same reasoning.)
 
-Consider the region corresponding to the box \((0,1)\) in any occurrence of \(m_1\),
+Consider the region corresponding to the box \(\boks{0,1}\) in any occurrence of \(m_1\)
 in a permutation. By \LemmaRef{lem:incdec} it must contain an increasing subsequence.
 If the region is empty then we have an occurrence of \(m_2\).
-If there is only one point in the region we can choose this to replace the \(1\) in
+If there is only one point in the region we can choose this point to replace the \(1\) in
 the mesh pattern to get the required shading. If there is more than one point then
@@
 \end{equation*}
-Where the two highlighted points are the original two points.
+Here the two highlighted points are the original two points.
 The other two points are an occurrence of the pattern
-we originally desired, and hence the two patterns are coincident.
+we \(m_2\), and hence the two patterns are coincident.
 It is also possible to calculate this coincidence by an extension of the Third
@@
 is discussed further in the future work section.
+Because of symmetries we have dealt with \(4\) out of the \(7\) \(\coinc[comp]{321}\)-equivalence classes.
 
-The other reasoning applies to the patterns
+Now consider the patterns
 \begin{equation*}
@@
 \end{equation*}
-which are coincident by experimentation.
+which are \(\coinc[comp]{321}\)-equivalent.
 
 In order to prove this coincidence we will proceed by mathematical
-induction on the number of points in region corresponding to the middle box.
+induction on the number of points in the region corresponding to the middle box.
 We call this number \(n\).
@@
         any points then we gain some extra shading on the mesh pattern
-        as any points in these boxes would create an occurrence of the dominating pattern \(\perm{3,2,1}\)
+        as any points in the boxes \(\boks{1,2}\) and \(\boks{1,3}\) would create an occurrence of the dominating pattern \(\perm{3,2,1}\)
      \begin{equation*}
@@
 This argument applies to another two pairs of classes. Therefore in
-total in \(\av{\{\perm{3,2,1}, \mperm{1,2}{R}\}}\) there are \(213\) coincidence
+total in \(\av{\{\perm{3,2,1}, \mperm{1,2}{R}\}}\) there are \(\coinc{321}\)-equivalence
 classes.

### 2017-10-12 writing/paper/sections/wilf.tex
Near final fixes

@@
 Since we are always considering Wilf-equivalences in a set \(\av{S}\) we
-must only use symmetries that preserve the dominating pattern(s), if we
+must only use symmetries that preserve the dominating pattern(s) in \(S\), if we
 were to allow other symmetries, then the equivalences calculated in the
@@
 The elements to the left of the maximum, \(A\), have the structure of a \(231\)
-avoiding permutation, and the elements to the right of the maximum, \(B\), have
+avoiding permutation, and the elements to the right of the maximum, \(B\), also have
 the structure of a \(\perm{2,3,1}\) avoiding permutation. Furthermore, all the
@@
 The pattern \(\textpattern{}{1}{0/0,1/0,0/1,1/1}\) is in its own
-Wilf-class since the only permutation containing this pattern is
+Wilf-class since the only permutation containing it is
 the permutation \(\perm{1}\). The avoiders of this pattern therefore
@@
 \nextvar[\varfine]
-It can be shown by use of \PropositionRef{prop:dom2} that the patterns
+It can be shown by use of the Second Dominating Pattern Rule that the patterns
 \textpattern{}{1}{0/0,1/1} and \(q_1 = \textpattern{}{1}{0/0,0/1,1/1}\) are
-coincident inside \(\av{\perm{2,3,1}}\). Consider the decomposition of a non-empty avoider of \(q_1\) in
+coincident under \(\perm{2,3,1}\). Consider the decomposition of a non-empty avoider of \(q_1\) in
 \(\av{\perm{2,3,1}}\) around the maximum:
@@
 \section{Wilf-classes with patterns of length 2}
-By the use of coincidence classes established in
-\SectionRef{sec:coincs} we know that there are at most 95 Wilf-equivalence classes.
+From \SectionRef{sec:coincs} we know there are \(95\) coincidence classes and
+therefore at most that many Wilf-equivalence classes.
 
@@
 Computing avoiders up to length 10 suggests that there are at least 23 Wilf-classes,
-of which 13 are non-trivial. Therefore there are Wilf-equivalences that are not
+of which 13 are not equal to coincidence classes (after taking symmetries). Therefore there seem to be Wilf-equivalences that are not
 explained by coincidences or symmetry.
 
-We will use two main methods of establishing Wilf-equivalence between mesh patterns of length \(2\)
-in \(\av{\perm{2,3,1}}\): the structural decomposition of avoiders, via generating functions; or the
+We will use two main methods of establishing these Wilf-equivalences: the structural decomposition of avoiders, via generating functions; or the
 structure of a general permutation containing the pattern, looking at a particular occurrence of the
@@
 
-If there is an occurrence of \(m_1\) there must be a lowest occurrence of \(m_1\), \ie the occurrence has the lowest possible values for any occurrence. Now consider the top
+If there is an occurrence of \(m_1\) there must be a lowest occurrence of \(m_1\), \ie, the occurrence has the lowest possible values for any occurrence. Now consider the top
 regions, labelled \(A \et B\), the subpermutation contained in these regions must avoid the permutation \(231\).

### 2017-10-13 writing/paper/sections/wilf.tex
USe Wilf-symbol

@@
 Throughout this section we will consider Wilf-equivalences of patterns
-whilst avoiding the \emph{dominating pattern} \(\perm{2,3,1}\). We will
+whilst avoiding the \emph{dominating pattern} \(\perm{2,3,1}\), if two patterns \(p_1 \et p_2\) are Wilf-equivalent we will denote this \(p_1\weqv{231}p_2\), similarly when computational methods indicate that two patterns are Wilf equivalent under \(231\) we write \(p_1\weqv[comp]{231}p_2\).
+We will
 use \(\mathcal{C}\) to denote \(\av{\perm{2,3,1}}\) and \(C(x)\) will
@@
 The pattern \(\textpattern{}{1}{0/0,1/0,0/1,1/1}\) is in its own
-Wilf-class since the only permutation containing it is
+\(\weqv{231}\)-equivalence class since the only permutation containing it is
 the permutation \(\perm{1}\). The avoiders of this pattern therefore
@@
 The pattern \(q_3 = \textpattern{}{1}{0/0,0/1,1/0}\) is the reverse-complement-inverse
-of \(q_2\) and hence the avoiders of \(q_2\) and \(q_3\) (\(\scriptvar_2\) and \(\scriptvar_3\)) are equinumerous.
+of \(q_2\) and hence the avoiders of \(q_2\) and \(q_3\) (\(\scriptvar_2\) and \(\scriptvar_3\)) are equinumerous, and so \(q_1 \weqv{231} q_2 \weqv{231}q_3\).
 All of these classes have the same generating function, namely
@@
 
-Computing avoiders up to length 10 suggests that there are at least 23 Wilf-classes,
+Computing avoiders up to length 10 suggests that there are at least 23 \(\weqv{231}\)-equivalence classes,
 of which 13 are not equal to coincidence classes (after taking symmetries). Therefore there seem to be Wilf-equivalences that are not
@@
 \nextvar
-The following patterns are experimentally Wilf-equivalent up to length 10 in
-\(\av{\perm{2,3,1}}\)
+The following patterns are in the same \(\weqv[comp]{231}\)-equivalence class
 \begin{gather}
@@
 
-First we prove the Wilf-equivalence between \(m_1\) and \(m_2\) shown in
-\eqref{eq:singleud}, by considering the form of a general permutation containing
+First we prove that \(m_1 \weqv{231} m_2\) by considering the form of a general permutation containing
 either of the two patterns. First looking at a general occurrence of \(m_1\) in a permutation
@@
 enumeration, \(\scriptvar_3\) and \(\scriptvar_4\) must also have
-the same enumeration and are therefore Wilf-equivalent.
+the same enumeration and are therefore \(m_3\weqv{m_4}\).
 
@@
 It is therefore obvious that avoiders of \(m_1\) and avoiders of \(m_4\)
-have the same enumeration, and therefore all four patterns are Wilf-equivalent
-in \(\av{\perm{2,3,1}}\) with generating function satisfying
+have the same enumeration, and therefore \(m_1 \weqv{231} m_2 \weqv{231}m_3\weqv{m_4}\) with generating function satisfying
 \begin{equation*}
@@
 \nextvar
-The following patterns are experimentally Wilf-equivalent up to length 10 in
-\(\av{\perm{2,3,1}}\)
+The patterns \(m_1 \et m_2\) are in the same \(\weqv[comp]{231}\)-equivalence class
 \begin{equation*}

### 2017-10-13 writing/paper/sections/conclfut.tex
Maybe final changes from Henning|

@@
  \begin{example}
-   We can establish the coincidence between the patterns
+   The coincidence of the patterns
    \begin{equation*}
@@
    \end{equation*}
-   That is not explained by the methods presented by \textcite{DBLP:journals/corr/ClaessonTU14}.
-
+	does not follow from the general methods presented by \textcite{DBLP:journals/corr/ClaessonTU14}, but is rather handled there as
+	a special case. We can do it as follows:
    Consider a permutation containing \(m_1\),

### 2017-10-13 writing/paper/sections/wilf.tex
Maybe final changes from Henning|

@@
 Wilf-equivalence is an important aspect to study in the field of permutation
-patterns.
+patterns. The original definition is as follows:
 
@@
 
-Since we are always considering Wilf-equivalences in a set \(\av{S}\) we
+Since we will consider Wilf-equivalences in a set \(\av{S}\) we
 must only use symmetries that preserve the dominating pattern(s) in \(S\), if we
@@
 
-Throughout this section we will consider Wilf-equivalences of patterns
+In the remainder of the paper we will consider Wilf-equivalences of patterns
 whilst avoiding the \emph{dominating pattern} \(\perm{2,3,1}\), if two patterns \(p_1 \et p_2\) are Wilf-equivalent we will denote this \(p_1\weqv{231}p_2\), similarly when computational methods indicate that two patterns are Wilf equivalent under \(231\) we write \(p_1\weqv[comp]{231}p_2\).
@@
 the Wilf-equivalence classes of the mesh patterns of length \(1\) inside
-\(\av{\perm{2,3,1}}\), this means that we are considering the set
-\(\av{\{\perm{2,3,1},p}\}\)
-where \(p\) is a mesh-pattern of length 1.
+\(\av{\perm{2,3,1}}\), this means that we are considering the \(\weqv{231}\)-equivalence classes of mesh patterns with underlying classical pattern \(1\).
 
@@
 
-Computing avoiders up to length 10 suggests that there are at least 23 \(\weqv{231}\)-equivalence classes,
+Computing avoiders up to length 10 gives 23 \(\weqv[comp]{231}\)-equivalence classes,
 of which 13 are not equal to coincidence classes (after taking symmetries). Therefore there seem to be Wilf-equivalences that are not
@@
 If there is an occurrence of \(m_1\) there must be a lowest occurrence of \(m_1\), \ie, the occurrence has the lowest possible values for any occurrence. Now consider the top
-regions, labelled \(A \et B\), the subpermutation contained in these regions must avoid the permutation \(231\).
-Also, the subpermutation in the region \(A\) must be a decreasing subsequence, otherwise an occurrence of \(231\) will be created with either of the points in the occurrence.
+regions, labelled \(A \et B\), the subpermutation contained in the union of these regions must avoid the permutation \(231\).
+Also, the subpermutation in the region \(A\) must be a decreasing subsequence, otherwise an occurrence of \(231\) will be created with either of the points in the occurrence of \(m_1\).
 Now, consider the region labelled \(C\), since we specified that we were focused on the lowest possible occurrence of \(m_1\) this region cannot contain an occurrence of either \(m_1\) or \(231\).
-We now have a structural decomposition of a container of \(m_1\) inside \(\av{231}\).
+This is a full structural decomposition of a container of \(m_1\) inside \(\av{231}\).
 
@@
 
-Similarly to a container of \(m_1\) we will look at the lowest possible occurrence of \(m_2\), similar to before the regions \(A^\prime \et B^\prime\) together contain an avoider of \(231\) and \(A\) contains a decreasing subsequence.
-Since we are considering the lowest occurrence of \(m_2\) so the region \(C^\prime\) does not contain an occurrence of either \(m_2\) or \(231\).
+Similarly to a container of \(m_1\) we will look at the lowest possible occurrence of \(m_2\), and as before the regions \(A^\prime \et B^\prime\) together contain an avoider of \(231\) and \(A\) contains a decreasing subsequence.
+Since we are considering the lowest occurrence of \(m_2\) the region \(C^\prime\) does not contain an occurrence of either \(m_2\) or \(231\).
 
@@
 element with largest value in this region, therefore the partial permutation
-in this region must avoid starting with the maximum.
+in this region must avoid starting with the maximum, \ie, be in \(\mathcal{\varmaxl}_1 = \av{\textpattern{}{1}{0/0,0/1,1/1}}\) described
+in \SectionRef{wilf1}.
 
@@
 this region, so this top right region must contain a sub-permutation that does not
-start with it's minimum, i.e., it is a member of \(\mathcal{\varmaxl}_3\) described
-in \SectionRef{wilf1}.
+start with it's minimum, i.e., it is a member of \(\mathcal{\varmaxl}_3 = \av{\av{\textpattern{}{1}{0/0,0/1,1/0}}}\).
 Since both \(\mathcal{\varmaxl}_1\) and \(\mathcal{\varmaxl}_3\) have the same
 enumeration, \(\scriptvar_3\) and \(\scriptvar_4\) must also have
-the same enumeration and are therefore \(m_3\weqv{m_4}\).
+the same enumeration and are therefore \(m_3\weqv{231}m_4\).
 
@@
 \end{equation*}
-It is therefore obvious that avoiders of \(m_1\) and avoiders of \(m_4\)
+It is therefore clear that avoiders of \(m_1\) and avoiders of \(m_4\)
 have the same enumeration, and therefore \(m_1 \weqv{231} m_2 \weqv{231}m_3\weqv{m_4}\) with generating function satisfying
@@
 In order to establish the remainder of the Wilf-equivalences of the form \(\av{\{\perm{2,3,1},p\}}\) where \(p\)
-is a mesh pattern we can use similar methods to allow us to consolidate experimental classes into
+is a mesh pattern of length \(2\) we can use similar methods to allow us to consolidate experimental classes into
 actual classes, these methods allow us to explain all 23 of the observed Wilf-classes seen in experimentation.

### 2017-10-14 writing/paper/sections/coincidences.tex
Final edits from Murray
Fix minor speling mistakes "patter"->"pattern"
"leftmostpoint"->"leftmost point"

Reflow text to line length 80

@@
 Coincidences among small mesh patterns have been considered by
-\textcite{DBLP:journals/corr/ClaessonTU14}, in which the authors use the Simultaneous Shading Lemma,
-a closure result and one special case to fully classify coincidences among mesh patterns
-of length 2.
+\textcite{DBLP:journals/corr/ClaessonTU14}, in which the authors use the
+Simultaneous Shading Lemma, a closure result and one special case to fully
+classify coincidences among mesh patterns of length 2.
 
-Recall that two patterns \(\lambda\) and \(\gamma\) are considered \emph{coincident} if the set of permutations
-that avoid \(\lambda\) is the same as the set of permutations that avoid \(\gamma\),
-\ie, \(\av{\lambda} = \av{\gamma}\). This is equivalent to \(\lambda\) and \(\gamma\) being contained in
-the same set of permutations, \ie, \(\cont{\lambda} = \cont{\gamma}\).
+Recall that two patterns \(\lambda\) and \(\gamma\) are considered
+\emph{coincident} if the set of permutations that avoid \(\lambda\) is the same
+as the set of permutations that avoid \(\gamma\), \ie, \(\av{\lambda} =
+\av{\gamma}\). This is equivalent to \(\lambda\) and \(\gamma\) being contained
+in the same set of permutations, \ie, \(\cont{\lambda} = \cont{\gamma}\).
 
@@
 \label{note:main}
-    If we find that \(\coinc[(i)]{\pi}\), for \(i = 1, 2\), or \(3\),
-    equals \(\coinc[comp]{\pi}\) then the equivalence relations \(\coinc[(i)]{\pi}\), \(\coinc{\pi}\),
-    and \(\coinc[comp]{\pi}\) are all equal.
+    If we find that \(\coinc[(i)]{\pi}\), for \(i = 1, 2\), or \(3\), equals
+    \(\coinc[comp]{\pi}\) then the equivalence relations \(\coinc[(i)]{\pi}\),
+    \(\coinc{\pi}\), and \(\coinc[comp]{\pi}\) are all equal.
 \end{note}
 
-In order to describe the rules it is useful to have a notion for inserting points, ascents, and
-descents into a mesh pattern.
+In order to describe the rules it is useful to have a notion for inserting
+points, ascents, and descents into a mesh pattern.
 \begin{definition}
 \label{def:ap}
-Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\). We define
-a mesh pattern \(p^\boks{i,j} = (\tau^\prime,R^\prime)\) of length \(n+1\) as the pattern where a
-point is \emph{inserted} into the box \(\boks{i,j}\) in \(G(p)\). Formally the new underlying
-classical pattern is defined by
+Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that
+\(\boks{i,j}\notin R\). We define a mesh pattern \(p^\boks{i,j} =
+(\tau^\prime,R^\prime)\) of length \(n+1\) as the pattern where a point is
+\emph{inserted} into the box \(\boks{i,j}\) in \(G(p)\). Formally the new
+underlying classical pattern is defined by
 \begin{equation*}
@@
 In addition, we give the following definitions:
-
 \begin{definition}
-Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\) and
-\(p^\boks{i,j} = (\tau^\prime,R^\prime)\) is as defined in \DefinitionRef{def:ap}. We define
-the following four modifications of \(p^\boks{i,j}\).
+Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that
+\(\boks{i,j}\notin R\) and \(p^\boks{i,j} = (\tau^\prime,R^\prime)\) is as
+defined in \DefinitionRef{def:ap}. We define the following four modifications of
+\(p^\boks{i,j}\).
 \begin{align*}
@@
 \end{definition}
-Informally, these are considering the topmost, rightmost, leftmost, or bottommost point in
-\(\boks{i,j}\). We collect the resulting mesh patterns in a set
+Informally, these are considering the topmost, rightmost, leftmost, or
+bottommost point in \(\boks{i,j}\). We collect the resulting mesh patterns in a
+set
 \begin{equation*}
@@
 \begin{definition}
-Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\). We define
-a mesh pattern \(p^{\boks{i,j}_a} = (\tau^\prime,R^\prime)\) (\(p^{\boks{i,j}_d}\)) of length \(n+2\)
-as the pattern where an ascent (descent) is \emph{inserted} into the box \(\boks{i,j}\) in \(G(p)\).
-Formally the new underlying classical pattern is defined by
+Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that
+\(\boks{i,j}\notin R\). We define a mesh pattern \(p^{\boks{i,j}_a} =
+(\tau^\prime,R^\prime)\) (\(p^{\boks{i,j}_d}\)) of length \(n+2\) as the pattern
+where an ascent (descent) is \emph{inserted} into the box \(\boks{i,j}\) in
+\(G(p)\). Formally the new underlying classical pattern is defined by
 \begin{equation*}
@@
 \end{equation*}
-\caption{The result of inserting an ascent into \(p=(231,\{\boks{0,0},\boks{1,0},\boks{1,3},\boks{3,0},\boks{3,1},\boks{3,3}\})\)}
+\caption{The result of inserting an ascent into
+\(p=(231,\{\boks{0,0},\boks{1,0},\boks{1,3},\boks{3,0},\boks{3,1},\boks{3,3}\})\)}
 \label{fig:adda}
@@
 
-We now attempt to fully classify coincidences in families characterised by avoidance
-of a classical pattern of length \(3\) and a mesh pattern of length \(2\), that is
-finding and explaining all coincidences between mesh patterns \(m\) and \(m^\prime\), \(m\coinc{\pi} m^\prime\),
-where \(\pi\) is a classical pattern of length \(3\).
+We now attempt to fully classify coincidences in families characterised by
+avoidance of a classical pattern of length \(3\) and a mesh pattern of length
+\(2\), that is finding and explaining all coincidences between mesh patterns
+\(m\) and \(m^\prime\), \(m\coinc{\pi} m^\prime\), where \(\pi\) is a classical
+pattern of length \(3\).
 
 It can be easily seen that in order to classify coincidences one need only
-consider coincidences within the family of mesh patterns with the same underlying
-classical pattern, this is due to the fact that \(\perm{2,1}\in\av{\mperm{1,2}{R}}\)
-and \(\perm{1,2}\in\av{\{\mperm{2,1}{R}\}}\) for all mesh-sets \(R\).
+consider coincidences within the family of mesh patterns with the same
+underlying classical pattern, this is due to the fact that
+\(\perm{2,1}\in\av{\mperm{1,2}{R}}\) and
+\(\perm{1,2}\in\av{\{\mperm{2,1}{R}\}}\) for all mesh-sets \(R\).
 
-We know that there are a total of \(512\) mesh-sets for each underlying classical
-pattern. By use of the previous results of \textcite{DBLP:journals/corr/ClaessonTU14}
-the number of coincidence classes can be reduced to \(220\).
+We know that there are a total of \(512\) mesh-sets for each underlying
+classical pattern. By use of the previous results of
+\textcite{DBLP:journals/corr/ClaessonTU14} the number of coincidence classes can
+be reduced to \(220\).
 
 \subsection{Coincidence classes of Av(\{321, (21, \textit{R})\})}
-Through experimentation, considering avoidance of permutations up to length \(11\), we discover
-that there are \(29\) \(\coinc[comp]{321}\)-equivalence classes where the underlying classical pattern of the mesh pattern is \(21\).
+Through experimentation, considering avoidance of permutations up to length
+\(11\), we discover that there are \(29\) \(\coinc[comp]{321}\)-equivalence
+classes where the underlying classical pattern of the mesh pattern is \(21\).
 
@@
 
-    Now we consider a permutation \(\omega \in \av{\pi}\),
-    containing an occurrence of \(m_1\). If there is a point in the
-    region corresponding to the box \(\boks{a,b}\), then that point, along with the
-    points of the occurrence of \(m_1\), form an occurrence of \(\sigma^\boks{a,b}\).
-    Then condition~\eqref{prop:dom1:cont} of the
-    proposition implies an occurrence of \(\pi\). Therefore there can be no
-    points in this region, which implies the occurrence of \(m_1\) is an occurrence of \(m_2\). Hence every occurrence of
+    Now we consider a permutation \(\omega \in \av{\pi}\), containing an
+    occurrence of \(m_1\). If there is a point in the region corresponding to
+    the box \(\boks{a,b}\), then that point, along with the points of the
+    occurrence of \(m_1\), form an occurrence of \(\sigma^\boks{a,b}\). Then
+    condition~\eqref{prop:dom1:cont} of the proposition implies an occurrence of
+    \(\pi\). Therefore there can be no points in this region, which implies the
+    occurrence of \(m_1\) is an occurrence of \(m_2\). Hence every occurrence of
     \(m_1\) is in fact an occurrence of \(m_2\), and we have that
@@
 After implementing the First Dominating Pattern Rule we find that there are
-\(29\) \(\coinc[(1)]{321}\)-equivalence classes, of mesh patterns where the underlying
-classical pattern is \(21\). By \NoteRef{note:main} there are
+\(29\) \(\coinc[(1)]{321}\)-equivalence classes, of mesh patterns where the
+underlying classical pattern is \(21\). By \NoteRef{note:main} there are
 therefore exactly \(29\) \(\coinc{321}\)-equivalence classes.
@@
 \subsection{Coincidence classes of Av(\{231, (21, \textit{R})\}).}
-By application of First Dominating Pattern Rule we obtain \(43\) \(\coinc[(1)]{231}\)-equivalence
-classes between mesh patterns with \(21\) as an underlying classical pattern.
-Experimentation shows that there are \(39 \coinc[comp]{231}\)-equivalence
-classes between mesh patterns with \(21\) as an underlying classical pattern, for example the following two patterns are in the same
-\(\coinc[comp]{231}\)-equivalence class, but this is not explained by the First Dominating Pattern Rule.
+By application of First Dominating Pattern Rule we obtain \(43\)
+\(\coinc[(1)]{231}\)-equivalence classes between mesh patterns with \(21\) as an
+underlying classical pattern. Experimentation shows that there are \(39
+\coinc[comp]{231}\)-equivalence classes between mesh patterns with \(21\) as an
+underlying classical pattern, for example the following two patterns are in the
+same \(\coinc[comp]{231}\)-equivalence class, but this is not explained by the
+First Dominating Pattern Rule.
 \begin{equation*}
@@
 
-Consider an occurrence of \(m_1\) in a permutation in \(\av{\perm{2,3,1}}\), consisting
-of elements \(x\) and \(y\). If the region corresponding
-to the box \(\boks{1,1}\) is empty we have an occurrence of \(m_2\). Otherwise, if there is any
-ascent in this box then we would have an occurrence of \(\perm{2,3,1}\),
-however, since the permutation is in \(\av{\perm{2,3,1}}\) this is not possible. This box
-must therefore contain a (non-empty) decreasing subsequence.
- This gives rise to the following lemma:
+Consider an occurrence of \(m_1\) in a permutation in \(\av{\perm{2,3,1}}\),
+consisting of elements \(x\) and \(y\). If the region corresponding to the box
+\(\boks{1,1}\) is empty we have an occurrence of \(m_2\). Otherwise, if there is
+any ascent in this box then we would have an occurrence of \(\perm{2,3,1}\),
+however, since the permutation is in \(\av{\perm{2,3,1}}\) this is not possible.
+This box must therefore contain a (non-empty) decreasing subsequence. This gives
+rise to the following lemma:
 
@@
 \end{equation*}
-we know that the region corresponding to the box \(\boks{1,1}\) contains a decreasing subsequence.
-If we let \(z\) be the topmost point in this decreasing subsequence, then \(xz\)
-is an occurrence of \(m_2\). This shows that our two example patterns are coincident.
-This example generalises into the following rule.
+we know that the region corresponding to the box \(\boks{1,1}\) contains a
+decreasing subsequence. If we let \(z\) be the topmost point in this decreasing
+subsequence, then \(xz\) is an occurrence of \(m_2\). This shows that our two
+example patterns are coincident. This example generalises into the following
+rule.
 
@@
   \item The mesh set \(R_2 = R_1 \cup \{\boks{a,b}\}\)
-        \item\label{prop:dom2:cond} Any one of the following four conditions hold \begin{enumerate}
+  \item\label{prop:dom2:cond} Any one of the following four conditions hold
+    \begin{enumerate}
     \item\label{prop:dom2:condc} \(\pi \preceq \sigma^{\boks{a,b}_a}\) and
@@
 \begin{proof}
-    By \NoteRef{not:downcmesh} we only need to show that an occurrence of \(m_1\)
-    implies an occurrence of \(m_2\).
-    We consider taking the first branch of every choice in condition~\eqref{prop:dom2:cond}.
-    Consider a permutation \(\omega \in \av{\pi}\). Suppose
-    \(\omega\) contains \(m_1\) and consider the region corresponding to
-    \(\boks{a,b}\) in \(R_1\). If the region is empty, the occurrence of \(m_1\) is trivially an occurrence
+    By \NoteRef{not:downcmesh} we only need to show that an occurrence of
+    \(m_1\) implies an occurrence of \(m_2\). We consider taking the first
+    branch of every choice in condition~\eqref{prop:dom2:cond}. Consider a
+    permutation \(\omega \in \av{\pi}\). Suppose \(\omega\) contains \(m_1\) and
+    consider the region corresponding to \(\boks{a,b}\) in \(R_1\). If the
+    region is empty, the occurrence of \(m_1\) is trivially an occurrence of
+    \(m_2\).
+
+    If the region is non-empty, then by \LemmaRef{lem:incdec} and
+    condition~\eqref{prop:dom2:condc} of the proposition it must contain a
+    decreasing subsequence. We can choose the topmost point in the region to
+    replace the corresponding point in the mesh pattern and the points from the
+    subsequence are now in the box southeast of the point.
+
+    The conditions on the mesh ensure that no elements of the permutation that
+    were inside a region corresponding to an unshaded box in the occurrence of
+    \(m_1\) would be in a region corresponding to a shaded box in an occurrence
     of \(m_2\).
 
-    If the region is non-empty, then by \LemmaRef{lem:incdec} and
-    condition~\eqref{prop:dom2:condc} of the proposition it must contain
-    a decreasing subsequence.
-    We can choose the topmost point in the region to replace the corresponding
-    point in the mesh pattern and the points from the subsequence are now in
-    the box southeast of the point.
-
-    The conditions on the mesh ensure that no elements of the permutation
-    that were inside a region corresponding to an unshaded box in the occurrence
-    of \(m_1\) would be in a region corresponding to a shaded box in an occurrence of
-    \(m_2\).
-
-    Hence there are no
-    points in the region corresponding to the box \(\boks{a,b}\) in the mesh
-    pattern, and therefore we can shade this region. This implies that every
-    occurrence of \(m_1\) in \(\av{\pi}\) can be modified into an occurrence of \(m_2\)
-    so \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\).
+    Hence there are no points in the region corresponding to the box
+    \(\boks{a,b}\) in the mesh pattern, and therefore we can shade this region.
+    This implies that every occurrence of \(m_1\) in \(\av{\pi}\) can be
+    modified into an occurrence of \(m_2\) so \(\av{\{\pi,m_2\}} \subseteq
+    \av{\{\pi,m_1\}}\).
 
@@
 
-This proposition essentially states that we slide all of the points in the box we desire
-to shade diagonally, and chose the topmost/bottommost point to replace the original
-point in the mesh pattern.
-\FigureRef{fig:d2} show how the cases apply to a general container of \(m_1\) to transform it into a container of \(m_2\), the circled points are the same points in the permutation.
+This proposition essentially states that we slide all of the points in the box
+we desire to shade diagonally, and chose the topmost/bottommost point to replace
+the original point in the mesh pattern. \FigureRef{fig:d2} show how the cases
+apply to a general container of \(m_1\) to transform it into a container of
+\(m_2\), the circled points are the same points in the permutation.
 \begin{figure}
@@
 After implementing the Second Dominating Pattern Rule we find that there are
-\(39\) \(\coinc[(2)]{231}\)-equivalence classes, of mesh patterns where the underlying
-classical pattern is \(21\). By \NoteRef{note:main} there are
+\(39\) \(\coinc[(2)]{231}\)-equivalence classes, of mesh patterns where the
+underlying classical pattern is \(21\). By \NoteRef{note:main} there are
 therefore exactly \(39\) \(\coinc{231}\)-equivalence classes.
@@
 \subsection{Coincidence classes of Av(\{231, (12, \textit{R})\}).}
-When considering the coincidence classes of \(\av{\{\perm{2,3,1},\mperm{1,2}{R}\}}\)
-we first apply the two Dominating Pattern rules previously established.
-Starting from \(220\) classes, we result in \(85 \coinc[(1)]{231}\)-equivalence classes
-reducing to \(59 \coinc[(2)]{231}\)-equivalence classes, considering mesh patterns with
-\(12\) as the underlying classical patterns. However there are fewer \(\coinc[comp]{231}\)-equivalence classes.
+When considering the coincidence classes of
+\(\av{\{\perm{2,3,1},\mperm{1,2}{R}\}}\) we first apply the two Dominating
+Pattern rules previously established. Starting from \(220\) classes, we result
+in \(85 \coinc[(1)]{231}\)-equivalence classes reducing to \(59
+\coinc[(2)]{231}\)-equivalence classes, considering mesh patterns with \(12\) as
+the underlying classical patterns. However there are fewer
+\(\coinc[comp]{231}\)-equivalence classes.
 
@@
 Consider an occurrence of \(m_1\) in a permutation, if the region corresponding
-to the box \(\boks{1,0}\) is empty then we have an occurrence of \(m_2\). Now look at
-the case when this region is not empty, and consider choosing the rightmost
-point in the region.
-This gives us an occurrence of the following mesh pattern.
+to the box \(\boks{1,0}\) is empty then we have an occurrence of \(m_2\). Now
+look at the case when this region is not empty, and consider choosing the
+rightmost point in the region. This gives us an occurrence of the following mesh
+pattern.
 \begin{equation*}
@@
 \end{equation*}
-By application of First Dominating Patter Rule we obtain the following
-mesh pattern
+By application of First Dominating Pattern Rule we obtain the following mesh
+pattern
 \begin{equation*}
@@
 \end{equation*}
-The highlighted points are an occurrence
-of the mesh pattern \(m_2\). This gives rise to the following
-rule:
+The highlighted points are an occurrence of the mesh pattern \(m_2\). This gives
+rise to the following rule:
 
@@
 \begin{proof}
-    \NoteRef{not:downcmesh} implies \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.
-    Now consider a permutation \(\omega \in \av{\pi}\) that contains an
-    occurrence of \(m_1\). If the region corresponding to the box \(\boks{a,b}\)
-    is empty then we have an occurrence of \(m_2\).
-    By condition~\ref{prop:dom3:condocc} of the proposition there exists
-    a pattern \(r\) in \(m_1^\boksall{a,b}\) such that \(m_2\) is a subpattern in some
-    pattern \(r'\) coincident with \(r\). If the region is non-empty then there exists
-    an occurrence of \(r\) in \(\omega\). Therefore there is an occurrence of \(r'\)
-    in \(\omega\), which implies that there is an occurrence of \(m_2\) in \(\omega\).
-    Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the
-    two patterns are coincident.
+    \NoteRef{not:downcmesh} implies \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\)
+as before. Now consider a permutation \(\omega \in \av{\pi}\) that contains an
+occurrence of \(m_1\). If the region corresponding to the box \(\boks{a,b}\) is
+empty then we have an occurrence of \(m_2\). By
+condition~\ref{prop:dom3:condocc} of the proposition there exists a pattern
+\(r\) in \(m_1^\boksall{a,b}\) such that \(m_2\) is a subpattern in some pattern
+\(r'\) coincident with \(r\). If the region is non-empty then there exists an
+occurrence of \(r\) in \(\omega\). Therefore, there is an occurrence of \(r'\) in
+\(\omega\), which implies that there is an occurrence of \(m_2\) in \(\omega\).
+Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the two patterns are
+coincident.
 \end{proof}
@@
 After implementing the Third Dominating Pattern Rule we find that there are
-\(56\) \(\coinc[(3)]{231}\)-equivalence classes, of mesh patterns where the underlying
-classical pattern is \(12\). By \NoteRef{note:main} there are
+\(56\) \(\coinc[(3)]{231}\)-equivalence classes, of mesh patterns where the
+underlying classical pattern is \(12\). By \NoteRef{note:main} there are
 therefore exactly \(56\) \(\coinc{231}\)-equivalence classes, since we have
-observed \(56\) \(\coinc[comp]{231}\)-equivalence classes through experimentation.
+observed \(56\) \(\coinc[comp]{231}\)-equivalence classes through
+experimentation.
 
 \subsection{Coincidence classes of Av(\{321, (12, \textit{R})\}).}
-When considering coincidences of mesh patterns with underlying classical
-pattern \(\perm{1,2}\) in \(\av{\perm{3,2,1}}\) application of the previously
+When considering coincidences of mesh patterns with underlying classical pattern
+\(\perm{1,2}\) in \(\av{\perm{3,2,1}}\) application of the previously
 established rules gives no coincidences. Through experimentation we discover
-that there are \(7\) \(\coinc[comp]{321}\)-equivalence classes are unexplained.
-Since the number of coincidences is
-so small we will reason for these coincidences without attempting to generalise
-into concrete rules.
+that there are \(7\) \(\coinc[comp]{321}\)-equivalence classes that are
+unexplained. Since the number of coincidences is so small we will reason for
+these coincidences without attempting to generalise into concrete rules.
 
-Intuitively, it is easy to see why our previous rules have no
-power here. It is impossible to add a single point to a mesh pattern \((12, R)\) and
-create an occurrence of \(\pi = \perm{3,2,1}\). It is also impossible to have
-a position where addition of an ascent, or descent, provides extra
-shading power.
+Intuitively, it is easy to see why our previous rules have no power here. It is
+impossible to add a single point to a mesh pattern \((12, R)\) and create an
+occurrence of \(\pi = \perm{3,2,1}\). It is also impossible to have a position
+where addition of an ascent, or descent, provides extra shading power.
 
@@
 \begin{equation*}
-    m_1 = \pattern{}{1,2}{0/0, 0/2, 1/1, 1/2, 2/1} \text { and } m_2 \pattern{}{1,2}{0/0, 0/1, 0/2, 1/1, 1/2, 2/1}
+    m_1 = \pattern{}{1,2}{0/0, 0/2, 1/1, 1/2, 2/1} \et m_2 \pattern{}{1,2}{0/0,
+    0/1, 0/2, 1/1, 1/2, 2/1}
 \end{equation*}
@@
 
-Consider the region corresponding to the box \(\boks{0,1}\) in any occurrence of \(m_1\)
-in a permutation. By \LemmaRef{lem:incdec} it must contain an increasing subsequence.
-If the region is empty then we have an occurrence of \(m_2\).
-If there is only one point in the region we can choose this point to replace the \(1\) in
-the mesh pattern to get the required shading. If there is more than one point then
-choosing the two leftmost points gives us the following mesh pattern.
+Consider the region corresponding to the box \(\boks{0,1}\) in any occurrence of
+\(m_1\) in a permutation. By \LemmaRef{lem:incdec} it must contain an increasing
+subsequence. If the region is empty then we have an occurrence of \(m_2\). If
+there is only one point in the region we can choose this point to replace the
+\(1\) in the mesh pattern to get the required shading. If there is more than one
+point then choosing the two leftmost points gives us the following mesh pattern.
 \begin{equation*}
@@
 \end{equation*}
-Here the two highlighted points are the original two points.
-The other two points are an occurrence of the pattern
-we \(m_2\), and hence the two patterns are coincident.
-It is also possible to calculate this coincidence by an extension of the Third
-Dominating rule, where we allow a sequence of point addition operations, this
-is discussed further in the future work section.
-Because of symmetries we have dealt with \(4\) out of the \(7\) \(\coinc[comp]{321}\)-equivalence classes.
+Here the two highlighted points are the original two points. The other two
+points are an occurrence of the pattern we \(m_2\), and hence the two patterns
+are coincident. It is also possible to calculate this coincidence by an
+extension of the Third Dominating rule, where we allow a sequence of point
+addition operations, this is discussed further in the future work section.
+Because of symmetries we have dealt with \(4\) out of the \(7\)
+\(\coinc[comp]{321}\)-equivalence classes.
 
@@
     \end{equation*}
-        Now we need to consider the box labelled \(X\). If this box is empty then
-        we have an occurrence of \(m_2\) and are done. If this box contains
-        any points then we gain some extra shading on the mesh pattern
-        as any points in the boxes \(\boks{1,2}\) and \(\boks{1,3}\) would create an occurrence of the dominating pattern \(\perm{3,2,1}\)
+        Now we need to consider the box labelled \(X\). If this box is empty
+        then we have an occurrence of \(m_2\) and are done. If this box contains
+        any points then we gain some extra shading on the mesh pattern as any
+        points in the boxes \(\boks{1,2}\) and \(\boks{1,3}\) would create an
+        occurrence of the dominating pattern \(\perm{3,2,1}\)
      \begin{equation*}
@@
 \(m_1\) so the two patterns are coincident.
-This argument applies to another two pairs of classes. Therefore in
-total in \(\av{\{\perm{3,2,1}, \mperm{1,2}{R}\}}\) there are \(\coinc{321}\)-equivalence
-classes.
+This argument applies to another two pairs of classes. Therefore we have
+explained all 7 of the coincidences in \(\av{\{\perm{3,2,1}, \mperm{1,2}{R}\}}\)
+and there are 213 \(\coinc{321}\)-equivalence classes.

### 2017-10-14 writing/paper/sections/conclfut.tex
Final edits from Murray
Fix minor speling mistakes "patter"->"pattern"
"leftmostpoint"->"leftmost point"

Reflow text to line length 80

@@
 
-It is interesting to consider the application of the Third Dominating rule, as well as the simple
-extension of allowing a sequence of point insertions, to mesh patterns without any dominating
-pattern in order to try to capture some of the coincidences described in
+It is interesting to consider the application of the Third Dominating rule, as
+well as the simple extension of allowing a sequence of point insertions, to mesh
+patterns without any dominating pattern in order to try to capture some of the
+coincidences described in
 \textcite{DBLP:journals/combinatorics/HilmarssonJSVU15} and
@@
    \end{equation*}
-	does not follow from the general methods presented by \textcite{DBLP:journals/corr/ClaessonTU14}, but is rather handled there as
-	a special case. We can do it as follows:
+	does not follow from the general methods presented by
+	\textcite{DBLP:journals/corr/ClaessonTU14}, but is rather handled there as a
+	special case. We can do it as follows:
    Consider a permutation containing \(m_1\),
@@
 
-It would be interesting to consider a systematic explanation of Wilf-equivalences among classes
-where \(\perm{3,2,1}\) is the dominating pattern, possibly using the  construction presented in
-\textcite[Sec.~11]{2015arXiv151203226B}, in order to directly reach enumeration and hopefully establish
-some of the non-trivial Wilf-equivalences between classes with different dominating patterns.
+It would be interesting to consider a systematic explanation of
+Wilf-equivalences among classes where \(\perm{3,2,1}\) is the dominating
+pattern, possibly using the  construction presented in
+\textcite[Sec.~11]{2015arXiv151203226B}, in order to directly reach enumeration
+and hopefully establish some of the non-trivial Wilf-equivalences between
+classes with different dominating patterns.

### 2017-10-14 writing/paper/sections/introduction.tex
Final edits from Murray
Fix minor speling mistakes "patter"->"pattern"
"leftmostpoint"->"leftmost point"

Reflow text to line length 80

@@
 stack sorting in \emph{The Art of Computer Programming}~\cite[p.~243,
-Ex.~5,6]{Knuth:1997:ACP:260999}. The original concept---a subsequence of
-symbols having a particular relative order, now known as classical
-patterns---has been expanded to a variety of definitions.
-\textcite{babstein2000} considered \emph{vincular} patterns (also known as
-\emph{generalised} or \emph{dashed} patterns) where two adjacent entries in the
-pattern can be required to be adjacent in the permutation. \textcite{MR2652101}
-look at classes of patterns where entries can also be required to be consecutive in
-value, these are called \emph{bivincular} patterns. \emph{Bruhat-restricted}
-patterns were studied by \textcite{MR2264071} to establish necessary conditions
-for a Schubert variety to be Gorenstein. These definitions are subsumed under
-the definition of \emph{mesh patterns}, introduced by
-\textcite{journals/combinatorics/BrandenC11} to capture explicit expansions for
-certain permutation statistics.
+Ex.~5,6]{Knuth:1997:ACP:260999}. The original concept---a subsequence of symbols
+having a particular relative order, now known as classical patterns---has been
+expanded to a variety of definitions. \textcite{babstein2000} considered
+\emph{vincular} patterns (also known as \emph{generalised} or \emph{dashed}
+patterns) where two adjacent entries in the pattern can be required to be
+adjacent in the permutation. \textcite{MR2652101} look at classes of patterns
+where entries can also be required to be consecutive in value, these are called
+\emph{bivincular} patterns. \emph{Bruhat-restricted} patterns were studied by
+\textcite{MR2264071} to establish necessary conditions for a Schubert variety to
+be Gorenstein. These definitions are subsumed under the definition of \emph{mesh
+patterns}, introduced by \textcite{journals/combinatorics/BrandenC11} to capture
+explicit expansions for certain permutation statistics.
 
@@
 avoiding \(\sigma\). Avoiding pairs of patterns of the same length with certain
-properties has been studied, \textcite{MR2178749} considered avoiding
-a pair of vincular patterns of length 3. \textcite{2015arXiv151203226B} study
-avoiding a vincular and a covincular pattern simultaneously in order to achieve
-several counting results. However, little work has been done on
-avoiding a mesh pattern and a classical pattern simultaneously.
+properties has been studied, \textcite{MR2178749} considered avoiding a pair of
+vincular patterns of length 3. \textcite{2015arXiv151203226B} study avoiding a
+vincular and a covincular pattern simultaneously in order to achieve several
+counting results. However, little work has been done on avoiding a mesh pattern
+and a classical pattern simultaneously.
 
@@
 begin by establishing coincidences between mesh patterns of length 2 while
-avoiding a dominating pattern by computational methods, which are then
-used to establish three ``Dominating Pattern Rules'' as well as some special cases
-that can be used to calculate coincidences.
+avoiding a dominating pattern by computational methods, which are then used to
+establish three ``Dominating Pattern Rules'' as well as some special cases that
+can be used to calculate coincidences.
 
- We then use these coincidence classes
-to calculate Wilf-equivalence classes showing some of the methods used.
+ We then use these coincidence classes to calculate Wilf-equivalence classes
+ showing some of the methods used.

### 2017-10-14 writing/paper/sections/meshpatterns.tex
Final edits from Murray
Fix minor speling mistakes "patter"->"pattern"
"leftmostpoint"->"leftmost point"

Reflow text to line length 80

@@
-A \emph{permutation} is a bijection from the set \(\nrange{n} = \setrange{1}{n}\) to itself. The
-set of all such bijections is denoted \(\mathfrak{S}_n\) and has \(n!\) elements. We can
-denote an individual permutation \(\pi\in\mathfrak{S}_n\) in \emph{one-line notation} by writing the
-entries of the permutation in order, therefore \(\pi = \perm{\pi(1),\pi(2),\dotsm,\pi(n)}\). The
-set \(\mathfrak{S}_0\) has exactly one element, the empty permutation \(\varepsilon\).
+A \emph{permutation} is a bijection from the set \(\nrange{n} =
+\setrange{1}{n}\) to itself. The set of all such bijections is denoted
+\(\mathfrak{S}_n\) and has \(n!\) elements. We can denote an individual
+permutation \(\pi\in\mathfrak{S}_n\) in \emph{one-line notation} by writing the
+entries of the permutation in order, therefore \(\pi =
+\perm{\pi(1),\pi(2),\dotsm,\pi(n)}\). The set \(\mathfrak{S}_0\) has exactly one
+element, the empty permutation \(\varepsilon\).
 
@@
 \begin{definition}
-    A permutation \(\pi\in\mathfrak{S}_n\) \emph{contains} the permutation \(\sigma\in\mathfrak{S}_k\) (\mbox{denoted}
-    \(\sigma \preceq \pi\)) if there is some sequence \( i_1,i_2,\dotsc,i_k\)
-    such that \(1\le i_1<i_2<\dotsm<i_k\le n\) and the sequence \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is order isomorphic to
+    A permutation \(\pi\in\mathfrak{S}_n\) \emph{contains} the permutation
+    \(\sigma\in\mathfrak{S}_k\) (\mbox{denoted} \(\sigma \preceq \pi\)) if there
+    is some sequence \( i_1,i_2,\dotsc,i_k\) such that \(1\le
+    i_1<i_2<\dotsm<i_k\le n\) and the sequence
+    \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is order isomorphic to
     \(\sigma(1)\sigma(2)\dotsm\sigma(k)\). If this is the case the sequence
-    \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is called an \emph{occurrence} of \(\sigma\)
-    in \(\pi\). If \(\pi\) does not contain \(\sigma\), we say that
-    \(\pi\) \emph{avoids} \(\sigma\). In this context \(\sigma\) is called a (\emph{classical})
-    \emph{permutation pattern}.
+    \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is called an \emph{occurrence} of
+    \(\sigma\) in \(\pi\). If \(\pi\) does not contain \(\sigma\), we say that
+    \(\pi\) \emph{avoids} \(\sigma\). In this context \(\sigma\) is called a
+    (\emph{classical}) \emph{permutation pattern}.
 \end{definition}
@@
 \label{ex:contexmpl}
-The permutation \(\pi = \perm{2,4,1,5,3}\) contains the pattern
-\(\sigma = \perm{2,3,1}\), since the second, fourth and fifth elements (\(453\)) are order
-isomorphic to \(\perm{231}\). The permutation also contains the occurrence \(241\) of the same pattern. The permutation
-\(\perm{2,4,1,5,3}\) avoids the pattern \(\perm{3,2,1}\).
+The permutation \(\pi = \perm{2,4,1,5,3}\) contains the pattern \(\sigma =
+\perm{2,3,1}\), since the second, fourth and fifth elements (\(453\)) are order
+isomorphic to \(\perm{231}\). The permutation also contains the occurrence
+\(241\) of the same pattern. The permutation \(\perm{2,4,1,5,3}\) avoids the
+pattern \(\perm{3,2,1}\).
 \end{example}
 
-We denote the set of permutations of length \(n\) avoiding a pattern \(\sigma\) as \(\Av_n(\sigma)\)
-and let \(\av{\sigma}=\bigcup_{i=0}^{\infty}\Av_i(\sigma)\).
+We denote the set of permutations of length \(n\) avoiding a pattern \(\sigma\)
+as \(\Av_n(\sigma)\) and let
+\(\av{\sigma}=\bigcup_{i=0}^{\infty}\Av_i(\sigma)\).
 
-We can display a permutation graphically in a \emph{plot}, where we display the points
-\(G(\pi) = \{(i,\pi(i))\mid i \in \nrange{n} \}\) in a Cartesian coordinate system. The plots of the permutations \(\pi =
-\perm{2,4,1,5,3}\) and \(\sigma = \perm{2,3,1}\) can be seen in \FigureRef{fig:plots}.
-\FigureRef{fig:containment} shows the containment of \(\sigma\) in \(\pi\) as in \ExampleRef{ex:contexmpl}.
+We can display a permutation graphically in a \emph{plot}, where we display the
+points \(G(\pi) = \{(i,\pi(i))\mid i \in \nrange{n} \}\) in a Cartesian
+coordinate system. The plots of the permutations \(\pi = \perm{2,4,1,5,3}\) and
+\(\sigma = \perm{2,3,1}\) can be seen in \FigureRef{fig:plots}.
+\FigureRef{fig:containment} shows the containment of \(\sigma\) in \(\pi\) as in
+\ExampleRef{ex:contexmpl}.
 
-The boxes in the plot of a permutation are denoted by \(\boks{i,j}\), where the point \((i,j)\) is
-the lower left corner of the box.
+The boxes in the plot of a permutation are denoted by \(\boks{i,j}\), where the
+point \((i,j)\) is the lower left corner of the box.
 
@@
 
-Formally defined by \textcite{journals/combinatorics/BrandenC11},
-an \emph{occurrence} of \(p\) in \(\pi\) is a subset \(\omega\) of the plot of
-\(\pi, G(\pi) = \{(i,\pi(i)\mid i\in\nrange{n}\}\) such that there are order-preserving
-injections \(\alpha,\beta:\nrange{k}\mapsto\nrange{n}\) satisfying the following two conditions.
+Formally defined by \textcite{journals/combinatorics/BrandenC11}, an
+\emph{occurrence} of \(p\) in \(\pi\) is a subset \(\omega\) of the plot of
+\(\pi, G(\pi) = \{(i,\pi(i)\mid i\in\nrange{n}\}\) such that there are
+order-preserving injections \(\alpha,\beta:\nrange{k}\mapsto\nrange{n}\)
+satisfying the following two conditions.
 Firstly, \(\omega\) is an occurrence of \(\tau\) in the classical sense
@@
 
-We denote the avoidance sets for mesh patterns in the same way as for
-classical patterns. Given a mesh pattern \(p=(\sigma,R)\) we say that \(\sigma\) is the \emph{underlying
-classical pattern} of \(p\).
+We denote the avoidance sets for mesh patterns in the same way as for classical
+patterns. Given a mesh pattern \(p=(\sigma,R)\) we say that \(\sigma\) is the
+\emph{underlying classical pattern} of \(p\).
 
-We define containment of a mesh pattern \(p\) in another mesh pattern \(q\) as above,
-with the additional condition that if \(\boks{i,j}\in R\text{ then } R_{ij}\) is
-contained in the mesh set of \(q\). More formally:
+We define containment of a mesh pattern \(p\) in another mesh pattern \(q\) as
+above, with the additional condition that if \(\boks{i,j}\in R\text{ then }
+R_{ij}\) is contained in the mesh set of \(q\). More formally:
 
 \begin{definition}
-A mesh pattern \(q=(\kappa,T)\) \emph{contains} a mesh pattern \(p=(\tau,R)\) as a \emph{subpattern}
-if \(\kappa\) contains \(p\) and \(\left(\bigcup_{\boks{i,j}\in R} R_{ij}\right)\subseteq T\).
+A mesh pattern \(q=(\kappa,T)\) \emph{contains} a mesh pattern \(p=(\tau,R)\) as
+a \emph{subpattern} if \(\kappa\) contains \(p\) and
+\(\left(\bigcup_{\boks{i,j}\in R} R_{ij}\right)\subseteq T\).
 \end{definition}

### 2017-10-14 writing/paper/sections/wilf.tex
Final edits from Murray
Fix minor speling mistakes "patter"->"pattern"
"leftmostpoint"->"leftmost point"

Reflow text to line length 80

@@
 
-When examining Wilf-equivalences we can use a number of symmetries to reduce
-the amount of work required. It can be seen that the reverse, complement
-and inverse operations (see \FigureRef{fig:symm}) preserve enumeration, and
-therefore classes related by these symmetries are trivially
-Wilf-equivalent.
+When examining Wilf-equivalences we can use a number of symmetries to reduce the
+amount of work required. It can be seen that the reverse, complement and inverse
+operations (see \FigureRef{fig:symm}) preserve enumeration, and therefore
+classes related by these symmetries are trivially Wilf-equivalent.
 \begin{figure}[!htb]
@@
 In the remainder of the paper we will consider Wilf-equivalences of patterns
-whilst avoiding the \emph{dominating pattern} \(\perm{2,3,1}\), if two patterns \(p_1 \et p_2\) are Wilf-equivalent we will denote this \(p_1\weqv{231}p_2\), similarly when computational methods indicate that two patterns are Wilf equivalent under \(231\) we write \(p_1\weqv[comp]{231}p_2\).
-We will
-use \(\mathcal{C}\) to denote \(\av{\perm{2,3,1}}\) and \(C(x)\) will
-be the usual Catalan generating function satisfying \(C(x) = 1 + xC(x)^2\).
-The fact that \(C(x)\) is the generating function for \(\mathcal{C}\)
-can be seen by structural decomposition around the maximum, as
+whilst avoiding the \emph{dominating pattern} \(\perm{2,3,1}\), if two patterns
+\(p_1 \et p_2\) are Wilf-equivalent we will denote this \(p_1\weqv{231}p_2\),
+similarly when computational methods indicate that two patterns are Wilf
+equivalent under \(231\) we write \(p_1\weqv[comp]{231}p_2\). We will use
+\(\mathcal{C}\) to denote \(\av{\perm{2,3,1}}\) and \(C(x)\) will be the usual
+Catalan generating function satisfying \(C(x) = 1 + xC(x)^2\). The fact that
+\(C(x)\) is the generating function for \(\mathcal{C}\) can be seen by
+structural decomposition around the maximum, as
 shown in \FigureRef{fig:decompmax}.
@@
 We can also decompose a permutation avoiding \(\perm{2,3,1}\) around the
-leftmostpoint, giving a similar figure.
+leftmost point, giving a similar figure.
 
@@
 \label{wilf1}
-When considering the mesh patterns of length \(2\) it will be useful to know
-the Wilf-equivalence classes of the mesh patterns of length \(1\) inside
-\(\av{\perm{2,3,1}}\), this means that we are considering the \(\weqv{231}\)-equivalence classes of mesh patterns with underlying classical pattern \(1\).
+When considering the mesh patterns of length \(2\) it will be useful to know the
+Wilf-equivalence classes of the mesh patterns of length \(1\) inside
+\(\av{\perm{2,3,1}}\), this means that we are considering the
+\(\weqv{231}\)-equivalence classes of mesh patterns with underlying classical
+pattern \(1\).
 
@@
 \textpattern{}{1}{0/0,1/1} and \(q_1 = \textpattern{}{1}{0/0,0/1,1/1}\) are
-coincident under \(\perm{2,3,1}\). Consider the decomposition of a non-empty avoider of \(q_1\) in
-\(\av{\perm{2,3,1}}\) around the maximum:
+coincident under \(\perm{2,3,1}\). Consider the decomposition of a non-empty
+avoider of \(q_1\) in \(\av{\perm{2,3,1}}\) around the maximum:
 \begin{equation*}
@@
 \end{equation*}
-The pattern \(q_3 = \textpattern{}{1}{0/0,0/1,1/0}\) is the reverse-complement-inverse
-of \(q_2\) and hence the avoiders of \(q_2\) and \(q_3\) (\(\scriptvar_2\) and \(\scriptvar_3\)) are equinumerous, and so \(q_1 \weqv{231} q_2 \weqv{231}q_3\).
-All of these classes have the same generating function, namely
+The pattern \(q_3 = \textpattern{}{1}{0/0,0/1,1/0}\) is the
+reverse-complement-inverse of \(q_2\) and hence the avoiders of \(q_2\) and
+\(q_3\) (\(\scriptvar_2\) and \(\scriptvar_3\)) are equinumerous, and so \(q_1
+\weqv{231} q_2 \weqv{231}q_3\). All of these classes have the same generating
+function, namely
 \begin{equation}
@@
 
-The only symmetry that we are able to consider is \emph{reverse-complement-inverse} as this is the only symmetry that preserves the
-\(\perm{2,3,1}\) pattern. Using this symmetry we can find 61 classes of trivial
-Wilf-equivalence, these equivalences being explained by either the patterns being
-coincident in \(\av{\perm{2,3,1}}\), or by one pattern being the reverse-complement-inverse of some other pattern.
+The only symmetry that we are able to consider is
+\emph{reverse-complement-inverse} as this is the only symmetry that preserves
+the \(\perm{2,3,1}\) pattern. Using this symmetry we can find 61 classes of
+trivial Wilf-equivalence, these equivalences being explained by either the
+patterns being coincident in \(\av{\perm{2,3,1}}\), or by one pattern being the
+reverse-complement-inverse of some other pattern.
 
-Computing avoiders up to length 10 gives 23 \(\weqv[comp]{231}\)-equivalence classes,
-of which 13 are not equal to coincidence classes (after taking symmetries). Therefore there seem to be Wilf-equivalences that are not
-explained by coincidences or symmetry.
+Computing avoiders up to length 10 gives 23 \(\weqv[comp]{231}\)-equivalence
+classes, of which 13 are not equal to coincidence classes (after taking
+symmetries). Therefore there seem to be Wilf-equivalences that are not explained
+by coincidences or symmetry.
 
-We will use two main methods of establishing these Wilf-equivalences: the structural decomposition of avoiders, via generating functions; or the
-structure of a general permutation containing the pattern, looking at a particular occurrence of the
-pattern in a permutation avoiding \(\perm{2,3,1}\). Sometimes it will be necessary to use both of
-these methods to consolidate a single Wilf-class.
+We will use two main methods of establishing these Wilf-equivalences: the
+structural decomposition of avoiders, via generating functions; or the structure
+of a general permutation containing the pattern, looking at a particular
+occurrence of the pattern in a permutation avoiding \(\perm{2,3,1}\). Sometimes
+it will be necessary to use both of these methods to consolidate a single
+Wilf-class.
 
@@
 
-If there is an occurrence of \(m_1\) there must be a lowest occurrence of \(m_1\), \ie, the occurrence has the lowest possible values for any occurrence. Now consider the top
-regions, labelled \(A \et B\), the subpermutation contained in the union of these regions must avoid the permutation \(231\).
-Also, the subpermutation in the region \(A\) must be a decreasing subsequence, otherwise an occurrence of \(231\) will be created with either of the points in the occurrence of \(m_1\).
-Now, consider the region labelled \(C\), since we specified that we were focused on the lowest possible occurrence of \(m_1\) this region cannot contain an occurrence of either \(m_1\) or \(231\).
-This is a full structural decomposition of a container of \(m_1\) inside \(\av{231}\).
+If there is an occurrence of \(m_1\) there must be a lowest occurrence of
+\(m_1\), \ie, the occurrence has the lowest possible values for any occurrence.
+Now consider the top regions, labelled \(A \et B\), the subpermutation contained
+in the union of these regions must avoid the permutation \(231\). Also, the
+subpermutation in the region \(A\) must be a decreasing subsequence, otherwise
+an occurrence of \(231\) will be created with either of the points in the
+occurrence of \(m_1\). Now, consider the region labelled \(C\), since we
+specified that we were focused on the lowest possible occurrence of \(m_1\) this
+region cannot contain an occurrence of either \(m_1\) or \(231\). This is a full
+structural decomposition of a container of \(m_1\) inside \(\av{231}\).
 
@@
 
-Similarly to a container of \(m_1\) we will look at the lowest possible occurrence of \(m_2\), and as before the regions \(A^\prime \et B^\prime\) together contain an avoider of \(231\) and \(A\) contains a decreasing subsequence.
-Since we are considering the lowest occurrence of \(m_2\) the region \(C^\prime\) does not contain an occurrence of either \(m_2\) or \(231\).
+Similarly to a container of \(m_1\) we will look at the lowest possible
+occurrence of \(m_2\), and as before the regions \(A^\prime \et B^\prime\)
+together contain an avoider of \(231\) and \(A\) contains a decreasing
+subsequence. Since we are considering the lowest occurrence of \(m_2\) the
+region \(C^\prime\) does not contain an occurrence of either \(m_2\) or \(231\).
 
-Since all the regions in both of these cases contain the same parts,  the classes defined by containment of \(m_1\) and \(m_2\) inside \(\av{\perm{2,3,1}}\) are
-equinumerous and therefore so are their avoiders.
+Since all the regions in both of these cases contain the same parts,  the
+classes defined by containment of \(m_1\) and \(m_2\) inside
+\(\av{\perm{2,3,1}}\) are equinumerous and therefore so are their avoiders.
 
-Consider the class, \(\scriptvar_3\), of permutations defined by avoiding \(\perm{2,3,1}\) and \(m_3\). We can decompose a member of this class around the maximum
+Consider the class, \(\scriptvar_3\), of permutations defined by avoiding
+\(\perm{2,3,1}\) and \(m_3\). We can decompose a member of this class around the
+maximum
 \begin{equation*}
@@
 \end{equation*}
-Only the first point in the top right region can
-create an occurrence of \(m_3\) if and only if it is the
-element with largest value in this region, therefore the partial permutation
-in this region must avoid starting with the maximum, \ie, be in \(\mathcal{\varmaxl}_1 = \av{\textpattern{}{1}{0/0,0/1,1/1}}\) described
+Only the first point in the top right region can create an occurrence of \(m_3\)
+if and only if it is the element with largest value in this region, therefore
+the partial permutation in this region must avoid starting with the maximum,
+\ie, be in \(\mathcal{\varmaxl}_1 = \av{\textpattern{}{1}{0/0,0/1,1/1}}\)
+described
 in \SectionRef{wilf1}.
@@
 \end{equation*}
-It is therefore clear that avoiders of \(m_1\) and avoiders of \(m_4\)
-have the same enumeration, and therefore \(m_1 \weqv{231} m_2 \weqv{231}m_3\weqv{m_4}\) with generating function satisfying
+It is therefore clear that avoiders of \(m_1\) and avoiders of \(m_4\) have the
+same enumeration, and therefore \(m_1 \weqv{231} m_2 \weqv{231}m_3\weqv{m_4}\)
+with generating function satisfying
 \begin{equation*}
@@
 \end{equation*}
-Let \(\scriptvar_1\) be the set of avoiders of \(m_1\) in \(\av{231}\). By structural decomposition
-around the leftmost point we have
+Let \(\scriptvar_1\) be the set of avoiders of \(m_1\) in \(\av{231}\). By
+structural decomposition around the leftmost point we have
 \begin{equation*}
@@
 \noindent
-Here \(\scriptvar^\prime_1\) is a permutation avoiding \(\perm{2,3,1}, m_1\)
-and \(\textpattern{}{1}{0/0,0/1,1/0}\).
-Now consider the decomposition of a permutation in \(\scriptvar^\prime_1\). It can
-once again be decomposed around the leftmost point
+Here \(\scriptvar^\prime_1\) is a permutation avoiding \(\perm{2,3,1}, m_1\) and
+\(\textpattern{}{1}{0/0,0/1,1/0}\). Now consider the decomposition of a
+permutation in \(\scriptvar^\prime_1\). It can once again be decomposed around
+the leftmost point
 \begin{equation*}
@@
 \end{equation*}
-This is a complete decomposition of avoiders of \(m_1\). Now we
-look at an avoider of \(m_2\), decomposed around the leftmost point
+This is a complete decomposition of avoiders of \(m_1\). Now we look at an
+avoider of \(m_2\), decomposed around the leftmost point
 \begin{equation*}
@@
 Where \(\scriptvar^\prime_2\) is a permutation avoiding \(\perm{2,3,1}, m_2\)
-and \(\textpattern{}{1}{0/0,0/1,1/1}\).
-Again we use the same method of decomposition of a permutation in \(\scriptvar^\prime_2\)
+and \(\textpattern{}{1}{0/0,0/1,1/1}\). Again we use the same method of
+decomposition of a permutation in \(\scriptvar^\prime_2\)
 \begin{equation*}
@@
 
-In order to establish the remainder of the Wilf-equivalences of the form \(\av{\{\perm{2,3,1},p\}}\) where \(p\)
-is a mesh pattern of length \(2\) we can use similar methods to allow us to consolidate experimental classes into
-actual classes, these methods allow us to explain all 23 of the observed Wilf-classes seen in experimentation.
+In order to establish the remainder of the Wilf-equivalences of the form
+\(\av{\{\perm{2,3,1},p\}}\) where \(p\) is a mesh pattern of length \(2\) we can
+use similar methods to allow us to consolidate experimental classes into actual
+classes, these methods allow us to explain all 23 of the observed Wilf-classes
+seen in experimentation.

### 2017-12-10 writing/paper/sections/coincidences.tex
Reviewer 1 comments

@@
 \(\av{\pi,p_1}\) and \(\av{\pi,p_2}\) are equal. If this is the case we say
-the two mesh patterns are \emph{coincidenc under} \(\pi\).
+the two mesh patterns are \emph{coincident under} \(\pi\).
 
@@
 \(\boks{i,j}\notin R\). We define a mesh pattern \(p^{\boks{i,j}_a} =
-(\tau^\prime,R^\prime)\) (\(p^{\boks{i,j}_d}\)) of length \(n+2\) as the pattern
-where an ascent (descent) is \emph{inserted} into the box \(\boks{i,j}\) in
+(\tau^\prime,R^\prime)\) (respectively, \(p^{\boks{i,j}_d}\)) of length \(n+2\) as the pattern
+where an ascent (respectively descent) is \emph{inserted} into the box \(\boks{i,j}\) in
 \(G(p)\). Formally the new underlying classical pattern is defined by
@@
 The ordering of the top branch determines whether an ascent(or descent) is added.
+
 The mesh becomes
@@
 \end{equation*}
+meaning that the ascent (or descent) is not allowed to be split in the box.
 \end{definition}
@@
 \coinc[(2)]{231}\)-equivalence classes, considering mesh patterns with \(12\) as
-the underlying classical patterns. However there are fewer
-\(\coinc[comp]{231}\)-equivalence classes.
+the underlying classical patterns. However there are
+\(56 \coinc[comp]{231}\)-equivalence classes.
 

### 2017-12-10 writing/paper/sections/wilf.tex
Reviewer 1 comments

@@
 \end{equation*}
-This can be explained succinctly by the fact that a permutation containing
-\(q_1\) starts with its maximum, and by not allowing the lower-left section of the
+This can be explained succinctly by the fact that a permutation contains
+\(q_1\) if and only if it starts with its maximum, and by not allowing the
+lower-left section of the
 \(\perm{2,3,1}\) avoider to be empty we prevent an occurrence from ever
@@
 occurrence of \(m_2\), and as before the regions \(A^\prime \et B^\prime\)
-together contain an avoider of \(231\) and \(A\) contains a decreasing
+together contain an avoider of \(231\) and \(A^\prime\) contains a decreasing
 subsequence. Since we are considering the lowest occurrence of \(m_2\) the
@@
 \end{equation*}
-An occurrence of \(m_4\) can never occur in the
-top right region. It could only occur between the maximum and the first point
+An occurrence of \(m_4\) can never occur entirely within the
+top right region, since the maximum is to the left and above any point in this region. It could only occur between the maximum and the first point
 in the region, if and only if this first point is the lowest valued element in
 this region, so this top right region must contain a sub-permutation that does not
-start with it's minimum, i.e., it is a member of \(\mathcal{\varmaxl}_3 = \av{\av{\textpattern{}{1}{0/0,0/1,1/0}}}\).
+start with it's minimum, i.e., it is a member of \(\mathcal{\varmaxl}_3 = \av{\textpattern{}{1}{0/0,0/1,1/0}}\).
 Since both \(\mathcal{\varmaxl}_1\) and \(\mathcal{\varmaxl}_3\) have the same

### 2017-12-20 writing/paper/sections/wilf.tex
Final fix for DMTCS

@@
 
-If there is an occurrence of \(m_1\) there must be a lowest occurrence of
-\(m_1\), \ie, the occurrence has the lowest possible values for any occurrence.
+If there is an occurrence of \(m_1\) there must be an occurrence of \(m_1\) that minimizes
+the values of the points \(xy\) in the occurrence, first with respect to \(x\) then with respect to \(y\).
 Now consider the top regions, labelled \(A \et B\), the subpermutation contained
